### 2021-03-08 doc/paper3/paper.tex
First draft ready

\documentclass[runningheads]{llncs}

\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\usepackage{amssymb}
\usepackage{mathtools}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage[capitalise]{cleveref}
\usepackage{tikz}
\usepackage[binary-units]{siunitx}
\usepackage{booktabs}
\usepackage{capt-of}

\crefname{enumi}{Condition}{Conditions}
\crefname{enumii}{Condition}{Conditions}

\begin{document}

%\title{Contribution Title\thanks{Supported by organization x.}}
\title{Weighted Model Counting Without Parameter Variables}

%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here

%\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
%Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
%Third Author\inst{3}\orcidID{2222--3333-4444-5555}}

%\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.

%\institute{Princeton University, Princeton NJ 08544, USA \and
%Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
%\email{lncs@springer.com}\\
%\url{http://www.springer.com/gp/computer-science/lncs} \and
%ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
%\email{\{abc,lncs\}@uni-heidelberg.de}}

\maketitle              % typeset the header of the contribution

% The abstract should briefly summarize the contents of the paper in
% 150--250 words.
% 1. Introduction. In one sentence, what’s the topic?
% 2. State the problem you tackle.
% 3. Summarize (in one sentence) why nobody else has adequately answered the research question yet.
% 4. Explain, in one sentence, how you tackled the research question.
% 5. In one sentence, how did you go about doing the research that follows from your big idea.
% 6. As a single sentence, what’s the key impact of your research?
\begin{abstract}
  Weighted model counting (WMC) is a powerful computational technique for a
  variety of problems, especially commonly used for probabilistic inference.
  However, the standard definition of WMC that puts weights on literals often
  necessitates WMC encodings to include additional variables and clauses just so
  each weight can be attached to a literal. This paper complements previous work
  by considering WMC instances in their full generality and using recent
  state-of-the-art WMC techniques based on pseudo-Boolean function manipulation,
  competitive with the more traditional WMC algorithms based on knowledge
  compilation and backtracking search. We present an algorithm that transforms
  WMC instances into a format based on pseudo-Boolean functions while
  eliminating around \SI{57}{\percent} of variables on average across various
  Bayesian network encodings. Moreover, we identify sufficient conditions for
  such a variable removal to be possible. Our experiments show significant
  improvement in WMC-based Bayesian network inference, outperforming the current
  state of the art.

  \keywords{Weighted model counting \and Probabilistic inference \and Bayesian
    networks.}
\end{abstract}

\section{Introduction}

% NOTE: What is the problem? Why is it interesting and important?
Weighted model counting (WMC), i.e., a generalisation of propositional model
counting that assigns weights to literals and computes the total weight of all
models of a propositional formula \cite{DBLP:journals/ai/ChaviraD08}, has
emerged as a powerful computational framework for problems in many domains,
e.g., probabilistic graphical models such as Bayesian networks and Markov
networks
\cite{DBLP:conf/ecai/BartKLM16,DBLP:conf/ijcai/ChaviraD05,DBLP:conf/sat/ChaviraD06,DBLP:conf/kr/Darwiche02,DBLP:conf/aaai/SangBK05},
neuro-symbolic artificial intelligence \cite{DBLP:conf/icml/XuZFLB18},
probabilistic programs \cite{DBLP:journals/pacmpl/HoltzenBM20}, and
probabilistic logic programs \cite{DBLP:journals/tplp/FierensBRSGTJR15}. It has
been extended to support continuous variables \cite{DBLP:conf/ijcai/BellePB15},
infinite domains \cite{DBLP:conf/aaai/Belle17}, first-order logic
\cite{DBLP:journals/cacm/GogateD16,DBLP:conf/ijcai/BroeckTMDR11}, and arbitrary
semirings \cite{DBLP:journals/ijar/BelleR20,DBLP:journals/japll/KimmigBR17}.
However, as the definition of WMC puts weights on literals, additional variables
often need to be added for the sole purpose of holding a weight
\cite{DBLP:conf/ecai/BartKLM16,DBLP:conf/ijcai/ChaviraD05,DBLP:conf/sat/ChaviraD06,DBLP:conf/kr/Darwiche02,DBLP:conf/aaai/SangBK05}.
As the parameterised complexity of model counting (and, by extension, WMC)
depends on the number of variables
\cite{DBLP:conf/focs/BacchusDP03,DBLP:journals/jda/SamerS10}, this can make WMC
unnecessarily slow and could be detrimental to WMC algorithms such as
\textsf{ADDMC} \cite{DBLP:conf/aaai/DudekPV20} that depend on variable ordering
heuristics.

% NOTE: What's wrong with previous proposed solutions? How does mine differ?
One approach to this problem considers weighted clauses and probabilistic
semantics based on Markov networks \cite{DBLP:conf/uai/GogateD10}. However, with
a new representation comes the need to invent new encodings and inference
algorithms. Our work is similar in that it introduces a new representation for
computational problems but can reuse recent WMC algorithms based on
pseudo-Boolean function manipulation, namely, \textsf{ADDMC}
\cite{DBLP:conf/aaai/DudekPV20} and \textsf{DPMC} \cite{DBLP:conf/cp/DudekPV20}.
Furthermore, we identify sufficient conditions for transforming a WMC instance
into our new format. As many WMC inference algorithms such as
\textsf{Ace}\footnote{\url{http://reasoning.cs.ucla.edu/ace/}}, \textsf{c2d}
\cite{DBLP:conf/ecai/Darwiche04}, and \textsf{miniC2D}
\cite{DBLP:conf/ijcai/OztokD15} work by compilation to tractable representations
such as arithmetic circuits, deterministic, decomposable negation normal form
\cite{DBLP:journals/jancl/Darwiche01}, and sentential decision diagrams (SDDs)
\cite{DBLP:conf/ijcai/Darwiche11}, another way to avoid parameter variables
could be via direct compilation to a more convenient representation. However,
while direct compilation of Bayesian networks to SDDs has been investigated
\cite{DBLP:conf/ecsqaru/ChoiKD13}, SDDs are not expressive enough to avoid the
issue. Furthermore, to the best of the authors' knowledge, neither
compilation/inference technique has a publicly available implementation.

% NOTE: What are the key components of my approach and results? Also include any
% specific limitations. What about your project is new or different from
% existing solutions? List of refutable contributions (with forward references
% to the rest of the paper) What is the outcome of the project?
In this work, we introduce a way to transform WMC problems into a new format
based on pseudo-Boolean functions---\emph{pseudo-Boolean projection} (PBP). We
formally show that every WMC problem instance has a corresponding PBP instance
and identify conditions under which this transformation can remove parameter
variables. Four out of the five known WMC encodings for Bayesian networks
\cite{DBLP:conf/ecai/BartKLM16,DBLP:conf/ijcai/ChaviraD05,DBLP:conf/sat/ChaviraD06,DBLP:conf/kr/Darwiche02,DBLP:conf/aaai/SangBK05}
can indeed be simplified in this manner. This transformation enables two
encodings that were previously incompatible with most WMC algorithms (due to
using a different definition of WMC
\cite{DBLP:conf/ijcai/ChaviraD05,DBLP:conf/sat/ChaviraD06}) to be run with
\textsf{ADDMC} and \textsf{DPMC} and results in a significant performance boost
for one other encoding, allowing it to surpass the current state of the art.
Finally, our theoretical contributions result in a convenient algebraic way of
reasoning about two-valued pseudo-Boolean functions and position WMC encodings
on common ground, identifying their key properties and assumptions.

% Specifically, direct compilation
% from Bayesian networks to SDDs \cite{DBLP:conf/ecsqaru/ChoiKD13} and from
% structured Bayesian networks (i.e., a generalisation of Bayesian networks) to
% probabilistic SDDs \cite{shen2020modeling} have been considered.

% * SDD compilation still has parameter variables
% 1) However, no conversion algorithms from either WMC or probabilistic models
% are provided.
% Since we present a WMC to PBP algorithm, we don't need to invent new
% encodings for each problem and we can benefit from all the good ideas that
% went into designing WMC encodings
% 3) Instead of combining techniques from older WMC algorithms like they did, we
% use a much more recent WMC solver that is shown to...
% 3) We benefit from a decade of research in WMC inference and, specifically, 
% 4) We are not restricted to clauses and probabilistic semantics
% 5) We identify conditions and describe an algorithm for transforming any WMC
% instance into PBP

\section{Weighted Model Counting}

We begin with an overview of some notation and terminology. Throughout the
paper, we use set-theoretic notation for many concepts in logic. A \emph{clause}
is a set of literals that are part of an implicit disjunction. Similarly, a
\emph{formula} in CNF is a set of clauses that are part of an implicit
conjunction. We identify a \emph{model} with a set of variables that correspond
to the positive literals in the model (and all other variables are the negative
literals of the model). We can then define the \emph{cardinality} of a model as
the cardinality of this set. For example, let $\phi = (\neg a \lor b) \land a$
be a propositional formula over variables $a$ and $b$. Then an equivalent
set-theoretic representation of $\phi$ is $\{ \{ \neg a, b \}, \{ a \} \}$. Any
subset of $\{ a, b \}$ is an interpretation of $\phi$, e.g., $\{ a, b \}$ is a
model of $\phi$ (written $\{ a, b \} \models \phi$) of cardinality two, while
$\emptyset$ is an interpretation but not a model. We can now formally define
WMC.

\begin{definition}[WMC]
  A \emph{WMC instance} is a tuple $(\phi, X_I, X_P, w)$, where $X_I$ is
  the set of \emph{indicator} variables, $X_P$ is the set of \emph{parameter}
  variables (with $X_I \cap X_P = \emptyset$), $\phi$ is a propositional formula
  in CNF over $X_I \cup X_P$, and $w\colon X_I \cup X_P \cup \{\neg x \mid x \in
  X_I \cup X_P\} \to \mathbb{R}$ is the weight function. Its \emph{answer} is
  $\sum_{Y \models \phi} \prod_{Y \models l} w(l)$.
\end{definition}

That is, the answer to a WMC instance is the sum of the weights of all models of
$\phi$, where the weight of a model is defined as the product of the weights of
all (positive and negative) literals in it. Our definition of WMC is largely
based on the standard definition \cite{DBLP:journals/ai/ChaviraD08}, but
explicitly partitions variables into indicator and parameter variables. In
practice, we identify this partition in one of two ways. If an encoding is
generated by \textsf{Ace}, then variable types are explicitly identified in the
LMAP file generated alongside the encoding. Otherwise, we define a variable $x$
to be a parameter variable if $w(x) = w(\neg x) = 1$. Next, we formally define a
variation of the WMC problem used by some of the Bayesian network encodings
\cite{DBLP:conf/ijcai/ChaviraD05,DBLP:conf/sat/ChaviraD06}.

\begin{definition}
  Let $\phi$ be a formula over a set of variables $X$. Then $Y \subseteq X$ is a
  \emph{minimum-cardinality model} of $\phi$ if $Y \models \phi$ and $|Y| \le
  |Z|$ for all $Z \models \phi$.
\end{definition}

\begin{definition}[Minimum-Cardinality WMC]
  A \emph{minimum-cardinality WMC} instance consists of the same tuple as a WMC
  instance, but its answer is defined to be $\sum_{Y \models \phi\text{, }|Y| =
    k} \prod_{Y \models l} w(l)$ (where $k = \min_{Y \models \phi} |Y|$) if
  $\phi$ is satisfiable, and zero otherwise.
\end{definition}

\subsection{Bayesian Network Encodings}

A \emph{Bayesian network} is a directed acyclic graph with random variables as
vertices and edges as conditional dependencies. As is common in related
literature
\cite{DBLP:conf/kr/Darwiche02,DBLP:conf/aaai/SangBK05}, we assume that each
variable has a finite number of values. We call a Bayesian network \emph{binary}
if every variable has two values. If all variables have finite numbers of
values, the probability function associated with each variable $v$ can be
represented as a \emph{conditional probability table} (CPT), i.e., a table
with a row for each combination of values that $v$ and its parent vertices can
take. Each row then also has a \emph{probability}, i.e., a number in $[0, 1]$.

WMC is a well-established technique for Bayesian network inference, particularly
effective on networks where most variables have only a few possible values
\cite{DBLP:conf/kr/Darwiche02}. Many ways of encoding a Bayesian network into a
WMC instance have been proposed. We will refer to them based on the initials of
the authors and the year of publication. Darwiche was the first to suggest the
\texttt{d02} \cite{DBLP:conf/kr/Darwiche02} encoding that, in many ways, remains
the foundation behind most other encodings. He also introduced the distinction
between \emph{indicator} and \emph{parameter variables}; the former represent
variable-value pairs in the Bayesian network, while the latter are associated
with probabilities in the CPTs. The encoding \texttt{sbk05}
\cite{DBLP:conf/aaai/SangBK05} is the only encoding that deviates from this
arrangement: for each variable in the Bayesian network, one indicator variable
acts simultaneously as a parameter variable. Chavira and Darwiche propose
\texttt{cd05} \cite{DBLP:conf/ijcai/ChaviraD05} where they shift from WMC to
minimum-cardinality WMC because that allows the encoding to have fewer variables
and clauses. In particular, they propose a way to use the same parameter
variable to represent all probabilities in a CPT that are equal and keep only
clauses that `imply' parameter variables (i.e., omit clauses where a parameter
variable implies indicator variables). In their next encoding, \texttt{cd06}
\cite{DBLP:conf/sat/ChaviraD06}, the same authors optimise the aforementioned
implication clauses, choosing the smallest sufficient selection of indicator
variables. A decade later, Bart et al. present \texttt{bklm16}
\cite{DBLP:conf/ecai/BartKLM16} that improves upon \texttt{cd06} in two ways.
First, they optimise the number of indicator variables used per Bayesian network
variable from a linear to a logarithmic amount. Second, they introduce a scaling
factor that can `absorb' one probability per Bayesian network variable. However,
for this work, we choose to disable the latter improvement since this scaling
factor is often small enough to be indistinguishable from zero without the use
of arbitrary precision arithmetic. Indeed, even a small Bayesian network with
seven mutually independent binary variables, 0.1 and 0.9 probabilities each, is
already big enough for the scaling factor to be exactly equal to zero (as
produced by the \texttt{bklm16}
encoder\footnote{\url{http://www.cril.univ-artois.fr/kc/bn2cnf.html}}).

%This additional condition on model
%cardinality becomes necessary because these encodings eliminate clauses of the
%form $p \Rightarrow i$, where $p \in X_P$ is a parameter variable, and $i \in
%X_I$ is an indicator variable. Nonetheless, our transformation algorithm still
%works on such encodings, although the experimental results are discouraging
%because they use approximately twice as many indicator variables. For
%instance, each binary variable of a Bayesian network is encoded using two
%indicator variables while one would suffice.

\section{Pseudo-Boolean Functions}

In this work, we propose a more expressive representation for WMC based on
pseudo-Boolean functions. A \emph{pseudo-Boolean function} is a function of the
form $\{ 0, 1 \}^n \to \mathbb{R}$ \cite{DBLP:journals/dam/BorosH02}.
Equivalently, let $X$ denote a set with $n$ elements (we will refer to them as
\emph{variables}), and $2^X$ denote its powerset. Then a pseudo-Boolean function
can have $2^X$ as its domain (then it is also known as a \emph{set function}).

Pseudo-Boolean functions, most commonly represented as
algebraic decision diagrams (ADDs) \cite{DBLP:journals/fmsd/BaharFGHMPS97}
(although a tensor-based approach has also been suggested
\cite{DBLP:journals/corr/abs-1908-04381,DBLP:conf/cp/DudekPV20}), have seen
extensive use in value iteration for Markov decision processes
\cite{DBLP:conf/uai/HoeySHB99}, both exact and approximate Bayesian network
inference \cite{DBLP:conf/ijcai/ChaviraD07,DBLP:conf/uai/GogateD11}, and
sum-product network \cite{DBLP:conf/uai/PoonD11} to Bayesian network conversion
\cite{DBLP:conf/icml/ZhaoMP15}. ADDs have been extended to compactly represent
additive and multiplicative structure \cite{DBLP:conf/ijcai/SannerM05},
sentences in first-order logic \cite{DBLP:journals/ai/SannerB09}, and continuous
variables \cite{DBLP:conf/uai/SannerDB11}, the last of which was also applied to
weighted model integration, i.e., the WMC extension for continuous variables
\cite{DBLP:conf/ijcai/BellePB15,DBLP:conf/ijcai/KolbMSBK18}.

Since two-valued pseudo-Boolean functions will be used extensively henceforth,
we introduce some new notation. For any propositional formula $\phi$ over $X$
and $p, q \in \mathbb{R}$, let $[\phi]^p_q\colon 2^X \to \mathbb{R}$ be the
pseudo-Boolean function defined as
\[
  [\phi]^p_q(Y) \coloneqq
  \begin{cases}
    p & \text{if } Y \models \phi \\
    q & \text{otherwise}
  \end{cases}
\]
for any $Y \subseteq X$. Next, we define some useful operations on
pseudo-Boolean functions. The definitions of multiplication and projection are
equivalent to those in previous work
\cite{DBLP:conf/aaai/DudekPV20,DBLP:conf/cp/DudekPV20}.

\begin{definition}[Operations] \label{def:operations}
  Let $f, g\colon 2^X \to \mathbb{R}$ be pseudo-Boolean functions, $x, y \in X$,
  $Y = \{y_i\}_{i=1}^n \subseteq X$, and $r \in \mathbb{R}$. Operations such as
  addition and multiplication are defined pointwise, i.e., $(f+g)(Y) \coloneqq
  f(Y)+g(Y)$, and likewise for multiplication. Note that properties such as
  associativity and commutativity are inherited from $\mathbb{R}$. By regarding
  a real number as a constant pseudo-Boolean function, we can reuse the same
  definitions to define \emph{scalar} operations as $(r+f)(Y) = r+f(Y)$, and $(r
  \cdot f)(Y) = r \cdot f(Y)$.

  \emph{Restrictions} $f|_{x=0}, f|_{x=1}\colon 2^X \to \mathbb{R}$ of $f$ are
  defined as $f|_{x=0}(Y) \coloneqq f(Y \setminus \{x\})$, and $f|_{x=1}(Y)
  \coloneqq f(Y \cup \{x\})$ for all $Y \subseteq X$.

  \emph{Projection} $\exists_x$ is an endomorphism $\exists_x\colon
  \mathbb{R}^{2^X} \to \mathbb{R}^{2^X}$ defined as $\exists_xf \coloneqq
  f|_{x=1} + f|_{x=0}$. Since projection is commutative (i.e.,
  $\exists_x\exists_yf = \exists_y\exists_xf$)
  \cite{DBLP:conf/aaai/DudekPV20,DBLP:conf/cp/DudekPV20}, we can define
  $\exists_Y\colon \mathbb{R}^{2^X} \to \mathbb{R}^{2^X}$ as $\exists_Y
  \coloneqq \exists_{y_1}\exists_{y_2}\dots\exists_{y_n}$. Throughout the paper,
  projection is assumed to have the lowest precedence (e.g., $\exists_x fg =
  \exists_x (fg)$).
\end{definition}

Below we list some properties of the operations on pseudo-Boolean functions
discussed in this section that can be conveniently represented using our syntax.
The proofs of all these properties follow directly from the definitions.

\begin{proposition}[Basic Properties] \label{prop:basic}
  For any propositional formulas $\phi$ and $\psi$, and $a, b, c, d \in
  \mathbb{R}$,
  \begin{itemize}
  \item $[\phi]^a_b = [\neg \phi]^b_a$;
  \item $c + [\phi]^a_b = [\phi]^{a+c}_{b+c}$;
  \item $c \cdot [\phi]^a_b = [\phi]^{ac}_{bc}$;
  \item $[\phi]^a_b \cdot [\phi]^c_d = [\phi]^{ac}_{bd}$;
  \item $[\phi]^1_0 \cdot [\psi]_0^1 = [\phi \land \psi]_0^1$.
  \end{itemize}
  And for any pair of pseudo-Boolean functions $f, g \colon 2^X \to \mathbb{R}$
  and $x \in  X$, $(fg)|_{x=i} = f|_{x=i} \cdot g|_{x=i}$ for $i = 0, 1$.
\end{proposition}

\begin{remark}
  Note that our definitions of binary operations assumed equal domains. For
  convenience, we can assume domains to shrink whenever a function is
  independent of some of the variables (i.e., $f|_{x=0} = f|_{x=1}$) and expand
  for binary operations to make the domains of both functions equal. For
  instance, let $[x]_0^1,[\neg x]_0^1\colon 2^{\{x\}} \to \mathbb{R}$ and
  $[y]_0^1\colon 2^{\{y\}} \to \mathbb{R}$ be pseudo-Boolean functions. Then
  $[x]_0^1 \cdot [\neg x]_0^1$ has $2^\emptyset$ as its domain. To multiply
  $[x]_0^1$ and $[y]_0^1$, we expand $[x]_0^1$ into $\left([x]_0^1\right)'\colon
  2^{\{x, y\}} \to \mathbb{R}$ which is defined as $\left([x]_0^1\right)'(Z)
  \coloneqq [x]_0^1(Z \cap \{ x \})$ for all $Z \subseteq \{ x, y \}$ (and
  equivalently for $[y]_0^1$).
\end{remark}

\section{Pseudo-Boolean Projection}

We introduce a new type of computational problem called \emph{pseudo-Boolean
  projection} based on two-valued pseudo-Boolean functions. While the same
computational framework can handle any pseudo-Boolean functions, two-valued
functions are particularly convenient because \textsf{DPMC} can be easily
adapted to use them as input, and they are easily representable in both text
files and using the syntax proposed in this paper.

\begin{definition}[PBP Instance] \label{def:new_wmc}
  A PBP instance is a tuple $(F, X, \omega)$, where $X$ is the set of variables,
  $F$ is a set of two-valued pseudo-Boolean functions $2^X \to \mathbb{R}$, and
  $\omega \in \mathbb{R}$ is the scaling factor. Its \emph{answer} is $\omega
  \cdot \left(\exists_X\prod_{f \in F}f\right)(\emptyset)$.
\end{definition}

Adding scaling factor $\omega$ to the definition allows us to remove clauses
that consist entirely of a single parameter variable. The idea of extracting
some of the structure of the WMC instance into an external multiplicative factor
was inspired by the \texttt{bklm16} encoding.

\subsection{From WMC to PBP}

\begin{algorithm}[t]
  \caption{WMC to PBP transformation}
  \label{alg:transformation}
  \SetKwFunction{rename}{rename}
  \SetKwProg{Fn}{Function}{:}{}
  \KwData{WMC (or minimum-cardinality WMC) instance $(\phi, X_I, X_P, w)$}
  \KwResult{PBP instance $(F, X_I, \omega)$}
  $F \gets \emptyset$\;
  $\omega \gets 1$\;
%  let $d\colon X_P \to \mathbb{N}$ be defined as $p \mapsto |\{ o \in X_P \mid o
%  \le p \}|$\;
  \ForEach{clause $c \in \phi$\label{line:foreach1start}}{
    \uIf{$c \cap X_P = \{ p \}$ for some $p$ \textnormal{\textbf{and}} $w(p) \ne
      1$}{
      \uIf{$|c| = 1$}{$\omega \gets \omega \times w(p)$\;}
      \Else{
        $F \gets F \cup \left\{ \left[ \bigwedge_{l \in c \setminus
              \{p\}} \neg l \right]^{w(p)}_1 \right\}$\;
      }
    }
    \ElseIf{$\{p \mid \neg p \in c\} \cap X_P  = \emptyset$}{
      $F \gets F \cup \{ [c]^1_0 \}$\; \label{line:foreach1end}
    }
  }
  \ForEach{$v \in X_I$ such that $\{[v]_1^p, [\neg v]_1^q\}
    \subseteq F$ for some $p$ and $q$ \label{line:foreach2start}}{
    $F \gets F \setminus \{ [v]_1^p, [\neg v]_1^q \} \cup \{ [v]_q^p
    \}$\; \label{line:foreach2end}
  }
%  replace every variable $v$ in $F$ with $\rename{$v$}$\;
%  \Return{$(F, X_I, \omega)$}\;
%  \Fn{\rename{$v$}}{
%    $S \gets \{u \in X_P \mid u \le v\}$\;
%    \lIf{$S = \emptyset$}{\Return{$v$}}
%    \Return{$v - d(\max S)$}\;
%  }
\end{algorithm}

In this section, we describe an algorithm for transforming WMC instances to the
PBP format while removing all parameter variables. The algorithm works on four
out of the five Bayesian network encodings. There is no obvious way to adjust it
to work with \texttt{sbk05} because the roles of indicator and parameter (i.e.,
`chance') variables overlap \cite{DBLP:conf/aaai/SangBK05}. The algorithm is
based on several observations that will be made more precise in
\cref{sec:proof}. First, all weights except for $\{w(p) \mid p \in X_P\}$ are
redundant as they either duplicate an already-defined weight or are equal to
one. Second, each clause has at most one parameter variable. Third, if the
parameter variable is negated, we can ignore the clause (this idea first appears
in the \texttt{cd05} paper \cite{DBLP:conf/ijcai/ChaviraD05}). Note that while
we formulate our algorithm as a sequel to the WMC encoding procedure primarily
because the implementations of Bayesian network WMC encodings are all
closed-source, as all transformations in the algorithm are local, it can be
efficiently incorporated into a WMC encoding algorithm with no slowdown.

The algorithm is listed as \cref{alg:transformation}. The main part of the
algorithms is the first loop that iterates over clauses. If a clause consists of
a single parameter variable, we incorporate it into $\omega$. If a clause is
of the form $\alpha \Rightarrow p$, where $p \in X_P$ and $\alpha$ is a
conjunction of literals over $X_I$, we transform it into a pseudo-Boolean
function $[\alpha]_1^{w(p)}$. If a clause (say, $c \in \phi$) has no parameter
variables, we reformulate it into a pseudo-Boolean function $[c]_0^1$. Finally,
if a clause has negative parameter literals, we skip it.

As all `weighted' pseudo-Boolean functions produced by the first loop are of the
form $[\alpha]_1^p$ (for some $p \in \mathbb{R}$ and formula $\alpha$), the
second loop merges two functions into one whenever $\alpha$ is a literal. Note
that taking into account the order in which clauses are typically generated by
encoding algorithms allows us to do this in linear time (i.e., the two mergeable
functions will be generated one after the other).

%\item The second \textbf{foreach} loop can be performed in constant time by
%  representing $\phi'$ as a list and assuming that the two 'clauses' are
%  adjacent in that list (and incorporating it into the first loop).
%\item The $d$ map is constructed in $\mathcal{O}(|X_P|\log|X_P|)$ time (we want
%  to use a data structure based on binary search trees rather than hashing).
%\item \texttt{rename} can be implemented in $\mathcal{O}(\log |X_P|)$ time.

\subsection{Correctness Proofs} \label{sec:proof}

In this section, we outline key properties that a (WMC or minimum-cardinality
WMC) encoding has to satisfy for \cref{alg:transformation} to output an
equivalent PBP instance. We divide the correctness proof into two theorems:
\cref{thm:correctness} for WMC encodings (i.e., \texttt{bklm16} and
\texttt{d02}) and \cref{thm:mccorrectness} for minimum-cardinality WMC encodings
(i.e., \texttt{cd05} and \texttt{cd06}). We begin by listing some properties of
pseudo-Boolean functions and establishing a canonical transformation from WMC
to PBP.

\begin{theorem}[Early Projection
  \cite{DBLP:conf/aaai/DudekPV20,DBLP:conf/cp/DudekPV20}] \label{thm:early}
  Let $X$ and $Y$ be sets of variables. For all pseudo-Boolean functions
  $f\colon 2^X \to \mathbb{R}$ and $g\colon 2^Y \to \mathbb{R}$, if $x \in X
  \setminus Y$, then $\exists_x (f \cdot g) = (\exists_x f) \cdot g$.
\end{theorem}

\begin{lemma} \label{lemma:sum}
  For any pseudo-Boolean function $f\colon 2^X \to \mathbb{R}$, we have that
  $(\exists_X f)(\emptyset) = \sum_{Y \subseteq X} f(Y)$.
\end{lemma}
\begin{proof}
  If $X = \{x\}$, then
  \[
    (\exists_xf)(\emptyset) = (f|_{x=1} + f|_{x=0})(\emptyset) =
    f|_{x=1}(\emptyset) + f|_{x=0}(\emptyset) = \sum_{Y \subseteq \{x\}} f(Y).
  \]
  This easily extends to $|X| > 1$ by the definition of projection on sets of
  variables.
\end{proof}

\begin{proposition} \label{prop:equivalence}
  Let $(\phi, X_I, X_P, w)$ be a WMC instance. Then
  \begin{equation}
  \left(\left\{[c]_0^1 \;\middle|\; c \in \phi\right\} \cup \left\{[x]_{w(\neg
        x)}^{w(x)} \;\middle|\; x \in X_I \cup X_P\right\}, X_I \cup X_P,
    1\right) \label{eq:new_wmc}
  \end{equation}
  is a PBP instance with the same answer.
\end{proposition}
\begin{proof}
  Let $f = \prod_{c \in \phi} [c]_0^1$, and $g = \prod_{x \in X_I \cup X_P}
  [x]_{w(\neg x)}^{w(x)}$. Then the WMC answer is
  \eqref{eq:new_wmc} is
  \[
    (\exists_{X_I \cup X_P} fg)(\emptyset) = \sum_{Y \subseteq X_I \cup X_P}
    (fg)(Y) = \sum_{Y \subseteq X_I \cup X_P} f(Y)g(Y)
  \]
  by \cref{lemma:sum}. Note that
  \[
    f(Y) =
    \begin{cases}
      1 & \text{if } Y \models \phi, \\
      0 & \text{otherwise},
    \end{cases}
    \quad
    \text{and}
    \quad
    g(Y) = \prod_{Y \models l} w(l),
  \]
  which means that $\sum_{Y \subseteq X_I \cup X_P} f(Y)g(Y) = \sum_{Y \models
    \phi} \prod_{Y \models l} w(l)$ as required.
\end{proof}

\begin{theorem}[Correctness for WMC] \label{thm:correctness}
  \Cref{alg:transformation}, when given a WMC instance $(\phi, X_I, X_P,
  w)$, returns PBP instance with the same answer, provided the following
  conditions are satisfied:
  \begin{enumerate}
  \item for all indicator variables $i \in X_I$, $w(i) = w(\neg i) =
    1$, \label{cond:one}
  \item and either
    \begin{enumerate}
    \item for all parameter variables $p \in X_P$, there is a non-empty family
      of literals $(l_i)_{i=1}^n$ such that \label{cond:d02}
      \begin{enumerate}
      \item $w(\neg p) = 1$,
      \item $l_i \in X_I$ or $\neg l_i \in X_I$ for all $i = 1, \dots, n$,
        \label{condition:equivalence1}
      \item and $\{ c \in \phi \mid p \in c \text{ or } \neg p \in c \} =
        \left\{p \lor \bigvee_{i=1}^n \neg l_i \right\} \cup \{ l_i \lor \neg p
        \mid i = 1, \dots, n \}$; \label{condition:equivalence2}
      \end{enumerate}
    \item or for all parameter variables $p \in X_P$, \label{cond:bklm16}
      \begin{enumerate}
      \item $w(p) + w(\neg p) = 1$,
      \item for any clause $c \in \phi$, $|c \cap X_P| \le 1$, \label{cond:2b2}
      \item there is no clause $c \in \phi$ such that $\neg p \in
        c$, \label{cond:2b3}
      \item if $\{ p \} \in \phi$, then there is no clause $c \in \phi$ such
        that $c \ne \{ p \}$ and $p \in c$, \label{cond:just_parameter}
        % 'prime implicants are logically disjoint'
      \item and for any $c, d \in \phi$ such that $c \ne d$, $p \in c$ and $p
        \in d$, $\bigwedge_{l \in c \setminus \{ p \}} \neg l \land \bigwedge_{l
          \in d \setminus \{ p \}} \neg l$ is false. \label{cond:disjoint}
      \end{enumerate}
    \end{enumerate}
  \end{enumerate}
\end{theorem}

\Cref{cond:one} applies to all four WMC encodings under consideration and
ensures that only parameter variables carry nontrivial weights. \Cref{cond:d02}
(for \texttt{d02}) simply states that each parameter variable is equivalent to a
conjunction of indicator literals. \Cref{cond:bklm16} is for encodings that have
implications rather than equivalences associated with parameter variables
(which, in this case, is \texttt{bklm16}). It ensures that each clause has at
most one positive parameter literal and no negative ones, and that at most one
implication clause per any parameter variable $p \in X_P$ can `force $p$ to be
positive'.

\begin{proof}
  By \cref{prop:equivalence},
  \begin{equation}
    \left(\left\{[c]_0^1 \;\middle|\; c \in \phi\right\} \cup \left\{[x]_{w(\neg
          x)}^{w(x)} \;\middle|\; x \in X_I \cup X_P\right\}, X_I \cup X_P,
      1\right) \label{eq:new_wmc2}
  \end{equation}
  is a PBP instance with the same answer as the given WMC instance. By
  \cref{def:new_wmc}, its answer is $\left(\exists_{X_I \cup X_P} \left(\prod_{c
        \in \phi} [c]_0^1 \right) \prod_{x \in X_I \cup X_P} [x]_{w(\neg
      x)}^{w(x)} \right)(\emptyset)$. Since both \cref{cond:d02,cond:bklm16}
  ensure that each clause in $\phi$ has at most one parameter variable, we can
  partition $\phi$ into $\phi_* \coloneqq \{c \in \phi \mid \mathtt{Vars}(c)
  \cap X_P = \emptyset \}$ and $\phi_p \coloneqq \{ c \in \phi \mid
  \mathtt{Vars}(c) \cap X_P = \{ p \} \}$ for all $p \in X_P$. We can then use
  \cref{thm:early} to reorder the answer into $\left(\exists_{X_I} \left(
      \prod_{x \in X_I} [x]_{w(\neg x)}^{w(x)} \right) \left( \prod_{c \in
        \phi_*} [c]_0^1 \right) \prod_{p \in X_P} \exists_p [p]_{w(\neg
      p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1 \right)(\emptyset)$.

  Let us first consider how the unfinished WMC instance $(F, X_I, \omega)$ after
  the loop on \crefrange{line:foreach1start}{line:foreach1end} differs from
  \eqref{eq:new_wmc2}. Note that \cref{alg:transformation} leaves each $c \in
  \phi_*$ unchanged, i.e., adds $[c]_0^1$ to $F$. We can then fix an arbitrary
  $p \in X_P$ and let $F_p$ be the set of functions added to $F$ as a
  replacement of $\phi_p$. It is sufficient to show that
  \begin{equation} \label{eq:to_show}
    \omega \prod_{f \in F_p} f = \exists_p [p]_{w(\neg p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1.
  \end{equation}
  Note that under \cref{cond:d02}, $\bigwedge_{c \in \phi_p} c \equiv p
  \Leftrightarrow \bigwedge_{i=1}^n l_i$ for some family of indicator variable
  literals $(l_i)_{i=1}^n$. Thus,
  \[
    \exists_p [p]_{w(\neg p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1 = \exists_p
    [p]_1^{w(p)} \left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1.
  \]
  If $w(p) = 1$, then
  \begin{equation} \label{eq:bigsums}
    \exists_p [p]_1^{w(p)} \left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1 = \left.\left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1\right|_{p=1} + \left.\left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1\right|_{p=0}.
  \end{equation}
  Since for any input, $\bigwedge_{i=1}^n l_i$ is either true or false, exactly
  one of the two summands in \cref{eq:bigsums} will be equal to one, and the
  other will be equal to zero, and so
  \[
    \left.\left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i
      \right]_0^1\right|_{p=1} + \left.\left[ p \Leftrightarrow \bigwedge_{i=1}^n
        l_i \right]_0^1\right|_{p=0} = 1,
  \]
  where $1$ is a pseudo-Boolean function that always returns one. On the other
  side of \cref{eq:to_show}, since $F_p = \emptyset$, and $\omega$ is unchanged,
  we get $\omega\prod_{f \in F_p} f = 1$, and so \cref{eq:to_show} is satisfied
  under \cref{cond:d02} when $w(p) = 1$.

  If $w(p) \ne 1$, then $F_p = \left\{ \left[ \bigwedge_{i = 1}^n l_i
    \right]_1^{w(p)} \right\}$, and $\omega = 1$, and so we want to show that
  $\left[ \bigwedge_{i = 1}^n l_i \right]_1^{w(p)} = \exists_p [p]_1^{w(p)}
  \left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1$, and indeed
  \[
    \exists_p [p]_1^{w(p)} \left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i
    \right]_0^1 = w(p) \cdot \left[ \bigwedge_{i=1}^n l_i \right]_0^1 + \left[
      \bigwedge_{i=1}^n l_i \right]_1^0 = \left[ \bigwedge_{i=1}^n l_i
    \right]_1^{w(p)}.
  \]
  This finishes the proof of the correctness of the first loop under
  \cref{cond:d02}.

  Now let us assume \cref{cond:bklm16}. We still want to prove
  \cref{eq:to_show}. If $w(p) = 1$, then $F_p = \emptyset$, and $\omega = 1$,
  and so the left-hand side of \cref{eq:to_show} is equal to one. Then the
  right-hand side is
  \[
    \exists_p [p]_0^1 \prod_{c \in \phi_p} [c]_0^1 = \exists_p \left[ p \land
      \bigwedge_{c \in \phi_p} c \right]_0^1 = \exists_p [p]_0^1 = 0 + 1 = 1
  \]
  since $p \in c$ for every clause $c \in \phi_p$.

  If $w(p) \ne 1$, and $\{ p \} \in \phi_p$, then, by
  \cref{cond:just_parameter}, $\phi_p = \{ \{ p \} \}$, and
  \cref{alg:transformation} produces $F_p = \emptyset$ and $\omega = w(p)$, and
  so
  \[
    \exists_p [p]_{w(\neg p)}^{w(p)} [p]_0^1 = \exists_p [p]^{w(p)}_0 = w(p) =
    \omega \prod_{f \in F_p} f.
  \]
  The only remaining case is when $w(p) \ne 1$ and $\{ p \} \not \in \phi_p$.
  Then $\omega = 1$, and $F_p = \left\{ \left[\bigwedge_{l \in c \setminus \{ p
        \}} \neg l\right]_1^{w(p)} \;\middle|\; c \in \phi_p \right\}$, so we
  need to show that
  \[
    \prod_{c \in \phi_p} \left[\bigwedge_{l \in c \setminus \{ p \}} \neg
      l\right]_1^{w(p)} = \exists_p [p]_{1-w(p)}^{w(p)} \prod_{c \in \phi_p}
    [c]_0^1.
  \]
  We can rearrange the right-hand side as
  % \begin{align*}
  %   \exists_p [p]_{1-w(p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1 &= \exists_p [p]_{1-w(p)}^{w(p)} \left[\bigwedge_{c \in \phi_p} c\right]_0^1 = \exists_p [p]_{1-w(p)}^{w(p)} \left[ p \lor \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^1 \\
  %                                                              &= w(p) + (1-w(p)) \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^1 = w(p) + \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^{1-w(p)} \\
  %                                                              &= \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_{w(p)}^1 = \left[ \neg \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_1^{w(p)} = \left[ \bigvee_{c \in \phi_p} \neg(c \setminus \{ p \}) \right]_1^{w(p)} \\
  %                                                              &= \left[ \bigvee_{c \in \phi_p} \neg \bigvee_{l \in c \setminus \{ p \}} l \right]_1^{w(p)} = \left[ \bigvee_{c \in \phi_p} \bigwedge_{l \in c \setminus \{ p \}} \neg l \right]_1^{w(p)}.
  % \end{align*}
  \begin{align*}
    \exists_p [p]_{1-w(p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1 &= \exists_p [p]_{1-w(p)}^{w(p)} \left[ p \lor \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^1 \\
                                                               &= w(p) + (1-w(p)) \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^1 \\
                                                               &= \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_{w(p)}^1 = \left[ \bigvee_{c \in \phi_p} \bigwedge_{l \in c \setminus \{ p \}} \neg l \right]_1^{w(p)}.
  \end{align*}
   By \cref{cond:disjoint}, $\bigwedge_{l \in c \setminus \{ p \}} \neg l$ can
   be true for at most one $c \in \phi_p$, and so
   \[
     \left[ \bigvee_{c \in \phi_p} \bigwedge_{l \in c \setminus \{ p \}} \neg l
     \right]_1^{w(p)} = \prod_{c \in \phi_p} \left[ \bigwedge_{l \in c \setminus
         \{ p \}} \neg l \right]_1^{w(p)}
   \]
   which is exactly what we needed to show. This ends the proof that the
   first loop of \cref{alg:transformation} preserves the answer under both
   \cref{cond:d02} and \cref{cond:bklm16}. Finally, the loop on
   \crefrange{line:foreach2start}{line:foreach2end} of \cref{alg:transformation}
   replaces $[v]_1^p[\neg v]_1^q$ with $[v]_q^p$ (for some $v \in X_I$ and $p, q
   \in \mathbb{R}$), but, of course, $[v]_1^p[\neg v]_1^q = [v]_1^p[v]_q^1 =
   [v]_q^p$, i.e., the answer is unchanged.
\end{proof}

\begin{theorem}[Minimum-Cardinality Correctness] \label{thm:mccorrectness}
  Let $(\phi, X_I, X_P, w)$ be a minimum-cardinality WMC instance that satisfies
  \cref{cond:one,cond:2b2,cond:2b3,cond:just_parameter,cond:disjoint} of
  \cref{thm:correctness} as well as the following:
  \begin{enumerate}
  \item for all parameter variables $p \in X_P$, $w(\neg p) = 1$.
  \item all models of $\{ c \in \phi \mid c \cap X_P = \emptyset \}$ (as subsets
    of $X_I$) have the same cardinality; \label{cond:22}
  \item $\min_{Z \subseteq X_P} |Z|$ such that $Y \cup Z \models \phi$ is the
    same for all $Y \models \{ c \in \phi \mid c \cap X_P = \emptyset
    \}$. \label{cond:23}
  \end{enumerate}
  Then \cref{alg:transformation}, when applied to $(\phi, X_I, X_P, w)$, outputs
  a PBP instance with the same answer.
\end{theorem}

In this case, we have to add some assumptions about the cardinality of models.
\Cref{cond:22} states that all models of the indicator-only part of the formula
have the same cardinality. Bayesian network encodings such as \texttt{cd05} and
\texttt{cd06} satisfy this condition by assigning an indicator variable to each
possible variable-value pair and requiring each random variable to be paired
with exactly one value. \Cref{cond:23} then says that the smallest number of
parameter variables needed to turn an indicator-only model into a full model is
the same for all indicator-only models. As some ideas duplicate between the
proofs of \cref{thm:correctness,thm:mccorrectness}, the following proof is
slightly less explicit and assumes that $\omega = 1$.
% We assume that $\omega = 1$ (i.e., $\{ p \} \not\in \phi$ for all parameter
% variables $p \in X_P$) as the proof can be easily extended to handle $\omega
% \neq 1$ similarly to the proof of \cref{thm:correctness}.

\begin{proof}
  Let $(F, X_I, \omega)$ be the tuple returned by \cref{alg:transformation} and
  note that
  \[
    F = \left\{ [c]_0^1 \mid c \in \phi \text{, } c \cap X_P = \emptyset
    \right\} \cup \left\{ \left[ \bigwedge_{l \in c \setminus \{ p \}} \neg l
      \right]_1^{w(p)} \;\middle|\; p \in X_P \text{, } p \in c \in \phi
      \text{, } c \ne \{ p \} \right\}.
  \]
  We split the proof into two parts. In the first part, we show that there is a
  bijection between minimum-cardinality models of $\phi$ and $Y \subseteq X_I$
  such that $\left(\prod_{f \in F} f\right)(Y) \ne 0$ (with the assumption that
  $w(p) \ne 0$ for all $p \in X_P$).\footnote{The assumption is only there to
    make the proof more convenient and can be omitted by reformulating the proof
    in a slightly more cumbersome way.} Let $Y \subseteq X_I$ and $Z \subseteq
  X_I \cup X_P$ be related via this bijection. Then in the second part we will
  show that
  \begin{equation} \label{eq:weights}
    \prod_{Z \models l} w(l) = \left(\prod_{f \in F} f\right)(Y).
  \end{equation}

  On the one hand, if $Z \subseteq X_I \cup X_P$ is a minimum-cardinality model
  of $\phi$, then $\left(\prod_{f \in F}\right)(Z \cap X_I) \ne 0$ under the
  given assumptions. On the other hand, if $Y \subseteq X_I$ is such that
  $\left(\prod_{f \in F}\right)(Y) \ne 0$, then $Y \models \{ c \in \phi \mid c
  \cap X_P = \emptyset \}$. Let $Y \subseteq Z \subseteq X_I \cup X_P$ be the
  smallest superset of $Y$ such that $Z \models \phi$ (it exists by
  \cref{cond:2b3} of \cref{thm:correctness}). We need to show that $Z$ has
  minimum cardinality. Let $Y'$ and $Z'$ be defined equivalently to $Y$ and $Z$.
  We will show that $|Z| = |Z'|$. Note that $|Y| = |Y'|$ by \cref{cond:22}, and
  $|Z \setminus Y| = |Z' \setminus Y'|$ by \cref{cond:23}. Combining that with
  the general property that $|Z| = |Y| + |Z \setminus Y|$ finishes the first
  part of the proof.

  For the second part, let us consider the multiplicative influence of a single
  parameter variable $p \in X_P$ on \cref{eq:weights}. If the left-hand side is
  multiplied by $w(p)$ (i.e., $p \in Z$), then there must be some clause $c \in
  \phi$ such that $Z \setminus \{ p \} \not\models c$. But then $Y \models
  \bigwedge_{l \in c \setminus \{ p \}} \neg l$, and so the right-hand side is
  multiplied by $w(p)$ as well (exactly once because of \cref{cond:disjoint} of
  \cref{thm:correctness}). This argument works in the other direction as well.
\end{proof}

\section{Experimental Evaluation}

We run a set of experiments, comparing all five original Bayesian network
encodings (\texttt{bklm16}, \texttt{cd05}, \texttt{cd06}, \texttt{d02}
\texttt{sbk05}) as well as the first four with \cref{alg:transformation} applied
afterwards.\footnote{Recall that \texttt{cd05} and \texttt{cd06} are
  incompatible with \textsf{DPMC}.} For each encoding \texttt{e}, we write
\texttt{e++} to denote the combination of encoding a Bayesian network as a WMC
instance using \texttt{e} and transforming it into a PBP instance using
\cref{alg:transformation}. Along with
\textsf{DPMC}\footnote{\url{https://github.com/vardigroup/DPMC}}, we also
include WMC algorithms used in the papers that introduce each encoding:
\textsf{Ace} for \texttt{cd05}, \texttt{cd06}, and \texttt{d02};
\textsf{Cachet}\footnote{\url{https://cs.rochester.edu/u/kautz/Cachet/}}
\cite{DBLP:conf/sat/SangBBKP04} for \texttt{sbk05}; and
\textsf{c2d}\footnote{\url{http://reasoning.cs.ucla.edu/c2d/}}
\cite{DBLP:conf/ecai/Darwiche04} with
\textsf{query-dnnf}\footnote{\url{http://www.cril.univ-artois.fr/kc/d-DNNF-reasoner.html}}
for \texttt{bklm16}. \textsf{Ace} is also used to encode Bayesian networks into
WMC instances for all encodings except for \texttt{bklm16} which uses another
encoder mentioned previously. We focus on the following questions:
\begin{itemize}
\item Can parameter variable elimination improve inference speed?
\item How does DPMC combined with encodings without (and with) parameter
  variables compare with other WMC algorithms and other encodings?
\item Which instances is our approach particularly successful on (compared to
  other algorithms and encodings and to the same encoding before our
  transformation)?
\item What proportion of variables is typically eliminated?
\item Do some encodings benefit from this transformation more than others?
\end{itemize}

\subsection{Setup}

\textsf{DPMC} is run with tree decomposition-based planning and ADD-based
execution---the best-performing combination in the original set of experiments
\cite{DBLP:conf/cp/DudekPV20}. We use a single iteration of \textsf{htd}
\cite{DBLP:conf/cpaior/AbseherMW17} to generate approximately optimal tree
decompositions---we found that this configuration is efficient enough to handle
huge instances, and yet the width of the returned decomposition is unlikely to
differ from optimal by more than one or two. We also enabled \textsf{DPMC}'s
greedy mode. This mode (which was not part of the original paper
\cite{DBLP:conf/cp/DudekPV20}) optimises the order in which pseudo-Boolean
functions are multiplied by prioritising functions with small representations.

For experimental data, we use Bayesian networks available with \textsf{Ace} and
\textsf{Cachet}. We split them into the following groups:
\begin{itemize}
\item DQMR (390 instances) and
\item Grid networks (450 instances) as described by Sang et al.
  \cite{DBLP:conf/aaai/SangBK05};
\item Mastermind (144 instances) and
\item Random Blocks (256 instances) by Chavira et al.
  \cite{DBLP:journals/ijar/ChaviraDJ06};
\item other binary Bayesian networks (50 instances) including Plan Recognition
  \cite{DBLP:conf/aaai/SangBK05}, Friends and Smokers, Students and Professors
  \cite{DBLP:journals/ijar/ChaviraDJ06}, and \texttt{tcc4f};
\item non-binary classic networks (176 instances): \texttt{alarm},
  \texttt{diabetes}, \texttt{hailfinder}, \texttt{mildew},
  \texttt{munin1}--\texttt{4}, \texttt{pathfinder}, \texttt{pigs}, and
  \texttt{water}.
\end{itemize}

To perform Bayesian network inference with \textsf{DPMC} (or with any
other WMC algorithm not based on compilation such as \textsf{Cachet}), one
needs to select a probability to compute
\cite{DBLP:conf/cp/DudekPV20,DBLP:conf/sat/SangBBKP04}. If a network comes
with an evidence file, we compute the probability of this evidence. Otherwise,
let $X$ be the variable last mentioned in the Bayesian network file. If
\textsf{true} is one of the values of $X$, then we compute $\Pr(X =
\textsf{true})$, otherwise we choose the first-mentioned value of $X$.

The experiments were run on a computing cluster with Intel Xeon E5-2630,
Intel Xeon E7-4820, and Intel Xeon Gold 6138 processors with a
\SI{1000}{\second} timeout separately on both encoding and inference, and a
\SI{32}{\gibi\byte} memory limit.\footnote{Each instance was run on the same
  processor across all algorithms and encodings.}

\subsection{Results}

\begin{figure}
  \centering
  \input{cumulative.tex}%
  \caption{Cumulative numbers of instances solved by each algorithm-encoding
    pair over time. The dotted line denotes the total number of instances used.}
  \label{fig:cumulative}
\end{figure}

\begin{figure}
  \centering
  \input{scatter.tex}%
  \caption{An instance-by-instance comparison between $\textsf{DPMC} +
    \texttt{bklm16++}$ (the best combination according to \cref{fig:cumulative})
  and the second and third best-performing combinations: $\textsf{Ace} +
  \texttt{cd06}$ and $\textsf{DPMC} + \texttt{bklm16}$.}
  \label{fig:scatter}
\end{figure}

\begin{figure}
  \centering
  \begin{minipage}{0.49\textwidth}
    \centering
    \input{box.tex}%
    \caption{Box plots of the numbers of variables in each encoding across all
      benchmark instances before and after applying \cref{alg:transformation}.
      Outliers and the top parts of some whiskers are
      omitted.}\label{fig:box}
  \end{minipage}\hfill
  \begin{minipage}{0.49\textwidth}
    \centering
    \captionof{table}{The numbers of instances (out of 1466) that each algorithm and
      encoding combination solved faster than any other combination and in
      total.}\label{tbl:performance}
    \begin{tabular}{lrr}
      \toprule
      Combination & Fastest & Solved \\
      \midrule
      $\textsf{Ace} + \texttt{cd05}$ & 27 & 1247 \\
      $\textsf{Ace} + \texttt{cd06}$ & 135 & 1340 \\
      $\textsf{Ace} + \texttt{d02}$ & 56 & 1060 \\
      $\textsf{DPMC} + \texttt{bklm16}$ & 241 & 1327 \\
      $\textsf{DPMC} + \texttt{bklm16++}$ & \textbf{992} & \textbf{1435} \\
      $\textsf{DPMC} + \texttt{cd05++}$ & \textcolor{gray}{0} & 867 \\
      $\textsf{DPMC} + \texttt{cd06++}$ & \textcolor{gray}{0} & 932 \\
      $\textsf{DPMC} + \texttt{d02}$ & 1 & 1267 \\
      $\textsf{DPMC} + \texttt{d02++}$ & 7 & 1272 \\
      $\textsf{DPMC} + \texttt{sbk05}$ & 31 & 1308 \\
      $\textsf{c2d} + \texttt{bklm16}$ & \textcolor{gray}{0} & 997 \\
      $\textsf{Cachet} + \texttt{sbk05}$ & 49 & 983 \\
      \bottomrule
    \end{tabular}
  \end{minipage}
\end{figure}

\Cref{fig:cumulative} shows $\textsf{DPMC}+\texttt{bklm16++}$ to be the
best-performing combination across all time limits up to \SI{1000}{\second} with
$\textsf{Ace} + \texttt{cd06}$ and $\textsf{DPMC}+\texttt{bklm16}$ not far
behind. Overall, $\textsf{DPMC}+\texttt{bklm16++}$ is 3.35 times faster than
$\textsf{DPMC}+\texttt{bklm16}$ and 2.96 times faster than
$\textsf{Ace}+\texttt{cd06}$. \Cref{tbl:performance} further shows that
$\textsf{DPMC}+\texttt{bklm16++}$ solves almost a hundred more instances than
any other combination, and is the fastest in \SI{69.1}{\percent} of them.

The scatter plots in \cref{fig:scatter} show that how $\textsf{DPMC} +
\texttt{bklm16++}$ (and perhaps \textsf{DPMC} more generally) compares to
$\textsf{Ace} + \texttt{cd06}$ depends significantly on the data set: the former
is a clear winner on DQMR and Grid instances, while the latter performs well on
Mastermind and Random Blocks. Perhaps because the underlying WMC algorithm
remains the same, the difference between $\textsf{DPMC} + \texttt{bklm16}$ with
and without applying \cref{alg:transformation} is quite noisy, i.e, with most
instances scattered around the line of equality. However, our transformation
does enable \textsf{DPMC} to solve many instances that were previously beyond
its reach.

We also record numbers of variables in each encoding before and after applying
\cref{alg:transformation}. \Cref{fig:box} shows a significant reduction in the
number of variables. For instance, the median number of variables in instances
encoded with \texttt{bklm16} was reduced four times: from 1499 to 376. While
\texttt{bklm16++} results in the overall lowest number of variables, the
difference between \texttt{bklm16++} and \texttt{d02++} seems small. Indeed, the
numbers of variables in these two encodings are equal for binary Bayesian
networks (i.e., most of our data). Nonetheless, \texttt{bklm16++} is still much
faster than \texttt{d02++} when run with \textsf{DPMC}.

Overall, transforming WMC instances to the PBP format allows us to significantly
simplify each instance. This transformation is particularly effective on
\texttt{bklm16}, allowing it to surpass \texttt{cd06} and become the new state
of the art. While there is a similarly significant reduction in the number of
variables for \texttt{d02}, the performance of $\textsf{DPMC}+\texttt{d02}$ is
virtually unaffected. Finally, while our transformation makes it possible to use
\texttt{cd05} and \texttt{cd06} with \textsf{DPMC}, the two combinations remain
inefficient.

\section{Conclusion} % So what?

In this paper, we showed how the number of variables in a WMC instance can be
significantly reduced by transforming it into a representation based on
two-valued pseudo-Boolean functions. In some cases, this led to significant
improvements in inference speed, allowing $\textsf{DPMC} + \texttt{bklm16++}$ to
overtake $\textsf{Ace}+\texttt{cd06}$ as the new state of the art WMC technique
for Bayesian network inference. Moreover, we identified key properties of
Bayesian network encodings that allow for parameter variable removal. However,
these properties were rather different for each encoding, and so an interesting
question for future work is whether they can be unified into a more abstract
and coherent list of conditions.

Bayesian network inference was chosen as the example application of WMC because
it is the first and the most studied one
\cite{DBLP:conf/ecai/BartKLM16,DBLP:conf/ijcai/ChaviraD05,DBLP:conf/sat/ChaviraD06,DBLP:conf/kr/Darwiche02,DBLP:conf/aaai/SangBK05}.
While the distinction between indicator and parameter variables is often not
explicitly described in other WMC encodings
\cite{DBLP:journals/tplp/FierensBRSGTJR15,DBLP:journals/pacmpl/HoltzenBM20,DBLP:conf/icml/XuZFLB18},
perhaps in some cases variables could still be partitioned in this way, allowing
for not just faster inference with \textsf{DPMC} or \textsf{ADDMC} but also for
well-established WMC encoding and inference techniques (such as in the
\texttt{cd05} and \texttt{cd06} papers
\cite{DBLP:conf/ijcai/ChaviraD05,DBLP:conf/sat/ChaviraD06}) to be transferred to
other application domains.

\bibliographystyle{splncs04}
\bibliography{paper}

\end{document}