### 2021-06-02 doc/long_talk/scatter1.tex
Slides are almost ready

% Created by tikzDevice version 0.12.3.1 on 2021-06-01 14:27:09
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (343.28,231.26);
\begin{scope}
\path[clip] (  0.00,  0.09) rectangle (343.28,231.18);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.09) rectangle (343.28,231.18);
\end{scope}
\begin{scope}
\path[clip] ( 40.51, 30.77) rectangle (235.42,225.68);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 40.51, 30.77) rectangle (235.42,225.68);
\definecolor{drawColor}{gray}{0.87}

\path[draw=drawColor,line width= 0.1pt,line join=round] ( 40.51, 39.63) --
	(235.42, 39.63);

\path[draw=drawColor,line width= 0.1pt,line join=round] ( 40.51,106.48) --
	(235.42,106.48);

\path[draw=drawColor,line width= 0.1pt,line join=round] ( 40.51,173.33) --
	(235.42,173.33);

\path[draw=drawColor,line width= 0.1pt,line join=round] ( 49.37, 30.77) --
	( 49.37,225.68);

\path[draw=drawColor,line width= 0.1pt,line join=round] (116.22, 30.77) --
	(116.22,225.68);

\path[draw=drawColor,line width= 0.1pt,line join=round] (183.07, 30.77) --
	(183.07,225.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 40.51, 73.06) --
	(235.42, 73.06);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 40.51,139.91) --
	(235.42,139.91);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 40.51,206.76) --
	(235.42,206.76);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 82.79, 30.77) --
	( 82.79,225.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] (149.64, 30.77) --
	(149.64,225.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] (216.49, 30.77) --
	(216.49,225.68);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.18,215.39) -- (136.04,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.18,218.25) -- (136.04,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.59,216.82) -- (136.63,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.61,214.80) -- (134.61,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.21,215.39) -- (137.06,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.21,218.25) -- (137.06,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.62,216.82) -- (137.65,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.64,214.80) -- (135.64,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.25,215.39) -- (137.10,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.25,218.25) -- (137.10,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.66,216.82) -- (137.69,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.68,214.80) -- (135.68,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.53,215.39) -- (140.39,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.53,218.25) -- (140.39,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.94,216.82) -- (140.98,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.96,214.80) -- (138.96,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.55,215.39) -- (136.40,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.55,218.25) -- (136.40,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.96,216.82) -- (136.99,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.97,214.80) -- (134.97,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.29,215.39) -- (137.14,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.29,218.25) -- (137.14,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.70,216.82) -- (137.73,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.71,214.80) -- (135.71,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.13,101.07) -- (137.99,103.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.13,103.92) -- (137.99,101.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.54,102.50) -- (138.58,102.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.56,100.48) -- (136.56,104.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.94,101.26) -- (136.79,104.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.94,104.11) -- (136.79,101.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.35,102.69) -- (137.39,102.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.37,100.67) -- (135.37,104.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.93,101.45) -- (138.79,104.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.93,104.30) -- (138.79,101.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.34,102.87) -- (139.38,102.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.36,100.86) -- (137.36,104.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.84, 97.15) -- (133.69,100.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.84,100.00) -- (133.69, 97.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.25, 98.57) -- (134.28, 98.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.26, 96.56) -- (132.26,100.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.59, 98.35) -- (136.44,101.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.59,101.20) -- (136.44, 98.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.00, 99.77) -- (137.03, 99.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.01, 97.76) -- (135.01,101.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.67,215.39) -- (136.52,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.67,218.25) -- (136.52,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.08,216.82) -- (137.11,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.09,214.80) -- (135.09,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.20,215.39) -- (140.05,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.20,218.25) -- (140.05,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.61,216.82) -- (140.64,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.62,214.80) -- (138.62,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.51,215.39) -- (136.36,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.51,218.25) -- (136.36,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.92,216.82) -- (136.95,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.93,214.80) -- (134.93,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.36,215.39) -- (137.22,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.36,218.25) -- (137.22,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.77,216.82) -- (137.81,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.79,214.80) -- (135.79,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.27,215.39) -- (136.12,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.27,218.25) -- (136.12,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.68,216.82) -- (136.71,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.69,214.80) -- (134.69,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.10,112.67) -- (139.96,115.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.10,115.53) -- (139.96,112.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.51,114.10) -- (140.55,114.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.53,112.08) -- (138.53,116.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.49,106.70) -- (138.34,109.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.49,109.55) -- (138.34,106.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.90,108.13) -- (138.93,108.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.91,106.11) -- (136.91,110.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.90,106.57) -- (138.75,109.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.90,109.42) -- (138.75,106.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.31,108.00) -- (139.34,108.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.33,105.98) -- (137.33,110.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.10,105.90) -- (137.95,108.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.10,108.75) -- (137.95,105.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.51,107.33) -- (138.54,107.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.52,105.31) -- (136.52,109.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.50,107.33) -- (140.36,110.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.50,110.19) -- (140.36,107.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.91,108.76) -- (140.95,108.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.93,106.74) -- (138.93,110.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.69,106.57) -- (139.54,109.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.69,109.42) -- (139.54,106.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.10,108.00) -- (140.14,108.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.12,105.98) -- (138.12,110.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.66,100.08) -- (137.51,102.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.66,102.94) -- (137.51,100.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.07,101.51) -- (138.10,101.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.09, 99.49) -- (136.09,103.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.92,101.07) -- (137.77,103.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.92,103.92) -- (137.77,101.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.33,102.50) -- (138.36,102.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.34,100.48) -- (136.34,104.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.94,101.07) -- (135.79,103.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.94,103.92) -- (135.79,101.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.35,102.50) -- (136.38,102.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.36,100.48) -- (134.36,104.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.23,104.46) -- (139.09,107.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.23,107.32) -- (139.09,104.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.64,105.89) -- (139.68,105.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.66,103.87) -- (137.66,107.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.31,102.17) -- (136.16,105.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.31,105.03) -- (136.16,102.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.72,103.60) -- (136.75,103.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.73,101.58) -- (134.73,105.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.44,106.17) -- (137.29,109.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.44,109.03) -- (137.29,106.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.85,107.60) -- (137.88,107.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.86,105.58) -- (135.86,109.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.07,106.44) -- (138.92,109.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.07,109.29) -- (138.92,106.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.48,107.86) -- (139.51,107.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.49,105.85) -- (137.49,109.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.29,107.94) -- (140.14,110.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.29,110.79) -- (140.14,107.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.70,109.37) -- (140.73,109.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.72,107.35) -- (138.72,111.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.83,107.58) -- (140.69,110.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.83,110.43) -- (140.69,107.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.24,109.01) -- (141.28,109.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.26,106.99) -- (139.26,111.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.85,107.33) -- (139.70,110.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.85,110.19) -- (139.70,107.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.26,108.76) -- (140.29,108.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.28,106.74) -- (138.28,110.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.69,112.24) -- (139.54,115.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.69,115.09) -- (139.54,112.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.10,113.66) -- (140.14,113.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.12,111.64) -- (138.12,115.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.17,111.04) -- (140.02,113.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.17,113.89) -- (140.02,111.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.57,112.46) -- (140.61,112.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.59,110.45) -- (138.59,114.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.97,114.75) -- (138.82,117.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.97,117.60) -- (138.82,114.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.38,116.18) -- (139.41,116.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.39,114.16) -- (137.39,118.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.50,112.67) -- (139.35,115.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.50,115.53) -- (139.35,112.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.91,114.10) -- (139.94,114.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.92,112.08) -- (137.92,116.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.95,116.50) -- (140.80,119.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.95,119.35) -- (140.80,116.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.36,117.93) -- (141.40,117.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.38,115.91) -- (139.38,119.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.19,111.42) -- (141.04,114.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.19,114.27) -- (141.04,111.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.59,112.84) -- (141.63,112.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.61,110.83) -- (139.61,114.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.10,111.60) -- (140.95,114.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.10,114.46) -- (140.95,111.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.51,113.03) -- (141.54,113.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.52,111.01) -- (139.52,115.05);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.35,105.05) -- (138.20,107.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.35,107.91) -- (138.20,105.05);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.76,106.48) -- (138.79,106.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.77,104.46) -- (136.77,108.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.17,107.70) -- (140.02,110.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.17,110.55) -- (140.02,107.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.57,109.13) -- (140.61,109.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.59,107.11) -- (138.59,111.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.35,112.59) -- (140.20,115.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.35,115.44) -- (140.20,112.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.76,114.01) -- (140.79,114.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.78,111.99) -- (138.78,116.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.63,105.48) -- (139.48,108.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.63,108.34) -- (139.48,105.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.04,106.91) -- (140.07,106.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.05,104.89) -- (138.05,108.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.81,121.10) -- (141.66,123.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.81,123.95) -- (141.66,121.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.22,122.53) -- (142.25,122.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.24,120.51) -- (140.24,124.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.37,116.43) -- (139.22,119.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.37,119.29) -- (139.22,116.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.78,117.86) -- (139.81,117.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.79,115.84) -- (137.79,119.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.35,109.84) -- (138.20,112.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.35,112.69) -- (138.20,109.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.76,111.26) -- (138.79,111.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.77,109.24) -- (136.77,113.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.69,109.19) -- (139.54,112.05);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.69,112.05) -- (139.54,109.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.10,110.62) -- (140.14,110.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.12,108.60) -- (138.12,112.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.24,111.23) -- (138.09,114.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.24,114.08) -- (138.09,111.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.65,112.65) -- (138.68,112.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.67,110.64) -- (136.67,114.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.80,215.39) -- (153.66,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.80,218.25) -- (153.66,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.21,216.82) -- (154.25,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.23,214.80) -- (152.23,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.64,215.39) -- (148.49,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.64,218.25) -- (148.49,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.04,216.82) -- (149.08,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.06,214.80) -- (147.06,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.10,215.39) -- (149.96,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.10,218.25) -- (149.96,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.51,216.82) -- (150.55,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.53,214.80) -- (148.53,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.45,215.39) -- (157.30,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.45,218.25) -- (157.30,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.86,216.82) -- (157.89,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.87,214.80) -- (155.87,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.66,215.39) -- (150.51,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.66,218.25) -- (150.51,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.06,216.82) -- (151.10,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.08,214.80) -- (149.08,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.72,215.39) -- (155.58,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.72,218.25) -- (155.58,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.13,216.82) -- (156.17,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.15,214.80) -- (154.15,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.93,187.58) -- (149.78,190.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.93,190.43) -- (149.78,187.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.34,189.01) -- (150.37,189.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.36,186.99) -- (148.36,191.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.84,155.90) -- (150.69,158.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.84,158.76) -- (150.69,155.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.24,157.33) -- (151.28,157.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.26,155.31) -- (149.26,159.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.32,152.68) -- (148.17,155.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.32,155.53) -- (148.17,152.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.73,154.11) -- (148.76,154.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.75,152.09) -- (146.75,156.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.91,156.31) -- (150.76,159.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.91,159.16) -- (150.76,156.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.32,157.74) -- (151.35,157.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.34,155.72) -- (149.34,159.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.40,153.70) -- (149.25,156.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.40,156.55) -- (149.25,153.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.80,155.12) -- (149.84,155.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.82,153.11) -- (147.82,157.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.31,215.39) -- (154.16,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.31,218.25) -- (154.16,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.71,216.82) -- (154.75,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.73,214.80) -- (152.73,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.07,215.39) -- (153.92,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.07,218.25) -- (153.92,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.48,216.82) -- (154.51,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.50,214.80) -- (152.50,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.33,215.39) -- (154.18,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.33,218.25) -- (154.18,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.74,216.82) -- (154.77,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.76,214.80) -- (152.76,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.18,215.39) -- (154.03,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.18,218.25) -- (154.03,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.58,216.82) -- (154.62,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.60,214.80) -- (152.60,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.62,215.39) -- (154.47,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.62,218.25) -- (154.47,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.03,216.82) -- (155.06,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.05,214.80) -- (153.05,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.58,215.39) -- (154.44,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.58,218.25) -- (154.44,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.99,216.82) -- (155.03,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.01,214.80) -- (153.01,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.61,215.39) -- (150.46,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.61,218.25) -- (150.46,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.02,216.82) -- (151.05,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.04,214.80) -- (149.04,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.95,215.39) -- (151.81,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.95,218.25) -- (151.81,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.36,216.82) -- (152.40,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.38,214.80) -- (150.38,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.60,215.39) -- (156.46,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.60,218.25) -- (156.46,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.01,216.82) -- (157.05,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.03,214.80) -- (155.03,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.19,215.39) -- (154.04,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.19,218.25) -- (154.04,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.60,216.82) -- (154.63,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.61,214.80) -- (152.61,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.03,215.39) -- (153.89,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.03,218.25) -- (153.89,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.44,216.82) -- (154.48,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.46,214.80) -- (152.46,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.82,155.61) -- (149.67,158.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.82,158.47) -- (149.67,155.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.23,157.04) -- (150.26,157.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.24,155.02) -- (148.24,159.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.88,158.55) -- (151.74,161.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.88,161.41) -- (151.74,158.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.29,159.98) -- (152.33,159.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.31,157.96) -- (150.31,162.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.56,154.94) -- (150.42,157.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.56,157.79) -- (150.42,154.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.97,156.37) -- (151.01,156.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.99,154.35) -- (148.99,158.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.40,156.62) -- (149.25,159.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.40,159.47) -- (149.25,156.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.80,158.04) -- (149.84,158.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.82,156.02) -- (147.82,160.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.90,155.64) -- (149.75,158.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.90,158.49) -- (149.75,155.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.31,157.07) -- (150.34,157.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.32,155.05) -- (148.32,159.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.74,215.39) -- (154.60,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.74,218.25) -- (154.60,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.15,216.82) -- (155.19,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.17,214.80) -- (153.17,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.20,215.39) -- (155.06,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.20,218.25) -- (155.06,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.61,216.82) -- (155.65,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.63,214.80) -- (153.63,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.07,215.39) -- (154.92,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.07,218.25) -- (154.92,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.48,216.82) -- (155.52,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.50,214.80) -- (153.50,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.91,215.39) -- (153.77,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.91,218.25) -- (153.77,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.32,216.82) -- (154.36,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.34,214.80) -- (152.34,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.91,215.39) -- (154.77,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.91,218.25) -- (154.77,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.32,216.82) -- (155.36,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.34,214.80) -- (153.34,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.52,215.39) -- (155.37,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.52,218.25) -- (155.37,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.93,216.82) -- (155.96,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.95,214.80) -- (153.95,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.80,215.39) -- (158.65,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.80,218.25) -- (158.65,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.21,216.82) -- (159.25,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.23,214.80) -- (157.23,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.39,215.39) -- (158.24,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.39,218.25) -- (158.24,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.80,216.82) -- (158.83,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (156.82,214.80) -- (156.82,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.52,215.39) -- (155.37,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.52,218.25) -- (155.37,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.93,216.82) -- (155.96,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.95,214.80) -- (153.95,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.41,215.39) -- (156.26,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.41,218.25) -- (156.26,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.82,216.82) -- (156.85,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.84,214.80) -- (154.84,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.91,215.39) -- (158.77,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.91,218.25) -- (158.77,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.32,216.82) -- (159.36,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.34,214.80) -- (157.34,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.28,215.39) -- (155.13,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.28,218.25) -- (155.13,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.69,216.82) -- (155.73,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.71,214.80) -- (153.71,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.76,160.28) -- (152.61,163.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.76,163.14) -- (152.61,160.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.17,161.71) -- (153.20,161.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.19,159.69) -- (151.19,163.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.52,165.91) -- (152.38,168.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.52,168.77) -- (152.38,165.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.93,167.34) -- (152.97,167.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.95,165.32) -- (150.95,169.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.90,159.28) -- (150.75,162.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.90,162.13) -- (150.75,159.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.30,160.71) -- (151.34,160.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.32,158.69) -- (149.32,162.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.85,157.91) -- (150.70,160.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.85,160.77) -- (150.70,157.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.26,159.34) -- (151.29,159.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.28,157.32) -- (149.28,161.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.88,215.39) -- (154.73,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.88,218.25) -- (154.73,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.29,216.82) -- (155.33,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.31,214.80) -- (153.31,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.32,215.39) -- (156.17,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.32,218.25) -- (156.17,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.73,216.82) -- (156.76,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.75,214.80) -- (154.75,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.05,215.39) -- (156.90,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.05,218.25) -- (156.90,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.45,216.82) -- (157.49,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.47,214.80) -- (155.47,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.23,215.39) -- (155.08,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.23,218.25) -- (155.08,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.64,216.82) -- (155.67,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.65,214.80) -- (153.65,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.10,215.39) -- (156.96,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.10,218.25) -- (156.96,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.51,216.82) -- (157.55,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.53,214.80) -- (155.53,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.96,215.39) -- (165.81,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.96,218.25) -- (165.81,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.36,216.82) -- (166.40,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.38,214.80) -- (164.38,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.86,215.39) -- (164.71,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.86,218.25) -- (164.71,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.27,216.82) -- (165.30,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.28,214.80) -- (163.28,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.38,215.39) -- (165.23,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.38,218.25) -- (165.23,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.79,216.82) -- (165.83,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.81,214.80) -- (163.81,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.04,215.39) -- (164.90,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.04,218.25) -- (164.90,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.45,216.82) -- (165.49,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.47,214.80) -- (163.47,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.77,215.39) -- (166.63,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.77,218.25) -- (166.63,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.18,216.82) -- (167.22,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.20,214.80) -- (165.20,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.17,215.39) -- (165.03,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.17,218.25) -- (165.03,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.58,216.82) -- (165.62,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.60,214.80) -- (163.60,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.06,215.39) -- (159.92,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.06,218.25) -- (159.92,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (156.47,216.82) -- (160.51,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.49,214.80) -- (158.49,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.37,195.92) -- (164.22,198.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.37,198.78) -- (164.22,195.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.78,197.35) -- (164.81,197.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.79,195.33) -- (162.79,199.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.14,215.39) -- (159.99,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.14,218.25) -- (159.99,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (156.55,216.82) -- (160.58,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.57,214.80) -- (158.57,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.79,215.39) -- (161.64,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.79,218.25) -- (161.64,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.20,216.82) -- (162.23,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.21,214.80) -- (160.21,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.48,215.39) -- (163.34,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.48,218.25) -- (163.34,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.89,216.82) -- (163.93,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.91,214.80) -- (161.91,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.60,215.39) -- (164.45,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.60,218.25) -- (164.45,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.01,216.82) -- (165.04,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.03,214.80) -- (163.03,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.61,215.39) -- (165.46,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.61,218.25) -- (165.46,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.02,216.82) -- (166.05,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.04,214.80) -- (164.04,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.04,215.39) -- (164.89,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.04,218.25) -- (164.89,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.45,216.82) -- (165.48,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.46,214.80) -- (163.46,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.48,215.39) -- (164.34,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.48,218.25) -- (164.34,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.89,216.82) -- (164.93,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.91,214.80) -- (162.91,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.31,215.39) -- (166.16,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.31,218.25) -- (166.16,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.72,216.82) -- (166.75,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.74,214.80) -- (164.74,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.69,215.39) -- (167.54,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.69,218.25) -- (167.54,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.10,216.82) -- (168.13,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.12,214.80) -- (166.12,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.44,215.39) -- (167.29,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.44,218.25) -- (167.29,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.85,216.82) -- (167.88,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.87,214.80) -- (165.87,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.10,215.39) -- (165.95,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.10,218.25) -- (165.95,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.51,216.82) -- (166.54,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.52,214.80) -- (164.52,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.10,215.39) -- (168.95,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.10,218.25) -- (168.95,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.51,216.82) -- (169.54,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.52,214.80) -- (167.52,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.82,215.39) -- (166.67,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.82,218.25) -- (166.67,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.23,216.82) -- (167.26,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.24,214.80) -- (165.24,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.42,215.39) -- (167.27,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.42,218.25) -- (167.27,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.82,216.82) -- (167.86,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.84,214.80) -- (165.84,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.81,188.01) -- (163.66,190.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.81,190.86) -- (163.66,188.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.21,189.44) -- (164.25,189.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.23,187.42) -- (162.23,191.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.37,195.89) -- (161.22,198.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.37,198.74) -- (161.22,195.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.78,197.32) -- (161.81,197.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.79,195.30) -- (159.79,199.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.85,185.93) -- (162.71,188.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.85,188.79) -- (162.71,185.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.26,187.36) -- (163.30,187.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.28,185.34) -- (161.28,189.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.47,194.82) -- (162.32,197.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.47,197.67) -- (162.32,194.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.88,196.24) -- (162.92,196.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.90,194.23) -- (160.90,198.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.91,215.39) -- (162.76,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.91,218.25) -- (162.76,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.31,216.82) -- (163.35,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.33,214.80) -- (161.33,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.48,215.39) -- (168.34,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.48,218.25) -- (168.34,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.89,216.82) -- (168.93,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.91,214.80) -- (166.91,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.02,215.39) -- (167.87,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.02,218.25) -- (167.87,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.43,216.82) -- (168.47,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.45,214.80) -- (166.45,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.47,215.39) -- (166.33,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.47,218.25) -- (166.33,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.88,216.82) -- (166.92,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.90,214.80) -- (164.90,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.31,215.39) -- (166.16,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.31,218.25) -- (166.16,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.72,216.82) -- (166.75,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.74,214.80) -- (164.74,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.57,215.39) -- (167.42,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.57,218.25) -- (167.42,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.98,216.82) -- (168.01,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.99,214.80) -- (165.99,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.26,215.39) -- (168.11,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.26,218.25) -- (168.11,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.67,216.82) -- (168.70,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.68,214.80) -- (166.68,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.01,215.39) -- (169.87,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.01,218.25) -- (169.87,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.42,216.82) -- (170.46,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.44,214.80) -- (168.44,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.35,215.39) -- (167.20,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.35,218.25) -- (167.20,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.76,216.82) -- (167.79,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.78,214.80) -- (165.78,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.91,215.39) -- (170.77,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.91,218.25) -- (170.77,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.32,216.82) -- (171.36,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.34,214.80) -- (169.34,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.68,215.39) -- (168.53,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.68,218.25) -- (168.53,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.09,216.82) -- (169.13,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.11,214.80) -- (167.11,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.34,215.39) -- (169.19,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.34,218.25) -- (169.19,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.75,216.82) -- (169.78,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.76,214.80) -- (167.76,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.84,215.39) -- (167.70,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.84,218.25) -- (167.70,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.25,216.82) -- (168.29,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.27,214.80) -- (166.27,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.10,215.39) -- (160.96,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.10,218.25) -- (160.96,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.51,216.82) -- (161.55,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.53,214.80) -- (159.53,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.81,192.67) -- (161.66,195.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.81,195.52) -- (161.66,192.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.22,194.09) -- (162.25,194.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.23,192.07) -- (160.23,196.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.37,187.64) -- (160.23,190.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.37,190.50) -- (160.23,187.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (156.78,189.07) -- (160.82,189.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.80,187.05) -- (158.80,191.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.12,187.48) -- (159.98,190.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.12,190.33) -- (159.98,187.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (156.53,188.90) -- (160.57,188.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.55,186.88) -- (158.55,190.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.31,215.39) -- (169.16,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.31,218.25) -- (169.16,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.72,216.82) -- (169.75,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.73,214.80) -- (167.73,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.86,215.39) -- (167.71,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.86,218.25) -- (167.71,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.26,216.82) -- (168.30,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.28,214.80) -- (166.28,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.92,215.39) -- (166.78,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.92,218.25) -- (166.78,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.33,216.82) -- (167.37,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.35,214.80) -- (165.35,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.78,215.39) -- (166.64,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.78,218.25) -- (166.64,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.19,216.82) -- (167.23,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.21,214.80) -- (165.21,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.87,215.39) -- (167.72,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.87,218.25) -- (167.72,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.28,216.82) -- (168.31,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.30,214.80) -- (166.30,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.39,215.39) -- (182.24,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.39,218.25) -- (182.24,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.80,216.82) -- (182.83,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.81,214.80) -- (180.81,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.94,215.39) -- (182.80,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.94,218.25) -- (182.80,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.35,216.82) -- (183.39,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.37,214.80) -- (181.37,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.39,215.39) -- (182.25,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.39,218.25) -- (182.25,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.80,216.82) -- (182.84,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.82,214.80) -- (180.82,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.02,215.39) -- (181.87,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.02,218.25) -- (181.87,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.42,216.82) -- (182.46,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.44,214.80) -- (180.44,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.46,215.39) -- (181.32,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.46,218.25) -- (181.32,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.87,216.82) -- (181.91,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.89,214.80) -- (179.89,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.05,215.39) -- (181.90,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.05,218.25) -- (181.90,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.46,216.82) -- (182.49,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.47,214.80) -- (180.47,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.47,215.39) -- (172.32,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.47,218.25) -- (172.32,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.88,216.82) -- (172.91,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.90,214.80) -- (170.90,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.88,215.39) -- (180.73,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.88,218.25) -- (180.73,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.28,216.82) -- (181.32,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.30,214.80) -- (179.30,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.37,215.39) -- (173.22,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.37,218.25) -- (173.22,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.78,216.82) -- (173.82,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.80,214.80) -- (171.80,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.52,215.39) -- (174.38,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.52,218.25) -- (174.38,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.93,216.82) -- (174.97,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.95,214.80) -- (172.95,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.60,215.39) -- (171.46,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.60,218.25) -- (171.46,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.01,216.82) -- (172.05,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.03,214.80) -- (170.03,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.34,215.39) -- (181.20,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.34,218.25) -- (181.20,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.75,216.82) -- (181.79,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.77,214.80) -- (179.77,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.07,215.39) -- (181.93,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.07,218.25) -- (181.93,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.48,216.82) -- (182.52,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.50,214.80) -- (180.50,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.40,215.39) -- (180.25,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.40,218.25) -- (180.25,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.81,216.82) -- (180.85,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.83,214.80) -- (178.83,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.76,215.39) -- (180.62,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.76,218.25) -- (180.62,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.17,216.82) -- (181.21,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.19,214.80) -- (179.19,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.41,215.39) -- (181.27,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.41,218.25) -- (181.27,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.82,216.82) -- (181.86,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.84,214.80) -- (179.84,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.84,215.39) -- (182.70,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.84,218.25) -- (182.70,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.25,216.82) -- (183.29,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.27,214.80) -- (181.27,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.97,215.39) -- (182.83,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.97,218.25) -- (182.83,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.38,216.82) -- (183.42,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.40,214.80) -- (181.40,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.87,215.39) -- (184.72,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.87,218.25) -- (184.72,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.28,216.82) -- (185.31,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (183.30,214.80) -- (183.30,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (182.43,215.39) -- (185.29,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (182.43,218.25) -- (185.29,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.84,216.82) -- (185.88,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (183.86,214.80) -- (183.86,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.63,215.39) -- (182.49,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.63,218.25) -- (182.49,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.04,216.82) -- (183.08,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.06,214.80) -- (181.06,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.94,215.39) -- (182.79,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.94,218.25) -- (182.79,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.35,216.82) -- (183.38,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.37,214.80) -- (181.37,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.05,215.39) -- (173.90,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.05,218.25) -- (173.90,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.46,216.82) -- (174.49,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.47,214.80) -- (172.47,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (173.79,215.39) -- (176.65,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (173.79,218.25) -- (176.65,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (173.20,216.82) -- (177.24,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (175.22,214.80) -- (175.22,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.94,215.39) -- (172.79,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.94,218.25) -- (172.79,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.35,216.82) -- (173.38,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.37,214.80) -- (171.37,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.69,215.39) -- (171.54,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.69,218.25) -- (171.54,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.10,216.82) -- (172.13,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.11,214.80) -- (170.11,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.75,215.39) -- (174.60,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.75,218.25) -- (174.60,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.16,216.82) -- (175.19,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (173.18,214.80) -- (173.18,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.30,215.39) -- (182.15,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.30,218.25) -- (182.15,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.71,216.82) -- (182.74,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.72,214.80) -- (180.72,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.06,215.39) -- (181.91,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.06,218.25) -- (181.91,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.47,216.82) -- (182.50,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.48,214.80) -- (180.48,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.85,215.39) -- (180.71,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.85,218.25) -- (180.71,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.26,216.82) -- (181.30,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.28,214.80) -- (179.28,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.09,215.39) -- (181.95,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.09,218.25) -- (181.95,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.50,216.82) -- (182.54,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.52,214.80) -- (180.52,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.28,215.39) -- (182.13,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.28,218.25) -- (182.13,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.68,216.82) -- (182.72,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.70,214.80) -- (180.70,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.36,215.39) -- (184.21,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.36,218.25) -- (184.21,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.77,216.82) -- (184.80,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (182.79,214.80) -- (182.79,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.70,215.39) -- (183.55,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.70,218.25) -- (183.55,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.11,216.82) -- (184.14,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (182.12,214.80) -- (182.12,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.52,215.39) -- (184.37,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.52,218.25) -- (184.37,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.93,216.82) -- (184.96,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (182.94,214.80) -- (182.94,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (182.73,215.39) -- (185.59,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (182.73,218.25) -- (185.59,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (182.14,216.82) -- (186.18,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (184.16,214.80) -- (184.16,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.51,215.39) -- (184.37,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.51,218.25) -- (184.37,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.92,216.82) -- (184.96,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (182.94,214.80) -- (182.94,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.51,215.39) -- (184.36,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.51,218.25) -- (184.36,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.92,216.82) -- (184.95,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (182.94,214.80) -- (182.94,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.23,215.39) -- (175.08,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.23,218.25) -- (175.08,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.64,216.82) -- (175.67,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (173.66,214.80) -- (173.66,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.80,215.39) -- (173.65,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.80,218.25) -- (173.65,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.21,216.82) -- (174.24,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.22,214.80) -- (172.22,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.01,215.39) -- (172.87,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.01,218.25) -- (172.87,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.42,216.82) -- (173.46,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.44,214.80) -- (171.44,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (173.47,215.39) -- (176.32,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (173.47,218.25) -- (176.32,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.88,216.82) -- (176.91,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (174.90,214.80) -- (174.90,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.48,215.39) -- (175.34,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.48,218.25) -- (175.34,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.89,216.82) -- (175.93,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (173.91,214.80) -- (173.91,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.48,215.39) -- (182.33,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.48,218.25) -- (182.33,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.89,216.82) -- (182.92,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.91,214.80) -- (180.91,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (182.61,215.39) -- (185.47,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (182.61,218.25) -- (185.47,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (182.02,216.82) -- (186.06,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (184.04,214.80) -- (184.04,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.15,215.39) -- (183.01,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.15,218.25) -- (183.01,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.56,216.82) -- (183.60,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.58,214.80) -- (181.58,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.13,215.39) -- (182.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.13,218.25) -- (182.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.54,216.82) -- (183.58,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.56,214.80) -- (181.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.73,215.39) -- (182.59,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.73,218.25) -- (182.59,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.14,216.82) -- (183.18,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.16,214.80) -- (181.16,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.36,215.39) -- (189.22,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.36,218.25) -- (189.22,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (185.77,216.82) -- (189.81,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.79,214.80) -- (187.79,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (185.37,215.39) -- (188.23,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (185.37,218.25) -- (188.23,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (184.78,216.82) -- (188.82,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.80,214.80) -- (186.80,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.11,215.39) -- (189.97,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.11,218.25) -- (189.97,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.52,216.82) -- (190.56,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.54,214.80) -- (188.54,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.77,215.39) -- (190.63,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.77,218.25) -- (190.63,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.18,216.82) -- (191.22,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (189.20,214.80) -- (189.20,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (185.34,215.39) -- (188.20,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (185.34,218.25) -- (188.20,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (184.75,216.82) -- (188.79,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.77,214.80) -- (186.77,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.09,215.39) -- (189.94,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.09,218.25) -- (189.94,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.49,216.82) -- (190.53,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.51,214.80) -- (188.51,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (174.24,215.39) -- (177.09,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (174.24,218.25) -- (177.09,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (173.65,216.82) -- (177.68,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (175.67,214.80) -- (175.67,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (183.38,215.39) -- (186.24,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (183.38,218.25) -- (186.24,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (182.79,216.82) -- (186.83,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (184.81,214.80) -- (184.81,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (175.48,215.39) -- (178.33,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (175.48,218.25) -- (178.33,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (174.89,216.82) -- (178.93,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.91,214.80) -- (176.91,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.41,215.39) -- (179.26,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.41,218.25) -- (179.26,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (175.81,216.82) -- (179.85,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.83,214.80) -- (177.83,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.07,215.39) -- (178.92,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.07,218.25) -- (178.92,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (175.48,216.82) -- (179.51,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.49,214.80) -- (177.49,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.61,215.39) -- (190.47,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.61,218.25) -- (190.47,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.02,216.82) -- (191.06,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (189.04,214.80) -- (189.04,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (185.51,215.39) -- (188.37,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (185.51,218.25) -- (188.37,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (184.92,216.82) -- (188.96,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.94,214.80) -- (186.94,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.26,215.39) -- (190.12,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.26,218.25) -- (190.12,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.67,216.82) -- (190.71,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.69,214.80) -- (188.69,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (185.51,215.39) -- (188.36,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (185.51,218.25) -- (188.36,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (184.92,216.82) -- (188.95,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.93,214.80) -- (186.93,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (185.80,215.39) -- (188.66,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (185.80,218.25) -- (188.66,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (185.21,216.82) -- (189.25,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.23,214.80) -- (187.23,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (189.78,215.39) -- (192.64,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (189.78,218.25) -- (192.64,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (189.19,216.82) -- (193.23,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (191.21,214.80) -- (191.21,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (189.58,215.39) -- (192.44,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (189.58,218.25) -- (192.44,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.99,216.82) -- (193.03,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (191.01,214.80) -- (191.01,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.78,215.39) -- (191.64,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.78,218.25) -- (191.64,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.19,216.82) -- (192.23,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (190.21,214.80) -- (190.21,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (190.43,215.39) -- (193.29,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (190.43,218.25) -- (193.29,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (189.84,216.82) -- (193.88,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (191.86,214.80) -- (191.86,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (189.01,215.39) -- (191.86,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (189.01,218.25) -- (191.86,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.42,216.82) -- (192.45,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (190.44,214.80) -- (190.44,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.98,215.39) -- (191.83,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.98,218.25) -- (191.83,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.38,216.82) -- (192.42,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (190.40,214.80) -- (190.40,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (175.92,215.39) -- (178.77,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (175.92,218.25) -- (178.77,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (175.33,216.82) -- (179.36,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.34,214.80) -- (177.34,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.57,215.39) -- (179.43,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.57,218.25) -- (179.43,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (175.98,216.82) -- (180.02,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.00,214.80) -- (178.00,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.92,215.39) -- (179.77,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.92,218.25) -- (179.77,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.33,216.82) -- (180.36,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.35,214.80) -- (178.35,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.14,215.39) -- (179.99,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.14,218.25) -- (179.99,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.55,216.82) -- (180.58,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.56,214.80) -- (178.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (175.92,215.39) -- (178.78,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (175.92,218.25) -- (178.78,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (175.33,216.82) -- (179.37,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.35,214.80) -- (177.35,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.61,215.39) -- (189.46,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.61,218.25) -- (189.46,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.01,216.82) -- (190.05,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.03,214.80) -- (188.03,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.75,215.39) -- (189.60,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.75,218.25) -- (189.60,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.16,216.82) -- (190.20,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.18,214.80) -- (188.18,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.67,215.39) -- (189.53,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.67,218.25) -- (189.53,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.08,216.82) -- (190.12,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.10,214.80) -- (188.10,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.38,215.39) -- (190.24,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.38,218.25) -- (190.24,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.79,216.82) -- (190.83,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.81,214.80) -- (188.81,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.32,215.39) -- (190.18,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.32,218.25) -- (190.18,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.73,216.82) -- (190.77,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.75,214.80) -- (188.75,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (193.98,215.39) -- (196.83,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (193.98,218.25) -- (196.83,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (193.39,216.82) -- (197.42,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (195.40,214.80) -- (195.40,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (196.77,215.39) -- (199.62,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (196.77,218.25) -- (199.62,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (196.17,216.82) -- (200.21,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.19,214.80) -- (198.19,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.63,215.39) -- (200.49,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.63,218.25) -- (200.49,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.04,216.82) -- (201.08,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (199.06,214.80) -- (199.06,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.80,215.39) -- (200.66,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.80,218.25) -- (200.66,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.21,216.82) -- (201.25,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (199.23,214.80) -- (199.23,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (194.11,215.39) -- (196.96,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (194.11,218.25) -- (196.96,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (193.52,216.82) -- (197.55,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (195.53,214.80) -- (195.53,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.95,215.39) -- (200.80,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.95,218.25) -- (200.80,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.36,216.82) -- (201.39,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (199.37,214.80) -- (199.37,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.10,215.39) -- (179.96,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.10,218.25) -- (179.96,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.51,216.82) -- (180.55,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.53,214.80) -- (178.53,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.45,215.39) -- (180.30,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.45,218.25) -- (180.30,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.86,216.82) -- (180.90,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.88,214.80) -- (178.88,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.62,215.39) -- (179.47,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.62,218.25) -- (179.47,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.03,216.82) -- (180.06,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.04,214.80) -- (178.04,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.82,215.39) -- (182.68,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.82,218.25) -- (182.68,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.23,216.82) -- (183.27,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.25,214.80) -- (181.25,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.93,215.39) -- (181.79,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.93,218.25) -- (181.79,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.34,216.82) -- (182.38,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.36,214.80) -- (180.36,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (190.30,215.39) -- (193.15,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (190.30,218.25) -- (193.15,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (189.71,216.82) -- (193.74,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (191.73,214.80) -- (191.73,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.34,215.39) -- (191.19,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.34,218.25) -- (191.19,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.75,216.82) -- (191.78,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (189.77,214.80) -- (189.77,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.85,215.39) -- (190.70,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.85,218.25) -- (190.70,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.26,216.82) -- (191.29,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (189.28,214.80) -- (189.28,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.42,215.39) -- (190.27,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.42,218.25) -- (190.27,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.83,216.82) -- (190.87,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.85,214.80) -- (188.85,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.84,215.39) -- (190.69,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.84,218.25) -- (190.69,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.25,216.82) -- (191.28,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (189.26,214.80) -- (189.26,218.84);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor,fill opacity=0.50] (139.30,215.39) --
	(142.15,215.39) --
	(142.15,218.25) --
	(139.30,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.30,215.39) --
	(139.15,215.39) --
	(139.15,218.25) --
	(136.30,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.98,215.39) --
	(139.83,215.39) --
	(139.83,218.25) --
	(136.98,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (135.93,215.39) --
	(138.79,215.39) --
	(138.79,218.25) --
	(135.93,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.77,215.39) --
	(140.63,215.39) --
	(140.63,218.25) --
	(137.77,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.65,215.39) --
	(140.51,215.39) --
	(140.51,218.25) --
	(137.65,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.06,215.39) --
	(141.91,215.39) --
	(141.91,218.25) --
	(139.06,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.41,119.58) --
	(140.26,119.58) --
	(140.26,122.43) --
	(137.41,122.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.04,215.39) --
	(140.89,215.39) --
	(140.89,218.25) --
	(138.04,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.50,116.37) --
	(139.35,116.37) --
	(139.35,119.22) --
	(136.50,119.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.59,156.75) --
	(140.45,156.75) --
	(140.45,159.60) --
	(137.59,159.60) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.66,215.39) --
	(143.51,215.39) --
	(143.51,218.25) --
	(140.66,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.20,215.39) --
	(140.05,215.39) --
	(140.05,218.25) --
	(137.20,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.59,215.39) --
	(142.45,215.39) --
	(142.45,218.25) --
	(139.59,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.44,215.39) --
	(141.30,215.39) --
	(141.30,218.25) --
	(138.44,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.44,215.39) --
	(141.30,215.39) --
	(141.30,218.25) --
	(138.44,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (154.80,215.39) --
	(157.66,215.39) --
	(157.66,218.25) --
	(154.80,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.34,215.39) --
	(143.19,215.39) --
	(143.19,218.25) --
	(140.34,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.90,215.39) --
	(143.75,215.39) --
	(143.75,218.25) --
	(140.90,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.05,215.39) --
	(144.91,215.39) --
	(144.91,218.25) --
	(142.05,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.70,215.39) --
	(142.55,215.39) --
	(142.55,218.25) --
	(139.70,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.28,215.39) --
	(144.13,215.39) --
	(144.13,218.25) --
	(141.28,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.23,132.18) --
	(144.09,132.18) --
	(144.09,135.03) --
	(141.23,135.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.32,215.39) --
	(145.17,215.39) --
	(145.17,218.25) --
	(142.32,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.16,215.39) --
	(144.01,215.39) --
	(144.01,218.25) --
	(141.16,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.16,171.06) --
	(143.01,171.06) --
	(143.01,173.91) --
	(140.16,173.91) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (143.21,215.39) --
	(146.06,215.39) --
	(146.06,218.25) --
	(143.21,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.61,215.39) --
	(147.46,215.39) --
	(147.46,218.25) --
	(144.61,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (146.59,215.39) --
	(149.44,215.39) --
	(149.44,218.25) --
	(146.59,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.69,215.39) --
	(147.54,215.39) --
	(147.54,218.25) --
	(144.69,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (146.11,215.39) --
	(148.97,215.39) --
	(148.97,218.25) --
	(146.11,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.22,215.39) --
	(147.07,215.39) --
	(147.07,218.25) --
	(144.22,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (180.26,215.39) --
	(183.12,215.39) --
	(183.12,218.25) --
	(180.26,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (146.53,215.39) --
	(149.38,215.39) --
	(149.38,218.25) --
	(146.53,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.03,215.39) --
	(147.89,215.39) --
	(147.89,218.25) --
	(145.03,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.48,215.39) --
	(148.33,215.39) --
	(148.33,218.25) --
	(145.48,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.18,215.39) --
	(148.03,215.39) --
	(148.03,218.25) --
	(145.18,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.28,215.39) --
	(148.14,215.39) --
	(148.14,218.25) --
	(145.28,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (143.47,139.37) --
	(146.32,139.37) --
	(146.32,142.22) --
	(143.47,142.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.29,139.66) --
	(147.14,139.66) --
	(147.14,142.52) --
	(144.29,142.52) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.44,215.39) --
	(148.30,215.39) --
	(148.30,218.25) --
	(145.44,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.48,141.86) --
	(147.33,141.86) --
	(147.33,144.71) --
	(144.48,144.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.78,199.65) --
	(147.63,199.65) --
	(147.63,202.50) --
	(144.78,202.50) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (158.04,215.39) --
	(160.89,215.39) --
	(160.89,218.25) --
	(158.04,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (163.60,215.39) --
	(166.45,215.39) --
	(166.45,218.25) --
	(163.60,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (155.73,215.39) --
	(158.59,215.39) --
	(158.59,218.25) --
	(155.73,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (161.34,215.39) --
	(164.19,215.39) --
	(164.19,218.25) --
	(161.34,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (160.80,215.39) --
	(163.65,215.39) --
	(163.65,218.25) --
	(160.80,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.31,215.39) --
	(147.16,215.39) --
	(147.16,218.25) --
	(144.31,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.23,215.39) --
	(140.08,215.39) --
	(140.08,218.25) --
	(137.23,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.53,215.39) --
	(140.39,215.39) --
	(140.39,218.25) --
	(137.53,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.44,215.39) --
	(140.30,215.39) --
	(140.30,218.25) --
	(137.44,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.36,215.39) --
	(141.21,215.39) --
	(141.21,218.25) --
	(138.36,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.47,215.39) --
	(140.33,215.39) --
	(140.33,218.25) --
	(137.47,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.64,215.39) --
	(141.49,215.39) --
	(141.49,218.25) --
	(138.64,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.03,132.34) --
	(141.88,132.34) --
	(141.88,135.19) --
	(139.03,135.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.88,136.32) --
	(139.74,136.32) --
	(139.74,139.18) --
	(136.88,139.18) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.35,124.28) --
	(140.20,124.28) --
	(140.20,127.13) --
	(137.35,127.13) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.50,119.84) --
	(141.35,119.84) --
	(141.35,122.70) --
	(138.50,122.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.33,215.39) --
	(144.18,215.39) --
	(144.18,218.25) --
	(141.33,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.67,215.39) --
	(144.52,215.39) --
	(144.52,218.25) --
	(141.67,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.93,215.39) --
	(142.78,215.39) --
	(142.78,218.25) --
	(139.93,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.96,215.39) --
	(144.82,215.39) --
	(144.82,218.25) --
	(141.96,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (143.25,215.39) --
	(146.10,215.39) --
	(146.10,218.25) --
	(143.25,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (170.49,215.39) --
	(173.35,215.39) --
	(173.35,218.25) --
	(170.49,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.53,215.39) --
	(144.39,215.39) --
	(144.39,218.25) --
	(141.53,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.27,215.39) --
	(147.13,215.39) --
	(147.13,218.25) --
	(144.27,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.64,215.39) --
	(145.49,215.39) --
	(145.49,218.25) --
	(142.64,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.39,215.39) --
	(148.24,215.39) --
	(148.24,218.25) --
	(145.39,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.21,215.39) --
	(145.06,215.39) --
	(145.06,218.25) --
	(142.21,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.77,215.39) --
	(145.62,215.39) --
	(145.62,218.25) --
	(142.77,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.21,160.09) --
	(145.06,160.09) --
	(145.06,162.94) --
	(142.21,162.94) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.32,215.39) --
	(145.17,215.39) --
	(145.17,218.25) --
	(142.32,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.76,140.76) --
	(144.62,140.76) --
	(144.62,143.61) --
	(141.76,143.61) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.49,215.39) --
	(144.34,215.39) --
	(144.34,218.25) --
	(141.49,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (155.36,215.39) --
	(158.22,215.39) --
	(158.22,218.25) --
	(155.36,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (149.39,215.39) --
	(152.24,215.39) --
	(152.24,218.25) --
	(149.39,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (154.29,215.39) --
	(157.14,215.39) --
	(157.14,218.25) --
	(154.29,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (153.11,215.39) --
	(155.97,215.39) --
	(155.97,218.25) --
	(153.11,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (156.51,215.39) --
	(159.36,215.39) --
	(159.36,218.25) --
	(156.51,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (148.06,215.39) --
	(150.91,215.39) --
	(150.91,218.25) --
	(148.06,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.92,215.39) --
	(141.77,215.39) --
	(141.77,218.25) --
	(138.92,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.41,215.39) --
	(143.26,215.39) --
	(143.26,218.25) --
	(140.41,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.22,215.39) --
	(142.07,215.39) --
	(142.07,218.25) --
	(139.22,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.58,215.39) --
	(143.44,215.39) --
	(143.44,218.25) --
	(140.58,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.47,215.39) --
	(141.33,215.39) --
	(141.33,218.25) --
	(138.47,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.30,117.64) --
	(141.15,117.64) --
	(141.15,120.50) --
	(138.30,120.50) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (135.97,117.94) --
	(138.82,117.94) --
	(138.82,120.80) --
	(135.97,120.80) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.64,181.79) --
	(141.49,181.79) --
	(141.49,184.64) --
	(138.64,184.64) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.97,215.39) --
	(143.82,215.39) --
	(143.82,218.25) --
	(140.97,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.10,116.50) --
	(140.95,116.50) --
	(140.95,119.35) --
	(138.10,119.35) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.50,215.39) --
	(148.35,215.39) --
	(148.35,218.25) --
	(145.50,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (143.51,215.39) --
	(146.36,215.39) --
	(146.36,218.25) --
	(143.51,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.08,215.39) --
	(146.93,215.39) --
	(146.93,218.25) --
	(144.08,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.77,215.39) --
	(148.63,215.39) --
	(148.63,218.25) --
	(145.77,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.41,215.39) --
	(147.26,215.39) --
	(147.26,218.25) --
	(144.41,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (154.17,215.39) --
	(157.02,215.39) --
	(157.02,218.25) --
	(154.17,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.78,215.39) --
	(142.63,215.39) --
	(142.63,218.25) --
	(139.78,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.53,215.39) --
	(144.39,215.39) --
	(144.39,218.25) --
	(141.53,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.90,215.39) --
	(142.76,215.39) --
	(142.76,218.25) --
	(139.90,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.68,215.39) --
	(143.53,215.39) --
	(143.53,218.25) --
	(140.68,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.81,215.39) --
	(141.66,215.39) --
	(141.66,218.25) --
	(138.81,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.53,215.39) --
	(140.39,215.39) --
	(140.39,218.25) --
	(137.53,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.21,215.39) --
	(144.06,215.39) --
	(144.06,218.25) --
	(141.21,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.19,188.53) --
	(142.05,188.53) --
	(142.05,191.38) --
	(139.19,191.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.42,215.39) --
	(144.27,215.39) --
	(144.27,218.25) --
	(141.42,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.57,154.29) --
	(142.42,154.29) --
	(142.42,157.14) --
	(139.57,157.14) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.02,215.39) --
	(147.87,215.39) --
	(147.87,218.25) --
	(145.02,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.96,215.39) --
	(148.81,215.39) --
	(148.81,218.25) --
	(145.96,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (147.75,215.39) --
	(150.60,215.39) --
	(150.60,218.25) --
	(147.75,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (147.90,215.39) --
	(150.75,215.39) --
	(150.75,218.25) --
	(147.90,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.76,215.39) --
	(148.61,215.39) --
	(148.61,218.25) --
	(145.76,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (173.21,215.39) --
	(176.07,215.39) --
	(176.07,218.25) --
	(173.21,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.92,215.39) --
	(145.77,215.39) --
	(145.77,218.25) --
	(142.92,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.40,215.39) --
	(145.26,215.39) --
	(145.26,218.25) --
	(142.40,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.89,215.39) --
	(145.75,215.39) --
	(145.75,218.25) --
	(142.89,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (143.06,215.39) --
	(145.91,215.39) --
	(145.91,218.25) --
	(143.06,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.65,215.39) --
	(148.51,215.39) --
	(148.51,218.25) --
	(145.65,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.80,141.43) --
	(143.66,141.43) --
	(143.66,144.28) --
	(140.80,144.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.83,215.39) --
	(144.68,215.39) --
	(144.68,218.25) --
	(141.83,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.67,215.39) --
	(144.52,215.39) --
	(144.52,218.25) --
	(141.67,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.81,215.39) --
	(144.66,215.39) --
	(144.66,218.25) --
	(141.81,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.03,215.39) --
	(144.88,215.39) --
	(144.88,218.25) --
	(142.03,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (156.81,215.39) --
	(159.66,215.39) --
	(159.66,218.25) --
	(156.81,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (160.07,215.39) --
	(162.92,215.39) --
	(162.92,218.25) --
	(160.07,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (156.42,215.39) --
	(159.28,215.39) --
	(159.28,218.25) --
	(156.42,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (158.37,215.39) --
	(161.22,215.39) --
	(161.22,218.25) --
	(158.37,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (158.19,215.39) --
	(161.04,215.39) --
	(161.04,218.25) --
	(158.19,218.25) --
	cycle;
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.29,136.70) -- (132.33,136.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.31,134.69) -- (130.31,138.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.90,109.37) -- (127.93,109.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.91,107.35) -- (125.91,111.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.04,115.49) -- (128.08,115.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.06,113.47) -- (126.06,117.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.79,134.22) -- (131.83,134.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.81,132.20) -- (129.81,136.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.24,125.55) -- (132.27,125.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.26,123.53) -- (130.26,127.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.57,130.80) -- (132.60,130.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.58,128.78) -- (130.58,132.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.75,135.92) -- (127.78,135.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.76,133.91) -- (125.76,137.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.97,119.19) -- (129.01,119.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.99,117.17) -- (126.99,121.21);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.26,135.75) -- (128.30,135.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.28,133.73) -- (126.28,137.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.05,125.86) -- (127.09,125.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.07,123.84) -- (125.07,127.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.62,137.48) -- (132.66,137.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.64,135.46) -- (130.64,139.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.67,133.59) -- (132.71,133.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.69,131.57) -- (130.69,135.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.07,134.09) -- (132.11,134.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.09,132.08) -- (130.09,136.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.56,133.11) -- (133.59,133.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.57,131.09) -- (131.57,135.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.20,135.28) -- (133.24,135.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.22,133.26) -- (131.22,137.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.26,131.49) -- (128.30,131.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.28,129.47) -- (126.28,133.51);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.04,113.98) -- (139.89,116.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.04,116.84) -- (139.89,113.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.45,115.41) -- (140.48,115.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.47,113.39) -- (138.47,117.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.50,114.97) -- (141.35,117.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.50,117.82) -- (141.35,114.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.91,116.40) -- (141.94,116.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.93,114.38) -- (139.93,118.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.74,119.90) -- (140.60,122.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.74,122.75) -- (140.60,119.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.15,121.32) -- (141.19,121.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.17,119.31) -- (139.17,123.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.32,113.83) -- (140.17,116.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.32,116.68) -- (140.17,113.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.73,115.25) -- (140.76,115.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.75,113.24) -- (138.75,117.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.47,117.46) -- (141.33,120.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.47,120.31) -- (141.33,117.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.88,118.88) -- (141.92,118.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.90,116.87) -- (139.90,120.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.01,114.30) -- (139.86,117.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.01,117.15) -- (139.86,114.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.42,115.72) -- (140.45,115.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.44,113.70) -- (138.44,117.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.46,114.06) -- (139.32,116.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.46,116.92) -- (139.32,114.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.87,115.49) -- (139.91,115.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.89,113.47) -- (137.89,117.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.23,109.94) -- (139.09,112.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.23,112.79) -- (139.09,109.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.64,111.37) -- (139.68,111.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.66,109.35) -- (137.66,113.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.85,110.35) -- (139.70,113.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.85,113.20) -- (139.70,110.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.26,111.77) -- (140.29,111.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.28,109.76) -- (138.28,113.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.73,108.75) -- (141.58,111.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.73,111.60) -- (141.58,108.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.13,110.18) -- (142.17,110.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.15,108.16) -- (140.15,112.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.68,108.52) -- (140.54,111.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.68,111.38) -- (140.54,108.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.09,109.95) -- (141.13,109.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.11,107.93) -- (139.11,111.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.92,116.23) -- (140.78,119.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.92,119.09) -- (140.78,116.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.33,117.66) -- (141.37,117.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.35,115.64) -- (139.35,119.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.89,118.36) -- (141.75,121.21);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.89,121.21) -- (141.75,118.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.30,119.78) -- (142.34,119.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.32,117.76) -- (140.32,121.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.07,112.32) -- (139.92,115.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.07,115.18) -- (139.92,112.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.48,113.75) -- (140.52,113.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.50,111.73) -- (138.50,115.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.07,111.79) -- (139.92,114.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.07,114.64) -- (139.92,111.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.48,113.21) -- (140.52,113.21);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.50,111.20) -- (138.50,115.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.13,114.37) -- (142.99,117.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.13,117.23) -- (142.99,114.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.54,115.80) -- (143.58,115.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.56,113.78) -- (141.56,117.82);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (214.20,216.82) -- (218.24,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (216.22,214.80) -- (216.22,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (209.19,216.82) -- (213.23,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (211.21,214.80) -- (211.21,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (208.62,216.82) -- (212.66,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (210.64,214.80) -- (210.64,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (213.61,216.82) -- (217.65,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (215.63,214.80) -- (215.63,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.73,216.82) -- (132.76,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.75,214.80) -- (130.75,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.30,206.31) -- (133.34,206.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.32,204.29) -- (131.32,208.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.56,204.05) -- (133.59,204.05);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.57,202.03) -- (131.57,206.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.35,205.62) -- (133.39,205.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.37,203.60) -- (131.37,207.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.18,206.47) -- (135.21,206.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.19,204.45) -- (133.19,208.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.89,204.42) -- (132.92,204.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.90,202.40) -- (130.90,206.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.00,205.41) -- (135.03,205.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.01,203.39) -- (133.01,207.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.03,204.52) -- (131.07,204.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.05,202.50) -- (129.05,206.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.62,205.41) -- (132.66,205.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.64,203.40) -- (130.64,207.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.25,205.79) -- (133.29,205.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.27,203.77) -- (131.27,207.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.15,192.95) -- (134.19,192.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.17,190.93) -- (132.17,194.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.35,204.31) -- (132.38,204.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.37,202.29) -- (130.37,206.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.92,206.20) -- (135.96,206.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.94,204.18) -- (133.94,208.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.99,205.97) -- (133.03,205.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.01,203.95) -- (131.01,207.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.30,216.82) -- (136.34,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.32,214.80) -- (134.32,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.91,203.78) -- (133.94,203.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.92,201.76) -- (131.92,205.80);

\path[fill=fillColor,fill opacity=0.50] (141.37,215.39) --
	(144.23,215.39) --
	(144.23,218.25) --
	(141.37,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.42,215.39) --
	(141.27,215.39) --
	(141.27,218.25) --
	(138.42,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.74,215.39) --
	(140.60,215.39) --
	(140.60,218.25) --
	(137.74,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.81,215.39) --
	(141.66,215.39) --
	(141.66,218.25) --
	(138.81,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.30,215.39) --
	(141.15,215.39) --
	(141.15,218.25) --
	(138.30,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.04,215.39) --
	(140.89,215.39) --
	(140.89,218.25) --
	(138.04,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.51,215.39) --
	(142.37,215.39) --
	(142.37,218.25) --
	(139.51,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.33,127.40) --
	(141.18,127.40) --
	(141.18,130.25) --
	(138.33,130.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.46,215.39) --
	(142.31,215.39) --
	(142.31,218.25) --
	(139.46,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.14,120.26) --
	(141.99,120.26) --
	(141.99,123.11) --
	(139.14,123.11) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.41,157.01) --
	(140.26,157.01) --
	(140.26,159.86) --
	(137.41,159.86) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.43,215.39) --
	(142.29,215.39) --
	(142.29,218.25) --
	(139.43,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.50,215.39) --
	(141.35,215.39) --
	(141.35,218.25) --
	(138.50,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.48,215.39) --
	(143.34,215.39) --
	(143.34,218.25) --
	(140.48,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.41,215.39) --
	(143.26,215.39) --
	(143.26,218.25) --
	(140.41,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.26,215.39) --
	(144.11,215.39) --
	(144.11,218.25) --
	(141.26,218.25) --
	cycle;

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.45,216.82) -- (184.48,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (182.47,214.80) -- (182.47,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (173.40,216.82) -- (177.44,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (175.42,214.80) -- (175.42,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (182.25,216.82) -- (186.29,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (184.27,214.80) -- (184.27,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.41,216.82) -- (180.45,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.43,214.80) -- (178.43,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (182.15,216.82) -- (186.19,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (184.17,214.80) -- (184.17,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.29,216.82) -- (182.33,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.31,214.80) -- (180.31,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.77,216.82) -- (168.80,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.79,214.80) -- (166.79,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.62,216.82) -- (143.65,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.64,214.80) -- (141.64,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.27,216.82) -- (141.31,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.29,214.80) -- (139.29,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (199.04,216.82) -- (203.08,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (201.06,214.80) -- (201.06,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (204.76,216.82) -- (208.80,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (206.78,214.80) -- (206.78,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (203.41,216.82) -- (207.44,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (205.42,214.80) -- (205.42,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (201.06,216.82) -- (205.10,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (203.08,214.80) -- (203.08,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (184.33,216.82) -- (188.37,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.35,214.80) -- (186.35,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.16,216.82) -- (202.20,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (200.18,214.80) -- (200.18,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.31,216.82) -- (139.34,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.33,214.80) -- (137.33,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.83,216.82) -- (138.86,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.84,214.80) -- (136.84,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.54,216.82) -- (139.58,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.56,214.80) -- (137.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.51,216.82) -- (138.54,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.52,214.80) -- (136.52,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.45,216.82) -- (141.48,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.47,214.80) -- (139.47,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.51,216.82) -- (139.55,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.53,214.80) -- (137.53,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.66,216.82) -- (137.69,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.68,214.80) -- (135.68,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.03,216.82) -- (139.07,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.05,214.80) -- (137.05,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.35,216.82) -- (173.38,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.37,214.80) -- (171.37,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (192.60,216.82) -- (196.63,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (194.61,214.80) -- (194.61,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (203.16,216.82) -- (207.20,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (205.18,214.80) -- (205.18,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.56,216.82) -- (202.59,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (200.58,214.80) -- (200.58,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (201.63,216.82) -- (205.67,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (203.65,214.80) -- (203.65,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.24,216.82) -- (185.27,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (183.25,214.80) -- (183.25,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.37,216.82) -- (146.40,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.38,214.80) -- (144.38,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.70,216.82) -- (146.73,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.71,214.80) -- (144.71,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.39,216.82) -- (147.43,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.41,214.80) -- (145.41,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.28,216.82) -- (147.31,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.29,214.80) -- (145.29,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.57,216.82) -- (146.61,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.59,214.80) -- (144.59,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.19,216.82) -- (145.23,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.21,214.80) -- (143.21,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.21,216.82) -- (144.25,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.23,214.80) -- (142.23,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.26,216.82) -- (147.29,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.27,214.80) -- (145.27,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.28,216.82) -- (145.32,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.30,214.80) -- (143.30,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.29,216.82) -- (143.32,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.31,214.80) -- (141.31,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.46,216.82) -- (183.50,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.48,214.80) -- (181.48,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.78,216.82) -- (146.81,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.80,214.80) -- (144.80,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.08,216.82) -- (150.12,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.10,214.80) -- (148.10,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.72,216.82) -- (152.75,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.74,214.80) -- (150.74,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.54,216.82) -- (150.58,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.56,214.80) -- (148.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.21,216.82) -- (153.24,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.22,214.80) -- (151.22,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.53,216.82) -- (145.56,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.55,214.80) -- (143.55,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.22,216.82) -- (142.25,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.24,214.80) -- (140.24,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.16,216.82) -- (143.19,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.18,214.80) -- (141.18,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.68,216.82) -- (145.72,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.70,214.80) -- (143.70,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (175.85,216.82) -- (179.88,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.87,214.80) -- (177.87,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.60,216.82) -- (192.64,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (190.62,214.80) -- (190.62,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (189.71,216.82) -- (193.75,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (191.73,214.80) -- (191.73,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.43,216.82) -- (180.46,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.44,214.80) -- (178.44,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (190.85,216.82) -- (194.88,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (192.87,214.80) -- (192.87,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.07,216.82) -- (190.10,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.09,214.80) -- (188.09,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.40,216.82) -- (150.44,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.42,214.80) -- (148.42,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.51,216.82) -- (152.55,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.53,214.80) -- (150.53,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.93,216.82) -- (150.96,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.95,214.80) -- (148.95,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.67,216.82) -- (153.70,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.69,214.80) -- (151.69,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.66,216.82) -- (149.69,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.67,214.80) -- (147.67,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (184.89,216.82) -- (188.92,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.90,214.80) -- (186.90,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.97,216.82) -- (145.00,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.98,214.80) -- (142.98,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.28,216.82) -- (144.32,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.30,214.80) -- (142.30,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.79,216.82) -- (149.82,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.81,214.80) -- (147.81,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.79,216.82) -- (138.83,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.81,214.80) -- (136.81,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.22,216.82) -- (135.26,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.24,214.80) -- (133.24,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.25,216.82) -- (138.29,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.27,214.80) -- (136.27,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.22,216.82) -- (135.26,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.24,214.80) -- (133.24,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.29,216.82) -- (138.32,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.31,214.80) -- (136.31,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.36,216.82) -- (142.39,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.37,214.80) -- (140.37,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.85,216.82) -- (137.88,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.86,214.80) -- (135.86,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.97,216.82) -- (141.01,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.99,214.80) -- (138.99,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.62,216.82) -- (143.65,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.64,214.80) -- (141.64,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.20,216.82) -- (140.23,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.21,214.80) -- (138.21,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.39,216.82) -- (141.43,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.41,214.80) -- (139.41,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.91,216.82) -- (140.95,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.93,214.80) -- (138.93,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.68,216.82) -- (138.72,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.70,214.80) -- (136.70,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.62,216.82) -- (141.66,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.64,214.80) -- (139.64,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.28,216.82) -- (161.31,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.29,214.80) -- (159.29,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.63,216.82) -- (136.67,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.65,214.80) -- (134.65,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.88,216.82) -- (136.91,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.89,214.80) -- (134.89,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.77,171.45) -- (137.81,171.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.79,169.44) -- (135.79,173.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.00,171.18) -- (138.03,171.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.01,169.16) -- (136.01,173.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.45,216.82) -- (141.48,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.47,214.80) -- (139.47,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.48,216.82) -- (141.51,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.50,214.80) -- (139.50,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.68,216.82) -- (142.72,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.70,214.80) -- (140.70,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.52,216.82) -- (142.56,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.54,214.80) -- (140.54,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.96,216.82) -- (146.00,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.98,214.80) -- (143.98,218.84);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.38,120.51) rectangle (140.23,123.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.38,120.51) -- (140.23,123.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.38,123.36) -- (140.23,120.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.10,119.04) rectangle (136.95,121.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.10,119.04) -- (136.95,121.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.10,121.89) -- (136.95,119.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.04,112.32) rectangle (134.89,115.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.04,112.32) -- (134.89,115.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.04,115.18) -- (134.89,112.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.21,113.51) rectangle (135.07,116.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.21,113.51) -- (135.07,116.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.21,116.36) -- (135.07,113.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.88,119.90) rectangle (137.73,122.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.88,119.90) -- (137.73,122.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.88,122.75) -- (137.73,119.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.67,111.88) rectangle (136.52,114.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.67,111.88) -- (136.52,114.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.67,114.73) -- (136.52,111.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.14, 91.01) rectangle (136.00, 93.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.14, 91.01) -- (136.00, 93.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.14, 93.86) -- (136.00, 91.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.54, 84.93) rectangle (134.39, 87.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.54, 84.93) -- (134.39, 87.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.54, 87.78) -- (134.39, 84.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.59, 85.50) rectangle (133.45, 88.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.59, 85.50) -- (133.45, 88.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.59, 88.35) -- (133.45, 85.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.43, 92.80) rectangle (136.28, 95.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.43, 92.80) -- (136.28, 95.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.43, 95.66) -- (136.28, 92.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.21, 97.40) rectangle (137.06,100.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.21, 97.40) -- (137.06,100.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.21,100.25) -- (137.06, 97.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.38,114.67) rectangle (138.23,117.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.38,114.67) -- (138.23,117.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.38,117.53) -- (138.23,114.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.77,116.23) rectangle (137.62,119.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.77,116.23) -- (137.62,119.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.77,119.09) -- (137.62,116.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.23,112.32) rectangle (136.08,115.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.23,112.32) -- (136.08,115.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.23,115.18) -- (136.08,112.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.03,117.15) rectangle (137.88,120.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.03,117.15) -- (137.88,120.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.03,120.00) -- (137.88,117.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.35,114.22) rectangle (136.20,117.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.35,114.22) -- (136.20,117.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.35,117.07) -- (136.20,114.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.31, 95.28) rectangle (136.16, 98.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.31, 95.28) -- (136.16, 98.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.31, 98.13) -- (136.16, 95.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.55, 88.51) rectangle (133.40, 91.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.55, 88.51) -- (133.40, 91.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.55, 91.37) -- (133.40, 88.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.95, 89.39) rectangle (129.80, 92.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.95, 89.39) -- (129.80, 92.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.95, 92.25) -- (129.80, 89.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.03, 92.46) rectangle (133.88, 95.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.03, 92.46) -- (133.88, 95.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.03, 95.32) -- (133.88, 92.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.49, 83.72) rectangle (128.35, 86.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.49, 83.72) -- (128.35, 86.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.49, 86.57) -- (128.35, 83.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.55, 85.50) rectangle (131.40, 88.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.55, 85.50) -- (131.40, 88.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.55, 88.35) -- (131.40, 85.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.21, 85.50) rectangle (131.06, 88.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.21, 85.50) -- (131.06, 88.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.21, 88.35) -- (131.06, 85.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.31, 99.24) rectangle (136.16,102.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.31, 99.24) -- (136.16,102.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.31,102.09) -- (136.16, 99.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.99, 85.50) rectangle (131.85, 88.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.99, 85.50) -- (131.85, 88.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.99, 88.35) -- (131.85, 85.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.42, 90.62) rectangle (132.28, 93.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.42, 90.62) -- (132.28, 93.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.42, 93.48) -- (132.28, 90.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.03, 96.89) rectangle (133.88, 99.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.03, 96.89) -- (133.88, 99.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.03, 99.75) -- (133.88, 96.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.56,128.45) rectangle (135.41,131.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.56,128.45) -- (135.41,131.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.56,131.30) -- (135.41,128.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.66,125.92) rectangle (131.52,128.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.66,125.92) -- (131.52,128.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.66,128.77) -- (131.52,125.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.72,129.80) rectangle (134.58,132.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.72,129.80) -- (134.58,132.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.72,132.65) -- (134.58,129.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.73,128.04) rectangle (135.58,130.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.73,128.04) -- (135.58,130.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.73,130.89) -- (135.58,128.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.98,130.22) rectangle (136.83,133.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.98,130.22) -- (136.83,133.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.98,133.07) -- (136.83,130.22);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.41,216.82) -- (142.45,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.43,214.80) -- (140.43,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.39,216.82) -- (136.42,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.41,214.80) -- (134.41,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.38,216.82) -- (139.41,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.39,214.80) -- (137.39,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.96,216.82) -- (136.99,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.97,214.80) -- (134.97,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.84,216.82) -- (139.88,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.86,214.80) -- (137.86,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.62,216.82) -- (137.65,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.64,214.80) -- (135.64,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.45,216.82) -- (135.48,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.46,214.80) -- (133.46,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.88,216.82) -- (136.91,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.89,214.80) -- (134.89,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.46,216.82) -- (133.49,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.47,214.80) -- (131.47,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.02,216.82) -- (132.05,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.03,214.80) -- (130.03,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.58,216.82) -- (135.61,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.59,214.80) -- (133.59,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.88,216.82) -- (136.91,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.89,214.80) -- (134.89,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.58,216.82) -- (139.61,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.59,214.80) -- (137.59,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.08,216.82) -- (137.11,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.09,214.80) -- (135.09,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.00,216.82) -- (139.03,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.02,214.80) -- (137.02,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.40,216.82) -- (135.44,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.42,214.80) -- (133.42,218.84);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.08, 68.39) rectangle (125.93, 71.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.08, 68.39) -- (125.93, 71.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.08, 71.24) -- (125.93, 68.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.72, 75.44) rectangle (131.57, 78.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.72, 75.44) -- (131.57, 78.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.72, 78.29) -- (131.57, 75.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.29,108.18) rectangle (140.14,111.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.29,108.18) -- (140.14,111.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.29,111.03) -- (140.14,108.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.39,143.87) rectangle (153.24,146.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.39,143.87) -- (153.24,146.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.39,146.73) -- (153.24,143.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.56,168.04) rectangle (155.42,170.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.56,168.04) -- (155.42,170.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.56,170.89) -- (155.42,168.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.34,193.96) rectangle (167.19,196.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.34,193.96) -- (167.19,196.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.34,196.81) -- (167.19,193.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.00,215.39) rectangle (166.85,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.00,215.39) -- (166.85,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.00,218.25) -- (166.85,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.88,215.39) rectangle (173.73,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.88,215.39) -- (173.73,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.88,218.25) -- (173.73,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.06,215.39) rectangle (180.91,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.06,215.39) -- (180.91,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.06,218.25) -- (180.91,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (184.80,215.39) rectangle (187.65,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (184.80,215.39) -- (187.65,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (184.80,218.25) -- (187.65,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.10,215.39) rectangle (190.95,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.10,215.39) -- (190.95,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.10,218.25) -- (190.95,215.39);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor,fill opacity=0.50] (226.56, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 83.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 83.12) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 83.12) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 83.12) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 83.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 83.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 83.12) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 83.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 83.12) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 83.12) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 84.50) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 83.12) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 83.12) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (135.13, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.28, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.13, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.26, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.40, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.43, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.54, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.09, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.83, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.91, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.67, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.46, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.23, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.85, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.15, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.34, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (132.64, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.53, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.99, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.26, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (131.57, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.87, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.86, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.31, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (131.68, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (128.37, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.91, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.13, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (159.77, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.85, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.06, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.99, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (131.73, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.21, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.04, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.54, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (134.24, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (128.37, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.76, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.06, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (158.41, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (133.55, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.91, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.43, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (133.10, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (130.90, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (128.44, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.50, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (164.86, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (133.15, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.93, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.06, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (149.89, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.76, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.09, 73.06) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.91, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (128.50, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (128.31, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.20, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.51, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (138.21, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (132.74, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.80, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.61, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (134.57, 73.06) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (128.56, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.01, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.61, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (131.87, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.93, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.46, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.58, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.93, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (128.75, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.50, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.83, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (155.05, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.84, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.59, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.75, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.11, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.26, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.18, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.59, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.99, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.78, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.43, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.38, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (149.86, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.91, 73.06) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.07, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.26, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.35, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (122.95, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.99, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.67, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (143.14, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (134.73, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.91, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.67, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (146.17, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.84, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.23, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.67, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.76, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.22, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.09, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.91, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (155.82, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.31, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.09, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.43, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (134.77, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.13, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.18, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.18, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (133.28, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.35, 73.06) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.75, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.91, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (163.67, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (137.99, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.17, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.73, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (160.83, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.53, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.26, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.38, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (143.12, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (133.90, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.06, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.54, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (184.52, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (133.59, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (128.87, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.26, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (136.88, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.35, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.35, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.71, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (132.41, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.13, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.99, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.31, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (178.02, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.64, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.92, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.06, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (133.81, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.67, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.06, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.92, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (128.31, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.40, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.71, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.21, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.29, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.92, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.85, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.78, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (170.20, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.73, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.99, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.64, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (154.59, 79.88) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (135.01, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.69, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.13, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (168.26, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (131.32, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.69, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.21, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (131.47, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.67, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (128.12, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.47, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (174.45, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (155.56, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (134.61, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (133.64, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (186.75, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (190.54, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (144.22, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.29, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (204.52, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (159.31, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (135.45, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (128.18, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (165.41, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (165.35, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (131.22, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.06, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (149.62, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.47, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.15, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.21, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (141.99, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.33, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.14, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.57, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.58, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.73, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.57, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.38, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.64, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.60, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.73, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.07, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (131.52, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.60, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.33, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.14, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (162.60, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (132.97, 79.88) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.85, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.50, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.80, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (164.85, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.14, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.76, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (134.41, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.20, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.91, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.38, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (137.53, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.57, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.43, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.69, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.46, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.67, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.43, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (128.31, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (150.59, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.26, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.23, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.43, 80.76) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (131.92, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.40, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.78, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.46, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (134.61, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.26, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.64, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (131.52, 89.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (178.70, 90.39) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (130.37, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.64, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (131.68, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.46, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (132.26, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (131.11, 89.48) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.15, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (133.94, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.99, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (131.01, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.64, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (198.67, 89.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (162.33, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (136.81, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.26, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (173.23, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (128.75, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (128.06, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (128.81, 90.39) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (152.50, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (135.60, 83.12) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.50, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (131.11, 81.59) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (181.20, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (138.34, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (133.01, 89.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (130.26, 83.12) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (134.61, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.64, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.05, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.47, 89.00) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (133.19, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (131.47, 84.50) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.47, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.35, 84.50) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (144.30, 88.00) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (136.05, 90.82) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (136.49, 90.82) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.81, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (166.42, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (205.42, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (151.77, 90.39) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.05, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (165.28, 90.39) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (140.51, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (136.31, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (130.26, 89.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (226.56, 88.51) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (157.23, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (143.14, 88.51) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (128.37, 84.50) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (203.67, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (173.21, 90.39) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.20, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.05, 90.82) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (207.49, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (158.60, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (137.29, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (132.07, 90.82) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (177.48, 79.88) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (130.42, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (132.55, 90.82) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.71, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (170.77, 89.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (140.10, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (128.56, 79.88) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.73, 79.88) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (204.88, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (187.83, 89.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (135.17, 89.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.87, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.52, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.47, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.71, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.35, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (140.97, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (130.48, 85.15) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (163.48, 79.88) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.31, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (142.16, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (128.68, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (132.22, 88.51) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.99, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (147.87, 89.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (130.75, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.99, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.60, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (131.17, 86.36) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (130.90, 87.47) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (130.20, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.99, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (165.50, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (130.48, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (128.44, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.84, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (139.55, 88.00) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (131.37, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.43, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.91, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (196.30, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (133.64, 88.51) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.78, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.21, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (155.37, 88.51) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.46, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.87, 90.39) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.84, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (143.44, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (130.80, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (130.58, 80.76) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.14, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (133.59, 84.50) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (130.58, 80.76) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.20, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (130.75, 86.36) circle (  1.43);
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor,fill opacity=0.50] (136.81, 96.78) --
	(138.73, 93.45) --
	(134.89, 93.45) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.27, 82.98) --
	(133.19, 79.65) --
	(129.35, 79.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.97, 79.08) --
	(136.90, 75.76) --
	(133.05, 75.76) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (132.02, 82.10) --
	(133.94, 78.77) --
	(130.10, 78.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.15, 81.16) --
	(132.07, 77.83) --
	(128.22, 77.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.11, 77.92) --
	(133.04, 74.59) --
	(129.19, 74.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.53, 82.98) --
	(132.45, 79.65) --
	(128.61, 79.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.78, 82.10) --
	(133.70, 78.77) --
	(129.85, 78.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.40, 76.66) --
	(131.33, 73.33) --
	(127.48, 73.33) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.75, 81.16) --
	(132.67, 77.83) --
	(128.82, 77.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (135.33, 91.22) --
	(137.25, 87.89) --
	(133.41, 87.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.27, 92.61) --
	(138.19, 89.28) --
	(134.35, 89.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.31, 95.76) --
	(138.23, 92.43) --
	(134.39, 92.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.56,106.68) --
	(142.49,103.35) --
	(138.64,103.35) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.07, 94.65) --
	(135.99, 91.33) --
	(132.15, 91.33) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.31,107.01) --
	(138.23,103.68) --
	(134.39,103.68) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.38,101.04) --
	(141.30, 97.71) --
	(137.46, 97.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.18,113.99) --
	(140.10,110.67) --
	(136.26,110.67) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.05, 98.93) --
	(137.97, 95.60) --
	(134.13, 95.60) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.77,100.79) --
	(138.69, 97.46) --
	(134.85, 97.46) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.90,117.39) --
	(146.82,114.06) --
	(142.97,114.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.05,125.58) --
	(145.97,122.26) --
	(142.12,122.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (149.34,141.53) --
	(151.26,138.20) --
	(147.42,138.20) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (147.67,121.65) --
	(149.59,118.32) --
	(145.75,118.32) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.86,138.63) --
	(143.78,135.30) --
	(139.94,135.30) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (156.19,115.61) --
	(158.11,112.29) --
	(154.27,112.29) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.64,132.60) --
	(147.56,129.28) --
	(143.72,129.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (149.17,120.21) --
	(151.09,116.88) --
	(147.25,116.88) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (143.23,110.98) --
	(145.16,107.65) --
	(141.31,107.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (151.16,134.34) --
	(153.08,131.01) --
	(149.24,131.01) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (168.93,121.47) --
	(170.85,118.14) --
	(167.01,118.14) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (146.75,131.80) --
	(148.67,128.47) --
	(144.82,128.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.17,127.09) --
	(146.10,123.76) --
	(142.25,123.76) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (158.46,117.63) --
	(160.39,114.30) --
	(156.54,114.30) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.37,120.79) --
	(144.30,117.46) --
	(140.45,117.46) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (170.80,137.85) --
	(172.73,134.52) --
	(168.88,134.52) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (158.13,123.28) --
	(160.05,119.95) --
	(156.21,119.95) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (152.65,137.01) --
	(154.57,133.68) --
	(150.73,133.68) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (164.49,141.09) --
	(166.41,137.76) --
	(162.57,137.76) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (153.27,113.89) --
	(155.19,110.56) --
	(151.35,110.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (182.71,128.90) --
	(184.63,125.57) --
	(180.79,125.57) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (182.16,143.57) --
	(184.08,140.25) --
	(180.24,140.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (181.62,149.73) --
	(183.54,146.41) --
	(179.70,146.41) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (161.02,136.05) --
	(162.95,132.72) --
	(159.10,132.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (161.62,148.63) --
	(163.54,145.30) --
	(159.69,145.30) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (171.54,144.78) --
	(173.46,141.46) --
	(169.62,141.46) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (176.92,119.47) --
	(178.84,116.14) --
	(175.00,116.14) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (159.67,132.94) --
	(161.59,129.61) --
	(157.75,129.61) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (156.03,137.34) --
	(157.96,134.01) --
	(154.11,134.01) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (184.97,154.10) --
	(186.89,150.77) --
	(183.05,150.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (173.80,145.39) --
	(175.73,142.06) --
	(171.88,142.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (177.80,171.08) --
	(179.73,167.75) --
	(175.88,167.75) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (196.62,156.25) --
	(198.54,152.92) --
	(194.70,152.92) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (185.09,156.91) --
	(187.01,153.58) --
	(183.17,153.58) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (185.49,150.07) --
	(187.41,146.75) --
	(183.57,146.75) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.87,160.49) --
	(190.79,157.16) --
	(186.94,157.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (161.98,148.06) --
	(163.90,144.73) --
	(160.06,144.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (166.52,159.24) --
	(168.44,155.92) --
	(164.60,155.92) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (180.70,169.65) --
	(182.62,166.32) --
	(178.78,166.32) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (167.52,176.56) --
	(169.44,173.23) --
	(165.59,173.23) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (190.91,162.92) --
	(192.83,159.59) --
	(188.99,159.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (181.01,170.31) --
	(182.93,166.98) --
	(179.09,166.98) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (175.20,169.04) --
	(177.12,165.71) --
	(173.28,165.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (161.62,168.57) --
	(163.54,165.25) --
	(159.70,165.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (183.91,179.13) --
	(185.83,175.80) --
	(181.99,175.80) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (198.77,181.72) --
	(200.69,178.40) --
	(196.84,178.40) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.76,148.61) --
	(190.68,145.28) --
	(186.84,145.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (215.62,172.21) --
	(217.54,168.88) --
	(213.69,168.88) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (189.81,171.14) --
	(191.73,167.81) --
	(187.89,167.81) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (185.86,154.01) --
	(187.79,150.68) --
	(183.94,150.68) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (179.23,184.52) --
	(181.15,181.19) --
	(177.31,181.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (217.06,177.23) --
	(218.98,173.90) --
	(215.14,173.90) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,181.11) --
	(228.48,177.78) --
	(224.64,177.78) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (194.20,172.40) --
	(196.12,169.07) --
	(192.28,169.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,187.50) --
	(228.48,184.17) --
	(224.64,184.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (204.60,172.98) --
	(206.52,169.65) --
	(202.68,169.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,174.52) --
	(228.48,171.19) --
	(224.64,171.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (185.97,178.05) --
	(187.90,174.72) --
	(184.05,174.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (205.50,172.05) --
	(207.42,168.72) --
	(203.58,168.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (184.28,168.64) --
	(186.20,165.31) --
	(182.36,165.31) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (204.66,182.01) --
	(206.58,178.68) --
	(202.74,178.68) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,194.03) --
	(228.48,190.70) --
	(224.64,190.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (200.42,193.76) --
	(202.34,190.43) --
	(198.50,190.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (207.77,179.94) --
	(209.69,176.61) --
	(205.85,176.61) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,186.79) --
	(228.48,183.46) --
	(224.64,183.46) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,192.99) --
	(228.48,189.66) --
	(224.64,189.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (214.14,204.87) --
	(216.06,201.54) --
	(212.22,201.54) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,189.81) --
	(228.48,186.48) --
	(224.64,186.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,190.09) --
	(228.48,186.76) --
	(224.64,186.76) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,186.08) --
	(228.48,182.75) --
	(224.64,182.75) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.42, 83.81) --
	(135.34, 80.48) --
	(131.50, 80.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.82, 86.72) --
	(133.75, 83.39) --
	(129.90, 83.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.81, 96.45) --
	(135.73, 93.12) --
	(131.89, 93.12) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 89.15) --
	(228.48, 85.82) --
	(224.64, 85.82) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.77, 90.73) --
	(135.69, 87.40) --
	(131.85, 87.40) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.24, 87.98) --
	(135.16, 84.66) --
	(131.32, 84.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (128.37, 73.75) --
	(130.29, 70.42) --
	(126.45, 70.42) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (132.97, 87.98) --
	(134.89, 84.66) --
	(131.05, 84.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (127.93, 75.27) --
	(129.85, 71.95) --
	(126.01, 71.95) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.68, 89.69) --
	(135.60, 86.37) --
	(131.76, 86.37) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.84, 92.16) --
	(138.76, 88.83) --
	(134.92, 88.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (128.31, 76.66) --
	(130.23, 73.33) --
	(126.39, 73.33) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.52, 80.16) --
	(131.44, 76.83) --
	(127.60, 76.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (135.09, 90.22) --
	(137.02, 86.89) --
	(133.17, 86.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (132.78, 92.61) --
	(134.70, 89.28) --
	(130.86, 89.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.03, 82.10) --
	(131.96, 78.77) --
	(128.11, 78.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.09, 90.22) --
	(138.01, 86.89) --
	(134.17, 86.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (128.56, 82.98) --
	(130.48, 79.65) --
	(126.64, 79.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.65, 92.61) --
	(136.57, 89.28) --
	(132.73, 89.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.64, 80.16) --
	(131.56, 76.83) --
	(127.72, 76.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (135.98, 88.58) --
	(137.90, 85.25) --
	(134.05, 85.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.34,117.71) --
	(138.26,114.38) --
	(134.42,114.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.07,110.21) --
	(135.99,106.89) --
	(132.15,106.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.05, 95.40) --
	(140.97, 92.07) --
	(137.13, 92.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.47,100.28) --
	(140.39, 96.96) --
	(136.55, 96.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (135.21, 98.34) --
	(137.13, 95.02) --
	(133.29, 95.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (135.01, 87.37) --
	(136.94, 84.04) --
	(133.09, 84.04) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (135.13,103.93) --
	(137.05,100.60) --
	(133.21,100.60) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.05,108.70) --
	(139.97,105.37) --
	(136.13,105.37) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.28,118.39) --
	(140.20,115.07) --
	(136.36,115.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.07, 99.21) --
	(135.99, 95.88) --
	(132.15, 95.88) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.53, 97.11) --
	(140.45, 93.78) --
	(136.61, 93.78) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.53,105.82) --
	(139.45,102.49) --
	(135.61,102.49) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.62,101.76) --
	(142.54, 98.43) --
	(138.70, 98.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.89, 93.46) --
	(139.81, 90.13) --
	(135.97, 90.13) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 94.27) --
	(228.48, 90.94) --
	(224.64, 90.94) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.84, 81.16) --
	(140.76, 77.83) --
	(136.92, 77.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.24, 87.98) --
	(136.16, 84.66) --
	(132.32, 84.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.56, 95.40) --
	(142.49, 92.07) --
	(138.64, 92.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.02,101.76) --
	(138.94, 98.43) --
	(135.10, 98.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.99, 92.61) --
	(140.91, 89.28) --
	(137.07, 89.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.02,100.79) --
	(142.94, 97.46) --
	(139.10, 97.46) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.45,106.68) --
	(146.37,103.35) --
	(142.53,103.35) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.45,107.96) --
	(147.37,104.63) --
	(143.53,104.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (147.17,127.61) --
	(149.09,124.29) --
	(145.24,124.29) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.86,102.22) --
	(139.78, 98.89) --
	(135.94, 98.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (135.13, 96.11) --
	(137.05, 92.78) --
	(133.21, 92.78) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (146.15,110.98) --
	(148.07,107.65) --
	(144.23,107.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.39, 96.11) --
	(139.31, 92.78) --
	(135.47, 92.78) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (166.63,111.47) --
	(168.55,108.14) --
	(164.71,108.14) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.73,127.05) --
	(146.66,123.72) --
	(142.81,123.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.02,114.19) --
	(142.94,110.87) --
	(139.10,110.87) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.89,108.84) --
	(142.81,105.52) --
	(138.97,105.52) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.61,138.01) --
	(144.53,134.68) --
	(140.69,134.68) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.61,103.93) --
	(143.53,100.60) --
	(139.69,100.60) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.83,118.02) --
	(139.75,114.69) --
	(135.90,114.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.70,115.25) --
	(141.62,111.92) --
	(137.78,111.92) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (148.10,115.25) --
	(150.02,111.92) --
	(146.18,111.92) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.51,122.12) --
	(143.43,118.79) --
	(139.59,118.79) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.43, 97.43) --
	(143.36, 94.10) --
	(139.51, 94.10) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.81, 99.76) --
	(140.73, 96.43) --
	(136.89, 96.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (148.84,147.94) --
	(150.76,144.62) --
	(146.92,144.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (146.19,115.43) --
	(148.11,112.10) --
	(144.26,112.10) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (155.22,133.84) --
	(157.14,130.51) --
	(153.30,130.51) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (149.19,120.21) --
	(151.11,116.88) --
	(147.27,116.88) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (146.06,113.48) --
	(147.98,110.15) --
	(144.14,110.15) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (159.21,131.26) --
	(161.13,127.93) --
	(157.29,127.93) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (152.53,133.37) --
	(154.45,130.04) --
	(150.61,130.04) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (149.90,130.63) --
	(151.83,127.30) --
	(147.98,127.30) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.64,141.88) --
	(141.56,138.55) --
	(137.72,138.55) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (147.85,148.71) --
	(149.78,145.38) --
	(145.93,145.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (150.21,157.06) --
	(152.14,153.73) --
	(148.29,153.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (174.40,141.50) --
	(176.32,138.17) --
	(172.48,138.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (162.93,133.45) --
	(164.86,130.12) --
	(161.01,130.12) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.90,135.78) --
	(146.82,132.45) --
	(142.97,132.45) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.00,151.74) --
	(146.92,148.41) --
	(143.08,148.41) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (158.43,154.98) --
	(160.35,151.65) --
	(156.50,151.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (156.96,146.12) --
	(158.88,142.79) --
	(155.04,142.79) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (154.58,129.04) --
	(156.50,125.71) --
	(152.66,125.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (149.44,137.44) --
	(151.36,134.11) --
	(147.52,134.11) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (172.46,150.40) --
	(174.38,147.07) --
	(170.54,147.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (159.52,153.06) --
	(161.45,149.73) --
	(157.60,149.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (163.76,143.86) --
	(165.69,140.53) --
	(161.84,140.53) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (156.23,147.24) --
	(158.15,143.91) --
	(154.31,143.91) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (150.98, 90.73) --
	(152.90, 87.40) --
	(149.05, 87.40) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.59,133.63) --
	(143.51,130.30) --
	(139.67,130.30) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (154.95,158.84) --
	(156.87,155.51) --
	(153.03,155.51) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (196.73,156.97) --
	(198.65,153.64) --
	(194.81,153.64) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (153.17,145.62) --
	(155.09,142.29) --
	(151.25,142.29) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (168.64,155.70) --
	(170.57,152.37) --
	(166.72,152.37) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (154.52,161.16) --
	(156.44,157.83) --
	(152.60,157.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (185.71,142.47) --
	(187.64,139.14) --
	(183.79,139.14) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (146.17,126.03) --
	(148.09,122.70) --
	(144.25,122.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (160.82,166.54) --
	(162.75,163.21) --
	(158.90,163.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (168.81,153.05) --
	(170.73,149.73) --
	(166.89,149.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (175.27,156.51) --
	(177.19,153.18) --
	(173.35,153.18) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (170.92,167.15) --
	(172.84,163.82) --
	(169.00,163.82) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (163.38,165.10) --
	(165.30,161.77) --
	(161.46,161.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (191.73,155.02) --
	(193.65,151.69) --
	(189.81,151.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,108.84) --
	(228.48,105.52) --
	(224.64,105.52) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (166.21,160.10) --
	(168.13,156.77) --
	(164.29,156.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (151.95,148.58) --
	(153.87,145.25) --
	(150.03,145.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (210.83,173.83) --
	(212.75,170.51) --
	(208.90,170.51) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (192.00,163.69) --
	(193.92,160.36) --
	(190.08,160.36) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (196.87,169.43) --
	(198.79,166.10) --
	(194.95,166.10) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (181.28,167.12) --
	(183.20,163.79) --
	(179.35,163.79) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (166.07,158.89) --
	(167.99,155.56) --
	(164.15,155.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (181.37,157.59) --
	(183.29,154.26) --
	(179.45,154.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (167.34,166.23) --
	(169.26,162.90) --
	(165.42,162.90) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (162.86,177.17) --
	(164.78,173.84) --
	(160.94,173.84) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (173.93,160.97) --
	(175.85,157.64) --
	(172.01,157.64) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (158.57,175.57) --
	(160.50,172.24) --
	(156.65,172.24) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (180.15,173.47) --
	(182.07,170.14) --
	(178.23,170.14) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (174.45,151.52) --
	(176.38,148.19) --
	(172.53,148.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (207.67,177.49) --
	(209.59,174.16) --
	(205.75,174.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (176.81,181.25) --
	(178.73,177.92) --
	(174.89,177.92) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (157.19,175.03) --
	(159.11,171.71) --
	(155.27,171.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (206.86,176.08) --
	(208.79,172.75) --
	(204.94,172.75) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (179.17,154.21) --
	(181.09,150.88) --
	(177.25,150.88) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (184.13,159.52) --
	(186.05,156.19) --
	(182.21,156.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (215.43,192.83) --
	(217.35,189.50) --
	(213.51,189.50) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (183.10,191.75) --
	(185.03,188.43) --
	(181.18,188.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,121.35) --
	(228.48,118.02) --
	(224.64,118.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (180.43,154.62) --
	(182.36,151.29) --
	(178.51,151.29) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (173.24,164.24) --
	(175.16,160.91) --
	(171.31,160.91) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (179.18,175.40) --
	(181.10,172.07) --
	(177.26,172.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.17,198.05) --
	(190.10,194.72) --
	(186.25,194.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (163.43,181.63) --
	(165.35,178.30) --
	(161.51,178.30) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (177.33,169.86) --
	(179.25,166.53) --
	(175.41,166.53) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (192.97,183.42) --
	(194.89,180.10) --
	(191.04,180.10) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (160.51,200.95) --
	(162.43,197.62) --
	(158.59,197.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (189.70,219.04) --
	(191.62,215.71) --
	(187.78,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (171.34,195.52) --
	(173.26,192.19) --
	(169.42,192.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (209.41,197.29) --
	(211.33,193.96) --
	(207.48,193.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (162.12,198.76) --
	(164.04,195.44) --
	(160.20,195.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,198.56) --
	(228.48,195.23) --
	(224.64,195.23) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (171.69,169.17) --
	(173.61,165.84) --
	(169.77,165.84) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (183.60,190.68) --
	(185.53,187.36) --
	(181.68,187.36) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (183.77,204.04) --
	(185.69,200.72) --
	(181.85,200.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (202.61,193.46) --
	(204.53,190.13) --
	(200.69,190.13) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (205.00,171.48) --
	(206.92,168.15) --
	(203.08,168.15) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (201.02,195.84) --
	(202.94,192.51) --
	(199.10,192.51) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (201.61,185.42) --
	(203.54,182.09) --
	(199.69,182.09) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,186.77) --
	(228.48,183.45) --
	(224.64,183.45) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,204.52) --
	(228.48,201.19) --
	(224.64,201.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (202.23,206.92) --
	(204.15,203.59) --
	(200.30,203.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (126.92, 73.75) --
	(128.84, 70.42) --
	(125.00, 70.42) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (127.40, 73.75) --
	(129.32, 70.42) --
	(125.48, 70.42) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (132.26, 86.72) --
	(134.19, 83.39) --
	(130.34, 83.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 73.75) --
	(228.48, 70.42) --
	(224.64, 70.42) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 76.66) --
	(228.48, 73.33) --
	(224.64, 73.33) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.09, 73.75) --
	(126.01, 70.42) --
	(122.17, 70.42) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.87, 77.92) --
	(131.79, 74.59) --
	(127.94, 74.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (128.56, 76.66) --
	(130.48, 73.33) --
	(126.64, 73.33) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (126.99, 75.27) --
	(128.91, 71.95) --
	(125.07, 71.95) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.22, 81.16) --
	(133.14, 77.83) --
	(129.30, 77.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (128.44, 76.66) --
	(130.36, 73.33) --
	(126.51, 73.33) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.78, 89.15) --
	(133.70, 85.82) --
	(129.85, 85.82) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (126.85, 76.66) --
	(128.77, 73.33) --
	(124.93, 73.33) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 77.92) --
	(228.48, 74.59) --
	(224.64, 74.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.37,103.31) --
	(133.29, 99.99) --
	(129.45, 99.99) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.20, 80.16) --
	(132.12, 76.83) --
	(128.28, 76.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (128.62, 77.92) --
	(130.54, 74.59) --
	(126.70, 74.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.37, 84.59) --
	(132.29, 81.26) --
	(128.45, 81.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.17, 77.92) --
	(131.09, 74.59) --
	(127.25, 74.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.15, 84.59) --
	(132.07, 81.26) --
	(128.22, 81.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.29, 86.72) --
	(131.21, 83.39) --
	(127.37, 83.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (132.12, 89.69) --
	(134.04, 86.37) --
	(130.20, 86.37) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.85, 81.16) --
	(132.77, 77.83) --
	(128.93, 77.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.42, 80.16) --
	(132.34, 76.83) --
	(128.50, 76.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.65, 91.22) --
	(136.57, 87.89) --
	(132.73, 87.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.01, 85.34) --
	(132.93, 82.01) --
	(129.09, 82.01) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.63, 82.98) --
	(133.55, 79.65) --
	(129.70, 79.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.81, 81.16) --
	(131.73, 77.83) --
	(127.89, 77.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.11, 87.37) --
	(133.04, 84.04) --
	(129.19, 84.04) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.82, 86.05) --
	(133.75, 82.72) --
	(129.90, 82.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.68, 83.81) --
	(133.60, 80.48) --
	(129.75, 80.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 84.59) --
	(228.48, 81.26) --
	(224.64, 81.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.11,100.02) --
	(136.03, 96.69) --
	(132.19, 96.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (132.07, 86.72) --
	(133.99, 83.39) --
	(130.15, 83.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 84.59) --
	(228.48, 81.26) --
	(224.64, 81.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.87, 92.61) --
	(133.80, 89.28) --
	(129.95, 89.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.90, 86.72) --
	(135.82, 83.39) --
	(131.98, 83.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (132.64, 85.34) --
	(134.56, 82.01) --
	(130.72, 82.01) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.15, 94.65) --
	(135.07, 91.33) --
	(131.23, 91.33) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.06, 85.34) --
	(134.98, 82.01) --
	(131.14, 82.01) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.42, 86.72) --
	(135.34, 83.39) --
	(131.50, 83.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (132.88, 94.27) --
	(134.80, 90.94) --
	(130.95, 90.94) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (135.83, 95.76) --
	(137.75, 92.43) --
	(133.91, 92.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.15, 86.72) --
	(135.07, 83.39) --
	(131.23, 83.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (135.83, 97.74) --
	(137.75, 94.41) --
	(133.91, 94.41) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.77, 98.34) --
	(136.70, 95.02) --
	(132.85, 95.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 86.72) --
	(228.48, 83.39) --
	(224.64, 83.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 86.72) --
	(228.48, 83.39) --
	(224.64, 83.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.93, 91.22) --
	(136.86, 87.89) --
	(133.01, 87.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (132.41,102.89) --
	(134.33, 99.56) --
	(130.49, 99.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 87.37) --
	(228.48, 84.04) --
	(224.64, 84.04) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.81, 95.03) --
	(136.74, 91.70) --
	(132.89, 91.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.26, 92.61) --
	(139.18, 89.28) --
	(135.34, 89.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.93, 87.98) --
	(136.86, 84.66) --
	(133.01, 84.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (135.98,105.82) --
	(137.90,102.49) --
	(134.05,102.49) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.02,104.52) --
	(138.94,101.20) --
	(135.10,101.20) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.53, 90.73) --
	(136.45, 87.40) --
	(132.61, 87.40) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.32, 92.61) --
	(136.24, 89.28) --
	(132.40, 89.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (135.41, 91.22) --
	(137.33, 87.89) --
	(133.49, 87.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.66,127.96) --
	(139.58,124.64) --
	(135.74,124.64) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.27,106.00) --
	(138.19,102.67) --
	(134.35,102.67) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.34, 90.73) --
	(138.26, 87.40) --
	(134.42, 87.40) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 90.22) --
	(228.48, 86.89) --
	(224.64, 86.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (235.42, 98.64) --
	(237.34, 95.31) --
	(233.49, 95.31) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (135.94,108.55) --
	(137.86,105.23) --
	(134.02,105.23) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.16,100.28) --
	(139.08, 96.96) --
	(135.23, 96.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.05,102.45) --
	(139.97, 99.12) --
	(136.13, 99.12) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.53, 92.16) --
	(139.45, 88.83) --
	(135.61, 88.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.81,104.13) --
	(138.73,100.80) --
	(134.89,100.80) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 92.61) --
	(228.48, 89.28) --
	(224.64, 89.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.23, 92.16) --
	(138.16, 88.83) --
	(134.31, 88.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.01, 95.76) --
	(141.93, 92.43) --
	(138.09, 92.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.33,101.28) --
	(139.25, 97.96) --
	(135.40, 97.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (152.84, 95.76) --
	(154.76, 92.43) --
	(150.92, 92.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.43,101.28) --
	(139.35, 97.96) --
	(135.51, 97.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.56,108.99) --
	(138.48,105.66) --
	(134.64,105.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.98,100.54) --
	(138.90, 97.21) --
	(135.06, 97.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.96,103.52) --
	(139.88,100.19) --
	(136.03,100.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.93, 97.11) --
	(140.85, 93.78) --
	(137.01, 93.78) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.44,107.96) --
	(141.36,104.63) --
	(137.52,104.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.28,107.01) --
	(140.20,103.68) --
	(136.36,103.68) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 95.03) --
	(228.48, 91.70) --
	(224.64, 91.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.02,100.79) --
	(140.94, 97.46) --
	(137.10, 97.46) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.45,138.39) --
	(138.37,135.06) --
	(134.53,135.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.28,103.52) --
	(140.20,100.19) --
	(136.36,100.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.25,121.47) --
	(143.18,118.14) --
	(139.33,118.14) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.37,119.05) --
	(142.30,115.72) --
	(138.45,115.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.81,113.58) --
	(142.73,110.26) --
	(138.89,110.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (167.06,123.38) --
	(168.98,120.06) --
	(165.14,120.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (148.99, 97.43) --
	(150.91, 94.10) --
	(147.07, 94.10) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (143.98,142.36) --
	(145.90,139.03) --
	(142.06,139.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 99.21) --
	(228.48, 95.88) --
	(224.64, 95.88) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (146.04,103.31) --
	(147.96, 99.99) --
	(144.12, 99.99) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (143.63,127.49) --
	(145.56,124.17) --
	(141.71,124.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.75,135.89) --
	(140.67,132.57) --
	(136.82,132.57) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (168.50,144.50) --
	(170.42,141.18) --
	(166.58,141.18) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 98.93) --
	(228.48, 95.60) --
	(224.64, 95.60) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 98.64) --
	(228.48, 95.31) --
	(224.64, 95.31) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (158.53,153.30) --
	(160.45,149.97) --
	(156.61,149.97) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.47,161.40) --
	(147.39,158.07) --
	(143.55,158.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (143.92,109.95) --
	(145.84,106.62) --
	(142.00,106.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (147.87,119.33) --
	(149.79,116.00) --
	(145.95,116.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (195.43,146.11) --
	(197.35,142.78) --
	(193.51,142.78) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (158.45,219.04) --
	(160.37,215.71) --
	(156.53,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (160.83,122.17) --
	(162.75,118.85) --
	(158.91,118.85) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (152.36,154.99) --
	(154.29,151.66) --
	(150.44,151.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (151.92,144.15) --
	(153.85,140.83) --
	(150.00,140.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (155.40,114.59) --
	(157.33,111.26) --
	(153.48,111.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (181.00,144.42) --
	(182.92,141.09) --
	(179.08,141.09) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (213.32,148.65) --
	(215.24,145.32) --
	(211.40,145.32) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (166.42,122.35) --
	(168.34,119.02) --
	(164.49,119.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (193.50,185.58) --
	(195.43,182.25) --
	(191.58,182.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,118.09) --
	(228.48,114.77) --
	(224.64,114.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (196.07,219.04) --
	(197.99,215.71) --
	(194.15,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (211.78,160.83) --
	(213.71,157.50) --
	(209.86,157.50) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (178.86,125.76) --
	(180.79,122.44) --
	(176.94,122.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (213.65,168.77) --
	(215.57,165.44) --
	(211.73,165.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,139.65) --
	(228.48,136.32) --
	(224.64,136.32) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (235.42,160.64) --
	(237.34,157.31) --
	(233.49,157.31) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,185.18) --
	(228.48,181.86) --
	(224.64,181.86) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,131.59) --
	(228.48,128.26) --
	(224.64,128.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,189.63) --
	(228.48,186.30) --
	(224.64,186.30) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,167.30) --
	(228.48,163.98) --
	(224.64,163.98) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,185.67) --
	(228.48,182.35) --
	(224.64,182.35) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,127.89) --
	(228.48,124.56) --
	(224.64,124.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (198.27,219.04) --
	(200.19,215.71) --
	(196.35,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,179.04) --
	(228.48,175.71) --
	(224.64,175.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (203.99,219.04) --
	(205.91,215.71) --
	(202.07,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.10, 92.46) rectangle (132.95, 95.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.10, 92.46) -- (132.95, 95.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.10, 95.32) -- (132.95, 92.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.56, 94.70) rectangle (135.41, 97.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.56, 94.70) -- (135.41, 97.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.56, 97.55) -- (135.41, 94.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) rectangle (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.75,215.39) rectangle (141.61,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.75,215.39) -- (141.61,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.75,218.25) -- (141.61,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.93,215.39) rectangle (153.78,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.93,215.39) -- (153.78,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.93,218.25) -- (153.78,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,127.52) rectangle (227.98,130.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,127.52) -- (227.98,130.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,130.37) -- (227.98,127.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,139.97) rectangle (227.98,142.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,139.97) -- (227.98,142.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,142.82) -- (227.98,139.97);
\definecolor{drawColor}{RGB}{152,152,152}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.51, 30.77) -- (235.42,225.68);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.95, 30.77) -- ( 41.95, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 44.19, 30.77) -- ( 44.19, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 46.13, 30.77) -- ( 46.13, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 47.84, 30.77) -- ( 47.84, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 49.37, 30.77) -- ( 49.37, 39.31);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 59.43, 30.77) -- ( 59.43, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 65.32, 30.77) -- ( 65.32, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 69.49, 30.77) -- ( 69.49, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 72.73, 30.77) -- ( 72.73, 36.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 75.38, 30.77) -- ( 75.38, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 77.62, 30.77) -- ( 77.62, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 79.56, 30.77) -- ( 79.56, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 81.27, 30.77) -- ( 81.27, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 82.79, 30.77) -- ( 82.79, 39.31);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 92.86, 30.77) -- ( 92.86, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 98.74, 30.77) -- ( 98.74, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (102.92, 30.77) -- (102.92, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (106.16, 30.77) -- (106.16, 36.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (108.80, 30.77) -- (108.80, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (111.04, 30.77) -- (111.04, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (112.98, 30.77) -- (112.98, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (114.69, 30.77) -- (114.69, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (116.22, 30.77) -- (116.22, 39.31);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (126.28, 30.77) -- (126.28, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (132.17, 30.77) -- (132.17, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (136.34, 30.77) -- (136.34, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (139.58, 30.77) -- (139.58, 36.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.23, 30.77) -- (142.23, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (144.47, 30.77) -- (144.47, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (146.41, 30.77) -- (146.41, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (148.12, 30.77) -- (148.12, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (149.64, 30.77) -- (149.64, 39.31);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (159.71, 30.77) -- (159.71, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (165.59, 30.77) -- (165.59, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (169.77, 30.77) -- (169.77, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (173.01, 30.77) -- (173.01, 36.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (175.65, 30.77) -- (175.65, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (177.89, 30.77) -- (177.89, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (179.83, 30.77) -- (179.83, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (181.54, 30.77) -- (181.54, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (183.07, 30.77) -- (183.07, 39.31);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (193.13, 30.77) -- (193.13, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (199.02, 30.77) -- (199.02, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (203.19, 30.77) -- (203.19, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (206.43, 30.77) -- (206.43, 36.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (209.08, 30.77) -- (209.08, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (211.32, 30.77) -- (211.32, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (213.26, 30.77) -- (213.26, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (214.97, 30.77) -- (214.97, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (216.49, 30.77) -- (216.49, 39.31);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (226.56, 30.77) -- (226.56, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (232.44, 30.77) -- (232.44, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 32.22) -- ( 43.36, 32.22);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 34.45) -- ( 43.36, 34.45);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 36.39) -- ( 43.36, 36.39);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 38.10) -- ( 43.36, 38.10);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 39.63) -- ( 49.05, 39.63);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 49.69) -- ( 43.36, 49.69);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 55.58) -- ( 43.36, 55.58);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 59.76) -- ( 43.36, 59.76);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 62.99) -- ( 46.20, 62.99);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 65.64) -- ( 43.36, 65.64);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 67.88) -- ( 43.36, 67.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 69.82) -- ( 43.36, 69.82);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 71.53) -- ( 43.36, 71.53);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 73.06) -- ( 49.05, 73.06);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 83.12) -- ( 43.36, 83.12);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 89.00) -- ( 43.36, 89.00);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 93.18) -- ( 43.36, 93.18);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 96.42) -- ( 46.20, 96.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 99.07) -- ( 43.36, 99.07);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,101.30) -- ( 43.36,101.30);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,103.24) -- ( 43.36,103.24);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,104.95) -- ( 43.36,104.95);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,106.48) -- ( 49.05,106.48);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,116.54) -- ( 43.36,116.54);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,122.43) -- ( 43.36,122.43);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,126.61) -- ( 43.36,126.61);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,129.84) -- ( 46.20,129.84);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,132.49) -- ( 43.36,132.49);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,134.73) -- ( 43.36,134.73);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,136.67) -- ( 43.36,136.67);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,138.38) -- ( 43.36,138.38);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,139.91) -- ( 49.05,139.91);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,149.97) -- ( 43.36,149.97);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,155.85) -- ( 43.36,155.85);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,160.03) -- ( 43.36,160.03);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,163.27) -- ( 46.20,163.27);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,165.92) -- ( 43.36,165.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,168.15) -- ( 43.36,168.15);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,170.09) -- ( 43.36,170.09);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,171.80) -- ( 43.36,171.80);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,173.33) -- ( 49.05,173.33);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,183.39) -- ( 43.36,183.39);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,189.28) -- ( 43.36,189.28);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,193.46) -- ( 43.36,193.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,196.69) -- ( 46.20,196.69);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,199.34) -- ( 43.36,199.34);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,201.58) -- ( 43.36,201.58);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,203.52) -- ( 43.36,203.52);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,205.23) -- ( 43.36,205.23);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,206.76) -- ( 49.05,206.76);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,216.82) -- ( 43.36,216.82);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,222.70) -- ( 43.36,222.70);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 30.77) rectangle (235.42,225.68);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.56, 70.03) {0.1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.56,136.88) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.56,203.73) {1000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 37.76, 73.06) --
	( 40.51, 73.06);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 37.76,139.91) --
	( 40.51,139.91);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 37.76,206.76) --
	( 40.51,206.76);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 82.79, 28.02) --
	( 82.79, 30.77);

\path[draw=drawColor,line width= 0.3pt,line join=round] (149.64, 28.02) --
	(149.64, 30.77);

\path[draw=drawColor,line width= 0.3pt,line join=round] (216.49, 28.02) --
	(216.49, 30.77);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 82.79, 19.76) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (149.64, 19.76) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (216.49, 19.76) {1000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (137.96,  7.72) {\textsf{Ace} + \texttt{cd06} time (s)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,128.22) {\textsf{ADDMC} + \texttt{cw} time (s)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (246.42, 71.76) rectangle (337.78,184.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (251.92,170.55) {Data set};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (251.92,149.53) rectangle (266.37,163.98);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor,fill opacity=0.50] (259.14,156.75) circle (  1.43);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (251.92,135.07) rectangle (266.37,149.53);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor,fill opacity=0.50] (259.14,144.52) --
	(261.06,141.19) --
	(257.22,141.19) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (251.92,120.62) rectangle (266.37,135.07);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor,fill opacity=0.50] (257.72,126.42) --
	(260.57,126.42) --
	(260.57,129.27) --
	(257.72,129.27) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (251.92,106.16) rectangle (266.37,120.62);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (257.13,113.39) -- (261.16,113.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (259.14,111.37) -- (259.14,115.41);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (251.92, 91.71) rectangle (266.37,106.16);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (257.72, 97.51) rectangle (260.57,100.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (257.72, 97.51) -- (260.57,100.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (257.72,100.36) -- (260.57, 97.51);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (251.92, 77.26) rectangle (266.37, 91.71);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (257.72, 83.06) -- (260.57, 85.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (257.72, 85.91) -- (260.57, 83.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (257.13, 84.48) -- (261.16, 84.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (259.14, 82.47) -- (259.14, 86.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (271.87,153.72) {DQMR};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (271.87,139.27) {Grid};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (271.87,124.81) {Mastermind};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (271.87,110.36) {Non-binary};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (271.87, 95.91) {Other binary};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (271.87, 81.45) {Random Blocks};
\end{scope}
\end{tikzpicture}