### 2021-06-02 doc/long_talk/scatter2.tex

% Created by tikzDevice version 0.12.3.1 on 2021-06-01 14:27:10
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (343.28,231.26);
\begin{scope}
\path[clip] (  0.00,  0.09) rectangle (343.28,231.18);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.09) rectangle (343.28,231.18);
\end{scope}
\begin{scope}
\path[clip] ( 40.51, 30.77) rectangle (235.42,225.68);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 40.51, 30.77) rectangle (235.42,225.68);
\definecolor{drawColor}{gray}{0.87}

\path[draw=drawColor,line width= 0.1pt,line join=round] ( 40.51, 39.63) --
	(235.42, 39.63);

\path[draw=drawColor,line width= 0.1pt,line join=round] ( 40.51,106.48) --
	(235.42,106.48);

\path[draw=drawColor,line width= 0.1pt,line join=round] ( 40.51,173.33) --
	(235.42,173.33);

\path[draw=drawColor,line width= 0.1pt,line join=round] ( 49.37, 30.77) --
	( 49.37,225.68);

\path[draw=drawColor,line width= 0.1pt,line join=round] (116.22, 30.77) --
	(116.22,225.68);

\path[draw=drawColor,line width= 0.1pt,line join=round] (183.07, 30.77) --
	(183.07,225.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 40.51, 73.06) --
	(235.42, 73.06);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 40.51,139.91) --
	(235.42,139.91);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 40.51,206.76) --
	(235.42,206.76);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 82.79, 30.77) --
	( 82.79,225.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] (149.64, 30.77) --
	(149.64,225.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] (216.49, 30.77) --
	(216.49,225.68);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.10,215.39) -- (131.96,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.10,218.25) -- (131.96,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.51,216.82) -- (132.55,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.53,214.80) -- (130.53,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.16,215.39) -- (132.01,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.16,218.25) -- (132.01,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.57,216.82) -- (132.60,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.58,214.80) -- (130.58,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.03,215.39) -- (133.88,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.03,218.25) -- (133.88,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.44,216.82) -- (134.47,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.45,214.80) -- (132.45,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.10,215.39) -- (135.96,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.10,218.25) -- (135.96,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.51,216.82) -- (136.55,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.53,214.80) -- (134.53,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.21,215.39) -- (132.06,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.21,218.25) -- (132.06,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.62,216.82) -- (132.66,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.64,214.80) -- (130.64,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.31,215.39) -- (134.16,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.31,218.25) -- (134.16,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.72,216.82) -- (134.75,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.74,214.80) -- (132.74,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.32,101.07) -- (132.17,103.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.32,103.92) -- (132.17,101.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.73,102.50) -- (132.76,102.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.75,100.48) -- (130.75,104.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.99,101.26) -- (131.85,104.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.99,104.11) -- (131.85,101.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.40,102.69) -- (132.44,102.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.42,100.67) -- (130.42,104.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.93,101.45) -- (133.79,104.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.93,104.30) -- (133.79,101.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.34,102.87) -- (134.38,102.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.36,100.86) -- (132.36,104.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.01, 97.15) -- (129.86,100.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.01,100.00) -- (129.86, 97.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.42, 98.57) -- (130.45, 98.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.44, 96.56) -- (128.44,100.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.80, 98.35) -- (130.65,101.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.80,101.20) -- (130.65, 98.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.21, 99.77) -- (131.25, 99.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.23, 97.76) -- (129.23,101.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.42,215.39) -- (132.28,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.42,218.25) -- (132.28,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.83,216.82) -- (132.87,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.85,214.80) -- (130.85,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.39,215.39) -- (135.24,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.39,218.25) -- (135.24,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.80,216.82) -- (135.83,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.81,214.80) -- (133.81,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.42,215.39) -- (132.28,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.42,218.25) -- (132.28,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.83,216.82) -- (132.87,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.85,214.80) -- (130.85,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.22,215.39) -- (134.07,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.22,218.25) -- (134.07,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.63,216.82) -- (134.66,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.64,214.80) -- (132.64,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.50,215.39) -- (134.35,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.50,218.25) -- (134.35,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.90,216.82) -- (134.94,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.92,214.80) -- (132.92,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.81,112.67) -- (137.66,115.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.81,115.53) -- (137.66,112.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.22,114.10) -- (138.25,114.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.23,112.08) -- (136.23,116.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.67,106.70) -- (136.52,109.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.67,109.55) -- (136.52,106.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.08,108.13) -- (137.11,108.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.09,106.11) -- (135.09,110.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.95,106.57) -- (134.80,109.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.95,109.42) -- (134.80,106.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.36,108.00) -- (135.39,108.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.37,105.98) -- (133.37,110.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.03,105.90) -- (133.88,108.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.03,108.75) -- (133.88,105.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.44,107.33) -- (134.47,107.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.45,105.31) -- (132.45,109.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.39,107.33) -- (135.24,110.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.39,110.19) -- (135.24,107.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.80,108.76) -- (135.83,108.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.81,106.74) -- (133.81,110.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.06,106.57) -- (135.91,109.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.06,109.42) -- (135.91,106.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.47,108.00) -- (136.51,108.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.49,105.98) -- (134.49,110.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.68,100.08) -- (134.53,102.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.68,102.94) -- (134.53,100.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.09,101.51) -- (135.12,101.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.10, 99.49) -- (133.10,103.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.47,101.07) -- (136.32,103.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.47,103.92) -- (136.32,101.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.88,102.50) -- (136.91,102.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.89,100.48) -- (134.89,104.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.36,101.07) -- (134.21,103.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.36,103.92) -- (134.21,101.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.77,102.50) -- (134.80,102.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.78,100.48) -- (132.78,104.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.06,104.46) -- (136.91,107.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.06,107.32) -- (136.91,104.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.47,105.89) -- (137.50,105.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.48,103.87) -- (135.48,107.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.56,102.17) -- (135.41,105.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.56,105.03) -- (135.41,102.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.97,103.60) -- (136.00,103.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.98,101.58) -- (133.98,105.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.52,106.17) -- (135.37,109.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.52,109.03) -- (135.37,106.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.92,107.60) -- (135.96,107.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.94,105.58) -- (133.94,109.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.45,106.44) -- (134.30,109.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.45,109.29) -- (134.30,106.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.86,107.86) -- (134.89,107.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.88,105.85) -- (132.88,109.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.02,107.94) -- (136.87,110.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.02,110.79) -- (136.87,107.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.43,109.37) -- (137.46,109.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.45,107.35) -- (135.45,111.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.08,107.58) -- (134.93,110.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.08,110.43) -- (134.93,107.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.49,109.01) -- (135.52,109.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.51,106.99) -- (133.51,111.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.63,107.33) -- (136.48,110.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.63,110.19) -- (136.48,107.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.04,108.76) -- (137.07,108.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.05,106.74) -- (135.05,110.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.47,112.24) -- (141.33,115.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.47,115.09) -- (141.33,112.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.88,113.66) -- (141.92,113.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.90,111.64) -- (139.90,115.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.91,111.04) -- (139.77,113.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.91,113.89) -- (139.77,111.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.32,112.46) -- (140.36,112.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.34,110.45) -- (138.34,114.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.10,114.75) -- (138.95,117.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.10,117.60) -- (138.95,114.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.51,116.18) -- (139.55,116.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.53,114.16) -- (137.53,118.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.13,112.67) -- (140.98,115.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.13,115.53) -- (140.98,112.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.54,114.10) -- (141.57,114.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.55,112.08) -- (139.55,116.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.39,116.50) -- (141.24,119.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.39,119.35) -- (141.24,116.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.80,117.93) -- (141.83,117.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.81,115.91) -- (139.81,119.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.83,111.42) -- (138.68,114.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.83,114.27) -- (138.68,111.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.24,112.84) -- (139.28,112.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.26,110.83) -- (137.26,114.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.42,111.60) -- (138.27,114.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.42,114.46) -- (138.27,111.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.83,113.03) -- (138.86,113.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.84,111.01) -- (136.84,115.05);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.47,105.05) -- (136.32,107.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.47,107.91) -- (136.32,105.05);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.88,106.48) -- (136.91,106.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.89,104.46) -- (134.89,108.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.27,107.70) -- (139.12,110.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.27,110.55) -- (139.12,107.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.68,109.13) -- (139.71,109.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.69,107.11) -- (137.69,111.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.73,112.59) -- (138.58,115.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.73,115.44) -- (138.58,112.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.14,114.01) -- (139.17,114.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.16,111.99) -- (137.16,116.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.47,105.48) -- (137.33,108.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.47,108.34) -- (137.33,105.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.88,106.91) -- (137.92,106.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.90,104.89) -- (135.90,108.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.53,121.10) -- (140.39,123.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.53,123.95) -- (140.39,121.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.94,122.53) -- (140.98,122.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.96,120.51) -- (138.96,124.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.95,116.43) -- (137.81,119.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.95,119.29) -- (137.81,116.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.36,117.86) -- (138.40,117.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.38,115.84) -- (136.38,119.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.84,109.84) -- (137.70,112.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.84,112.69) -- (137.70,109.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.25,111.26) -- (138.29,111.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.27,109.24) -- (136.27,113.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.73,109.19) -- (137.59,112.05);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.73,112.05) -- (137.59,109.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.14,110.62) -- (138.18,110.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.16,108.60) -- (136.16,112.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.97,111.23) -- (138.82,114.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.97,114.08) -- (138.82,111.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.38,112.65) -- (139.41,112.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.39,110.64) -- (137.39,114.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.33,187.58) -- (173.18,190.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.33,190.43) -- (173.18,187.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.74,189.01) -- (173.77,189.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.76,186.99) -- (171.76,191.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.07,155.90) -- (171.93,158.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.07,158.76) -- (171.93,155.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.48,157.33) -- (172.52,157.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.50,155.31) -- (170.50,159.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.16,152.68) -- (170.01,155.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.16,155.53) -- (170.01,152.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.57,154.11) -- (170.60,154.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.59,152.09) -- (168.59,156.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.77,156.31) -- (173.62,159.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.77,159.16) -- (173.62,156.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.18,157.74) -- (174.21,157.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.19,155.72) -- (172.19,159.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.12,153.70) -- (169.98,156.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.12,156.55) -- (169.98,153.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.53,155.12) -- (170.57,155.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.55,153.11) -- (168.55,157.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.59,215.39) -- (201.44,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.59,218.25) -- (201.44,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.00,216.82) -- (202.04,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (200.02,214.80) -- (200.02,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.49,215.39) -- (201.34,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.49,218.25) -- (201.34,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.90,216.82) -- (201.93,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (199.91,214.80) -- (199.91,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (201.11,215.39) -- (203.97,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (201.11,218.25) -- (203.97,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (200.52,216.82) -- (204.56,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (202.54,214.80) -- (202.54,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.62,215.39) -- (201.47,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.62,218.25) -- (201.47,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.03,216.82) -- (202.06,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (200.04,214.80) -- (200.04,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (199.02,215.39) -- (201.88,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (199.02,218.25) -- (201.88,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.43,216.82) -- (202.47,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (200.45,214.80) -- (200.45,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.58,215.39) -- (200.43,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.58,218.25) -- (200.43,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (196.98,216.82) -- (201.02,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (199.00,214.80) -- (199.00,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.36,155.61) -- (173.22,158.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.36,158.47) -- (173.22,155.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.77,157.04) -- (173.81,157.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.79,155.02) -- (171.79,159.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.84,158.55) -- (173.70,161.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.84,161.41) -- (173.70,158.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.25,159.98) -- (174.29,159.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.27,157.96) -- (172.27,162.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.78,154.94) -- (173.63,157.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.78,157.79) -- (173.63,154.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.18,156.37) -- (174.22,156.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.20,154.35) -- (172.20,158.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.67,156.62) -- (172.52,159.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.67,159.47) -- (172.52,156.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.08,158.04) -- (173.11,158.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.10,156.02) -- (171.10,160.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.54,155.64) -- (172.39,158.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.54,158.49) -- (172.39,155.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.95,157.07) -- (172.98,157.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.97,155.05) -- (170.97,159.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (196.80,215.39) -- (199.65,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (196.80,218.25) -- (199.65,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (196.21,216.82) -- (200.24,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.23,214.80) -- (198.23,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (196.61,215.39) -- (199.47,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (196.61,218.25) -- (199.47,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (196.02,216.82) -- (200.06,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.04,214.80) -- (198.04,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.67,215.39) -- (201.53,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.67,218.25) -- (201.53,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.08,216.82) -- (202.12,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (200.10,214.80) -- (200.10,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.33,215.39) -- (200.19,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.33,218.25) -- (200.19,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (196.74,216.82) -- (200.78,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.76,214.80) -- (198.76,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.23,215.39) -- (200.08,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.23,218.25) -- (200.08,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (196.64,216.82) -- (200.67,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.65,214.80) -- (198.65,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.37,160.28) -- (175.23,163.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.37,163.14) -- (175.23,160.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.78,161.71) -- (175.82,161.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (173.80,159.69) -- (173.80,163.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.39,165.91) -- (175.25,168.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.39,168.77) -- (175.25,165.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.80,167.34) -- (175.84,167.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (173.82,165.32) -- (173.82,169.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.69,159.28) -- (175.54,162.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.69,162.13) -- (175.54,159.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.10,160.71) -- (176.13,160.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (174.11,158.69) -- (174.11,162.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.75,157.91) -- (174.60,160.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.75,160.77) -- (174.60,157.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.16,159.34) -- (175.19,159.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (173.18,157.32) -- (173.18,161.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (213.38,215.39) -- (216.23,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (213.38,218.25) -- (216.23,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (212.79,216.82) -- (216.82,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (214.80,214.80) -- (214.80,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (212.80,215.39) -- (215.66,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (212.80,218.25) -- (215.66,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (212.21,216.82) -- (216.25,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (214.23,214.80) -- (214.23,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (212.25,215.39) -- (215.11,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (212.25,218.25) -- (215.11,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (211.66,216.82) -- (215.70,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (213.68,214.80) -- (213.68,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (212.15,215.39) -- (215.00,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (212.15,218.25) -- (215.00,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (211.56,216.82) -- (215.59,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (213.58,214.80) -- (213.58,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (212.74,215.39) -- (215.59,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (212.74,218.25) -- (215.59,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (212.15,216.82) -- (216.18,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (214.17,214.80) -- (214.17,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (213.25,215.39) -- (216.10,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (213.25,218.25) -- (216.10,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (212.66,216.82) -- (216.69,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (214.68,214.80) -- (214.68,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (199.41,215.39) -- (202.26,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (199.41,218.25) -- (202.26,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.82,216.82) -- (202.85,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (200.83,214.80) -- (200.83,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (199.56,195.92) -- (202.41,198.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (199.56,198.78) -- (202.41,195.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.96,197.35) -- (203.00,197.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (200.98,195.33) -- (200.98,199.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.78,215.39) -- (200.64,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.78,218.25) -- (200.64,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.19,216.82) -- (201.23,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (199.21,214.80) -- (199.21,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (195.83,215.39) -- (198.68,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (195.83,218.25) -- (198.68,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (195.24,216.82) -- (199.27,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.26,214.80) -- (197.26,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (204.81,215.39) -- (207.67,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (204.81,218.25) -- (207.67,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (204.22,216.82) -- (208.26,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (206.24,214.80) -- (206.24,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (212.05,215.39) -- (214.91,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (212.05,218.25) -- (214.91,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (211.46,216.82) -- (215.50,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (213.48,214.80) -- (213.48,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (213.08,215.39) -- (215.93,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (213.08,218.25) -- (215.93,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (212.49,216.82) -- (216.52,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (214.50,214.80) -- (214.50,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (212.29,215.39) -- (215.15,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (212.29,218.25) -- (215.15,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (211.70,216.82) -- (215.74,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (213.72,214.80) -- (213.72,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (211.65,215.39) -- (214.50,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (211.65,218.25) -- (214.50,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (211.06,216.82) -- (215.09,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (213.07,214.80) -- (213.07,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (214.63,215.39) -- (217.48,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (214.63,218.25) -- (217.48,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (214.04,216.82) -- (218.07,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (216.05,214.80) -- (216.05,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (200.65,188.01) -- (203.50,190.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (200.65,190.86) -- (203.50,188.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (200.05,189.44) -- (204.09,189.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (202.07,187.42) -- (202.07,191.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.12,195.89) -- (200.97,198.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.12,198.74) -- (200.97,195.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.52,197.32) -- (201.56,197.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (199.54,195.30) -- (199.54,199.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.68,185.93) -- (201.53,188.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.68,188.79) -- (201.53,185.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.09,187.36) -- (202.12,187.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (200.11,185.34) -- (200.11,189.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.71,194.82) -- (200.56,197.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.71,197.67) -- (200.56,194.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.12,196.24) -- (201.15,196.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (199.13,194.23) -- (199.13,198.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (199.19,215.39) -- (202.04,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (199.19,218.25) -- (202.04,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.60,216.82) -- (202.63,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (200.61,214.80) -- (200.61,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (202.43,215.39) -- (205.28,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (202.43,218.25) -- (205.28,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (201.84,216.82) -- (205.87,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (203.86,214.80) -- (203.86,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (203.61,192.67) -- (206.46,195.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (203.61,195.52) -- (206.46,192.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (203.02,194.09) -- (207.06,194.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (205.04,192.07) -- (205.04,196.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (200.26,187.64) -- (203.12,190.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (200.26,190.50) -- (203.12,187.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (199.67,189.07) -- (203.71,189.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (201.69,187.05) -- (201.69,191.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (200.39,187.48) -- (203.24,190.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (200.39,190.33) -- (203.24,187.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (199.80,188.90) -- (203.83,188.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (201.82,186.88) -- (201.82,190.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (148.45,119.58) --
	(151.30,119.58) --
	(151.30,122.43) --
	(148.45,122.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.06,116.37) --
	(141.91,116.37) --
	(141.91,119.22) --
	(139.06,119.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (146.40,156.75) --
	(149.25,156.75) --
	(149.25,159.60) --
	(146.40,159.60) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (174.58,132.18) --
	(177.43,132.18) --
	(177.43,135.03) --
	(174.58,135.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (182.22,171.06) --
	(185.08,171.06) --
	(185.08,173.91) --
	(182.22,173.91) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (158.02,139.37) --
	(160.87,139.37) --
	(160.87,142.22) --
	(158.02,142.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (204.49,139.66) --
	(207.34,139.66) --
	(207.34,142.52) --
	(204.49,142.52) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,141.86) --
	(227.98,141.86) --
	(227.98,144.71) --
	(225.13,144.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,199.65) --
	(227.98,199.65) --
	(227.98,202.50) --
	(225.13,202.50) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (213.33,132.34) --
	(216.18,132.34) --
	(216.18,135.19) --
	(213.33,135.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (166.63,136.32) --
	(169.48,136.32) --
	(169.48,139.18) --
	(166.63,139.18) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.87,124.28) --
	(145.73,124.28) --
	(145.73,127.13) --
	(142.87,127.13) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (146.11,119.84) --
	(148.97,119.84) --
	(148.97,122.70) --
	(146.11,122.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (211.17,215.39) --
	(214.02,215.39) --
	(214.02,218.25) --
	(211.17,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (161.07,160.09) --
	(163.92,160.09) --
	(163.92,162.94) --
	(161.07,162.94) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (154.80,140.76) --
	(157.66,140.76) --
	(157.66,143.61) --
	(154.80,143.61) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (198.07,215.39) --
	(200.92,215.39) --
	(200.92,218.25) --
	(198.07,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (150.63,117.64) --
	(153.49,117.64) --
	(153.49,120.50) --
	(150.63,120.50) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (150.67,117.94) --
	(153.52,117.94) --
	(153.52,120.80) --
	(150.67,120.80) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,181.79) --
	(227.98,181.79) --
	(227.98,184.64) --
	(225.13,184.64) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.60,116.50) --
	(148.45,116.50) --
	(148.45,119.35) --
	(145.60,119.35) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (208.53,215.39) --
	(211.38,215.39) --
	(211.38,218.25) --
	(208.53,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (210.25,188.53) --
	(213.11,188.53) --
	(213.11,191.38) --
	(210.25,191.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (194.87,154.29) --
	(197.72,154.29) --
	(197.72,157.14) --
	(194.87,157.14) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (152.10,141.43) --
	(154.96,141.43) --
	(154.96,144.28) --
	(152.10,144.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (156.92,215.39) --
	(159.77,215.39) --
	(159.77,218.25) --
	(156.92,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.33,215.39) --
	(191.19,215.39) --
	(191.19,218.25) --
	(188.33,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.92,136.70) -- (171.95,136.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.93,134.69) -- (169.93,138.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.25,109.37) -- (166.28,109.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.27,107.35) -- (164.27,111.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.40,115.49) -- (166.44,115.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.42,113.47) -- (164.42,117.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.25,134.22) -- (169.28,134.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.26,132.20) -- (167.26,136.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.75,125.55) -- (167.78,125.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.77,123.53) -- (165.77,127.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.40,130.80) -- (167.44,130.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.42,128.78) -- (165.42,132.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.71,135.92) -- (169.75,135.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.73,133.91) -- (167.73,137.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.15,119.19) -- (169.18,119.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.16,117.17) -- (167.16,121.21);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.26,135.75) -- (169.30,135.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.28,133.73) -- (167.28,137.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.27,125.86) -- (166.30,125.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.29,123.84) -- (164.29,127.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.79,137.48) -- (167.82,137.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.80,135.46) -- (165.80,139.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.60,133.59) -- (168.63,133.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.62,131.57) -- (166.62,135.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.17,134.09) -- (168.21,134.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.19,132.08) -- (166.19,136.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.09,133.11) -- (168.12,133.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.11,131.09) -- (166.11,135.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.19,135.28) -- (168.23,135.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.21,133.26) -- (166.21,137.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.64,131.49) -- (164.67,131.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.65,129.47) -- (162.65,133.51);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.17,113.98) -- (142.02,116.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.17,116.84) -- (142.02,113.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.57,115.41) -- (142.61,115.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.59,113.39) -- (140.59,117.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.27,114.97) -- (142.13,117.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.27,117.82) -- (142.13,114.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.68,116.40) -- (142.72,116.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.70,114.38) -- (140.70,118.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.73,119.90) -- (143.58,122.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.73,122.75) -- (143.58,119.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.14,121.32) -- (144.17,121.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.16,119.31) -- (142.16,123.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.86,113.83) -- (140.72,116.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.86,116.68) -- (140.72,113.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.27,115.25) -- (141.31,115.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.29,113.24) -- (139.29,117.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.75,117.46) -- (142.60,120.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.75,120.31) -- (142.60,117.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.16,118.88) -- (143.19,118.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.18,116.87) -- (141.18,120.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.36,114.30) -- (141.21,117.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.36,117.15) -- (141.21,114.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.77,115.72) -- (141.80,115.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.78,113.70) -- (139.78,117.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.98,114.06) -- (136.83,116.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.98,116.92) -- (136.83,114.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.39,115.49) -- (137.42,115.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.41,113.47) -- (135.41,117.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.66,109.94) -- (139.51,112.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.66,112.79) -- (139.51,109.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.07,111.37) -- (140.10,111.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.09,109.35) -- (138.09,113.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.30,110.35) -- (139.15,113.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.30,113.20) -- (139.15,110.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.71,111.77) -- (139.74,111.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.73,109.76) -- (137.73,113.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.27,108.75) -- (139.12,111.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.27,111.60) -- (139.12,108.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.68,110.18) -- (139.71,110.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.69,108.16) -- (137.69,112.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.45,108.52) -- (138.30,111.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.45,111.38) -- (138.30,108.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.86,109.95) -- (138.90,109.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.88,107.93) -- (136.88,111.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.17,116.23) -- (140.02,119.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.17,119.09) -- (140.02,116.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.57,117.66) -- (140.61,117.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.59,115.64) -- (138.59,119.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.65,118.36) -- (140.51,121.21);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.65,121.21) -- (140.51,118.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.06,119.78) -- (141.10,119.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.08,117.76) -- (139.08,121.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.63,112.32) -- (138.48,115.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.63,115.18) -- (138.48,112.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.03,113.75) -- (139.07,113.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.05,111.73) -- (137.05,115.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.30,111.79) -- (139.15,114.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.30,114.64) -- (139.15,111.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.71,113.21) -- (139.74,113.21);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.73,111.20) -- (137.73,115.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.96,114.37) -- (144.82,117.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.96,117.23) -- (144.82,114.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.37,115.80) -- (145.41,115.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.39,113.78) -- (143.39,117.82);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,206.31) -- (228.57,206.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,204.29) -- (226.56,208.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,204.05) -- (228.57,204.05);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,202.03) -- (226.56,206.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,205.62) -- (228.57,205.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,203.60) -- (226.56,207.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,206.47) -- (228.57,206.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,204.45) -- (226.56,208.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,204.42) -- (228.57,204.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,202.40) -- (226.56,206.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,205.41) -- (228.57,205.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,203.39) -- (226.56,207.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,204.52) -- (228.57,204.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,202.50) -- (226.56,206.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,205.41) -- (228.57,205.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,203.40) -- (226.56,207.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,205.79) -- (228.57,205.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,203.77) -- (226.56,207.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,192.95) -- (228.57,192.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,190.93) -- (226.56,194.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,204.31) -- (228.57,204.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,202.29) -- (226.56,206.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,206.20) -- (228.57,206.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,204.18) -- (226.56,208.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,205.97) -- (228.57,205.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,203.95) -- (226.56,207.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,203.78) -- (228.57,203.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,201.76) -- (226.56,205.80);

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (152.41,127.40) --
	(155.27,127.40) --
	(155.27,130.25) --
	(152.41,130.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.40,120.26) --
	(145.26,120.26) --
	(145.26,123.11) --
	(142.40,123.11) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (148.16,157.01) --
	(151.01,157.01) --
	(151.01,159.86) --
	(148.16,159.86) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (225.13,215.39) --
	(227.98,215.39) --
	(227.98,218.25) --
	(225.13,218.25) --
	cycle;

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,171.45) -- (228.57,171.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,169.44) -- (226.56,173.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,171.18) -- (228.57,171.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,169.16) -- (226.56,173.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.24,120.51) rectangle (171.09,123.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.24,120.51) -- (171.09,123.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.24,123.36) -- (171.09,120.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.31,119.04) rectangle (171.16,121.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.31,119.04) -- (171.16,121.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.31,121.89) -- (171.16,119.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.85,112.32) rectangle (167.70,115.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.85,112.32) -- (167.70,115.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.85,115.18) -- (167.70,112.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.66,113.51) rectangle (168.52,116.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.66,113.51) -- (168.52,116.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.66,116.36) -- (168.52,113.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.08,119.90) rectangle (171.93,122.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.08,119.90) -- (171.93,122.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.08,122.75) -- (171.93,119.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.94,111.88) rectangle (166.79,114.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.94,111.88) -- (166.79,114.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.94,114.73) -- (166.79,111.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.74, 91.01) rectangle (132.59, 93.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.74, 91.01) -- (132.59, 93.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.74, 93.86) -- (132.59, 91.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.21, 84.93) rectangle (128.07, 87.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.21, 84.93) -- (128.07, 87.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.21, 87.78) -- (128.07, 84.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.97, 85.50) rectangle (128.83, 88.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.97, 85.50) -- (128.83, 88.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.97, 88.35) -- (128.83, 85.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.42, 92.80) rectangle (132.28, 95.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.42, 92.80) -- (132.28, 95.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.42, 95.66) -- (132.28, 92.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.77, 97.40) rectangle (131.63,100.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.77, 97.40) -- (131.63,100.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.77,100.25) -- (131.63, 97.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.17,114.67) rectangle (170.03,117.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.17,114.67) -- (170.03,117.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.17,117.53) -- (170.03,114.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.11,116.23) rectangle (168.97,119.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.11,116.23) -- (168.97,119.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.11,119.09) -- (168.97,116.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.72,112.32) rectangle (167.57,115.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.72,112.32) -- (167.57,115.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.72,115.18) -- (167.57,112.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.91,117.15) rectangle (172.77,120.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.91,117.15) -- (172.77,120.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.91,120.00) -- (172.77,117.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.18,114.22) rectangle (169.03,117.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.18,114.22) -- (169.03,117.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.18,117.07) -- (169.03,114.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.67, 95.28) rectangle (142.52, 98.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.67, 95.28) -- (142.52, 98.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.67, 98.13) -- (142.52, 95.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.53, 88.51) rectangle (144.39, 91.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.53, 88.51) -- (144.39, 91.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.53, 91.37) -- (144.39, 88.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.33, 89.39) rectangle (139.19, 92.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.33, 89.39) -- (139.19, 92.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.33, 92.25) -- (139.19, 89.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.35, 92.46) rectangle (144.20, 95.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.35, 92.46) -- (144.20, 95.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.35, 95.32) -- (144.20, 92.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.29, 83.72) rectangle (140.14, 86.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.29, 83.72) -- (140.14, 86.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.29, 86.57) -- (140.14, 83.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.03, 85.50) rectangle (141.88, 88.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.03, 85.50) -- (141.88, 88.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.03, 88.35) -- (141.88, 85.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.65, 85.50) rectangle (140.51, 88.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.65, 85.50) -- (140.51, 88.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.65, 88.35) -- (140.51, 85.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.29, 99.24) rectangle (146.14,102.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.29, 99.24) -- (146.14,102.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.29,102.09) -- (146.14, 99.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.01, 85.50) rectangle (139.86, 88.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.01, 85.50) -- (139.86, 88.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.01, 88.35) -- (139.86, 85.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.50, 90.62) rectangle (140.36, 93.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.50, 90.62) -- (140.36, 93.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.50, 93.48) -- (140.36, 90.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.01, 96.89) rectangle (142.86, 99.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.01, 96.89) -- (142.86, 99.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.01, 99.75) -- (142.86, 96.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.43,128.45) rectangle (166.28,131.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.43,128.45) -- (166.28,131.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.43,131.30) -- (166.28,128.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.75,125.92) rectangle (162.60,128.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.75,125.92) -- (162.60,128.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.75,128.77) -- (162.60,125.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.02,129.80) rectangle (166.88,132.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.02,129.80) -- (166.88,132.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.02,132.65) -- (166.88,129.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.39,128.04) rectangle (167.24,130.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.39,128.04) -- (167.24,130.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.39,130.89) -- (167.24,128.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.70,130.22) rectangle (160.55,133.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.70,130.22) -- (160.55,133.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.70,133.07) -- (160.55,130.22);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.91,216.82) -- (172.95,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.93,214.80) -- (170.93,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (213.09,216.82) -- (217.13,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (215.11,214.80) -- (215.11,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (197.64,216.82) -- (201.68,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (199.66,214.80) -- (199.66,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.46,216.82) -- (181.50,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.48,214.80) -- (179.48,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.22,216.82) -- (165.26,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.24,214.80) -- (163.24,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (224.54,216.82) -- (228.57,216.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (226.56,214.80) -- (226.56,218.84);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.37, 68.39) rectangle (114.22, 71.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.37, 68.39) -- (114.22, 71.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.37, 71.24) -- (114.22, 68.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.56, 75.44) rectangle (120.41, 78.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.56, 75.44) -- (120.41, 78.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.56, 78.29) -- (120.41, 75.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.43,108.18) rectangle (139.29,111.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.43,108.18) -- (139.29,111.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.43,111.03) -- (139.29,108.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.21,143.87) rectangle (163.06,146.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.21,143.87) -- (163.06,146.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.21,146.73) -- (163.06,143.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (174.81,168.04) rectangle (177.66,170.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (174.81,168.04) -- (177.66,170.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (174.81,170.89) -- (177.66,168.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.90,193.96) rectangle (201.75,196.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.90,193.96) -- (201.75,196.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (198.90,196.81) -- (201.75,193.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) rectangle (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) rectangle (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) rectangle (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) rectangle (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) rectangle (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor,fill opacity=0.50] (124.01, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.18, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.66, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.84, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.67, 83.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.67, 83.12) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.58, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.66, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.59, 83.12) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.07, 83.12) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.26, 83.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.51, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.58, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.43, 83.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.58, 83.12) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (122.67, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.84, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.07, 83.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.34, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.22, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.75, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.34, 83.12) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.09, 83.12) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.75, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.49, 84.50) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.75, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.09, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.59, 83.12) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.75, 83.12) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.18, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.07, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.37, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (115.63, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.47, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.34, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.99, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.86, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.73, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (115.63, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.65, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (115.32, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.47, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.07, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.22, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.47, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.20, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.93, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.93, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.34, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.93, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.65, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.93, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.93, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.60, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.07, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.99, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.65, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.07, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.20, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.93, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.12, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.47, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.60, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.47, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.86, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.25, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.65, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.34, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.20, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.20, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (140.86, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.65, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.93, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.34, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.73, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.79, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.73, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.34, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (115.93, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.07, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.73, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.62, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.73, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.86, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (115.32, 73.06) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.93, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.20, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.20, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.36, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (115.48, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.86, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.34, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.60, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.34, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.93, 73.06) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.51, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (115.93, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.20, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.07, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.07, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.20, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.65, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.79, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.86, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.34, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.79, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.93, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.79, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.60, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.20, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.79, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.22, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.93, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.65, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.93, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.99, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.99, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.34, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.47, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.65, 73.06) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.93, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.47, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.20, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.65, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.47, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.07, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.07, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.73, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.36, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.60, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.20, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.34, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.60, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.73, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.47, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (115.78, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.47, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.93, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.99, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.12, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (115.78, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.86, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.99, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.93, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.22, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.86, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.47, 74.44) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.22, 73.06) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.20, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.99, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.22, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.11, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.46, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.87, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.99, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.34, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.11, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.87, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.03, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.87, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.69, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.74, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.14, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.66, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.92, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.92, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.69, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.50, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.34, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.34, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.46, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.57, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.22, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.25, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.34, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.57, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.46, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.34, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.92, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.25, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.69, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.25, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.12, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.25, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.69, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.46, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.25, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.57, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.92, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.92, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.69, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.62, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.80, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.57, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.37, 79.88) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.46, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.80, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.92, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.99, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.57, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.62, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.46, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.25, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.50, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.03, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.22, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.12, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.34, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.11, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.69, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.92, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.80, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.11, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.69, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.92, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.46, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.46, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.11, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.57, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.11, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.92, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.57, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.34, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.57, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.37, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.14, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.92, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.22, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.80, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.80, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.69, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.57, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.46, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.68, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.80, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.87, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.03, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.62, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.69, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.11, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.46, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.22, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.87, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.80, 79.88) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.36, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.36, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.99, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (140.86, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (122.20, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.46, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.11, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.46, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.37, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.60, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.34, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.62, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.46, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.11, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.03, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.03, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.34, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.34, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.50, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.46, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.99, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.69, 80.76) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.62, 75.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.46, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.11, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.34, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.25, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.46, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.92, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.75, 89.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.26, 90.39) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.47, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.41, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.01, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.50, 82.37) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.00, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.43, 89.48) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.79, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.90, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.03, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.09, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.31, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.83, 89.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.00, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.79, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.90, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.25, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.31, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (151.47, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.07, 90.39) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.83, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.23, 83.12) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.00, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.21, 81.59) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.23, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.31, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.67, 89.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.07, 83.12) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.92, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.03, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.49, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.58, 89.00) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.00, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.91, 84.50) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.03, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.38, 84.50) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.38, 88.00) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.51, 90.82) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.07, 90.82) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.00, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.68, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.79, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.35, 90.39) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.00, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.43, 90.39) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.47, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.31, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.35, 89.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.59, 88.51) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.03, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.22, 88.51) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.90, 84.50) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.25, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.28, 90.39) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.61, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (122.01, 90.82) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.03, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.79, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.61, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.14, 90.82) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.99, 79.88) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.92, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.07, 90.82) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.14, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.35, 89.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.03, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.47, 79.88) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (122.11, 79.88) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.68, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.75, 89.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.38, 89.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.51, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.14, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.68, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.47, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.71, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.36, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.07, 85.15) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (144.65, 79.88) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.14, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.03, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.69, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.22, 88.51) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.00, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.49, 89.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.79, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.14, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.47, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.21, 86.36) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.66, 87.47) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.14, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.36, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.00, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.51, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.36, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.21, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.43, 88.00) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.51, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.00, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.00, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.25, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.91, 88.51) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.68, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.79, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.58, 88.51) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.69, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.59, 90.39) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.03, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.10, 77.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.79, 76.86) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.99, 80.76) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.68, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (122.39, 84.50) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.71, 80.76) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.91, 78.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (124.99, 86.36) circle (  1.43);
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor,fill opacity=0.50] (128.75, 96.78) --
	(130.67, 93.45) --
	(126.82, 93.45) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.10, 82.98) --
	(123.03, 79.65) --
	(119.18, 79.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.01, 79.08) --
	(123.93, 75.76) --
	(120.09, 75.76) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.20, 82.10) --
	(124.12, 78.77) --
	(120.28, 78.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.68, 81.16) --
	(122.60, 77.83) --
	(118.76, 77.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (154.35, 77.92) --
	(156.27, 74.59) --
	(152.43, 74.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.71, 82.98) --
	(123.63, 79.65) --
	(119.79, 79.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.01, 82.10) --
	(123.93, 78.77) --
	(120.09, 78.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.69, 76.66) --
	(121.61, 73.33) --
	(117.77, 73.33) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.01, 81.16) --
	(123.93, 77.83) --
	(120.09, 77.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (191.39, 91.22) --
	(193.31, 87.89) --
	(189.47, 87.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (179.56, 92.61) --
	(181.48, 89.28) --
	(177.64, 89.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 95.76) --
	(228.48, 92.43) --
	(224.64, 92.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,106.68) --
	(228.48,103.35) --
	(224.64,103.35) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 94.65) --
	(228.48, 91.33) --
	(224.64, 91.33) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.19,107.01) --
	(139.11,103.68) --
	(135.27,103.68) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,101.04) --
	(228.48, 97.71) --
	(224.64, 97.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (163.99,113.99) --
	(165.91,110.67) --
	(162.07,110.67) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (210.55, 98.93) --
	(212.47, 95.60) --
	(208.63, 95.60) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,100.79) --
	(228.48, 97.46) --
	(224.64, 97.46) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (210.09,117.39) --
	(212.01,114.06) --
	(208.16,114.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,125.58) --
	(228.48,122.26) --
	(224.64,122.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,141.53) --
	(228.48,138.20) --
	(224.64,138.20) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,121.65) --
	(228.48,118.32) --
	(224.64,118.32) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,138.63) --
	(228.48,135.30) --
	(224.64,135.30) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (201.43,115.61) --
	(203.35,112.29) --
	(199.51,112.29) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (182.56,132.60) --
	(184.48,129.28) --
	(180.64,129.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,120.21) --
	(228.48,116.88) --
	(224.64,116.88) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.49,110.98) --
	(147.41,107.65) --
	(143.57,107.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,134.34) --
	(228.48,131.01) --
	(224.64,131.01) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,121.47) --
	(228.48,118.14) --
	(224.64,118.14) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,131.80) --
	(228.48,128.47) --
	(224.64,128.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,127.09) --
	(228.48,123.76) --
	(224.64,123.76) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,117.63) --
	(228.48,114.30) --
	(224.64,114.30) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,120.79) --
	(228.48,117.46) --
	(224.64,117.46) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,137.85) --
	(228.48,134.52) --
	(224.64,134.52) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,123.28) --
	(228.48,119.95) --
	(224.64,119.95) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,137.01) --
	(228.48,133.68) --
	(224.64,133.68) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,141.09) --
	(228.48,137.76) --
	(224.64,137.76) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,113.89) --
	(228.48,110.56) --
	(224.64,110.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,128.90) --
	(228.48,125.57) --
	(224.64,125.57) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,143.57) --
	(228.48,140.25) --
	(224.64,140.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,149.73) --
	(228.48,146.41) --
	(224.64,146.41) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,136.05) --
	(228.48,132.72) --
	(224.64,132.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,148.63) --
	(228.48,145.30) --
	(224.64,145.30) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,144.78) --
	(228.48,141.46) --
	(224.64,141.46) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,119.47) --
	(228.48,116.14) --
	(224.64,116.14) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,132.94) --
	(228.48,129.61) --
	(224.64,129.61) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,137.34) --
	(228.48,134.01) --
	(224.64,134.01) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,154.10) --
	(228.48,150.77) --
	(224.64,150.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,145.39) --
	(228.48,142.06) --
	(224.64,142.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,171.08) --
	(228.48,167.75) --
	(224.64,167.75) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,156.25) --
	(228.48,152.92) --
	(224.64,152.92) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,156.91) --
	(228.48,153.58) --
	(224.64,153.58) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,150.07) --
	(228.48,146.75) --
	(224.64,146.75) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,160.49) --
	(228.48,157.16) --
	(224.64,157.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,148.06) --
	(228.48,144.73) --
	(224.64,144.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,159.24) --
	(228.48,155.92) --
	(224.64,155.92) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,169.65) --
	(228.48,166.32) --
	(224.64,166.32) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,176.56) --
	(228.48,173.23) --
	(224.64,173.23) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,162.92) --
	(228.48,159.59) --
	(224.64,159.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,170.31) --
	(228.48,166.98) --
	(224.64,166.98) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,169.04) --
	(228.48,165.71) --
	(224.64,165.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,168.57) --
	(228.48,165.25) --
	(224.64,165.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,179.13) --
	(228.48,175.80) --
	(224.64,175.80) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,181.72) --
	(228.48,178.40) --
	(224.64,178.40) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,148.61) --
	(228.48,145.28) --
	(224.64,145.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,172.21) --
	(228.48,168.88) --
	(224.64,168.88) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,171.14) --
	(228.48,167.81) --
	(224.64,167.81) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,154.01) --
	(228.48,150.68) --
	(224.64,150.68) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,184.52) --
	(228.48,181.19) --
	(224.64,181.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,177.23) --
	(228.48,173.90) --
	(224.64,173.90) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,181.11) --
	(228.48,177.78) --
	(224.64,177.78) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,172.40) --
	(228.48,169.07) --
	(224.64,169.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,187.50) --
	(228.48,184.17) --
	(224.64,184.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,172.98) --
	(228.48,169.65) --
	(224.64,169.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,174.52) --
	(228.48,171.19) --
	(224.64,171.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,178.05) --
	(228.48,174.72) --
	(224.64,174.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,172.05) --
	(228.48,168.72) --
	(224.64,168.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,168.64) --
	(228.48,165.31) --
	(224.64,165.31) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,182.01) --
	(228.48,178.68) --
	(224.64,178.68) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,194.03) --
	(228.48,190.70) --
	(224.64,190.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,193.76) --
	(228.48,190.43) --
	(224.64,190.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,179.94) --
	(228.48,176.61) --
	(224.64,176.61) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,186.79) --
	(228.48,183.46) --
	(224.64,183.46) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,192.99) --
	(228.48,189.66) --
	(224.64,189.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,204.87) --
	(228.48,201.54) --
	(224.64,201.54) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,189.81) --
	(228.48,186.48) --
	(224.64,186.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,190.09) --
	(228.48,186.76) --
	(224.64,186.76) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,186.08) --
	(228.48,182.75) --
	(224.64,182.75) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (127.93, 83.81) --
	(129.85, 80.48) --
	(126.01, 80.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.34, 86.72) --
	(126.26, 83.39) --
	(122.42, 83.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.73, 96.45) --
	(133.65, 93.12) --
	(129.80, 93.12) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (143.01, 89.15) --
	(144.93, 85.82) --
	(141.09, 85.82) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (151.29, 90.73) --
	(153.21, 87.40) --
	(149.37, 87.40) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (154.36, 87.98) --
	(156.28, 84.66) --
	(152.44, 84.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.09, 73.75) --
	(138.01, 70.42) --
	(134.17, 70.42) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.53, 87.98) --
	(139.45, 84.66) --
	(135.61, 84.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.11, 75.27) --
	(121.03, 71.95) --
	(117.18, 71.95) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (154.91, 89.69) --
	(156.83, 86.37) --
	(152.99, 86.37) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (160.75, 92.16) --
	(162.67, 88.83) --
	(158.83, 88.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.86, 76.66) --
	(124.78, 73.33) --
	(120.94, 73.33) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.48, 80.16) --
	(132.40, 76.83) --
	(128.55, 76.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (205.31, 90.22) --
	(207.23, 86.89) --
	(203.39, 86.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (127.60, 92.61) --
	(129.52, 89.28) --
	(125.68, 89.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (187.94, 82.10) --
	(189.86, 78.77) --
	(186.02, 78.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (128.87, 90.22) --
	(130.79, 86.89) --
	(126.95, 86.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (155.52, 82.98) --
	(157.44, 79.65) --
	(153.60, 79.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (125.31, 92.61) --
	(127.23, 89.28) --
	(123.38, 89.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (128.37, 80.16) --
	(130.29, 76.83) --
	(126.45, 76.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (180.89, 88.58) --
	(182.81, 85.25) --
	(178.97, 85.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (162.74,117.71) --
	(164.66,114.38) --
	(160.82,114.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,110.21) --
	(228.48,106.89) --
	(224.64,106.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.48, 95.40) --
	(142.40, 92.07) --
	(138.56, 92.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (161.58,100.28) --
	(163.50, 96.96) --
	(159.66, 96.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (192.84, 98.34) --
	(194.76, 95.02) --
	(190.92, 95.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (199.95, 87.37) --
	(201.87, 84.04) --
	(198.03, 84.04) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,103.93) --
	(228.48,100.60) --
	(224.64,100.60) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,108.70) --
	(228.48,105.37) --
	(224.64,105.37) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (213.42,118.39) --
	(215.34,115.07) --
	(211.50,115.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 99.21) --
	(228.48, 95.88) --
	(224.64, 95.88) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (151.79, 97.11) --
	(153.71, 93.78) --
	(149.87, 93.78) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (216.21,105.82) --
	(218.13,102.49) --
	(214.29,102.49) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (168.97,101.76) --
	(170.89, 98.43) --
	(167.05, 98.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (189.29, 93.46) --
	(191.21, 90.13) --
	(187.37, 90.13) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 94.27) --
	(228.48, 90.94) --
	(224.64, 90.94) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.63, 81.16) --
	(133.55, 77.83) --
	(129.70, 77.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (132.45, 87.98) --
	(134.38, 84.66) --
	(130.53, 84.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 95.40) --
	(228.48, 92.07) --
	(224.64, 92.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (185.81,101.76) --
	(187.73, 98.43) --
	(183.89, 98.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (215.70, 92.61) --
	(217.62, 89.28) --
	(213.78, 89.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (163.81,100.79) --
	(165.74, 97.46) --
	(161.89, 97.46) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.40,106.68) --
	(190.32,103.35) --
	(186.48,103.35) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,107.96) --
	(228.48,104.63) --
	(224.64,104.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,127.61) --
	(228.48,124.29) --
	(224.64,124.29) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (153.56,102.22) --
	(155.49, 98.89) --
	(151.64, 98.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (196.06, 96.11) --
	(197.98, 92.78) --
	(194.14, 92.78) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,110.98) --
	(228.48,107.65) --
	(224.64,107.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (172.04, 96.11) --
	(173.96, 92.78) --
	(170.11, 92.78) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,111.47) --
	(228.48,108.14) --
	(224.64,108.14) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (202.33,127.05) --
	(204.25,123.72) --
	(200.41,123.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.66,114.19) --
	(144.58,110.87) --
	(140.74,110.87) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,108.84) --
	(228.48,105.52) --
	(224.64,105.52) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,138.01) --
	(228.48,134.68) --
	(224.64,134.68) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,103.93) --
	(228.48,100.60) --
	(224.64,100.60) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,118.02) --
	(228.48,114.69) --
	(224.64,114.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (197.58,115.25) --
	(199.50,111.92) --
	(195.66,111.92) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,115.25) --
	(228.48,111.92) --
	(224.64,111.92) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,122.12) --
	(228.48,118.79) --
	(224.64,118.79) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (193.52, 97.43) --
	(195.44, 94.10) --
	(191.60, 94.10) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 99.76) --
	(228.48, 96.43) --
	(224.64, 96.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,147.94) --
	(228.48,144.62) --
	(224.64,144.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,115.43) --
	(228.48,112.10) --
	(224.64,112.10) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,133.84) --
	(228.48,130.51) --
	(224.64,130.51) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,120.21) --
	(228.48,116.88) --
	(224.64,116.88) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,113.48) --
	(228.48,110.15) --
	(224.64,110.15) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,131.26) --
	(228.48,127.93) --
	(224.64,127.93) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,133.37) --
	(228.48,130.04) --
	(224.64,130.04) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,130.63) --
	(228.48,127.30) --
	(224.64,127.30) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,141.88) --
	(228.48,138.55) --
	(224.64,138.55) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,148.71) --
	(228.48,145.38) --
	(224.64,145.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,157.06) --
	(228.48,153.73) --
	(224.64,153.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,141.50) --
	(228.48,138.17) --
	(224.64,138.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,133.45) --
	(228.48,130.12) --
	(224.64,130.12) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,135.78) --
	(228.48,132.45) --
	(224.64,132.45) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,151.74) --
	(228.48,148.41) --
	(224.64,148.41) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,154.98) --
	(228.48,151.65) --
	(224.64,151.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,146.12) --
	(228.48,142.79) --
	(224.64,142.79) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,129.04) --
	(228.48,125.71) --
	(224.64,125.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,137.44) --
	(228.48,134.11) --
	(224.64,134.11) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,150.40) --
	(228.48,147.07) --
	(224.64,147.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,153.06) --
	(228.48,149.73) --
	(224.64,149.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,143.86) --
	(228.48,140.53) --
	(224.64,140.53) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,147.24) --
	(228.48,143.91) --
	(224.64,143.91) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 90.73) --
	(228.48, 87.40) --
	(224.64, 87.40) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,133.63) --
	(228.48,130.30) --
	(224.64,130.30) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,158.84) --
	(228.48,155.51) --
	(224.64,155.51) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,156.97) --
	(228.48,153.64) --
	(224.64,153.64) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,145.62) --
	(228.48,142.29) --
	(224.64,142.29) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,155.70) --
	(228.48,152.37) --
	(224.64,152.37) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,161.16) --
	(228.48,157.83) --
	(224.64,157.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,142.47) --
	(228.48,139.14) --
	(224.64,139.14) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,126.03) --
	(228.48,122.70) --
	(224.64,122.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,166.54) --
	(228.48,163.21) --
	(224.64,163.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,153.05) --
	(228.48,149.73) --
	(224.64,149.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,156.51) --
	(228.48,153.18) --
	(224.64,153.18) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,167.15) --
	(228.48,163.82) --
	(224.64,163.82) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,165.10) --
	(228.48,161.77) --
	(224.64,161.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,155.02) --
	(228.48,151.69) --
	(224.64,151.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,108.84) --
	(228.48,105.52) --
	(224.64,105.52) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,160.10) --
	(228.48,156.77) --
	(224.64,156.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,148.58) --
	(228.48,145.25) --
	(224.64,145.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,173.83) --
	(228.48,170.51) --
	(224.64,170.51) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,163.69) --
	(228.48,160.36) --
	(224.64,160.36) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,169.43) --
	(228.48,166.10) --
	(224.64,166.10) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,167.12) --
	(228.48,163.79) --
	(224.64,163.79) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,158.89) --
	(228.48,155.56) --
	(224.64,155.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,157.59) --
	(228.48,154.26) --
	(224.64,154.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,166.23) --
	(228.48,162.90) --
	(224.64,162.90) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,177.17) --
	(228.48,173.84) --
	(224.64,173.84) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,160.97) --
	(228.48,157.64) --
	(224.64,157.64) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,175.57) --
	(228.48,172.24) --
	(224.64,172.24) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,173.47) --
	(228.48,170.14) --
	(224.64,170.14) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,151.52) --
	(228.48,148.19) --
	(224.64,148.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,177.49) --
	(228.48,174.16) --
	(224.64,174.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,181.25) --
	(228.48,177.92) --
	(224.64,177.92) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,175.03) --
	(228.48,171.71) --
	(224.64,171.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,176.08) --
	(228.48,172.75) --
	(224.64,172.75) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,154.21) --
	(228.48,150.88) --
	(224.64,150.88) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,159.52) --
	(228.48,156.19) --
	(224.64,156.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,192.83) --
	(228.48,189.50) --
	(224.64,189.50) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,191.75) --
	(228.48,188.43) --
	(224.64,188.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,121.35) --
	(228.48,118.02) --
	(224.64,118.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,154.62) --
	(228.48,151.29) --
	(224.64,151.29) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,164.24) --
	(228.48,160.91) --
	(224.64,160.91) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,175.40) --
	(228.48,172.07) --
	(224.64,172.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,198.05) --
	(228.48,194.72) --
	(224.64,194.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,181.63) --
	(228.48,178.30) --
	(224.64,178.30) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,169.86) --
	(228.48,166.53) --
	(224.64,166.53) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,183.42) --
	(228.48,180.10) --
	(224.64,180.10) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,200.95) --
	(228.48,197.62) --
	(224.64,197.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,195.52) --
	(228.48,192.19) --
	(224.64,192.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,197.29) --
	(228.48,193.96) --
	(224.64,193.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,198.76) --
	(228.48,195.44) --
	(224.64,195.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,198.56) --
	(228.48,195.23) --
	(224.64,195.23) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,169.17) --
	(228.48,165.84) --
	(224.64,165.84) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,190.68) --
	(228.48,187.36) --
	(224.64,187.36) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,204.04) --
	(228.48,200.72) --
	(224.64,200.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,193.46) --
	(228.48,190.13) --
	(224.64,190.13) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,171.48) --
	(228.48,168.15) --
	(224.64,168.15) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,195.84) --
	(228.48,192.51) --
	(224.64,192.51) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,185.42) --
	(228.48,182.09) --
	(224.64,182.09) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,186.77) --
	(228.48,183.45) --
	(224.64,183.45) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,204.52) --
	(228.48,201.19) --
	(224.64,201.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,206.92) --
	(228.48,203.59) --
	(224.64,203.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.07, 73.75) --
	(118.00, 70.42) --
	(114.15, 70.42) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.74, 73.75) --
	(120.67, 70.42) --
	(116.82, 70.42) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.30, 86.72) --
	(124.22, 83.39) --
	(120.38, 83.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.47, 73.75) --
	(119.39, 70.42) --
	(115.55, 70.42) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.11, 76.66) --
	(121.03, 73.33) --
	(117.18, 73.33) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.07, 73.75) --
	(118.00, 70.42) --
	(114.15, 70.42) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.41, 77.92) --
	(123.33, 74.59) --
	(119.49, 74.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.58, 76.66) --
	(122.50, 73.33) --
	(118.65, 73.33) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.58, 75.27) --
	(122.50, 71.95) --
	(118.65, 71.95) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.68, 81.16) --
	(122.60, 77.83) --
	(118.76, 77.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.22, 76.66) --
	(121.15, 73.33) --
	(117.30, 73.33) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.84, 89.15) --
	(125.76, 85.82) --
	(121.92, 85.82) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.46, 76.66) --
	(121.38, 73.33) --
	(117.54, 73.33) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.14, 77.92) --
	(122.06, 74.59) --
	(118.22, 74.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.22,103.31) --
	(139.14, 99.99) --
	(135.30, 99.99) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.66, 80.16) --
	(125.59, 76.83) --
	(121.74, 76.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.20, 77.92) --
	(124.12, 74.59) --
	(120.28, 74.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (180.96, 84.59) --
	(182.88, 81.26) --
	(179.04, 81.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.66, 77.92) --
	(125.59, 74.59) --
	(121.74, 74.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (161.68, 84.59) --
	(163.60, 81.26) --
	(159.76, 81.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (162.89, 86.72) --
	(164.81, 83.39) --
	(160.97, 83.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.78, 89.69) --
	(141.71, 86.37) --
	(137.86, 86.37) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (196.33, 81.16) --
	(198.25, 77.83) --
	(194.41, 77.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.75, 80.16) --
	(126.67, 76.83) --
	(122.83, 76.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (147.13, 91.22) --
	(149.05, 87.89) --
	(145.21, 87.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.84, 85.34) --
	(125.76, 82.01) --
	(121.92, 82.01) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.56, 82.98) --
	(190.48, 79.65) --
	(186.64, 79.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (174.73, 81.16) --
	(176.65, 77.83) --
	(172.81, 77.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.06, 87.37) --
	(134.98, 84.04) --
	(131.14, 84.04) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.19, 86.05) --
	(135.12, 82.72) --
	(131.27, 82.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.31, 83.81) --
	(125.23, 80.48) --
	(121.39, 80.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.96, 84.59) --
	(146.88, 81.26) --
	(143.04, 81.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (176.30,100.02) --
	(178.22, 96.69) --
	(174.37, 96.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (126.43, 86.72) --
	(128.35, 83.39) --
	(124.50, 83.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.32, 84.59) --
	(146.24, 81.26) --
	(142.40, 81.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (175.70, 92.61) --
	(177.62, 89.28) --
	(173.78, 89.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (209.00, 86.72) --
	(210.92, 83.39) --
	(207.08, 83.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (159.44, 85.34) --
	(161.36, 82.01) --
	(157.52, 82.01) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (152.76, 94.65) --
	(154.68, 91.33) --
	(150.83, 91.33) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.75, 85.34) --
	(126.67, 82.01) --
	(122.83, 82.01) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (169.77, 86.72) --
	(171.69, 83.39) --
	(167.85, 83.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (179.21, 94.27) --
	(181.13, 90.94) --
	(177.29, 90.94) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (193.61, 95.76) --
	(195.53, 92.43) --
	(191.69, 92.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (156.57, 86.72) --
	(158.49, 83.39) --
	(154.65, 83.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.15, 97.74) --
	(143.07, 94.41) --
	(139.23, 94.41) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (206.23, 98.34) --
	(208.16, 95.02) --
	(204.31, 95.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.11, 86.72) --
	(131.03, 83.39) --
	(127.19, 83.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (165.21, 86.72) --
	(167.13, 83.39) --
	(163.28, 83.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (166.79, 91.22) --
	(168.71, 87.89) --
	(164.86, 87.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (187.14,102.89) --
	(189.06, 99.56) --
	(185.22, 99.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (196.45, 87.37) --
	(198.37, 84.04) --
	(194.53, 84.04) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (173.51, 95.03) --
	(175.43, 91.70) --
	(171.59, 91.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 92.61) --
	(228.48, 89.28) --
	(224.64, 89.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 87.98) --
	(228.48, 84.66) --
	(224.64, 84.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (180.17,105.82) --
	(182.09,102.49) --
	(178.24,102.49) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,104.52) --
	(228.48,101.20) --
	(224.64,101.20) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (128.06, 90.73) --
	(129.98, 87.40) --
	(126.13, 87.40) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (166.14, 92.61) --
	(168.06, 89.28) --
	(164.22, 89.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (195.34, 91.22) --
	(197.26, 87.89) --
	(193.42, 87.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,127.96) --
	(228.48,124.64) --
	(224.64,124.64) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (149.95,106.00) --
	(151.87,102.67) --
	(148.03,102.67) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (205.29, 90.73) --
	(207.21, 87.40) --
	(203.37, 87.40) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (205.26, 90.22) --
	(207.18, 86.89) --
	(203.34, 86.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 98.64) --
	(228.48, 95.31) --
	(224.64, 95.31) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (213.54,108.55) --
	(215.46,105.23) --
	(211.62,105.23) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (214.13,100.28) --
	(216.05, 96.96) --
	(212.21, 96.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,102.45) --
	(228.48, 99.12) --
	(224.64, 99.12) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (163.26, 92.16) --
	(165.18, 88.83) --
	(161.34, 88.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (206.57,104.13) --
	(208.49,100.80) --
	(204.65,100.80) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (209.44, 92.61) --
	(211.36, 89.28) --
	(207.52, 89.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (194.06, 92.16) --
	(195.98, 88.83) --
	(192.14, 88.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (180.11, 95.76) --
	(182.03, 92.43) --
	(178.19, 92.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (173.63,101.28) --
	(175.55, 97.96) --
	(171.71, 97.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 95.76) --
	(228.48, 92.43) --
	(224.64, 92.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (179.32,101.28) --
	(181.25, 97.96) --
	(177.40, 97.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (179.93,108.99) --
	(181.85,105.66) --
	(178.01,105.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,100.54) --
	(228.48, 97.21) --
	(224.64, 97.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.42,103.52) --
	(138.34,100.19) --
	(134.49,100.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (191.88, 97.11) --
	(193.80, 93.78) --
	(189.96, 93.78) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,107.96) --
	(228.48,104.63) --
	(224.64,104.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (209.77,107.01) --
	(211.69,103.68) --
	(207.84,103.68) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 95.03) --
	(228.48, 91.70) --
	(224.64, 91.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,100.79) --
	(228.48, 97.46) --
	(224.64, 97.46) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,138.39) --
	(228.48,135.06) --
	(224.64,135.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (208.11,103.52) --
	(210.03,100.19) --
	(206.19,100.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,121.47) --
	(228.48,118.14) --
	(224.64,118.14) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (211.63,119.05) --
	(213.55,115.72) --
	(209.71,115.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,113.58) --
	(228.48,110.26) --
	(224.64,110.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,123.38) --
	(228.48,120.06) --
	(224.64,120.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 97.43) --
	(228.48, 94.10) --
	(224.64, 94.10) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,142.36) --
	(228.48,139.03) --
	(224.64,139.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 99.21) --
	(228.48, 95.88) --
	(224.64, 95.88) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,103.31) --
	(228.48, 99.99) --
	(224.64, 99.99) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,127.49) --
	(228.48,124.17) --
	(224.64,124.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,135.89) --
	(228.48,132.57) --
	(224.64,132.57) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,144.50) --
	(228.48,141.18) --
	(224.64,141.18) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 98.93) --
	(228.48, 95.60) --
	(224.64, 95.60) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56, 98.64) --
	(228.48, 95.31) --
	(224.64, 95.31) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,153.30) --
	(228.48,149.97) --
	(224.64,149.97) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,161.40) --
	(228.48,158.07) --
	(224.64,158.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,109.95) --
	(228.48,106.62) --
	(224.64,106.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,119.33) --
	(228.48,116.00) --
	(224.64,116.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,146.11) --
	(228.48,142.78) --
	(224.64,142.78) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,122.17) --
	(228.48,118.85) --
	(224.64,118.85) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,154.99) --
	(228.48,151.66) --
	(224.64,151.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,144.15) --
	(228.48,140.83) --
	(224.64,140.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,114.59) --
	(228.48,111.26) --
	(224.64,111.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,144.42) --
	(228.48,141.09) --
	(224.64,141.09) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,148.65) --
	(228.48,145.32) --
	(224.64,145.32) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,122.35) --
	(228.48,119.02) --
	(224.64,119.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,185.58) --
	(228.48,182.25) --
	(224.64,182.25) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,118.09) --
	(228.48,114.77) --
	(224.64,114.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,160.83) --
	(228.48,157.50) --
	(224.64,157.50) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,125.76) --
	(228.48,122.44) --
	(224.64,122.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,168.77) --
	(228.48,165.44) --
	(224.64,165.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,139.65) --
	(228.48,136.32) --
	(224.64,136.32) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,160.64) --
	(228.48,157.31) --
	(224.64,157.31) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,185.18) --
	(228.48,181.86) --
	(224.64,181.86) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,131.59) --
	(228.48,128.26) --
	(224.64,128.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,189.63) --
	(228.48,186.30) --
	(224.64,186.30) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,167.30) --
	(228.48,163.98) --
	(224.64,163.98) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,185.67) --
	(228.48,182.35) --
	(224.64,182.35) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,127.89) --
	(228.48,124.56) --
	(224.64,124.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,179.04) --
	(228.48,175.71) --
	(224.64,175.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (226.56,219.04) --
	(228.48,215.71) --
	(224.64,215.71) --
	cycle;

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.64, 92.46) rectangle (135.50, 95.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.64, 92.46) -- (135.50, 95.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.64, 95.32) -- (135.50, 92.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.17, 94.70) rectangle (171.02, 97.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.17, 94.70) -- (171.02, 97.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.17, 97.55) -- (171.02, 94.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) rectangle (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) rectangle (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) rectangle (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,215.39) -- (227.98,218.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (225.13,218.25) -- (227.98,215.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.12,127.52) rectangle (150.97,130.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.12,127.52) -- (150.97,130.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.12,130.37) -- (150.97,127.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.06,139.97) rectangle (160.91,142.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.06,139.97) -- (160.91,142.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.06,142.82) -- (160.91,139.97);
\definecolor{drawColor}{RGB}{152,152,152}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.51, 30.77) -- (235.42,225.68);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.95, 30.77) -- ( 41.95, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 44.19, 30.77) -- ( 44.19, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 46.13, 30.77) -- ( 46.13, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 47.84, 30.77) -- ( 47.84, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 49.37, 30.77) -- ( 49.37, 39.31);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 59.43, 30.77) -- ( 59.43, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 65.32, 30.77) -- ( 65.32, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 69.49, 30.77) -- ( 69.49, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 72.73, 30.77) -- ( 72.73, 36.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 75.38, 30.77) -- ( 75.38, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 77.62, 30.77) -- ( 77.62, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 79.56, 30.77) -- ( 79.56, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 81.27, 30.77) -- ( 81.27, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 82.79, 30.77) -- ( 82.79, 39.31);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 92.86, 30.77) -- ( 92.86, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 98.74, 30.77) -- ( 98.74, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (102.92, 30.77) -- (102.92, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (106.16, 30.77) -- (106.16, 36.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (108.80, 30.77) -- (108.80, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (111.04, 30.77) -- (111.04, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (112.98, 30.77) -- (112.98, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (114.69, 30.77) -- (114.69, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (116.22, 30.77) -- (116.22, 39.31);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (126.28, 30.77) -- (126.28, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (132.17, 30.77) -- (132.17, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (136.34, 30.77) -- (136.34, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (139.58, 30.77) -- (139.58, 36.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.23, 30.77) -- (142.23, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (144.47, 30.77) -- (144.47, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (146.41, 30.77) -- (146.41, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (148.12, 30.77) -- (148.12, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (149.64, 30.77) -- (149.64, 39.31);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (159.71, 30.77) -- (159.71, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (165.59, 30.77) -- (165.59, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (169.77, 30.77) -- (169.77, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (173.01, 30.77) -- (173.01, 36.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (175.65, 30.77) -- (175.65, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (177.89, 30.77) -- (177.89, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (179.83, 30.77) -- (179.83, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (181.54, 30.77) -- (181.54, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (183.07, 30.77) -- (183.07, 39.31);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (193.13, 30.77) -- (193.13, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (199.02, 30.77) -- (199.02, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (203.19, 30.77) -- (203.19, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (206.43, 30.77) -- (206.43, 36.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (209.08, 30.77) -- (209.08, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (211.32, 30.77) -- (211.32, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (213.26, 30.77) -- (213.26, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (214.97, 30.77) -- (214.97, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (216.49, 30.77) -- (216.49, 39.31);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (226.56, 30.77) -- (226.56, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (232.44, 30.77) -- (232.44, 33.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 32.22) -- ( 43.36, 32.22);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 34.45) -- ( 43.36, 34.45);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 36.39) -- ( 43.36, 36.39);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 38.10) -- ( 43.36, 38.10);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 39.63) -- ( 49.05, 39.63);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 49.69) -- ( 43.36, 49.69);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 55.58) -- ( 43.36, 55.58);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 59.76) -- ( 43.36, 59.76);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 62.99) -- ( 46.20, 62.99);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 65.64) -- ( 43.36, 65.64);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 67.88) -- ( 43.36, 67.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 69.82) -- ( 43.36, 69.82);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 71.53) -- ( 43.36, 71.53);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 73.06) -- ( 49.05, 73.06);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 83.12) -- ( 43.36, 83.12);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 89.00) -- ( 43.36, 89.00);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 93.18) -- ( 43.36, 93.18);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 96.42) -- ( 46.20, 96.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 99.07) -- ( 43.36, 99.07);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,101.30) -- ( 43.36,101.30);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,103.24) -- ( 43.36,103.24);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,104.95) -- ( 43.36,104.95);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,106.48) -- ( 49.05,106.48);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,116.54) -- ( 43.36,116.54);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,122.43) -- ( 43.36,122.43);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,126.61) -- ( 43.36,126.61);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,129.84) -- ( 46.20,129.84);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,132.49) -- ( 43.36,132.49);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,134.73) -- ( 43.36,134.73);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,136.67) -- ( 43.36,136.67);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,138.38) -- ( 43.36,138.38);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,139.91) -- ( 49.05,139.91);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,149.97) -- ( 43.36,149.97);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,155.85) -- ( 43.36,155.85);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,160.03) -- ( 43.36,160.03);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,163.27) -- ( 46.20,163.27);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,165.92) -- ( 43.36,165.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,168.15) -- ( 43.36,168.15);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,170.09) -- ( 43.36,170.09);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,171.80) -- ( 43.36,171.80);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,173.33) -- ( 49.05,173.33);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,183.39) -- ( 43.36,183.39);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,189.28) -- ( 43.36,189.28);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,193.46) -- ( 43.36,193.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,196.69) -- ( 46.20,196.69);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,199.34) -- ( 43.36,199.34);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,201.58) -- ( 43.36,201.58);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,203.52) -- ( 43.36,203.52);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,205.23) -- ( 43.36,205.23);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,206.76) -- ( 49.05,206.76);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,216.82) -- ( 43.36,216.82);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,222.70) -- ( 43.36,222.70);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 30.77) rectangle (235.42,225.68);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.56, 70.03) {0.1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.56,136.88) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.56,203.73) {1000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 37.76, 73.06) --
	( 40.51, 73.06);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 37.76,139.91) --
	( 40.51,139.91);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 37.76,206.76) --
	( 40.51,206.76);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 82.79, 28.02) --
	( 82.79, 30.77);

\path[draw=drawColor,line width= 0.3pt,line join=round] (149.64, 28.02) --
	(149.64, 30.77);

\path[draw=drawColor,line width= 0.3pt,line join=round] (216.49, 28.02) --
	(216.49, 30.77);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 82.79, 19.76) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (149.64, 19.76) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (216.49, 19.76) {1000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (137.96,  7.72) {\textsf{ADDMC} + \texttt{sbk05} time (s)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,128.22) {\textsf{ADDMC} + \texttt{cw} time (s)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (246.42, 71.76) rectangle (337.78,184.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (251.92,170.55) {Data set};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (251.92,149.53) rectangle (266.37,163.98);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor,fill opacity=0.50] (259.14,156.75) circle (  1.43);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (251.92,135.07) rectangle (266.37,149.53);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor,fill opacity=0.50] (259.14,144.52) --
	(261.06,141.19) --
	(257.22,141.19) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (251.92,120.62) rectangle (266.37,135.07);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor,fill opacity=0.50] (257.72,126.42) --
	(260.57,126.42) --
	(260.57,129.27) --
	(257.72,129.27) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (251.92,106.16) rectangle (266.37,120.62);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (257.13,113.39) -- (261.16,113.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (259.14,111.37) -- (259.14,115.41);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (251.92, 91.71) rectangle (266.37,106.16);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (257.72, 97.51) rectangle (260.57,100.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (257.72, 97.51) -- (260.57,100.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (257.72,100.36) -- (260.57, 97.51);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (251.92, 77.26) rectangle (266.37, 91.71);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (257.72, 83.06) -- (260.57, 85.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (257.72, 85.91) -- (260.57, 83.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (257.13, 84.48) -- (261.16, 84.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (259.14, 82.47) -- (259.14, 86.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (271.87,153.72) {DQMR};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (271.87,139.27) {Grid};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (271.87,124.81) {Mastermind};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (271.87,110.36) {Non-binary};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (271.87, 95.91) {Other binary};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (343.28,231.26);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (271.87, 81.45) {Random Blocks};
\end{scope}
\end{tikzpicture}