### 2021-06-13 doc/SAT_long_talk/scatter1.tex
Started working on slides for SAT

% Created by tikzDevice version 0.12.3.1 on 2021-06-13 19:02:01
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (303.53,224.04);
\begin{scope}
\path[clip] (  0.00, 16.35) rectangle (303.53,207.69);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00, 16.35) rectangle (303.53,207.69);
\end{scope}
\begin{scope}
\path[clip] ( 40.51, 47.03) rectangle (195.67,202.19);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 40.51, 47.03) rectangle (195.67,202.19);
\definecolor{drawColor}{gray}{0.87}

\path[draw=drawColor,line width= 0.1pt,line join=round] ( 40.51, 54.09) --
	(195.67, 54.09);

\path[draw=drawColor,line width= 0.1pt,line join=round] ( 40.51,107.30) --
	(195.67,107.30);

\path[draw=drawColor,line width= 0.1pt,line join=round] ( 40.51,160.52) --
	(195.67,160.52);

\path[draw=drawColor,line width= 0.1pt,line join=round] ( 47.56, 47.03) --
	( 47.56,202.19);

\path[draw=drawColor,line width= 0.1pt,line join=round] (100.78, 47.03) --
	(100.78,202.19);

\path[draw=drawColor,line width= 0.1pt,line join=round] (154.00, 47.03) --
	(154.00,202.19);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 40.51, 80.69) --
	(195.67, 80.69);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 40.51,133.91) --
	(195.67,133.91);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 40.51,187.13) --
	(195.67,187.13);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 74.17, 47.03) --
	( 74.17,202.19);

\path[draw=drawColor,line width= 0.3pt,line join=round] (127.39, 47.03) --
	(127.39,202.19);

\path[draw=drawColor,line width= 0.3pt,line join=round] (180.61, 47.03) --
	(180.61,202.19);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.28, 91.57) -- (117.14, 94.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.28, 94.42) -- (117.14, 91.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.69, 93.00) -- (117.73, 93.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.71, 90.98) -- (115.71, 95.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.63, 91.57) -- (118.48, 94.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.63, 94.42) -- (118.48, 91.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.04, 93.00) -- (119.07, 93.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.06, 90.98) -- (117.06, 95.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.83, 93.41) -- (116.68, 96.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.83, 96.26) -- (116.68, 93.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.24, 94.84) -- (117.27, 94.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.26, 92.82) -- (115.26, 96.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.63, 94.07) -- (119.48, 96.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.63, 96.92) -- (119.48, 94.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.04, 95.50) -- (120.08, 95.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.06, 93.48) -- (118.06, 97.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.77, 93.06) -- (118.62, 95.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.77, 95.92) -- (118.62, 93.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.18, 94.49) -- (119.21, 94.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.20, 92.47) -- (117.20, 96.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.41, 91.57) -- (117.26, 94.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.41, 94.42) -- (117.26, 91.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.82, 93.00) -- (117.85, 93.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.84, 90.98) -- (115.84, 95.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.48, 91.96) -- (113.34, 94.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.48, 94.82) -- (113.34, 91.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (109.89, 93.39) -- (113.93, 93.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.91, 91.37) -- (111.91, 95.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.90, 93.74) -- (117.75, 96.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.90, 96.60) -- (117.75, 93.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.31, 95.17) -- (118.35, 95.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.33, 93.15) -- (116.33, 97.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.86, 94.69) -- (116.72, 97.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.86, 97.55) -- (116.72, 94.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.27, 96.12) -- (117.31, 96.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.29, 94.10) -- (115.29, 98.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.83, 91.17) -- (115.68, 94.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.83, 94.02) -- (115.68, 91.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.24, 92.59) -- (116.27, 92.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.26, 90.57) -- (114.26, 94.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.33, 92.34) -- (114.19, 95.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.33, 95.19) -- (114.19, 92.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.74, 93.77) -- (114.78, 93.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.76, 91.75) -- (112.76, 95.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.88, 91.96) -- (118.73, 94.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.88, 94.82) -- (118.73, 91.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.29, 93.39) -- (119.33, 93.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.31, 91.37) -- (117.31, 95.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.28, 92.71) -- (117.14, 95.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.28, 95.56) -- (117.14, 92.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.69, 94.13) -- (117.73, 94.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.71, 92.12) -- (115.71, 96.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.46, 91.96) -- (116.31, 94.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.46, 94.82) -- (116.31, 91.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.87, 93.39) -- (116.91, 93.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.89, 91.37) -- (114.89, 95.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.32, 93.74) -- (119.17, 96.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.32, 96.60) -- (119.17, 93.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.72, 95.17) -- (119.76, 95.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.74, 93.15) -- (117.74, 97.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.34, 92.71) -- (119.20, 95.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.34, 95.56) -- (119.20, 92.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.75, 94.13) -- (119.79, 94.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.77, 92.12) -- (117.77, 96.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.21, 97.15) -- (119.06,100.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.21,100.00) -- (119.06, 97.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.62, 98.58) -- (119.65, 98.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.64, 96.56) -- (117.64,100.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.86, 95.57) -- (119.72, 98.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.86, 98.43) -- (119.72, 95.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.27, 97.00) -- (120.31, 97.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.29, 94.98) -- (118.29, 99.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.70, 94.39) -- (116.55, 97.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.70, 97.24) -- (116.55, 94.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.11, 95.81) -- (117.14, 95.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.12, 93.80) -- (115.12, 97.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.50, 94.99) -- (120.36, 97.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.50, 97.85) -- (120.36, 94.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.91, 96.42) -- (120.95, 96.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.93, 94.40) -- (118.93, 98.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.28, 96.65) -- (117.14, 99.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.28, 99.50) -- (117.14, 96.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.69, 98.07) -- (117.73, 98.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.71, 96.06) -- (115.71,100.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.28, 93.74) -- (117.14, 96.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.28, 96.60) -- (117.14, 93.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.69, 95.17) -- (117.73, 95.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.71, 93.15) -- (115.71, 97.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.09,108.64) -- (114.94,111.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.09,111.49) -- (114.94,108.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.50,110.06) -- (115.53,110.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.51,108.05) -- (113.51,112.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.76, 93.74) -- (116.62, 96.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.76, 96.60) -- (116.62, 93.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.17, 95.17) -- (117.21, 95.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.19, 93.15) -- (115.19, 97.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.29, 93.74) -- (116.14, 96.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.29, 96.60) -- (116.14, 93.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.70, 95.17) -- (116.73, 95.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.72, 93.15) -- (114.72, 97.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.21, 95.57) -- (119.06, 98.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.21, 98.43) -- (119.06, 95.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.62, 97.00) -- (119.65, 97.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.64, 94.98) -- (117.64, 99.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.01, 98.09) -- (119.87,100.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.01,100.95) -- (119.87, 98.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.42, 99.52) -- (120.46, 99.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.44, 97.50) -- (118.44,101.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.69, 94.69) -- (118.54, 97.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.69, 97.55) -- (118.54, 94.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.10, 96.12) -- (119.13, 96.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.11, 94.10) -- (117.11, 98.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.01, 93.06) -- (115.86, 95.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.01, 95.92) -- (115.86, 93.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.42, 94.49) -- (116.45, 94.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.44, 92.47) -- (114.44, 96.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.31, 95.29) -- (117.17, 98.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.31, 98.14) -- (117.17, 95.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.72, 96.71) -- (117.76, 96.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.74, 94.70) -- (115.74, 98.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.90, 93.06) -- (115.76, 95.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.90, 95.92) -- (115.76, 93.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.31, 94.49) -- (116.35, 94.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.33, 92.47) -- (114.33, 96.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.63, 97.39) -- (117.48,100.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.63,100.25) -- (117.48, 97.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.04, 98.82) -- (118.07, 98.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.05, 96.80) -- (116.05,100.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.86,106.33) -- (120.71,109.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.86,109.18) -- (120.71,106.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.27,107.76) -- (121.30,107.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.29,105.74) -- (119.29,109.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.06, 97.15) -- (116.91,100.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.06,100.00) -- (116.91, 97.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.47, 98.58) -- (117.50, 98.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.49, 96.56) -- (115.49,100.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.73, 96.12) -- (116.58, 98.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.73, 98.98) -- (116.58, 96.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.14, 97.55) -- (117.17, 97.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.16, 95.53) -- (115.16, 99.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.49, 96.90) -- (118.34, 99.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.49, 99.76) -- (118.34, 96.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.90, 98.33) -- (118.93, 98.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.91, 96.31) -- (116.91,100.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.72, 98.54) -- (120.57,101.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.72,101.39) -- (120.57, 98.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.13, 99.97) -- (121.16, 99.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.14, 97.95) -- (119.14,101.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.41, 95.29) -- (117.26, 98.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.41, 98.14) -- (117.26, 95.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.82, 96.71) -- (117.85, 96.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.84, 94.70) -- (115.84, 98.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.42,103.00) -- (119.28,105.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.42,105.86) -- (119.28,103.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.83,104.43) -- (119.87,104.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.85,102.41) -- (117.85,106.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.13, 99.38) -- (118.98,102.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.13,102.23) -- (118.98, 99.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.54,100.81) -- (119.57,100.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.55, 98.79) -- (117.55,102.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.36, 97.63) -- (120.21,100.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.36,100.49) -- (120.21, 97.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.77, 99.06) -- (120.80, 99.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.79, 97.04) -- (118.79,101.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.99,107.19) -- (119.84,110.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.99,110.04) -- (119.84,107.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.40,108.61) -- (120.43,108.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.41,106.59) -- (118.41,110.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.17, 97.15) -- (130.02,100.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.17,100.00) -- (130.02, 97.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.58, 98.58) -- (130.61, 98.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.59, 96.56) -- (128.59,100.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.53, 96.12) -- (117.39, 98.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.53, 98.98) -- (117.39, 96.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.94, 97.55) -- (117.98, 97.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.96, 95.53) -- (115.96, 99.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.25, 95.57) -- (116.11, 98.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.25, 98.43) -- (116.11, 95.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.66, 97.00) -- (116.70, 97.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.68, 94.98) -- (114.68, 99.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.90, 96.90) -- (122.76, 99.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.90, 99.76) -- (122.76, 96.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.31, 98.33) -- (123.35, 98.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.33, 96.31) -- (121.33,100.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.53, 96.90) -- (117.39, 99.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.53, 99.76) -- (117.39, 96.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.94, 98.33) -- (117.98, 98.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.96, 96.31) -- (115.96,100.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.40,100.16) -- (121.26,103.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.40,103.02) -- (121.26,100.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.81,101.59) -- (121.85,101.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.83, 99.57) -- (119.83,103.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.21,130.78) -- (131.07,133.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.21,133.63) -- (131.07,130.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.62,132.21) -- (131.66,132.21);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.64,130.19) -- (129.64,134.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.91,132.43) -- (129.77,135.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.91,135.28) -- (129.77,132.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.32,133.85) -- (130.36,133.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.34,131.84) -- (128.34,135.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.31,130.82) -- (125.16,133.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.31,133.67) -- (125.16,130.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.72,132.25) -- (125.75,132.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.74,130.23) -- (123.74,134.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.51,130.29) -- (129.37,133.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.51,133.14) -- (129.37,130.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.92,131.72) -- (129.96,131.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.94,129.70) -- (127.94,133.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.30,130.89) -- (133.16,133.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.30,133.74) -- (133.16,130.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.71,132.31) -- (133.75,132.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.73,130.30) -- (131.73,134.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.51,130.85) -- (133.36,133.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.51,133.70) -- (133.36,130.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.92,132.27) -- (133.95,132.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.93,130.26) -- (131.93,134.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.00,131.79) -- (125.85,134.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.00,134.65) -- (125.85,131.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.41,133.22) -- (126.45,133.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.43,131.20) -- (124.43,135.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.00,130.87) -- (126.85,133.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.00,133.73) -- (126.85,130.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.41,132.30) -- (127.45,132.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.43,130.28) -- (125.43,134.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.99,130.87) -- (126.84,133.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.99,133.73) -- (126.84,130.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.40,132.30) -- (127.43,132.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.41,130.28) -- (125.41,134.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.45,131.27) -- (126.31,134.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.45,134.12) -- (126.31,131.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.86,132.69) -- (126.90,132.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.88,130.68) -- (124.88,134.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.64,130.19) -- (124.49,133.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.64,133.04) -- (124.49,130.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.05,131.62) -- (125.08,131.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.07,129.60) -- (123.07,133.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.90,130.70) -- (131.75,133.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.90,133.55) -- (131.75,130.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.30,132.13) -- (132.34,132.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.32,130.11) -- (130.32,134.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.80,132.26) -- (133.65,135.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.80,135.12) -- (133.65,132.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.21,133.69) -- (134.24,133.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.23,131.67) -- (132.23,135.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.80,131.73) -- (131.65,134.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.80,134.59) -- (131.65,131.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.21,133.16) -- (132.24,133.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.22,131.14) -- (130.22,135.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.76,130.59) -- (128.62,133.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.76,133.45) -- (128.62,130.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.17,132.02) -- (129.21,132.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.19,130.00) -- (127.19,134.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.14,134.41) -- (129.99,137.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.14,137.26) -- (129.99,134.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.55,135.83) -- (130.58,135.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.56,133.82) -- (128.56,137.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.70,134.43) -- (131.55,137.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.70,137.28) -- (131.55,134.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.11,135.85) -- (132.14,135.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.12,133.84) -- (130.12,137.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.65,138.50) -- (125.51,141.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.65,141.35) -- (125.51,138.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.06,139.93) -- (126.10,139.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.08,137.91) -- (124.08,141.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.53,134.94) -- (128.38,137.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.53,137.80) -- (128.38,134.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.93,136.37) -- (128.97,136.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.95,134.35) -- (126.95,138.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.98,131.68) -- (129.83,134.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.98,134.54) -- (129.83,131.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.39,133.11) -- (130.42,133.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.41,131.09) -- (128.41,135.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.41,132.22) -- (133.26,135.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.41,135.07) -- (133.26,132.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.81,133.64) -- (133.85,133.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.83,131.62) -- (131.83,135.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.59,131.96) -- (128.44,134.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.59,134.82) -- (128.44,131.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.99,133.39) -- (129.03,133.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.01,131.37) -- (127.01,135.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.49,131.65) -- (128.34,134.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.49,134.50) -- (128.34,131.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.90,133.07) -- (128.93,133.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.92,131.05) -- (126.92,135.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.57,133.56) -- (128.43,136.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.57,136.42) -- (128.43,133.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.98,134.99) -- (129.02,134.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.00,132.97) -- (127.00,137.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.20,131.63) -- (129.06,134.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.20,134.49) -- (129.06,131.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.61,133.06) -- (129.65,133.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.63,131.04) -- (127.63,135.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.91,131.57) -- (125.76,134.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.91,134.42) -- (125.76,131.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.32,133.00) -- (126.36,133.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.34,130.98) -- (124.34,135.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.04,133.65) -- (126.90,136.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.04,136.50) -- (126.90,133.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.45,135.07) -- (127.49,135.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.47,133.06) -- (125.47,137.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.87,131.44) -- (133.72,134.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.87,134.30) -- (133.72,131.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.28,132.87) -- (134.31,132.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.30,130.85) -- (132.30,134.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.80,134.90) -- (129.65,137.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.80,137.75) -- (129.65,134.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.21,136.32) -- (130.24,136.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.22,134.30) -- (128.22,138.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.77,132.60) -- (129.62,135.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.77,135.45) -- (129.62,132.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.17,134.03) -- (130.21,134.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.19,132.01) -- (128.19,136.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.84,132.71) -- (132.69,135.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.84,135.57) -- (132.69,132.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.25,134.14) -- (133.29,134.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.27,132.12) -- (131.27,136.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.64,134.10) -- (132.49,136.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.64,136.95) -- (132.49,134.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.05,135.53) -- (133.09,135.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.07,133.51) -- (131.07,137.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.90,135.14) -- (131.75,137.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.90,137.99) -- (131.75,135.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.30,136.56) -- (132.34,136.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.32,134.55) -- (130.32,138.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.16,139.31) -- (131.02,142.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.16,142.16) -- (131.02,139.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.57,140.74) -- (131.61,140.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.59,138.72) -- (129.59,142.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.63,136.71) -- (133.48,139.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.63,139.56) -- (133.48,136.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.04,138.13) -- (134.08,138.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.06,136.11) -- (132.06,140.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.06,135.54) -- (132.92,138.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.06,138.40) -- (132.92,135.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.47,136.97) -- (133.51,136.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.49,134.95) -- (131.49,138.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.20,136.83) -- (132.06,139.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.20,139.68) -- (132.06,136.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.61,138.25) -- (132.65,138.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.63,136.23) -- (130.63,140.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.18,137.00) -- (131.04,139.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.18,139.85) -- (131.04,137.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.59,138.43) -- (131.63,138.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.61,136.41) -- (129.61,140.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.95,138.11) -- (132.80,140.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.95,140.96) -- (132.80,138.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.36,139.54) -- (133.39,139.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.38,137.52) -- (131.38,141.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.50,136.69) -- (128.36,139.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.50,139.54) -- (128.36,136.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.91,138.12) -- (128.95,138.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.93,136.10) -- (126.93,140.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.69,137.93) -- (128.55,140.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.69,140.78) -- (128.55,137.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.10,139.36) -- (129.14,139.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.12,137.34) -- (127.12,141.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.19,135.75) -- (130.04,138.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.19,138.61) -- (130.04,135.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.60,137.18) -- (130.63,137.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.61,135.16) -- (128.61,139.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.64,133.34) -- (126.49,136.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.64,136.19) -- (126.49,133.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.05,134.77) -- (127.08,134.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.07,132.75) -- (125.07,136.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.79,136.97) -- (132.65,139.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.79,139.82) -- (132.65,136.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.20,138.39) -- (133.24,138.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.22,136.38) -- (131.22,140.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.15,136.87) -- (132.01,139.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.15,139.72) -- (132.01,136.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.56,138.29) -- (132.60,138.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.58,136.27) -- (130.58,140.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.04,135.08) -- (131.89,137.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.04,137.93) -- (131.89,135.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.45,136.51) -- (132.48,136.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.46,134.49) -- (130.46,138.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.42,138.62) -- (131.27,141.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.42,141.47) -- (131.27,138.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.83,140.04) -- (131.86,140.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.85,138.02) -- (129.85,142.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.69,141.02) -- (133.54,143.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.69,143.88) -- (133.54,141.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.09,142.45) -- (134.13,142.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.11,140.43) -- (132.11,144.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.77,152.24) -- (137.63,155.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.77,155.10) -- (137.63,152.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.18,153.67) -- (138.22,153.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.20,151.65) -- (136.20,155.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.76,153.17) -- (137.61,156.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.76,156.02) -- (137.61,153.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.17,154.60) -- (138.20,154.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.19,152.58) -- (136.19,156.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.20,153.43) -- (139.06,156.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.20,156.28) -- (139.06,153.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.61,154.85) -- (139.65,154.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.63,152.83) -- (137.63,156.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.02,152.69) -- (138.87,155.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.02,155.54) -- (138.87,152.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.42,154.12) -- (139.46,154.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.44,152.10) -- (137.44,156.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.93,155.89) -- (137.78,158.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.93,158.75) -- (137.78,155.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.34,157.32) -- (138.37,157.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.36,155.30) -- (136.36,159.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.99,154.48) -- (138.84,157.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.99,157.33) -- (138.84,154.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.40,155.91) -- (139.44,155.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.42,153.89) -- (137.42,157.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.08,154.01) -- (134.93,156.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.08,156.87) -- (134.93,154.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.49,155.44) -- (135.52,155.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.51,153.42) -- (133.51,157.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.82,155.04) -- (133.68,157.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.82,157.89) -- (133.68,155.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.23,156.47) -- (134.27,156.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.25,154.45) -- (132.25,158.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.14,154.48) -- (134.99,157.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.14,157.33) -- (134.99,154.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.55,155.91) -- (135.59,155.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.57,153.89) -- (133.57,157.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.49,155.12) -- (137.34,157.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.49,157.97) -- (137.34,155.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.90,156.55) -- (137.94,156.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.92,154.53) -- (135.92,158.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.73,156.09) -- (137.59,158.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.73,158.94) -- (137.59,156.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.14,157.51) -- (138.18,157.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.16,155.49) -- (136.16,159.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.39,155.61) -- (140.24,158.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.39,158.47) -- (140.24,155.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.80,157.04) -- (140.83,157.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.81,155.02) -- (138.81,159.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.33,152.31) -- (138.18,155.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.33,155.16) -- (138.18,152.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.74,153.73) -- (138.77,153.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.75,151.72) -- (136.75,155.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.62,155.76) -- (140.48,158.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.62,158.61) -- (140.48,155.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.03,157.18) -- (141.07,157.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.05,155.17) -- (139.05,159.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.70,152.55) -- (135.55,155.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.70,155.41) -- (135.55,152.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.11,153.98) -- (136.14,153.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.12,151.96) -- (134.12,156.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.60,154.41) -- (139.45,157.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.60,157.27) -- (139.45,154.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.01,155.84) -- (140.05,155.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.03,153.82) -- (138.03,157.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.22,154.02) -- (138.08,156.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.22,156.88) -- (138.08,154.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.63,155.45) -- (138.67,155.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.65,153.43) -- (136.65,157.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.98,154.63) -- (140.84,157.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.98,157.49) -- (140.84,154.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.39,156.06) -- (141.43,156.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.41,154.04) -- (139.41,158.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.42,157.93) -- (141.28,160.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.42,160.78) -- (141.28,157.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.83,159.35) -- (141.87,159.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.85,157.34) -- (139.85,161.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.32,154.86) -- (140.18,157.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.32,157.71) -- (140.18,154.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.73,156.29) -- (140.77,156.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.75,154.27) -- (138.75,158.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.53,153.97) -- (138.38,156.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.53,156.82) -- (138.38,153.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.94,155.40) -- (138.97,155.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.95,153.38) -- (136.95,157.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.26,155.60) -- (140.12,158.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.26,158.45) -- (140.12,155.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.67,157.03) -- (140.71,157.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.69,155.01) -- (138.69,159.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.53,154.33) -- (135.39,157.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.53,157.19) -- (135.39,154.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.94,155.76) -- (135.98,155.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.96,153.74) -- (133.96,157.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.03,155.00) -- (136.88,157.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.03,157.86) -- (136.88,155.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.44,156.43) -- (137.47,156.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.46,154.41) -- (135.46,158.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.44,153.98) -- (134.30,156.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.44,156.83) -- (134.30,153.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.85,155.41) -- (134.89,155.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.87,153.39) -- (132.87,157.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.15,156.26) -- (136.00,159.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.15,159.11) -- (136.00,156.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.55,157.69) -- (136.59,157.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.57,155.67) -- (134.57,159.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.08,154.01) -- (135.94,156.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.08,156.86) -- (135.94,154.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.49,155.44) -- (136.53,155.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.51,153.42) -- (134.51,157.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.24,155.99) -- (138.09,158.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.24,158.85) -- (138.09,155.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.65,157.42) -- (138.68,157.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.67,155.40) -- (136.67,159.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.69,158.81) -- (139.54,161.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.69,161.66) -- (139.54,158.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.10,160.24) -- (140.13,160.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.11,158.22) -- (138.11,162.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.67,157.81) -- (139.53,160.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.67,160.67) -- (139.53,157.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.08,159.24) -- (140.12,159.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.10,157.22) -- (138.10,161.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.33,156.22) -- (138.18,159.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.33,159.07) -- (138.18,156.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.74,157.64) -- (138.77,157.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.75,155.63) -- (136.75,159.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.37,157.22) -- (139.23,160.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.37,160.07) -- (139.23,157.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.78,158.64) -- (139.82,158.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.80,156.63) -- (137.80,160.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.15,156.73) -- (138.00,159.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.15,159.58) -- (138.00,156.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.56,158.16) -- (138.59,158.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.57,156.14) -- (136.57,160.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.52,158.71) -- (139.37,161.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.52,161.56) -- (139.37,158.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.93,160.13) -- (139.96,160.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.94,158.12) -- (137.94,162.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.73,155.82) -- (139.58,158.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.73,158.68) -- (139.58,155.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.14,157.25) -- (140.17,157.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.16,155.23) -- (138.16,159.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.74,157.37) -- (140.59,160.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.74,160.23) -- (140.59,157.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.15,158.80) -- (141.18,158.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.17,156.78) -- (139.17,160.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.22,156.78) -- (138.08,159.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.22,159.64) -- (138.08,156.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.63,158.21) -- (138.67,158.21);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.65,156.19) -- (136.65,160.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.62,162.47) -- (142.47,165.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.62,165.32) -- (142.47,162.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.03,163.90) -- (143.06,163.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.04,161.88) -- (141.04,165.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.07,157.62) -- (139.92,160.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.07,160.47) -- (139.92,157.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.47,159.04) -- (140.51,159.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.49,157.03) -- (138.49,161.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.11,156.93) -- (133.96,159.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.11,159.78) -- (133.96,156.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.52,158.35) -- (134.55,158.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.53,156.34) -- (132.53,160.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.87,155.33) -- (134.73,158.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.87,158.18) -- (134.73,155.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.28,156.76) -- (135.32,156.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.30,154.74) -- (133.30,158.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.99,155.50) -- (135.84,158.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.99,158.36) -- (135.84,155.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.40,156.93) -- (136.43,156.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.42,154.91) -- (134.42,158.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.97,156.76) -- (136.82,159.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.97,159.61) -- (136.82,156.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.37,158.18) -- (137.41,158.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.39,156.16) -- (135.39,160.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.52,155.54) -- (141.37,158.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.52,158.39) -- (141.37,155.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.93,156.96) -- (141.97,156.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.95,154.95) -- (139.95,158.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.88,158.45) -- (141.74,161.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.88,161.30) -- (141.74,158.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.29,159.88) -- (142.33,159.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.31,157.86) -- (140.31,161.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.34,158.61) -- (144.20,161.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.34,161.46) -- (144.20,158.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.75,160.03) -- (144.79,160.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.77,158.02) -- (142.77,162.05);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.60,160.33) -- (137.45,163.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.60,163.19) -- (137.45,160.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.00,161.76) -- (138.04,161.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.02,159.74) -- (136.02,163.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.49,156.49) -- (140.34,159.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.49,159.35) -- (140.34,156.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.90,157.92) -- (140.93,157.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.91,155.90) -- (138.91,159.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.44,175.49) -- (151.29,178.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.44,178.34) -- (151.29,175.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.85,176.92) -- (151.88,176.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.86,174.90) -- (149.86,178.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.17,174.87) -- (150.02,177.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.17,177.72) -- (150.02,174.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.58,176.29) -- (150.61,176.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.60,174.28) -- (148.60,178.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.62,177.44) -- (151.48,180.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.62,180.30) -- (151.48,177.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.03,178.87) -- (152.07,178.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.05,176.85) -- (150.05,180.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.89,173.07) -- (149.74,175.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.89,175.92) -- (149.74,173.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.30,174.50) -- (150.33,174.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.31,172.48) -- (148.31,176.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.28,174.53) -- (150.14,177.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.28,177.39) -- (150.14,174.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.69,175.96) -- (150.73,175.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.71,173.94) -- (148.71,177.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.28,174.38) -- (150.14,177.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.28,177.24) -- (150.14,174.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.69,175.81) -- (150.73,175.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.71,173.79) -- (148.71,177.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.66,174.07) -- (144.51,176.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.66,176.92) -- (144.51,174.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.07,175.49) -- (145.10,175.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.08,173.48) -- (143.08,177.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.62,177.73) -- (149.48,180.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.62,180.58) -- (149.48,177.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.03,179.15) -- (150.07,179.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.05,177.14) -- (148.05,181.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.95,175.18) -- (146.80,178.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.95,178.04) -- (146.80,175.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.35,176.61) -- (147.39,176.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.37,174.59) -- (145.37,178.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.65,171.83) -- (143.50,174.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.65,174.68) -- (143.50,171.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.06,173.25) -- (144.09,173.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.07,171.24) -- (142.07,175.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.80,176.79) -- (144.65,179.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.80,179.64) -- (144.65,176.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.20,178.22) -- (145.24,178.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.22,176.20) -- (143.22,180.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.99,174.49) -- (148.84,177.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.99,177.35) -- (148.84,174.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.40,175.92) -- (149.44,175.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.42,173.90) -- (147.42,177.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.11,174.48) -- (152.96,177.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.11,177.33) -- (152.96,174.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.52,175.90) -- (153.55,175.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.53,173.89) -- (151.53,177.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.04,174.03) -- (151.89,176.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.04,176.88) -- (151.89,174.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.45,175.45) -- (152.48,175.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.46,173.44) -- (150.46,177.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.34,175.55) -- (152.19,178.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.34,178.40) -- (152.19,175.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.75,176.98) -- (152.78,176.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.77,174.96) -- (150.77,178.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.59,176.75) -- (148.44,179.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.59,179.60) -- (148.44,176.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.00,178.17) -- (149.03,178.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.01,176.16) -- (147.01,180.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.84,176.62) -- (153.70,179.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.84,179.47) -- (153.70,176.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.25,178.04) -- (154.29,178.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.27,176.03) -- (152.27,180.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.29,173.64) -- (152.14,176.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.29,176.50) -- (152.14,173.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.70,175.07) -- (152.73,175.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.72,173.05) -- (150.72,177.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.09,176.52) -- (149.94,179.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.09,179.37) -- (149.94,176.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.50,177.94) -- (150.53,177.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.51,175.93) -- (148.51,179.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.59,175.51) -- (152.44,178.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.59,178.36) -- (152.44,175.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.99,176.94) -- (153.03,176.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.01,174.92) -- (151.01,178.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.27,180.37) -- (153.12,183.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.27,183.22) -- (153.12,180.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.68,181.80) -- (153.71,181.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.70,179.78) -- (151.70,183.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.07,176.54) -- (151.93,179.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.07,179.39) -- (151.93,176.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.48,177.96) -- (152.52,177.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.50,175.95) -- (150.50,179.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.29,175.70) -- (149.14,178.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.29,178.56) -- (149.14,175.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.69,177.13) -- (149.73,177.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.71,175.11) -- (147.71,179.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.62,175.95) -- (144.48,178.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.62,178.81) -- (144.48,175.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.03,177.38) -- (145.07,177.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.05,175.36) -- (143.05,179.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.29,176.22) -- (145.15,179.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.29,179.08) -- (145.15,176.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.70,177.65) -- (145.74,177.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.72,175.63) -- (143.72,179.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.72,173.55) -- (145.58,176.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.72,176.40) -- (145.58,173.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.13,174.98) -- (146.17,174.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.15,172.96) -- (144.15,176.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.84,176.58) -- (145.69,179.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.84,179.44) -- (145.69,176.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.25,178.01) -- (146.28,178.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.26,175.99) -- (144.26,180.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.88,174.91) -- (150.73,177.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.88,177.77) -- (150.73,174.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.29,176.34) -- (151.32,176.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.31,174.32) -- (149.31,178.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.05,175.45) -- (151.90,178.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.05,178.31) -- (151.90,175.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.46,176.88) -- (152.49,176.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.47,174.86) -- (150.47,178.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.92,174.52) -- (150.78,177.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.92,177.37) -- (150.78,174.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.33,175.95) -- (151.37,175.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.35,173.93) -- (149.35,177.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.47,175.78) -- (149.33,178.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.47,178.63) -- (149.33,175.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.88,177.20) -- (149.92,177.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.90,175.19) -- (147.90,179.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.46,177.35) -- (149.32,180.21);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.46,180.21) -- (149.32,177.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.87,178.78) -- (149.91,178.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.89,176.76) -- (147.89,180.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.58,179.65) -- (152.43,182.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.58,182.50) -- (152.43,179.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.99,181.07) -- (153.02,181.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.00,179.06) -- (151.00,183.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.27,178.96) -- (153.13,181.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.27,181.81) -- (153.13,178.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.68,180.39) -- (153.72,180.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.70,178.37) -- (151.70,182.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.07,179.64) -- (151.92,182.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.07,182.49) -- (151.92,179.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.48,181.06) -- (152.51,181.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.49,179.05) -- (150.49,183.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.42,176.26) -- (151.27,179.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.42,179.11) -- (151.27,176.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.82,177.68) -- (151.86,177.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.84,175.67) -- (149.84,179.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.27,181.05) -- (152.13,183.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.27,183.90) -- (152.13,181.05);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.68,182.48) -- (152.72,182.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.70,180.46) -- (150.70,184.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.23,177.15) -- (153.09,180.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.23,180.00) -- (153.09,177.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.64,178.57) -- (153.68,178.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.66,176.56) -- (151.66,180.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.40,174.13) -- (144.26,176.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.40,176.98) -- (144.26,174.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.81,175.55) -- (144.85,175.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.83,173.54) -- (142.83,177.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.06,178.39) -- (142.91,181.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.06,181.24) -- (142.91,178.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.47,179.81) -- (143.50,179.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.49,177.80) -- (141.49,181.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.80,179.61) -- (143.65,182.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.80,182.47) -- (143.65,179.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.21,181.04) -- (144.24,181.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.23,179.02) -- (142.23,183.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.00,175.42) -- (145.85,178.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.00,178.27) -- (145.85,175.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.41,176.85) -- (146.45,176.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.43,174.83) -- (144.43,178.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.27,176.00) -- (147.12,178.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.27,178.86) -- (147.12,176.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.68,177.43) -- (147.71,177.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.69,175.41) -- (145.69,179.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.54,176.14) -- (151.39,178.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.54,178.99) -- (151.39,176.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.95,177.57) -- (151.98,177.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.97,175.55) -- (149.97,179.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.72,176.19) -- (151.57,179.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.72,179.04) -- (151.57,176.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.13,177.61) -- (152.16,177.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.14,175.60) -- (150.14,179.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.82,178.31) -- (150.67,181.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.82,181.16) -- (150.67,178.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.23,179.74) -- (151.27,179.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.25,177.72) -- (149.25,181.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.89,178.43) -- (149.74,181.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.89,181.29) -- (149.74,178.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.30,179.86) -- (150.33,179.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.31,177.84) -- (148.31,181.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.88,182.26) -- (152.74,185.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.88,185.11) -- (152.74,182.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.29,183.68) -- (153.33,183.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.31,181.66) -- (151.31,185.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.18,186.74) -- (156.03,189.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.18,189.60) -- (156.03,186.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.59,188.17) -- (156.63,188.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.61,186.15) -- (154.61,190.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.99,193.71) -- (155.84,196.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.99,196.56) -- (155.84,193.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.40,195.14) -- (156.43,195.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.41,193.12) -- (154.41,197.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.91,186.98) -- (155.76,189.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.91,189.83) -- (155.76,186.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.32,188.40) -- (156.35,188.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.33,186.39) -- (154.33,190.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.84,182.60) -- (156.69,185.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.84,185.45) -- (156.69,182.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.25,184.03) -- (157.28,184.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.26,182.01) -- (155.26,186.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.63,184.24) -- (157.48,187.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.63,187.09) -- (157.48,184.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.04,185.66) -- (158.07,185.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (156.06,183.65) -- (156.06,187.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.55,184.44) -- (157.41,187.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.55,187.30) -- (157.41,184.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.96,185.87) -- (158.00,185.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.98,183.85) -- (155.98,187.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.03,182.02) -- (146.88,184.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.03,184.88) -- (146.88,182.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.44,183.45) -- (147.47,183.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.45,181.43) -- (145.45,185.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.43,180.15) -- (155.28,183.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.43,183.01) -- (155.28,180.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.84,181.58) -- (155.87,181.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.85,179.56) -- (153.85,183.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.43,180.07) -- (146.28,182.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.43,182.92) -- (146.28,180.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.84,181.50) -- (146.87,181.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.85,179.48) -- (144.85,183.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.01,185.11) -- (146.86,187.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.01,187.96) -- (146.86,185.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.42,186.54) -- (147.45,186.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.44,184.52) -- (145.44,188.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.80,183.02) -- (147.66,185.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.80,185.87) -- (147.66,183.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.21,184.45) -- (148.25,184.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.23,182.43) -- (146.23,186.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.71,180.80) -- (155.56,183.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.71,183.66) -- (155.56,180.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.11,182.23) -- (156.15,182.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.13,180.21) -- (154.13,184.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.80,186.88) -- (155.65,189.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.80,189.73) -- (155.65,186.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.21,188.30) -- (156.24,188.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.22,186.29) -- (154.22,190.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.49,182.23) -- (154.34,185.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.49,185.08) -- (154.34,182.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.90,183.66) -- (154.93,183.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.92,181.64) -- (152.92,185.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.54,185.96) -- (154.40,188.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.54,188.82) -- (154.40,185.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.95,187.39) -- (154.99,187.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.97,185.37) -- (152.97,189.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.39,186.54) -- (156.25,189.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.39,189.39) -- (156.25,186.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.80,187.96) -- (156.84,187.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.82,185.94) -- (154.82,189.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (156.95,186.47) -- (159.80,189.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (156.95,189.32) -- (159.80,186.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (156.36,187.90) -- (160.39,187.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.37,185.88) -- (158.37,189.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.91,193.71) -- (158.76,196.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.91,196.56) -- (158.76,193.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.32,195.14) -- (159.35,195.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.33,193.12) -- (157.33,197.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (156.52,193.71) -- (159.38,196.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (156.52,196.56) -- (159.38,193.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.93,195.14) -- (159.97,195.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.95,193.12) -- (157.95,197.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.65,193.71) -- (158.50,196.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.65,196.56) -- (158.50,193.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.06,195.14) -- (159.10,195.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.08,193.12) -- (157.08,197.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.94,193.71) -- (161.79,196.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.94,196.56) -- (161.79,193.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.35,195.14) -- (162.39,195.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.37,193.12) -- (160.37,197.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.58,193.71) -- (158.43,196.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.58,196.56) -- (158.43,193.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.99,195.14) -- (159.02,195.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.00,193.12) -- (157.00,197.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.18,185.27) -- (148.04,188.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.18,188.12) -- (148.04,185.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.59,186.70) -- (148.63,186.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.61,184.68) -- (146.61,188.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.20,181.12) -- (149.06,183.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.20,183.97) -- (149.06,181.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.61,182.54) -- (149.65,182.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.63,180.53) -- (147.63,184.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.42,184.12) -- (147.28,186.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.42,186.97) -- (147.28,184.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.83,185.54) -- (147.87,185.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.85,183.52) -- (145.85,187.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.15,181.58) -- (148.00,184.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.15,184.43) -- (148.00,181.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.56,183.00) -- (148.59,183.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.58,180.99) -- (146.58,185.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.62,184.34) -- (149.48,187.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.62,187.19) -- (149.48,184.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.03,185.77) -- (150.07,185.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.05,183.75) -- (148.05,187.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.41,186.33) -- (161.26,189.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.41,189.18) -- (161.26,186.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.82,187.75) -- (161.85,187.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.84,185.74) -- (159.84,189.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.50,193.71) -- (157.35,196.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.50,196.56) -- (157.35,193.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.91,195.14) -- (157.94,195.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.93,193.12) -- (155.93,197.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.29,187.19) -- (158.14,190.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.29,190.04) -- (158.14,187.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.70,188.62) -- (158.73,188.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (156.72,186.60) -- (156.72,190.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.49,193.71) -- (156.34,196.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.49,196.56) -- (156.34,193.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.90,195.14) -- (156.93,195.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.92,193.12) -- (154.92,197.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.04,186.66) -- (159.90,189.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.04,189.51) -- (159.90,186.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (156.45,188.09) -- (160.49,188.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.47,186.07) -- (158.47,190.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.81,193.71) -- (160.66,196.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.81,196.56) -- (160.66,193.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.22,195.14) -- (161.25,195.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.24,193.12) -- (159.24,197.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.60,186.14) -- (166.45,189.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.60,189.00) -- (166.45,186.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.01,187.57) -- (167.04,187.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.03,185.55) -- (165.03,189.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.76,185.40) -- (164.62,188.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.76,188.25) -- (164.62,185.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.17,186.82) -- (165.21,186.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.19,184.80) -- (163.19,188.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.17,185.28) -- (164.02,188.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.17,188.13) -- (164.02,185.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.58,186.71) -- (164.61,186.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.59,184.69) -- (162.59,188.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.38,187.34) -- (162.24,190.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.38,190.19) -- (162.24,187.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.79,188.77) -- (162.83,188.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.81,186.75) -- (160.81,190.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.30,187.16) -- (166.15,190.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.30,190.01) -- (166.15,187.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.71,188.59) -- (166.74,188.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.72,186.57) -- (164.72,190.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.73,185.56) -- (148.59,188.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.73,188.42) -- (148.59,185.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.14,186.99) -- (149.18,186.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.16,184.97) -- (147.16,189.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.50,182.02) -- (151.35,184.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.50,184.87) -- (151.35,182.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.91,183.44) -- (151.94,183.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.93,181.43) -- (149.93,185.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.82,186.24) -- (149.68,189.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.82,189.10) -- (149.68,186.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.23,187.67) -- (150.27,187.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.25,185.65) -- (148.25,189.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.90,181.26) -- (149.75,184.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.90,184.11) -- (149.75,181.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.31,182.68) -- (150.35,182.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.33,180.66) -- (148.33,184.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.52,185.10) -- (149.38,187.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.52,187.95) -- (149.38,185.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.93,186.53) -- (149.97,186.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.95,184.51) -- (147.95,188.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.08,182.82) -- (156.94,185.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.08,185.67) -- (156.94,182.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.49,184.25) -- (157.53,184.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.51,182.23) -- (155.51,186.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.87,193.71) -- (158.72,196.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.87,196.56) -- (158.72,193.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.28,195.14) -- (159.31,195.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.30,193.12) -- (157.30,197.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.87,185.71) -- (156.72,188.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.87,188.56) -- (156.72,185.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.28,187.14) -- (157.31,187.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.29,185.12) -- (155.29,189.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.85,188.79) -- (161.70,191.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.85,191.64) -- (161.70,188.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.26,190.22) -- (162.29,190.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.28,188.20) -- (160.28,192.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.89,188.04) -- (156.75,190.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.89,190.89) -- (156.75,188.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.30,189.46) -- (157.34,189.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.32,187.45) -- (155.32,191.48);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor,fill opacity=0.50] (118.09, 99.18) --
	(120.94, 99.18) --
	(120.94,102.03) --
	(118.09,102.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.34,100.72) --
	(122.20,100.72) --
	(122.20,103.57) --
	(119.34,103.57) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.43,108.55) --
	(118.28,108.55) --
	(118.28,111.40) --
	(115.43,111.40) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.04, 97.15) --
	(119.89, 97.15) --
	(119.89,100.00) --
	(117.04,100.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.84, 98.97) --
	(120.69, 98.97) --
	(120.69,101.82) --
	(117.84,101.82) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.43,105.52) --
	(118.28,105.52) --
	(118.28,108.38) --
	(115.43,108.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.13, 99.78) --
	(118.98, 99.78) --
	(118.98,102.63) --
	(116.13,102.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.23, 96.90) --
	(118.08, 96.90) --
	(118.08, 99.76) --
	(115.23, 99.76) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.09, 96.90) --
	(116.94, 96.90) --
	(116.94, 99.76) --
	(114.09, 99.76) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.34, 98.75) --
	(119.20, 98.75) --
	(119.20,101.61) --
	(116.34,101.61) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.91,101.25) --
	(119.77,101.25) --
	(119.77,104.10) --
	(116.91,104.10) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.49, 99.18) --
	(118.34, 99.18) --
	(118.34,102.03) --
	(115.49,102.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.96,108.27) --
	(116.81,108.27) --
	(116.81,111.12) --
	(113.96,111.12) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.69, 99.78) --
	(122.54, 99.78) --
	(122.54,102.63) --
	(119.69,102.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.79, 99.38) --
	(119.64, 99.38) --
	(119.64,102.23) --
	(116.79,102.23) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.26, 98.75) --
	(120.11, 98.75) --
	(120.11,101.61) --
	(117.26,101.61) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (127.04,117.01) --
	(129.90,117.01) --
	(129.90,119.86) --
	(127.04,119.86) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.52,111.31) --
	(124.38,111.31) --
	(124.38,114.16) --
	(121.52,114.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.92,108.91) --
	(122.78,108.91) --
	(122.78,111.76) --
	(119.92,111.76) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.64,110.09) --
	(123.49,110.09) --
	(123.49,112.94) --
	(120.64,112.94) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.19,109.85) --
	(124.05,109.85) --
	(124.05,112.70) --
	(121.19,112.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.47,110.48) --
	(124.32,110.48) --
	(124.32,113.34) --
	(121.47,113.34) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.18,110.56) --
	(122.03,110.56) --
	(122.03,113.41) --
	(119.18,113.41) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.09,109.51) --
	(123.94,109.51) --
	(123.94,112.37) --
	(121.09,112.37) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.87,111.66) --
	(124.73,111.66) --
	(124.73,114.52) --
	(121.87,114.52) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.91,118.99) --
	(123.77,118.99) --
	(123.77,121.84) --
	(120.91,121.84) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.42,114.23) --
	(124.27,114.23) --
	(124.27,117.08) --
	(121.42,117.08) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.81,113.94) --
	(124.66,113.94) --
	(124.66,116.80) --
	(121.81,116.80) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.68,113.71) --
	(126.54,113.71) --
	(126.54,116.56) --
	(123.68,116.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.71,118.30) --
	(126.56,118.30) --
	(126.56,121.15) --
	(123.71,121.15) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (125.54,117.14) --
	(128.39,117.14) --
	(128.39,119.99) --
	(125.54,119.99) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.07,113.59) --
	(126.92,113.59) --
	(126.92,116.45) --
	(124.07,116.45) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (146.09,141.63) --
	(148.95,141.63) --
	(148.95,144.49) --
	(146.09,144.49) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.04,128.41) --
	(126.90,128.41) --
	(126.90,131.27) --
	(124.04,131.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.46,123.69) --
	(127.31,123.69) --
	(127.31,126.54) --
	(124.46,126.54) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.65,123.81) --
	(126.51,123.81) --
	(126.51,126.66) --
	(123.65,126.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (125.76,126.41) --
	(128.62,126.41) --
	(128.62,129.26) --
	(125.76,129.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.71,125.38) --
	(125.57,125.38) --
	(125.57,128.24) --
	(122.71,128.24) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (128.62,121.87) --
	(131.48,121.87) --
	(131.48,124.72) --
	(128.62,124.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.82,119.46) --
	(124.68,119.46) --
	(124.68,122.31) --
	(121.82,122.31) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.39,123.95) --
	(125.24,123.95) --
	(125.24,126.81) --
	(122.39,126.81) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.71,118.38) --
	(123.57,118.38) --
	(123.57,121.23) --
	(120.71,121.23) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.55,120.77) --
	(126.41,120.77) --
	(126.41,123.63) --
	(123.55,123.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.82,140.25) --
	(137.68,140.25) --
	(137.68,143.11) --
	(134.82,143.11) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.27,142.29) --
	(137.13,142.29) --
	(137.13,145.14) --
	(134.27,145.14) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.55,139.69) --
	(136.40,139.69) --
	(136.40,142.54) --
	(133.55,142.54) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.95,143.87) --
	(134.80,143.87) --
	(134.80,146.72) --
	(131.95,146.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.01,142.80) --
	(133.86,142.80) --
	(133.86,145.65) --
	(131.01,145.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.44,102.24) --
	(123.29,102.24) --
	(123.29,105.09) --
	(120.44,105.09) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.50, 99.97) --
	(120.36, 99.97) --
	(120.36,102.83) --
	(117.50,102.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.60, 99.78) --
	(117.45, 99.78) --
	(117.45,102.63) --
	(114.60,102.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.40,107.89) --
	(119.25,107.89) --
	(119.25,110.74) --
	(116.40,110.74) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.36,101.07) --
	(121.21,101.07) --
	(121.21,103.93) --
	(118.36,103.93) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.20, 99.97) --
	(121.06, 99.97) --
	(121.06,102.83) --
	(118.20,102.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.14,101.59) --
	(120.99,101.59) --
	(120.99,104.44) --
	(118.14,104.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.79,101.59) --
	(122.64,101.59) --
	(122.64,104.44) --
	(119.79,104.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.09, 99.58) --
	(120.94, 99.58) --
	(120.94,102.43) --
	(118.09,102.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.65,103.72) --
	(121.50,103.72) --
	(121.50,106.58) --
	(118.65,106.58) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.91,101.75) --
	(118.76,101.75) --
	(118.76,104.61) --
	(115.91,104.61) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.41,101.42) --
	(120.26,101.42) --
	(120.26,104.27) --
	(117.41,104.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.66,103.58) --
	(123.51,103.58) --
	(123.51,106.44) --
	(120.66,106.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.83,103.00) --
	(122.68,103.00) --
	(122.68,105.86) --
	(119.83,105.86) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.32,103.58) --
	(125.18,103.58) --
	(125.18,106.44) --
	(122.32,106.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.46,103.00) --
	(123.31,103.00) --
	(123.31,105.86) --
	(120.46,105.86) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.18,138.57) --
	(142.03,138.57) --
	(142.03,141.42) --
	(139.18,141.42) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.21,112.80) --
	(124.07,112.80) --
	(124.07,115.65) --
	(121.21,115.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.59,112.98) --
	(122.44,112.98) --
	(122.44,115.84) --
	(119.59,115.84) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.42,116.37) --
	(124.27,116.37) --
	(124.27,119.22) --
	(121.42,119.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.11,113.59) --
	(123.96,113.59) --
	(123.96,116.45) --
	(121.11,116.45) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.86,113.94) --
	(121.72,113.94) --
	(121.72,116.80) --
	(118.86,116.80) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.84,115.25) --
	(123.69,115.25) --
	(123.69,118.10) --
	(120.84,118.10) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.55,114.50) --
	(122.40,114.50) --
	(122.40,117.36) --
	(119.55,117.36) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.91,118.42) --
	(124.76,118.42) --
	(124.76,121.27) --
	(121.91,121.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.05,112.73) --
	(121.91,112.73) --
	(121.91,115.59) --
	(119.05,115.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.62,111.80) --
	(121.48,111.80) --
	(121.48,114.65) --
	(118.62,114.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.09,127.09) --
	(132.95,127.09) --
	(132.95,129.94) --
	(130.09,129.94) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (127.96,122.97) --
	(130.82,122.97) --
	(130.82,125.82) --
	(127.96,125.82) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (128.34,122.84) --
	(131.19,122.84) --
	(131.19,125.69) --
	(128.34,125.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.08,123.51) --
	(131.94,123.51) --
	(131.94,126.36) --
	(129.08,126.36) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (128.55,126.07) --
	(131.40,126.07) --
	(131.40,128.92) --
	(128.55,128.92) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (125.80,116.46) --
	(128.65,116.46) --
	(128.65,119.32) --
	(125.80,119.32) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.36,104.53) --
	(120.21,104.53) --
	(120.21,107.38) --
	(117.36,107.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.91,106.10) --
	(123.77,106.10) --
	(123.77,108.96) --
	(120.91,108.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.34,104.53) --
	(121.19,104.53) --
	(121.19,107.38) --
	(118.34,107.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.10,106.55) --
	(122.95,106.55) --
	(122.95,109.40) --
	(120.10,109.40) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.48,104.13) --
	(120.33,104.13) --
	(120.33,106.99) --
	(117.48,106.99) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.38,105.88) --
	(126.24,105.88) --
	(126.24,108.73) --
	(123.38,108.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.69,105.04) --
	(121.54,105.04) --
	(121.54,107.89) --
	(118.69,107.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.58,105.04) --
	(121.43,105.04) --
	(121.43,107.89) --
	(118.58,107.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.47,107.59) --
	(122.32,107.59) --
	(122.32,110.44) --
	(119.47,110.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.38,105.04) --
	(122.24,105.04) --
	(122.24,107.89) --
	(119.38,107.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.73,108.27) --
	(123.59,108.27) --
	(123.59,111.12) --
	(120.73,111.12) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.21,112.21) --
	(124.07,112.21) --
	(124.07,115.06) --
	(121.21,115.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.00,107.98) --
	(124.86,107.98) --
	(124.86,110.84) --
	(122.00,110.84) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.06,106.10) --
	(122.91,106.10) --
	(122.91,108.96) --
	(120.06,108.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.07,112.14) --
	(126.92,112.14) --
	(126.92,115.00) --
	(124.07,115.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (127.13,132.27) --
	(129.98,132.27) --
	(129.98,135.13) --
	(127.13,135.13) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.05,106.66) --
	(121.91,106.66) --
	(121.91,109.51) --
	(119.05,109.51) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.94,108.17) --
	(118.79,108.17) --
	(118.79,111.03) --
	(115.94,111.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.77,107.69) --
	(120.62,107.69) --
	(120.62,110.54) --
	(117.77,110.54) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.57,109.26) --
	(123.42,109.26) --
	(123.42,112.11) --
	(120.57,112.11) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.04,108.73) --
	(119.89,108.73) --
	(119.89,111.58) --
	(117.04,111.58) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.76,106.10) --
	(121.61,106.10) --
	(121.61,108.96) --
	(118.76,108.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.16,109.51) --
	(122.01,109.51) --
	(122.01,112.37) --
	(119.16,112.37) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.84,107.79) --
	(121.69,107.79) --
	(121.69,110.64) --
	(118.84,110.64) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.80,114.72) --
	(118.65,114.72) --
	(118.65,117.57) --
	(115.80,117.57) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.78,109.51) --
	(121.63,109.51) --
	(121.63,112.37) --
	(118.78,112.37) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.95,121.84) --
	(126.80,121.84) --
	(126.80,124.69) --
	(123.95,124.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.04,122.94) --
	(126.90,122.94) --
	(126.90,125.80) --
	(124.04,125.80) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (125.89,126.21) --
	(128.75,126.21) --
	(128.75,129.06) --
	(125.89,129.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.53,126.27) --
	(126.38,126.27) --
	(126.38,129.12) --
	(123.53,129.12) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.72,128.28) --
	(126.58,128.28) --
	(126.58,131.13) --
	(123.72,131.13) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.70,193.71) --
	(145.55,193.71) --
	(145.55,196.56) --
	(142.70,196.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.37,183.07) --
	(126.22,183.07) --
	(126.22,185.92) --
	(123.37,185.92) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.36,171.74) --
	(122.22,171.74) --
	(122.22,174.60) --
	(119.36,174.60) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.42,154.12) --
	(125.27,154.12) --
	(125.27,156.97) --
	(122.42,156.97) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.89,160.65) --
	(126.74,160.65) --
	(126.74,163.50) --
	(123.89,163.50) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.18,173.50) --
	(127.03,173.50) --
	(127.03,176.36) --
	(124.18,176.36) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.00,113.53) --
	(120.85,113.53) --
	(120.85,116.39) --
	(118.00,116.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.84,116.04) --
	(124.69,116.04) --
	(124.69,118.89) --
	(121.84,118.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.47,114.99) --
	(123.33,114.99) --
	(123.33,117.84) --
	(120.47,117.84) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.97,116.56) --
	(125.82,116.56) --
	(125.82,119.41) --
	(122.97,119.41) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.77,193.71) --
	(125.63,193.71) --
	(125.63,196.56) --
	(122.77,196.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.56,193.71) --
	(133.42,193.71) --
	(133.42,196.56) --
	(130.56,196.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.50,193.71) --
	(132.35,193.71) --
	(132.35,196.56) --
	(129.50,196.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.45,193.71) --
	(136.30,193.71) --
	(136.30,196.56) --
	(133.45,196.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.90,193.71) --
	(136.75,193.71) --
	(136.75,196.56) --
	(133.90,196.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.86,193.71) --
	(137.71,193.71) --
	(137.71,196.56) --
	(134.86,196.56) --
	cycle;
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (105.68, 86.83) -- (109.72, 86.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (107.70, 84.81) -- (107.70, 88.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (105.55, 82.80) -- (109.59, 82.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (107.57, 80.78) -- (107.57, 84.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (103.60, 81.80) -- (107.64, 81.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (105.62, 79.78) -- (105.62, 83.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (106.12, 83.73) -- (110.15, 83.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.14, 81.71) -- (108.14, 85.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (105.29, 83.73) -- (109.33, 83.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (107.31, 81.71) -- (107.31, 85.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (106.00, 84.58) -- (110.03, 84.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.01, 82.56) -- (108.01, 86.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (103.60, 82.80) -- (107.64, 82.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (105.62, 80.78) -- (105.62, 84.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (106.48, 83.73) -- (110.51, 83.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.50, 81.71) -- (108.50, 85.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (105.62, 83.73) -- (109.65, 83.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (107.64, 81.71) -- (107.64, 85.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (106.83, 83.73) -- (110.86, 83.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.85, 81.71) -- (108.85, 85.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (105.43, 82.80) -- (109.46, 82.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (107.44, 80.78) -- (107.44, 84.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (105.75, 82.80) -- (109.78, 82.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (107.76, 80.78) -- (107.76, 84.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (105.68, 83.73) -- (109.72, 83.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (107.70, 81.71) -- (107.70, 85.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (104.27, 81.80) -- (108.30, 81.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (106.28, 79.78) -- (106.28, 83.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (106.24, 83.73) -- (110.28, 83.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.26, 81.71) -- (108.26, 85.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (105.93, 84.58) -- (109.97, 84.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (107.95, 82.56) -- (107.95, 86.60);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.77,101.25) -- (137.62,104.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.77,104.10) -- (137.62,101.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.18,102.67) -- (138.21,102.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.20,100.66) -- (136.20,104.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.01,100.35) -- (119.87,103.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.01,103.20) -- (119.87,100.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.42,101.78) -- (120.46,101.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.44, 99.76) -- (118.44,103.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.15,100.16) -- (119.01,103.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.15,103.02) -- (119.01,100.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.56,101.59) -- (119.60,101.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.58, 99.57) -- (117.58,103.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.02,100.90) -- (118.87,103.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.02,103.75) -- (118.87,100.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.43,102.32) -- (119.46,102.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.45,100.31) -- (117.45,104.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.76,100.54) -- (119.61,103.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.76,103.39) -- (119.61,100.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.17,101.96) -- (120.20,101.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.19, 99.95) -- (118.19,103.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.86,101.07) -- (119.72,103.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.86,103.93) -- (119.72,101.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.27,102.50) -- (120.31,102.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.29,100.48) -- (118.29,104.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.69, 98.75) -- (117.54,101.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.69,101.61) -- (117.54, 98.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.10,100.18) -- (118.13,100.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.12, 98.16) -- (116.12,102.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.52, 99.38) -- (118.37,102.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.52,102.23) -- (118.37, 99.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.93,100.81) -- (118.96,100.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.94, 98.79) -- (116.94,102.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.38,100.72) -- (120.24,103.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.38,103.57) -- (120.24,100.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.79,102.14) -- (120.83,102.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.81,100.13) -- (118.81,104.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.43, 99.78) -- (118.28,102.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.43,102.63) -- (118.28, 99.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.84,101.20) -- (118.87,101.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.86, 99.19) -- (116.86,103.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.52, 99.97) -- (118.37,102.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.52,102.83) -- (118.37, 99.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.93,101.40) -- (118.96,101.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.94, 99.38) -- (116.94,103.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.15,100.16) -- (119.01,103.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.15,103.02) -- (119.01,100.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.56,101.59) -- (119.60,101.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.58, 99.57) -- (117.58,103.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.71, 99.97) -- (118.57,102.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.71,102.83) -- (118.57, 99.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.12,101.40) -- (119.16,101.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.14, 99.38) -- (117.14,103.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.57,100.35) -- (118.43,103.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.57,103.20) -- (118.43,100.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.98,101.78) -- (119.02,101.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.00, 99.76) -- (117.00,103.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.85,100.35) -- (118.71,103.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.85,103.20) -- (118.71,100.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.26,101.78) -- (119.30,101.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.28, 99.76) -- (117.28,103.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.47,100.54) -- (119.33,103.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.47,103.39) -- (119.33,100.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.88,101.96) -- (119.92,101.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.90, 99.95) -- (117.90,103.98);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,157.19) -- (190.63,157.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,155.18) -- (188.62,159.21);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,156.54) -- (190.63,156.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,154.52) -- (188.62,158.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,157.32) -- (190.63,157.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,155.30) -- (188.62,159.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,158.13) -- (190.63,158.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,156.11) -- (188.62,160.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,158.10) -- (190.63,158.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,156.09) -- (188.62,160.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,158.54) -- (190.63,158.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,156.53) -- (188.62,160.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,157.92) -- (190.63,157.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,155.90) -- (188.62,159.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,158.07) -- (190.63,158.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,156.05) -- (188.62,160.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (175.14,158.28) -- (179.18,158.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.16,156.26) -- (177.16,160.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,156.44) -- (190.63,156.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,154.42) -- (188.62,158.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,156.06) -- (190.63,156.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,154.04) -- (188.62,158.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,157.17) -- (190.63,157.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,155.15) -- (188.62,159.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.98,157.55) -- (181.01,157.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.00,155.53) -- (179.00,159.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.50,159.16) -- (176.54,159.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (174.52,157.14) -- (174.52,161.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.91,159.34) -- (176.95,159.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (174.93,157.33) -- (174.93,161.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (174.33,159.31) -- (178.37,159.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (176.35,157.29) -- (176.35,161.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.68, 96.12) -- (112.72, 96.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.70, 94.10) -- (110.70, 98.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60, 96.71) -- (190.63, 96.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62, 94.70) -- (188.62, 98.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60, 97.28) -- (190.63, 97.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62, 95.26) -- (188.62, 99.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60, 97.28) -- (190.63, 97.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62, 95.26) -- (188.62, 99.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60, 97.82) -- (190.63, 97.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62, 95.80) -- (188.62, 99.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60, 97.28) -- (190.63, 97.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62, 95.26) -- (188.62, 99.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (109.30, 97.00) -- (113.34, 97.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.32, 94.98) -- (111.32, 99.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.08, 97.28) -- (112.12, 97.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.10, 95.26) -- (110.10, 99.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.20, 98.82) -- (114.23, 98.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.21, 96.80) -- (112.21,100.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (109.53, 96.12) -- (113.57, 96.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.55, 94.10) -- (111.55, 98.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.78, 97.28) -- (112.82, 97.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.80, 95.26) -- (110.80, 99.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (109.89, 96.71) -- (113.93, 96.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.91, 94.70) -- (111.91, 98.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.41, 97.28) -- (114.45, 97.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.43, 95.26) -- (112.43, 99.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.73, 97.28) -- (112.77, 97.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.75, 95.26) -- (110.75, 99.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.49, 97.82) -- (112.52, 97.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.50, 95.80) -- (110.50, 99.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.68, 96.42) -- (112.72, 96.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.70, 94.40) -- (110.70, 98.44);

\path[fill=fillColor,fill opacity=0.50] (119.79, 99.58) --
	(122.64, 99.58) --
	(122.64,102.43) --
	(119.79,102.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.67,101.25) --
	(120.52,101.25) --
	(120.52,104.10) --
	(117.67,104.10) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.99,100.16) --
	(118.84,100.16) --
	(118.84,103.02) --
	(115.99,103.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.37,101.42) --
	(119.22,101.42) --
	(119.22,104.27) --
	(116.37,104.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.41,101.07) --
	(120.26,101.07) --
	(120.26,103.93) --
	(117.41,103.93) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.50,101.07) --
	(120.36,101.07) --
	(120.36,103.93) --
	(117.50,103.93) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.24,101.59) --
	(120.09,101.59) --
	(120.09,104.44) --
	(117.24,104.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.21,102.24) --
	(119.06,102.24) --
	(119.06,105.09) --
	(116.21,105.09) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.50,102.24) --
	(119.35,102.24) --
	(119.35,105.09) --
	(116.50,105.09) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.29, 99.58) --
	(119.14, 99.58) --
	(119.14,102.43) --
	(116.29,102.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.46,101.25) --
	(118.31,101.25) --
	(118.31,104.10) --
	(115.46,104.10) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.24,102.08) --
	(120.09,102.08) --
	(120.09,104.93) --
	(117.24,104.93) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.19, 99.97) --
	(120.04, 99.97) --
	(120.04,102.83) --
	(117.19,102.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.84,102.55) --
	(120.69,102.55) --
	(120.69,105.40) --
	(117.84,105.40) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.04,102.24) --
	(119.89,102.24) --
	(119.89,105.09) --
	(117.04,105.09) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.67,102.40) --
	(120.52,102.40) --
	(120.52,105.25) --
	(117.67,105.25) --
	cycle;

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.21,175.48) -- (152.25,175.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.23,173.47) -- (150.23,177.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.25,174.94) -- (149.29,174.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.27,172.92) -- (147.27,176.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.53,174.86) -- (155.57,174.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.55,172.84) -- (153.55,176.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.19,174.16) -- (150.23,174.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.21,172.15) -- (148.21,176.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.12,174.04) -- (151.15,174.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.14,172.02) -- (149.14,176.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.51,173.64) -- (149.54,173.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.53,171.63) -- (147.53,175.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.32,174.82) -- (142.36,174.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.34,172.80) -- (140.34,176.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,175.39) -- (190.63,175.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,173.37) -- (188.62,177.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,174.39) -- (190.63,174.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,172.37) -- (188.62,176.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.81,174.16) -- (121.85,174.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.83,172.15) -- (119.83,176.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.29,174.72) -- (119.33,174.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.31,172.70) -- (117.31,176.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.42,173.83) -- (165.46,173.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.44,171.81) -- (163.44,175.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.32,173.77) -- (170.35,173.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.33,171.75) -- (168.33,175.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.45,174.85) -- (170.49,174.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.47,172.83) -- (168.47,176.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.08,172.88) -- (167.12,172.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.10,170.87) -- (165.10,174.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.71,175.10) -- (154.74,175.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.73,173.08) -- (152.73,177.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.88,147.82) -- (165.92,147.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.90,145.80) -- (163.90,149.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.07,123.69) -- (118.10,123.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.08,121.67) -- (116.08,125.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.53,113.02) -- (117.57,113.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.55,111.00) -- (115.55,115.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.80,112.29) -- (116.84,112.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.82,110.27) -- (114.82,114.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.30,120.19) -- (115.34,120.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.32,118.17) -- (113.32,122.21);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.56,114.90) -- (117.60,114.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.58,112.88) -- (115.58,116.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.99,124.21) -- (120.02,124.21);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.01,122.19) -- (118.01,126.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,117.51) -- (190.63,117.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,115.50) -- (188.62,119.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.84,110.42) -- (115.87,110.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.85,108.41) -- (113.85,112.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.01,113.02) -- (118.04,113.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.02,111.00) -- (116.02,115.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.96,131.87) -- (120.00,131.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.98,129.85) -- (117.98,133.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.35,147.26) -- (138.39,147.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.37,145.24) -- (136.37,149.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.02,159.25) -- (166.06,159.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.04,157.23) -- (164.04,161.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.20,158.47) -- (167.24,158.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.22,156.46) -- (165.22,160.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.69,160.65) -- (166.72,160.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.70,158.63) -- (164.70,162.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.68,157.40) -- (165.71,157.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.70,155.38) -- (163.70,159.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.73,133.69) -- (155.76,133.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.75,131.67) -- (153.75,135.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.99,132.43) -- (133.03,132.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.01,130.42) -- (131.01,134.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.48,132.17) -- (125.51,132.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.50,130.15) -- (123.50,134.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.83,131.17) -- (125.86,131.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.85,129.15) -- (123.85,133.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.75,130.52) -- (123.79,130.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.77,128.51) -- (121.77,132.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.13,131.39) -- (125.17,131.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.15,129.37) -- (123.15,133.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.44,130.78) -- (122.48,130.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.46,128.77) -- (120.46,132.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.10,130.49) -- (123.13,130.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.11,128.48) -- (121.11,132.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.86,131.74) -- (124.90,131.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.88,129.73) -- (122.88,133.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.81,131.08) -- (122.85,131.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.83,129.07) -- (120.83,133.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.53,131.66) -- (123.56,131.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.54,129.64) -- (121.54,133.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,134.58) -- (190.63,134.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,132.57) -- (188.62,136.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,136.65) -- (190.63,136.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,134.64) -- (188.62,138.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,134.77) -- (190.63,134.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,132.75) -- (188.62,136.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,134.74) -- (190.63,134.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,132.72) -- (188.62,136.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,134.41) -- (190.63,134.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,132.39) -- (188.62,136.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.32,132.11) -- (156.35,132.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.34,130.10) -- (154.34,134.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.78,132.60) -- (124.81,132.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.79,130.59) -- (122.79,134.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.14,132.52) -- (126.17,132.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.16,130.51) -- (124.16,134.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.73,132.74) -- (127.77,132.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.75,130.73) -- (125.75,134.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.26,132.30) -- (126.30,132.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.28,130.28) -- (124.28,134.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.98,130.24) -- (128.02,130.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.00,128.22) -- (126.00,132.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.81,130.35) -- (136.85,130.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.83,128.34) -- (134.83,132.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.49,131.63) -- (125.53,131.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.51,129.61) -- (123.51,133.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.54,131.26) -- (123.58,131.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.56,129.24) -- (121.56,133.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.92,129.59) -- (121.96,129.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.94,127.57) -- (119.94,131.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.87,131.77) -- (123.90,131.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.88,129.75) -- (121.88,133.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.09,133.74) -- (161.12,133.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.11,131.72) -- (159.11,135.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.91,133.58) -- (153.94,133.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.93,131.56) -- (151.93,135.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.58,133.03) -- (155.62,133.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.60,131.02) -- (153.60,135.05);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.03,132.14) -- (143.06,132.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.04,130.12) -- (141.04,134.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (156.43,132.56) -- (160.46,132.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.45,130.55) -- (158.45,134.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.25,137.82) -- (157.28,137.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.26,135.81) -- (155.26,139.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.41,137.14) -- (126.45,137.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.43,135.12) -- (124.43,139.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.33,135.70) -- (130.37,135.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.35,133.68) -- (128.35,137.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.40,135.89) -- (127.43,135.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (125.41,133.87) -- (125.41,137.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.58,135.01) -- (131.62,135.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (129.60,132.99) -- (129.60,137.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.47,137.25) -- (126.50,137.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.49,135.23) -- (124.49,139.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.05,137.84) -- (161.08,137.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.06,135.82) -- (159.06,139.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,134.57) -- (190.63,134.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,132.56) -- (188.62,136.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.33,132.85) -- (123.37,132.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.35,130.83) -- (121.35,134.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.14,132.42) -- (124.18,132.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.16,130.40) -- (122.16,134.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,133.82) -- (190.63,133.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,131.80) -- (188.62,135.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,141.44) -- (190.63,141.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,139.43) -- (188.62,143.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,138.96) -- (190.63,138.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,136.94) -- (188.62,140.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,137.54) -- (190.63,137.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,135.52) -- (188.62,139.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,140.05) -- (190.63,140.05);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,138.03) -- (188.62,142.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,140.72) -- (190.63,140.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,138.70) -- (188.62,142.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.73,128.95) -- (123.77,128.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.75,126.93) -- (121.75,130.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.69,128.61) -- (115.72,128.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.70,126.59) -- (113.70,130.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.43,128.88) -- (118.47,128.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.45,126.86) -- (116.45,130.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.69,128.97) -- (115.72,128.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.70,126.95) -- (113.70,130.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.76,126.42) -- (115.80,126.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.78,124.40) -- (113.78,128.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.24,126.57) -- (114.28,126.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.26,124.56) -- (112.26,128.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.01,128.41) -- (121.04,128.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.03,126.39) -- (119.03,130.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.99,128.91) -- (120.02,128.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.01,126.90) -- (118.01,130.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.66,131.72) -- (114.70,131.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.68,129.70) -- (112.68,133.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.88,129.00) -- (119.92,129.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.90,126.99) -- (117.90,131.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.25,128.22) -- (122.29,128.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.27,126.20) -- (120.27,130.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.29,128.88) -- (119.33,128.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.31,126.86) -- (117.31,130.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.28,128.44) -- (118.32,128.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.30,126.42) -- (116.30,130.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.88,128.70) -- (119.92,128.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.90,126.68) -- (117.90,130.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.22,127.60) -- (118.25,127.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.24,125.58) -- (116.24,129.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.46,128.57) -- (119.49,128.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.47,126.55) -- (117.47,130.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.77,121.06) -- (134.80,121.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.78,119.05) -- (132.78,123.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,105.15) -- (190.63,105.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,103.13) -- (188.62,107.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.80,103.98) -- (115.83,103.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.82,101.96) -- (113.82,106.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60, 99.75) -- (190.63, 99.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62, 97.73) -- (188.62,101.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60, 99.97) -- (190.63, 99.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62, 97.95) -- (188.62,101.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.34,105.96) -- (117.37,105.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.36,103.94) -- (115.36,107.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,100.18) -- (190.63,100.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62, 98.16) -- (188.62,102.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.38,100.60) -- (116.42,100.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.40, 98.58) -- (114.40,102.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,100.39) -- (190.63,100.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62, 98.38) -- (188.62,102.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.27,103.18) -- (117.31,103.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.29,101.16) -- (115.29,105.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.50, 99.52) -- (115.53, 99.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.51, 97.50) -- (113.51,101.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.56,107.98) -- (119.60,107.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.58,105.96) -- (117.58,109.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.59,107.30) -- (121.62,107.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.61,105.28) -- (119.61,109.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.17,108.19) -- (121.21,108.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.19,106.17) -- (119.19,110.21);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.01,100.18) -- (119.05,100.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.03, 98.16) -- (117.03,102.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.12,108.51) -- (122.16,108.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.14,106.49) -- (120.14,110.53);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.63, 87.84) rectangle (116.48, 90.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.63, 87.84) -- (116.48, 90.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.63, 90.69) -- (116.48, 87.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.72, 89.86) rectangle (115.58, 92.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.72, 89.86) -- (115.58, 92.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.72, 92.71) -- (115.58, 89.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.90, 88.89) rectangle (115.76, 91.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.90, 88.89) -- (115.76, 91.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.90, 91.75) -- (115.76, 88.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.04, 94.69) rectangle (115.90, 97.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.04, 94.69) -- (115.90, 97.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.04, 97.55) -- (115.90, 94.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.18, 88.38) rectangle (116.04, 91.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.18, 88.38) -- (116.04, 91.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.18, 91.23) -- (116.04, 88.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.25, 88.38) rectangle (116.11, 91.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.25, 88.38) -- (116.11, 91.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.25, 91.23) -- (116.11, 88.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.89, 87.84) rectangle (114.75, 90.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.89, 87.84) -- (114.75, 90.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.89, 90.69) -- (114.75, 87.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.09, 87.28) rectangle (114.94, 90.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.09, 87.28) -- (114.94, 90.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.09, 90.13) -- (114.94, 87.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.54, 88.38) rectangle (114.39, 91.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.54, 88.38) -- (114.39, 91.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.54, 91.23) -- (114.39, 88.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.17, 88.89) rectangle (114.02, 91.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.17, 88.89) -- (114.02, 91.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.17, 91.75) -- (114.02, 88.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.42, 89.38) rectangle (114.27, 92.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.42, 89.38) -- (114.27, 92.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.42, 92.24) -- (114.27, 89.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.97, 89.38) rectangle (115.83, 92.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.97, 89.38) -- (115.83, 92.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.97, 92.24) -- (115.83, 89.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.35, 91.57) rectangle (115.21, 94.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.35, 91.57) -- (115.21, 94.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.35, 94.42) -- (115.21, 91.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.83, 90.31) rectangle (115.68, 93.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.83, 90.31) -- (115.68, 93.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.83, 93.16) -- (115.68, 90.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.72, 88.89) rectangle (115.58, 91.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.72, 88.89) -- (115.58, 91.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.72, 91.75) -- (115.58, 88.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.58, 90.31) rectangle (114.43, 93.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.58, 90.31) -- (114.43, 93.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.58, 93.16) -- (114.43, 90.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.58, 93.41) rectangle (114.43, 96.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.58, 93.41) -- (114.43, 96.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.58, 96.26) -- (114.43, 93.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.78, 93.41) rectangle (111.63, 96.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.78, 93.41) -- (111.63, 96.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.78, 96.26) -- (111.63, 93.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.36, 94.07) rectangle (111.21, 96.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.36, 94.07) -- (111.21, 96.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.36, 96.92) -- (111.21, 94.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (109.80, 93.41) rectangle (112.66, 96.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (109.80, 93.41) -- (112.66, 96.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (109.80, 96.26) -- (112.66, 93.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.31, 93.06) rectangle (111.16, 95.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.31, 93.06) -- (111.16, 95.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.31, 95.92) -- (111.16, 93.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (107.76, 93.06) rectangle (110.61, 95.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (107.76, 93.06) -- (110.61, 95.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (107.76, 95.92) -- (110.61, 93.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (109.03, 92.71) rectangle (111.88, 95.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (109.03, 92.71) -- (111.88, 95.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (109.03, 95.56) -- (111.88, 92.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.66, 92.71) rectangle (113.51, 95.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.66, 92.71) -- (113.51, 95.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.66, 95.56) -- (113.51, 92.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.78, 91.57) rectangle (111.63, 94.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.78, 91.57) -- (111.63, 94.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.78, 94.42) -- (111.63, 91.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.31, 90.31) rectangle (113.16, 93.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.31, 90.31) -- (113.16, 93.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.31, 93.16) -- (113.16, 90.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (109.99, 92.71) rectangle (112.84, 95.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (109.99, 92.71) -- (112.84, 95.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (109.99, 95.56) -- (112.84, 92.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.25, 93.41) rectangle (114.10, 96.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.25, 93.41) -- (114.10, 96.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.25, 96.26) -- (114.10, 93.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (109.57, 92.71) rectangle (112.42, 95.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (109.57, 92.71) -- (112.42, 95.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (109.57, 95.56) -- (112.42, 92.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.50,101.25) rectangle (114.35,104.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.50,101.25) -- (114.35,104.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.50,104.10) -- (114.35,101.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.22, 91.96) rectangle (113.07, 94.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.22, 91.96) -- (113.07, 94.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.22, 94.82) -- (113.07, 91.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.13, 93.06) rectangle (113.98, 95.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.13, 93.06) -- (113.98, 95.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.13, 95.92) -- (113.98, 93.06);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.59,108.71) -- (116.63,108.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.61,106.70) -- (114.61,110.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.66,109.50) -- (117.70,109.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.68,107.49) -- (115.68,111.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.37,111.11) -- (117.41,111.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.39,109.09) -- (115.39,113.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.24,109.60) -- (117.27,109.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.26,107.58) -- (115.26,111.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.47,110.51) -- (117.50,110.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.49,108.49) -- (115.49,112.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.27,109.02) -- (117.31,109.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.29,107.00) -- (115.29,111.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (107.34,108.08) -- (111.37,108.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (109.35,106.07) -- (109.35,110.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.78,107.87) -- (114.82,107.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.80,105.85) -- (112.80,109.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.18,109.69) -- (115.22,109.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.20,107.68) -- (113.20,111.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (109.71,107.76) -- (113.75,107.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.73,105.74) -- (111.73,109.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.97,108.51) -- (113.01,108.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.99,106.49) -- (110.99,110.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.52,111.68) -- (116.56,111.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.54,109.66) -- (114.54,113.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (111.57,110.24) -- (115.61,110.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.59,108.23) -- (113.59,112.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.11,115.08) -- (117.14,115.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.12,113.06) -- (115.12,117.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.09,112.14) -- (116.13,112.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.11,110.12) -- (114.11,114.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (113.21,109.69) -- (117.24,109.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.22,107.68) -- (115.22,111.71);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (104.04, 79.27) rectangle (106.89, 82.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (104.04, 79.27) -- (106.89, 82.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (104.04, 82.12) -- (106.89, 79.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.61, 83.95) rectangle (115.47, 86.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.61, 83.95) -- (115.47, 86.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.61, 86.81) -- (115.47, 83.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.15, 99.38) rectangle (123.01,102.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.15, 99.38) -- (123.01,102.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.15,102.23) -- (123.01, 99.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.41,118.30) rectangle (129.27,121.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.41,118.30) -- (129.27,121.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.41,121.15) -- (129.27,118.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.37,131.42) rectangle (134.22,134.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.37,131.42) -- (134.22,134.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.37,134.27) -- (134.22,131.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.10,145.40) rectangle (140.95,148.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.10,145.40) -- (140.95,148.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.10,148.25) -- (140.95,145.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.67,155.04) rectangle (141.53,157.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.67,155.04) -- (141.53,157.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.67,157.89) -- (141.53,155.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.49,167.44) rectangle (149.34,170.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.49,167.44) -- (149.34,170.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.49,170.30) -- (149.34,167.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.24,177.50) rectangle (154.09,180.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.24,177.50) -- (154.09,180.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.24,180.35) -- (154.09,177.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.79,185.01) rectangle (157.65,187.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.79,185.01) -- (157.65,187.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.79,187.86) -- (157.65,185.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.87,186.06) rectangle (161.72,188.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.87,186.06) -- (161.72,188.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.87,188.91) -- (161.72,186.06);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor,fill opacity=0.50] (188.62,153.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,136.00) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,138.50) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,158.63) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,161.09) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,153.48) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,157.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,140.33) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,139.63) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,148.24) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,154.05) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,170.95) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,155.41) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,159.97) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,155.03) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,161.93) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,195.14) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,195.14) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,195.14) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,179.40) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,150.04) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,121.24) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,139.89) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,118.82) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,140.88) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,132.65) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,155.03) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,122.97) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,133.50) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,129.33) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.77, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.84, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (105.99, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.07, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.20, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.05, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.89, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.57, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.01, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.64, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.35, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.35, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (112.04, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.70, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.70, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.44, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (112.64, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.62, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.38, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.23, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.64, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.25, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.50, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.57, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (114.33, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (112.60, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.69, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.35, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (133.11, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.62, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.14, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.20, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.20, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (122.79, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.19, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.02, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (114.58, 86.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (116.86, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.71, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.64, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (133.27, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (113.55, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.70, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.84, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (115.87, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.87, 95.50) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.46, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.57, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (136.01, 87.49) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (115.19, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.50, 80.69) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.51, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (125.43, 90.81) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.08, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.15, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.46, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.90, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.13, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.18, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.44, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (115.42, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (112.55, 91.74) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.14, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.64, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (115.12, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (112.26, 92.17) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.51, 80.69) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.71, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (113.78, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.37, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.25, 94.84) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.26, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.20, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.57, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.31, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.70, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (133.02, 94.49) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.28, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.10, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.68, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.82, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.98, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.57, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.83, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.35, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.24, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.43, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.98, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.05, 88.11) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.10, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.31, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.32, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.31, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.31, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.83, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.99, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (115.09, 88.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.13, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.90, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (126.78, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.32, 95.17) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.57, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.90, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.44, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.14, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.13, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (105.99, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.15, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.44, 95.81) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.35, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.38, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (113.93, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.60, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.38, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.91, 80.69) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (113.32,100.60) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.56, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.98, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.24, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (136.49, 94.49) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.39, 90.81) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (112.13, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.85, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (133.25, 91.74) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.30, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (115.58, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.57, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (122.25, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (112.84, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.83, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.30, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (152.51, 95.17) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (115.74, 86.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.30, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.95, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (114.51, 90.32) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.98, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (113.40, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.50, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (112.30, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.45, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.02, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.51, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (146.97, 92.59) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.57, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.35, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.63, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (112.96, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (112.04, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.67, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.35, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.99, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.85, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.64, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.18, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.91, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.95, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.98, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.98, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (153.11, 99.06) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.00, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.15, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.57, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (131.57, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (112.92, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.23, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.30, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (141.74, 97.55) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.35, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.11, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.64, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.09, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.26, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.67, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.95, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (148.70, 99.97) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (132.71, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (114.33, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (114.75, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (157.47,103.34) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (155.98,114.29) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (121.56, 91.28) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.65, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (168.80,118.78) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (132.30, 98.82) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (115.16, 89.27) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (117.11, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (138.49, 97.55) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (141.72,101.59) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (113.40, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.57, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.51, 86.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.91, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.05, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.14, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.31, 94.49) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.68, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (123.88, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.25, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.19, 80.69) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.55, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.71, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.02, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (113.12, 88.11) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.32, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.79, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.84, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (114.11, 90.81) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.99, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.98, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.50, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (134.42, 93.00) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.28, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.05, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.46, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.79, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.90, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.95, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (114.61,105.01) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.25, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.99, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.73, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (114.72, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.89, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.19, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.19, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.02, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.01, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.76, 86.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.44, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.56, 92.17) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.79, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.19, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.05, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (112.30, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.04, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.77, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.44, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (113.12, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.50, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.63, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.43, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (147.05, 88.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (112.00, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.73, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.63, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (158.00,103.98) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (112.60, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.70, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.50, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (113.32, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.11, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.43, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (115.09, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (161.89,101.96) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (133.79, 88.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (122.14, 87.49) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.31, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (141.84, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.99, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.25, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.64, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (130.25, 88.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.45, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.08, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.25, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (150.14, 94.84) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.79, 86.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.73, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.14, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (115.68, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.28, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.38, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.19, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (112.72, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.35, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.11, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.77, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.93, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.18, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.51, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.40, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (138.22, 97.55) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (171.13,116.88) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (130.84, 96.71) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.68, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (138.79, 98.07) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.79, 94.49) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (114.00, 90.32) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.85, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,125.45) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (129.52, 96.42) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.92, 96.42) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.57, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (170.90,110.42) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (141.24, 97.00) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.50, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.67, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (170.80,110.06) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (134.38,113.02) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (113.89,100.60) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.30, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (148.35, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (113.12, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.11, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.91, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (142.78,115.25) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.03, 99.97) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.90, 86.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.41, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (169.78,111.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (154.68,108.30) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.91, 86.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.20, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.09, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.95, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.90, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.50, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.40, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.38, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.84, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.44, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.45, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.24, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.84, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.20, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (119.33, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.25, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.18, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.56, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.70, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.44, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.57, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.44, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (142.93, 89.27) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.73,100.81) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.89, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.64, 90.81) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (115.84, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (110.25, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (105.92, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.44, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (165.82, 89.81) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.84, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.11, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.38, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (130.13, 86.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.51, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.52, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.26, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (120.14, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.60, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.11, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.57, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.56, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (109.63, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.83, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.64, 83.73) circle (  1.43);
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor,fill opacity=0.50] (110.50, 84.01) --
	(112.43, 80.69) --
	(108.58, 80.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.11, 87.60) --
	(116.03, 84.27) --
	(112.19, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (109.41, 84.01) --
	(111.33, 80.69) --
	(107.49, 80.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.93, 85.94) --
	(115.85, 82.62) --
	(112.01, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.18, 85.94) --
	(113.10, 82.62) --
	(109.26, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (108.01, 82.91) --
	(109.93, 79.58) --
	(106.09, 79.58) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (108.85, 82.91) --
	(110.77, 79.58) --
	(106.93, 79.58) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.00, 86.80) --
	(113.92, 83.47) --
	(110.08, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.05, 85.02) --
	(111.97, 81.69) --
	(108.13, 81.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.43, 87.60) --
	(114.35, 84.27) --
	(110.51, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.54, 89.04) --
	(118.46, 85.72) --
	(114.62, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.80, 85.02) --
	(117.73, 81.69) --
	(113.88, 81.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.54, 88.34) --
	(116.46, 85.02) --
	(112.62, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.94, 89.04) --
	(121.86, 85.72) --
	(118.02, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.47, 90.33) --
	(119.39, 87.00) --
	(115.55, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.47, 85.94) --
	(116.39, 82.62) --
	(112.55, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.36, 86.80) --
	(116.29, 83.47) --
	(112.44, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.36, 90.33) --
	(120.29, 87.00) --
	(116.44, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.80,101.28) --
	(118.72, 97.95) --
	(114.88, 97.95) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.45, 88.34) --
	(118.37, 85.02) --
	(114.53, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.67, 90.92) --
	(125.59, 87.59) --
	(121.75, 87.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (126.11, 97.05) --
	(128.03, 93.73) --
	(124.18, 93.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.40,103.23) --
	(126.32, 99.90) --
	(122.48, 99.90) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.22,110.93) --
	(125.14,107.61) --
	(121.30,107.61) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.91, 87.60) --
	(118.84, 84.27) --
	(114.99, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (127.59, 95.61) --
	(129.52, 92.28) --
	(125.67, 92.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.67, 94.39) --
	(135.59, 91.06) --
	(131.75, 91.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (125.56, 97.05) --
	(127.49, 93.73) --
	(123.64, 93.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.11, 90.33) --
	(120.03, 87.00) --
	(116.19, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.49, 95.22) --
	(126.41, 91.89) --
	(122.57, 91.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.65,100.55) --
	(141.57, 97.22) --
	(137.73, 97.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (125.14,100.55) --
	(127.06, 97.22) --
	(123.22, 97.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.19, 98.64) --
	(123.11, 95.31) --
	(119.27, 95.31) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.68,110.19) --
	(132.61,106.87) --
	(128.76,106.87) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.17, 89.04) --
	(123.09, 85.72) --
	(119.25, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.78, 99.22) --
	(139.70, 95.89) --
	(135.85, 95.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (126.36,101.28) --
	(128.28, 97.95) --
	(124.44, 97.95) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.13,115.72) --
	(131.05,112.39) --
	(127.21,112.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.81,119.04) --
	(136.73,115.72) --
	(132.88,115.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (128.03,103.23) --
	(129.95, 99.90) --
	(126.11, 99.90) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.37,104.00) --
	(147.29,100.67) --
	(143.45,100.67) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.67,122.25) --
	(146.59,118.93) --
	(142.75,118.93) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (158.39,122.33) --
	(160.32,119.00) --
	(156.47,119.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.00,112.37) --
	(132.92,109.05) --
	(129.08,109.05) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.87,139.55) --
	(140.79,136.22) --
	(136.95,136.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (143.87,122.41) --
	(145.79,119.08) --
	(141.95,119.08) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.77,109.05) --
	(131.69,105.72) --
	(127.85,105.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (125.63, 97.39) --
	(127.55, 94.06) --
	(123.71, 94.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.13,110.41) --
	(133.05,107.08) --
	(129.21,107.08) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.89,111.91) --
	(138.81,108.59) --
	(134.97,108.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (158.65,124.86) --
	(160.57,121.53) --
	(156.73,121.53) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (154.87,121.46) --
	(156.79,118.13) --
	(152.95,118.13) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (150.21,109.64) --
	(152.13,106.31) --
	(148.29,106.31) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.67,130.07) --
	(144.60,126.74) --
	(140.75,126.74) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (151.84,103.02) --
	(153.76, 99.70) --
	(149.92, 99.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (157.52,122.85) --
	(159.44,119.53) --
	(155.60,119.53) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.84,114.66) --
	(135.76,111.33) --
	(131.92,111.33) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (150.20,124.42) --
	(152.13,121.09) --
	(148.28,121.09) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.32,124.70) --
	(146.25,121.37) --
	(142.40,121.37) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.31,127.18) --
	(133.23,123.85) --
	(129.39,123.85) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (173.82,147.84) --
	(175.74,144.51) --
	(171.90,144.51) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (164.19,127.21) --
	(166.12,123.88) --
	(162.27,123.88) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (154.57,120.78) --
	(156.49,117.45) --
	(152.65,117.45) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (135.22,118.31) --
	(137.14,114.98) --
	(133.30,114.98) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (165.00,127.79) --
	(166.92,124.46) --
	(163.07,124.46) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (154.09,126.16) --
	(156.01,122.83) --
	(152.17,122.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (154.66,123.80) --
	(156.58,120.47) --
	(152.74,120.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (155.94,124.92) --
	(157.86,121.59) --
	(154.02,121.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (154.74,117.53) --
	(156.66,114.20) --
	(152.82,114.20) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (157.78,130.40) --
	(159.70,127.07) --
	(155.86,127.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (161.37,124.03) --
	(163.29,120.70) --
	(159.45,120.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (181.10,131.27) --
	(183.03,127.95) --
	(179.18,127.95) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (171.77,134.92) --
	(173.69,131.60) --
	(169.85,131.60) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (163.78,124.23) --
	(165.70,120.90) --
	(161.86,120.90) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (174.86,146.39) --
	(176.78,143.07) --
	(172.94,143.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (180.90,137.19) --
	(182.82,133.86) --
	(178.97,133.86) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (161.44,131.54) --
	(163.36,128.21) --
	(159.52,128.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (151.01,139.34) --
	(152.93,136.01) --
	(149.09,136.01) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (167.61,119.69) --
	(169.54,116.36) --
	(165.69,116.36) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.07,132.01) --
	(146.99,128.68) --
	(143.15,128.68) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,144.76) --
	(190.54,141.43) --
	(186.69,141.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (170.48,140.30) --
	(172.40,136.97) --
	(168.56,136.97) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,147.36) --
	(190.54,144.03) --
	(186.69,144.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (179.05,135.59) --
	(180.97,132.26) --
	(177.12,132.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (173.93,153.77) --
	(175.85,150.45) --
	(172.01,150.45) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,162.28) --
	(190.54,158.96) --
	(186.69,158.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (156.63,135.72) --
	(158.55,132.39) --
	(154.71,132.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,146.84) --
	(190.54,143.51) --
	(186.69,143.51) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (177.89,137.16) --
	(179.81,133.83) --
	(175.97,133.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (178.74,126.85) --
	(180.67,123.52) --
	(176.82,123.52) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.40, 85.02) --
	(112.33, 81.69) --
	(108.48, 81.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (108.14, 84.01) --
	(110.06, 80.69) --
	(106.21, 80.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.60, 87.60) --
	(113.52, 84.27) --
	(109.68, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 84.01) --
	(190.54, 80.69) --
	(186.69, 80.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.91, 90.33) --
	(113.83, 87.00) --
	(109.99, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (108.20, 84.01) --
	(110.12, 80.69) --
	(106.28, 80.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.85, 87.60) --
	(112.77, 84.27) --
	(108.93, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.60, 86.80) --
	(112.53, 83.47) --
	(108.68, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (108.56, 85.02) --
	(110.48, 81.69) --
	(106.63, 81.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.73, 85.02) --
	(113.65, 81.69) --
	(109.81, 81.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.80, 88.34) --
	(114.72, 85.02) --
	(110.88, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (108.50, 85.94) --
	(110.42, 82.62) --
	(106.58, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.55, 88.34) --
	(114.47, 85.02) --
	(110.63, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.28, 90.92) --
	(113.20, 87.59) --
	(109.35, 87.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.75, 88.34) --
	(112.67, 85.02) --
	(108.83, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (109.73, 85.02) --
	(111.65, 81.69) --
	(107.81, 81.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.65, 87.60) --
	(112.57, 84.27) --
	(108.73, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (108.79, 85.94) --
	(110.71, 82.62) --
	(106.87, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.26, 87.60) --
	(114.18, 84.27) --
	(110.34, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.32, 88.34) --
	(113.24, 85.02) --
	(109.40, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (109.68, 86.80) --
	(111.60, 83.47) --
	(107.76, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.23, 88.34) --
	(113.15, 85.02) --
	(109.31, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.55, 86.80) --
	(113.47, 83.47) --
	(109.63, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.62, 89.70) --
	(118.55, 86.38) --
	(114.70, 86.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.27, 90.33) --
	(118.19, 87.00) --
	(114.35, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.55, 90.92) --
	(115.47, 87.59) --
	(111.63, 87.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.84, 93.50) --
	(114.76, 90.17) --
	(110.92, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.70,101.28) --
	(115.63, 97.95) --
	(111.78, 97.95) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.18, 92.02) --
	(118.10, 88.70) --
	(114.26, 88.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.62, 95.99) --
	(118.55, 92.66) --
	(114.70, 92.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.02, 91.49) --
	(116.94, 88.16) --
	(113.10, 88.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.11, 88.34) --
	(116.03, 85.02) --
	(112.19, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.34, 89.04) --
	(119.26, 85.72) --
	(115.41, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.09, 86.80) --
	(117.01, 83.47) --
	(113.17, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.42, 87.60) --
	(118.34, 84.27) --
	(114.50, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 91.49) --
	(190.54, 88.16) --
	(186.69, 88.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.39, 89.04) --
	(117.31, 85.72) --
	(113.47, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.12, 93.50) --
	(117.05, 90.17) --
	(113.20, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.03, 90.92) --
	(121.95, 87.59) --
	(118.11, 87.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.96, 90.92) --
	(116.88, 87.59) --
	(113.03, 87.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.75, 89.70) --
	(116.67, 86.38) --
	(112.83, 86.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.31, 90.33) --
	(119.23, 87.00) --
	(115.39, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.99, 92.02) --
	(121.91, 88.70) --
	(118.06, 88.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.16, 97.71) --
	(120.08, 94.39) --
	(116.24, 94.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.15, 92.54) --
	(125.07, 89.21) --
	(121.23, 89.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.42, 89.04) --
	(118.34, 85.72) --
	(114.50, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.98, 93.03) --
	(120.90, 89.70) --
	(117.06, 89.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.22, 92.54) --
	(117.14, 89.21) --
	(113.30, 89.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.87, 97.05) --
	(119.80, 93.73) --
	(115.95, 93.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.67, 92.54) --
	(124.60, 89.21) --
	(120.75, 89.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.23, 92.54) --
	(119.15, 89.21) --
	(115.30, 89.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.61, 93.50) --
	(119.53, 90.17) --
	(115.69, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.87, 95.61) --
	(119.80, 92.28) --
	(115.95, 92.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (128.04, 95.22) --
	(129.96, 91.89) --
	(126.12, 91.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.72, 90.33) --
	(116.64, 87.00) --
	(112.79, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.54, 89.70) --
	(116.46, 86.38) --
	(112.62, 86.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.14, 93.50) --
	(121.07, 90.17) --
	(117.22, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.49, 94.39) --
	(123.41, 91.06) --
	(119.56, 91.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.91, 93.50) --
	(125.83, 90.17) --
	(121.99, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.75, 98.93) --
	(123.67, 95.60) --
	(119.83, 95.60) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (125.04,100.03) --
	(126.96, 96.71) --
	(123.12, 96.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.15, 95.61) --
	(123.07, 92.28) --
	(119.23, 92.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.00,100.80) --
	(125.92, 97.47) --
	(122.08, 97.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (132.88, 98.03) --
	(134.81, 94.70) --
	(130.96, 94.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.65, 93.50) --
	(118.58, 90.17) --
	(114.73, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.87, 93.50) --
	(119.80, 90.17) --
	(115.95, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.37,108.17) --
	(123.29,104.85) --
	(119.45,104.85) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.31, 95.22) --
	(120.24, 91.89) --
	(116.39, 91.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (126.59,103.42) --
	(128.51,100.10) --
	(124.67,100.10) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.58, 94.39) --
	(121.51, 91.06) --
	(117.66, 91.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.16,114.59) --
	(133.08,111.26) --
	(129.24,111.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (127.66,108.17) --
	(129.58,104.85) --
	(125.74,104.85) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.17,102.18) --
	(144.10, 98.86) --
	(140.25, 98.86) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.86,106.50) --
	(123.79,103.17) --
	(119.94,103.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.05,100.55) --
	(122.97, 97.22) --
	(119.13, 97.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.07, 95.22) --
	(121.00, 91.89) --
	(117.15, 91.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (125.65,100.80) --
	(127.57, 97.47) --
	(123.72, 97.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.19,110.52) --
	(131.11,107.19) --
	(127.27,107.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.91,109.17) --
	(133.83,105.84) --
	(129.99,105.84) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.29, 95.22) --
	(126.21, 91.89) --
	(122.37, 91.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.92,124.73) --
	(146.84,121.41) --
	(143.00,121.41) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (127.18,122.82) --
	(129.10,119.49) --
	(125.26,119.49) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.17,124.45) --
	(143.10,121.12) --
	(139.25,121.12) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.84, 96.71) --
	(126.76, 93.38) --
	(122.92, 93.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.08, 98.03) --
	(146.01, 94.70) --
	(142.16, 94.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (127.19,104.72) --
	(129.11,101.39) --
	(125.27,101.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.53,107.51) --
	(126.45,104.18) --
	(122.61,104.18) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (135.58,102.40) --
	(137.50, 99.07) --
	(133.66, 99.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.82,103.81) --
	(131.74,100.48) --
	(127.90,100.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (128.75,110.30) --
	(130.67,106.97) --
	(126.83,106.97) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (132.30,110.30) --
	(134.22,106.97) --
	(130.38,106.97) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.62,101.74) --
	(136.54, 98.41) --
	(132.70, 98.41) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.29,127.79) --
	(132.22,124.46) --
	(128.37,124.46) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.31,110.83) --
	(136.23,107.50) --
	(132.39,107.50) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.13,108.04) --
	(126.05,104.72) --
	(122.20,104.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.19,108.04) --
	(140.11,104.72) --
	(136.27,104.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.40,131.40) --
	(142.32,128.07) --
	(138.48,128.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.30,125.48) --
	(140.23,122.15) --
	(136.38,122.15) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.90,117.06) --
	(140.82,113.73) --
	(136.98,113.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,126.46) --
	(190.54,123.13) --
	(186.69,123.13) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (151.73,122.25) --
	(153.65,118.93) --
	(149.81,118.93) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (135.50,122.10) --
	(137.42,118.77) --
	(133.58,118.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (153.60,133.11) --
	(155.52,129.78) --
	(151.68,129.78) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.11,130.03) --
	(140.04,126.70) --
	(136.19,126.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (151.93,142.14) --
	(153.85,138.81) --
	(150.01,138.81) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (151.48,121.42) --
	(153.40,118.09) --
	(149.56,118.09) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (162.23,134.31) --
	(164.15,130.98) --
	(160.31,130.98) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (147.65,123.28) --
	(149.57,119.95) --
	(145.73,119.95) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.12,118.63) --
	(142.04,115.30) --
	(138.20,115.30) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.47,118.42) --
	(136.39,115.09) --
	(132.55,115.09) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.74,121.66) --
	(132.67,118.34) --
	(128.82,118.34) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.92,125.45) --
	(139.84,122.13) --
	(136.00,122.13) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (141.34,114.44) --
	(143.26,111.11) --
	(139.42,111.11) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (152.15,137.77) --
	(154.07,134.45) --
	(150.23,134.45) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (173.30,124.45) --
	(175.22,121.12) --
	(171.38,121.12) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (156.98,132.82) --
	(158.90,129.49) --
	(155.06,129.49) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.89,120.65) --
	(138.81,117.32) --
	(134.97,117.32) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (164.71,145.01) --
	(166.63,141.68) --
	(162.79,141.68) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (154.66,129.07) --
	(156.58,125.74) --
	(152.74,125.74) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.85,113.33) --
	(135.77,110.00) --
	(131.93,110.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,143.05) --
	(190.54,139.72) --
	(186.69,139.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (173.20,159.10) --
	(175.12,155.77) --
	(171.28,155.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,115.02) --
	(190.54,111.70) --
	(186.69,111.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (154.18,132.20) --
	(156.10,128.88) --
	(152.26,128.88) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.82,126.54) --
	(144.74,123.21) --
	(140.89,123.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,156.09) --
	(190.54,152.76) --
	(186.69,152.76) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.92,157.29) --
	(146.85,153.96) --
	(143.00,153.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (158.27,120.52) --
	(160.19,117.19) --
	(156.35,117.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (132.61,129.55) --
	(134.53,126.22) --
	(130.69,126.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (152.79,116.75) --
	(154.71,113.43) --
	(150.87,113.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (159.82,122.06) --
	(161.74,118.73) --
	(157.90,118.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,154.01) --
	(190.54,150.68) --
	(186.69,150.68) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (147.41,128.48) --
	(149.33,125.16) --
	(145.49,125.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (178.88,138.18) --
	(180.80,134.85) --
	(176.96,134.85) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.84,127.60) --
	(136.76,124.27) --
	(132.91,124.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,158.89) --
	(190.54,155.56) --
	(186.69,155.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,168.92) --
	(190.54,165.60) --
	(186.69,165.60) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (169.70,144.80) --
	(171.62,141.47) --
	(167.78,141.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,168.96) --
	(190.54,165.63) --
	(186.69,165.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.71,137.56) --
	(141.63,134.23) --
	(137.79,134.23) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (149.82,135.72) --
	(151.74,132.39) --
	(147.90,132.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (151.31,151.91) --
	(153.23,148.58) --
	(149.39,148.58) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (159.90,157.32) --
	(161.83,153.99) --
	(157.98,153.99) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (177.37,154.24) --
	(179.29,150.91) --
	(175.45,150.91) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (159.13,135.52) --
	(161.06,132.20) --
	(157.21,132.20) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (173.04,152.77) --
	(174.96,149.44) --
	(171.12,149.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (176.98,165.16) --
	(178.90,161.83) --
	(175.06,161.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (153.25,125.54) --
	(155.17,122.21) --
	(151.33,122.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (179.35,139.83) --
	(181.27,136.50) --
	(177.43,136.50) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,149.10) --
	(190.54,145.77) --
	(186.69,145.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (109.52, 86.80) --
	(111.44, 83.47) --
	(107.60, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (108.79, 85.94) --
	(110.71, 82.62) --
	(106.87, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 85.94) --
	(190.54, 82.62) --
	(186.69, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.25, 93.03) --
	(112.17, 89.70) --
	(108.33, 89.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (109.02, 85.94) --
	(110.94, 82.62) --
	(107.10, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (108.07, 85.94) --
	(110.00, 82.62) --
	(106.15, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 85.94) --
	(190.54, 82.62) --
	(186.69, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.78, 85.02) --
	(142.70, 81.69) --
	(138.86, 81.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 85.94) --
	(190.54, 82.62) --
	(186.69, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (108.73, 85.94) --
	(110.65, 82.62) --
	(106.81, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.05, 87.60) --
	(111.97, 84.27) --
	(108.13, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (109.52, 86.80) --
	(111.44, 83.47) --
	(107.60, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (109.46, 86.80) --
	(111.38, 83.47) --
	(107.54, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.64, 94.81) --
	(113.56, 91.48) --
	(109.72, 91.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (109.08, 99.22) --
	(111.00, 95.89) --
	(107.15, 95.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.60, 86.80) --
	(112.53, 83.47) --
	(108.68, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.80, 86.80) --
	(112.72, 83.47) --
	(108.88, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (109.41, 86.80) --
	(111.33, 83.47) --
	(107.49, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.64, 87.60) --
	(113.56, 84.27) --
	(109.72, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 87.60) --
	(190.54, 84.27) --
	(186.69, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.46, 89.04) --
	(113.38, 85.72) --
	(109.54, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.70, 93.50) --
	(112.62, 90.17) --
	(108.78, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.09, 88.34) --
	(113.01, 85.02) --
	(109.17, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.75, 89.04) --
	(112.67, 85.72) --
	(108.83, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 87.60) --
	(190.54, 84.27) --
	(186.69, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.69, 90.92) --
	(113.61, 87.59) --
	(109.77, 87.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.32, 88.34) --
	(113.24, 85.02) --
	(109.40, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.13, 89.04) --
	(113.06, 85.72) --
	(109.21, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.34, 89.04) --
	(114.26, 85.72) --
	(110.42, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.51, 87.60) --
	(113.43, 84.27) --
	(109.58, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.60, 88.34) --
	(114.52, 85.02) --
	(110.67, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.00, 92.02) --
	(114.92, 88.70) --
	(111.08, 88.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.72, 89.70) --
	(114.64, 86.38) --
	(110.80, 86.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.44, 89.70) --
	(115.36, 86.38) --
	(111.52, 86.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.64, 89.04) --
	(113.56, 85.72) --
	(109.72, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.09, 89.04) --
	(114.01, 85.72) --
	(110.16, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.17, 90.33) --
	(114.09, 87.00) --
	(110.25, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.04, 92.02) --
	(114.96, 88.70) --
	(111.12, 88.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.21, 92.02) --
	(114.14, 88.70) --
	(110.29, 88.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 88.34) --
	(190.54, 85.02) --
	(186.69, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.65, 92.54) --
	(117.57, 89.21) --
	(113.72, 89.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.11, 89.70) --
	(116.03, 86.38) --
	(112.19, 86.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.60, 90.33) --
	(114.52, 87.00) --
	(110.67, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.84, 94.81) --
	(114.76, 91.48) --
	(110.92, 91.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.89, 89.04) --
	(115.81, 85.72) --
	(111.97, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.72, 90.33) --
	(114.64, 87.00) --
	(110.80, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.78, 89.04) --
	(113.70, 85.72) --
	(109.86, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.28,100.80) --
	(115.20, 97.47) --
	(111.36, 97.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 89.04) --
	(190.54, 85.72) --
	(186.69, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.15, 89.04) --
	(116.07, 85.72) --
	(112.23, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.58, 98.93) --
	(117.50, 95.60) --
	(113.66, 95.60) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.51, 90.33) --
	(115.44, 87.00) --
	(111.59, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 90.33) --
	(190.54, 87.00) --
	(186.69, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.97, 94.39) --
	(115.89, 91.06) --
	(112.04, 91.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.04, 94.39) --
	(115.96, 91.06) --
	(112.12, 91.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.21, 90.33) --
	(114.14, 87.00) --
	(110.29, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.70, 90.92) --
	(115.63, 87.59) --
	(111.78, 87.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 91.49) --
	(190.54, 88.16) --
	(186.69, 88.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.32, 90.33) --
	(115.24, 87.00) --
	(111.40, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.40, 91.49) --
	(116.32, 88.16) --
	(112.48, 88.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.99, 91.49) --
	(117.91, 88.16) --
	(114.07, 88.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.09, 92.02) --
	(117.01, 88.70) --
	(113.17, 88.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.92, 90.92) --
	(116.84, 87.59) --
	(113.00, 87.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.40,104.89) --
	(116.32,101.57) --
	(112.48,101.57) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.82, 93.50) --
	(113.74, 90.17) --
	(109.90, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.82, 92.02) --
	(115.74, 88.70) --
	(111.90, 88.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.75, 91.49) --
	(116.67, 88.16) --
	(112.83, 88.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.75, 92.02) --
	(116.67, 88.70) --
	(112.83, 88.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.44, 91.49) --
	(115.36, 88.16) --
	(111.52, 88.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.65, 91.49) --
	(116.57, 88.16) --
	(112.73, 88.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.74, 93.03) --
	(118.66, 89.70) --
	(114.82, 89.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.12, 93.03) --
	(118.04, 89.70) --
	(114.19, 89.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.40, 93.50) --
	(116.32, 90.17) --
	(112.48, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 91.49) --
	(190.54, 88.16) --
	(186.69, 88.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.45, 92.02) --
	(117.37, 88.70) --
	(113.53, 88.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 93.03) --
	(190.54, 89.70) --
	(186.69, 89.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.93, 92.54) --
	(115.85, 89.21) --
	(112.01, 89.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.29, 92.54) --
	(117.21, 89.21) --
	(113.37, 89.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.89, 93.50) --
	(116.81, 90.17) --
	(112.97, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 93.95) --
	(190.54, 90.63) --
	(186.69, 90.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.32, 93.03) --
	(117.24, 89.70) --
	(113.40, 89.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.87,100.55) --
	(117.79, 97.22) --
	(113.95, 97.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.99,104.18) --
	(117.91,100.85) --
	(114.07,100.85) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.54, 93.03) --
	(118.46, 89.70) --
	(114.62, 89.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.05, 93.95) --
	(117.98, 90.63) --
	(114.13, 90.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.77, 94.39) --
	(118.69, 91.06) --
	(114.85, 91.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.55, 93.03) --
	(117.47, 89.70) --
	(113.63, 89.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.08, 93.03) --
	(119.01, 89.70) --
	(115.16, 89.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.45, 93.95) --
	(118.37, 90.63) --
	(114.53, 90.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.02, 93.95) --
	(117.94, 90.63) --
	(114.10, 90.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.42, 94.81) --
	(118.34, 91.48) --
	(114.50, 91.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.22, 95.22) --
	(121.14, 91.89) --
	(117.29, 91.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.42, 94.81) --
	(119.34, 91.48) --
	(115.50, 91.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.05, 94.39) --
	(117.98, 91.06) --
	(114.13, 91.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 93.50) --
	(190.54, 90.17) --
	(186.69, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.89,102.61) --
	(118.81, 99.28) --
	(114.96, 99.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 93.50) --
	(190.54, 90.17) --
	(186.69, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.60, 94.81) --
	(118.52, 91.48) --
	(114.67, 91.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 93.03) --
	(190.54, 89.70) --
	(186.69, 89.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.24, 93.95) --
	(120.16, 90.63) --
	(116.32, 90.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.55, 95.61) --
	(119.48, 92.28) --
	(115.63, 92.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.03, 94.81) --
	(119.95, 91.48) --
	(116.11, 91.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.42, 95.61) --
	(118.34, 92.28) --
	(114.50, 92.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.55, 95.99) --
	(119.48, 92.66) --
	(115.63, 92.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.85, 94.39) --
	(119.77, 91.06) --
	(115.93, 91.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.85, 96.35) --
	(121.78, 93.03) --
	(117.93, 93.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.45, 93.95) --
	(118.37, 90.63) --
	(114.53, 90.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.89, 97.39) --
	(118.81, 94.06) --
	(114.96, 94.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 94.81) --
	(190.54, 91.48) --
	(186.69, 91.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.22, 94.81) --
	(121.14, 91.48) --
	(117.29, 91.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.11, 96.71) --
	(119.03, 93.38) --
	(115.19, 93.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.82, 97.05) --
	(119.74, 93.73) --
	(115.90, 93.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.53,106.04) --
	(126.45,102.71) --
	(122.61,102.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.19, 95.22) --
	(120.11, 91.89) --
	(116.27, 91.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.77, 95.61) --
	(117.69, 92.28) --
	(113.85, 92.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.17, 96.71) --
	(119.09, 93.38) --
	(115.25, 93.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.36, 98.34) --
	(119.28, 95.01) --
	(115.44, 95.01) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.69, 95.99) --
	(120.61, 92.66) --
	(116.77, 92.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.23, 95.61) --
	(119.15, 92.28) --
	(115.30, 92.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,100.55) --
	(190.54, 97.22) --
	(186.69, 97.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 95.61) --
	(190.54, 92.28) --
	(186.69, 92.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.20, 96.71) --
	(119.12, 93.38) --
	(115.28, 93.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.24, 97.05) --
	(120.16, 93.73) --
	(116.32, 93.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.56, 97.39) --
	(121.48, 94.06) --
	(117.64, 94.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.39,100.03) --
	(120.31, 96.71) --
	(116.47, 96.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.06, 97.71) --
	(119.98, 94.39) --
	(116.14, 94.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 99.22) --
	(190.54, 95.89) --
	(186.69, 95.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.34, 95.61) --
	(120.26, 92.28) --
	(116.42, 92.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.64, 97.71) --
	(120.56, 94.39) --
	(116.72, 94.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.17, 97.05) --
	(121.09, 93.73) --
	(117.25, 93.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.83, 97.71) --
	(125.75, 94.39) --
	(121.91, 94.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.44, 99.22) --
	(120.36, 95.89) --
	(116.52, 95.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.98, 97.05) --
	(120.90, 93.73) --
	(117.06, 93.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.81,100.55) --
	(120.73, 97.22) --
	(116.89, 97.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.61, 99.22) --
	(122.53, 95.89) --
	(118.68, 95.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.03,104.18) --
	(121.95,100.85) --
	(118.11,100.85) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.77, 97.71) --
	(122.69, 94.39) --
	(118.85, 94.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.46,117.76) --
	(125.38,114.43) --
	(121.54,114.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.31,102.18) --
	(126.23, 98.86) --
	(122.39, 98.86) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.23,131.31) --
	(126.15,127.98) --
	(122.31,127.98) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 99.50) --
	(190.54, 96.17) --
	(186.69, 96.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.07,109.17) --
	(124.99,105.84) --
	(121.15,105.84) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 99.50) --
	(190.54, 96.17) --
	(186.69, 96.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.47,102.18) --
	(133.39, 98.86) --
	(129.55, 98.86) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.40,103.81) --
	(125.32,100.48) --
	(121.48,100.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.40,101.51) --
	(140.32, 98.18) --
	(136.48, 98.18) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.99,108.17) --
	(132.91,104.85) --
	(129.07,104.85) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62, 99.50) --
	(190.54, 96.17) --
	(186.69, 96.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (126.52,107.64) --
	(128.45,104.31) --
	(124.60,104.31) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,100.03) --
	(190.54, 96.71) --
	(186.69, 96.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (127.53,116.44) --
	(129.45,113.11) --
	(125.60,113.11) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.06,112.55) --
	(134.98,109.22) --
	(131.13,109.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.51,107.37) --
	(125.43,104.04) --
	(121.59,104.04) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.25,123.46) --
	(126.17,120.13) --
	(122.33,120.13) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.14,105.56) --
	(122.06,102.24) --
	(118.22,102.24) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.88,139.26) --
	(123.80,135.93) --
	(119.96,135.93) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (127.13,109.29) --
	(129.05,105.96) --
	(125.21,105.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.16,106.04) --
	(126.08,102.71) --
	(122.23,102.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (159.20,137.64) --
	(161.12,134.32) --
	(157.27,134.32) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (125.83,132.56) --
	(127.75,129.23) --
	(123.91,129.23) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.20,143.14) --
	(147.12,139.82) --
	(143.27,139.82) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (147.22,112.10) --
	(149.14,108.77) --
	(145.30,108.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (128.69,122.78) --
	(130.61,119.45) --
	(126.77,119.45) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.48,110.30) --
	(131.40,106.97) --
	(127.55,106.97) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.90,108.81) --
	(132.82,105.48) --
	(128.98,105.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.77,119.15) --
	(131.69,115.82) --
	(127.85,115.82) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.77,132.22) --
	(140.69,128.89) --
	(136.85,128.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (150.26,150.73) --
	(152.18,147.40) --
	(148.34,147.40) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,118.53) --
	(190.54,115.20) --
	(186.69,115.20) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.50,114.88) --
	(146.42,111.55) --
	(142.58,111.55) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (166.35,147.34) --
	(168.27,144.02) --
	(164.43,144.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (159.30,160.17) --
	(161.22,156.84) --
	(157.38,156.84) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.27,124.76) --
	(144.20,121.44) --
	(140.35,121.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (161.07,153.26) --
	(163.00,149.93) --
	(159.15,149.93) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (135.21,131.82) --
	(137.13,128.50) --
	(133.29,128.50) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (150.78,141.19) --
	(152.70,137.87) --
	(148.85,137.87) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (177.95,182.15) --
	(179.87,178.82) --
	(176.03,178.82) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,141.73) --
	(190.54,138.40) --
	(186.69,138.40) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (165.19,143.67) --
	(167.12,140.35) --
	(163.27,140.35) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.26,117.76) --
	(138.18,114.43) --
	(134.33,114.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (166.17,166.04) --
	(168.09,162.71) --
	(164.25,162.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,180.65) --
	(190.54,177.32) --
	(186.69,177.32) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (153.58,127.55) --
	(155.50,124.22) --
	(151.66,124.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (157.38,132.63) --
	(159.30,129.31) --
	(155.46,129.31) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (167.58,197.36) --
	(169.50,194.03) --
	(165.66,194.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (170.23,149.98) --
	(172.15,146.65) --
	(168.30,146.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (165.54,145.91) --
	(167.46,142.58) --
	(163.62,142.58) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (179.79,144.41) --
	(181.71,141.08) --
	(177.86,141.08) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,197.36) --
	(190.54,194.03) --
	(186.69,194.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,169.55) --
	(190.54,166.22) --
	(186.69,166.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (154.03,176.67) --
	(155.95,173.34) --
	(152.11,173.34) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,138.70) --
	(190.54,135.37) --
	(186.69,135.37) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,141.47) --
	(190.54,138.14) --
	(186.69,138.14) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (170.16,175.25) --
	(172.08,171.92) --
	(168.23,171.92) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (170.61,163.06) --
	(172.53,159.74) --
	(168.69,159.74) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,197.36) --
	(190.54,194.03) --
	(186.69,194.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,181.08) --
	(190.54,177.75) --
	(186.69,177.75) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,197.36) --
	(190.54,194.03) --
	(186.69,194.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,171.27) --
	(190.54,167.94) --
	(186.69,167.94) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (166.45,197.36) --
	(168.37,194.03) --
	(164.53,194.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,197.36) --
	(190.54,194.03) --
	(186.69,194.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,188.01) --
	(190.54,184.69) --
	(186.69,184.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,197.36) --
	(190.54,194.03) --
	(186.69,194.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,178.99) --
	(190.54,175.67) --
	(186.69,175.67) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (167.78,197.36) --
	(169.70,194.03) --
	(165.86,194.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,183.54) --
	(190.54,180.21) --
	(186.69,180.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (171.42,197.36) --
	(173.34,194.03) --
	(169.50,194.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,158.67) --
	(190.54,155.34) --
	(186.69,155.34) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (167.77,197.36) --
	(169.69,194.03) --
	(165.85,194.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,197.36) --
	(190.54,194.03) --
	(186.69,194.03) --
	cycle;

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.77, 94.39) rectangle (120.62, 97.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.77, 94.39) -- (120.62, 97.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.77, 97.24) -- (120.62, 94.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.12, 95.57) rectangle (122.97, 98.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.12, 95.57) -- (122.97, 98.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.12, 98.43) -- (122.97, 95.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,150.84) rectangle (190.04,153.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,150.84) -- (190.04,153.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,153.70) -- (190.04,150.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.85,119.10) rectangle (126.70,121.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.85,119.10) -- (126.70,121.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (123.85,121.95) -- (126.70,119.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.45,161.49) rectangle (134.30,164.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.45,161.49) -- (134.30,164.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.45,164.34) -- (134.30,161.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,115.20) rectangle (190.04,118.05);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,115.20) -- (190.04,118.05);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,118.05) -- (190.04,115.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,122.49) rectangle (190.04,125.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,122.49) -- (190.04,125.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,125.34) -- (190.04,122.49);
\definecolor{drawColor}{RGB}{152,152,152}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.51, 47.03) -- (195.67,202.19);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.66, 47.03) -- ( 41.66, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 43.44, 47.03) -- ( 43.44, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 44.98, 47.03) -- ( 44.98, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 46.35, 47.03) -- ( 46.35, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 47.56, 47.03) -- ( 47.56, 55.57);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 55.57, 47.03) -- ( 55.57, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 60.26, 47.03) -- ( 60.26, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 63.58, 47.03) -- ( 63.58, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 66.16, 47.03) -- ( 66.16, 52.72);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 68.27, 47.03) -- ( 68.27, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 70.05, 47.03) -- ( 70.05, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 71.59, 47.03) -- ( 71.59, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 72.95, 47.03) -- ( 72.95, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 74.17, 47.03) -- ( 74.17, 55.57);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 82.18, 47.03) -- ( 82.18, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 86.87, 47.03) -- ( 86.87, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 90.19, 47.03) -- ( 90.19, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 92.77, 47.03) -- ( 92.77, 52.72);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 94.88, 47.03) -- ( 94.88, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 96.66, 47.03) -- ( 96.66, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 98.20, 47.03) -- ( 98.20, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 99.56, 47.03) -- ( 99.56, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (100.78, 47.03) -- (100.78, 55.57);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (108.79, 47.03) -- (108.79, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (113.48, 47.03) -- (113.48, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (116.80, 47.03) -- (116.80, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (119.38, 47.03) -- (119.38, 52.72);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (121.49, 47.03) -- (121.49, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (123.27, 47.03) -- (123.27, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (124.81, 47.03) -- (124.81, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (126.17, 47.03) -- (126.17, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (127.39, 47.03) -- (127.39, 55.57);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (135.40, 47.03) -- (135.40, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (140.08, 47.03) -- (140.08, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (143.41, 47.03) -- (143.41, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (145.99, 47.03) -- (145.99, 52.72);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (148.09, 47.03) -- (148.09, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (149.87, 47.03) -- (149.87, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (151.42, 47.03) -- (151.42, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (152.78, 47.03) -- (152.78, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (154.00, 47.03) -- (154.00, 55.57);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (162.01, 47.03) -- (162.01, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (166.69, 47.03) -- (166.69, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (170.02, 47.03) -- (170.02, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (172.60, 47.03) -- (172.60, 52.72);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (174.70, 47.03) -- (174.70, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (176.48, 47.03) -- (176.48, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (178.03, 47.03) -- (178.03, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (179.39, 47.03) -- (179.39, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (180.61, 47.03) -- (180.61, 55.57);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (188.62, 47.03) -- (188.62, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (193.30, 47.03) -- (193.30, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 48.18) -- ( 43.36, 48.18);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 49.96) -- ( 43.36, 49.96);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 51.51) -- ( 43.36, 51.51);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 52.87) -- ( 43.36, 52.87);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 54.09) -- ( 49.05, 54.09);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 62.10) -- ( 43.36, 62.10);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 66.78) -- ( 43.36, 66.78);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 70.11) -- ( 43.36, 70.11);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 72.68) -- ( 46.20, 72.68);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 74.79) -- ( 43.36, 74.79);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 76.57) -- ( 43.36, 76.57);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 78.12) -- ( 43.36, 78.12);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 79.48) -- ( 43.36, 79.48);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 80.69) -- ( 49.05, 80.69);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 88.70) -- ( 43.36, 88.70);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 93.39) -- ( 43.36, 93.39);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 96.71) -- ( 43.36, 96.71);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 99.29) -- ( 46.20, 99.29);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,101.40) -- ( 43.36,101.40);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,103.18) -- ( 43.36,103.18);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,104.72) -- ( 43.36,104.72);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,106.08) -- ( 43.36,106.08);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,107.30) -- ( 49.05,107.30);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,115.31) -- ( 43.36,115.31);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,120.00) -- ( 43.36,120.00);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,123.32) -- ( 43.36,123.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,125.90) -- ( 46.20,125.90);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,128.01) -- ( 43.36,128.01);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,129.79) -- ( 43.36,129.79);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,131.33) -- ( 43.36,131.33);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,132.69) -- ( 43.36,132.69);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,133.91) -- ( 49.05,133.91);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,141.92) -- ( 43.36,141.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,146.61) -- ( 43.36,146.61);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,149.93) -- ( 43.36,149.93);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,152.51) -- ( 46.20,152.51);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,154.62) -- ( 43.36,154.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,156.40) -- ( 43.36,156.40);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,157.94) -- ( 43.36,157.94);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,159.30) -- ( 43.36,159.30);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,160.52) -- ( 49.05,160.52);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,168.53) -- ( 43.36,168.53);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,173.21) -- ( 43.36,173.21);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,176.54) -- ( 43.36,176.54);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,179.12) -- ( 46.20,179.12);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,181.22) -- ( 43.36,181.22);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,183.01) -- ( 43.36,183.01);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,184.55) -- ( 43.36,184.55);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,185.91) -- ( 43.36,185.91);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,187.13) -- ( 49.05,187.13);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,195.14) -- ( 43.36,195.14);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,199.82) -- ( 43.36,199.82);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 47.03) rectangle (195.67,202.19);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.56, 77.66) {0.1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.56,130.88) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.56,184.10) {1000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 37.76, 80.69) --
	( 40.51, 80.69);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 37.76,133.91) --
	( 40.51,133.91);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 37.76,187.13) --
	( 40.51,187.13);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 74.17, 44.28) --
	( 74.17, 47.03);

\path[draw=drawColor,line width= 0.3pt,line join=round] (127.39, 44.28) --
	(127.39, 47.03);

\path[draw=drawColor,line width= 0.3pt,line join=round] (180.61, 44.28) --
	(180.61, 47.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 74.17, 36.02) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (127.39, 36.02) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (180.61, 36.02) {1000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (118.09, 23.99) {\textsf{Ace} + \texttt{cd06} time (s)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,124.61) {\textsf{DPMC} + \texttt{bklm16++} time (s)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (206.67, 68.14) rectangle (298.03,181.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (212.17,145.91) rectangle (226.62,160.37);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor,fill opacity=0.50] (219.39,153.14) circle (  1.43);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (212.17,131.46) rectangle (226.62,145.91);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor,fill opacity=0.50] (219.39,140.90) --
	(221.32,137.58) --
	(217.47,137.58) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (212.17,117.00) rectangle (226.62,131.46);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor,fill opacity=0.50] (217.97,122.80) --
	(220.82,122.80) --
	(220.82,125.66) --
	(217.97,125.66) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (212.17,102.55) rectangle (226.62,117.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (217.38,109.78) -- (221.41,109.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (219.39,107.76) -- (219.39,111.79);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (212.17, 88.10) rectangle (226.62,102.55);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (217.97, 93.90) rectangle (220.82, 96.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (217.97, 93.90) -- (220.82, 96.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (217.97, 96.75) -- (220.82, 93.90);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (212.17, 73.64) rectangle (226.62, 88.10);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (217.97, 79.44) -- (220.82, 82.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (217.97, 82.30) -- (220.82, 79.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (217.38, 80.87) -- (221.41, 80.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (219.39, 78.85) -- (219.39, 82.89);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (232.12,150.11) {DQMR};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (232.12,135.65) {Grid};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (232.12,121.20) {Mastermind};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (232.12,106.75) {Non-binary};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (232.12, 92.29) {Other binary};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (232.12, 77.84) {Random Blocks};
\end{scope}
\end{tikzpicture}