### 2021-06-13 doc/SAT_long_talk/scatter2.tex

% Created by tikzDevice version 0.12.3.1 on 2021-06-13 19:02:04
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (303.53,224.04);
\begin{scope}
\path[clip] (  0.00, 16.35) rectangle (303.53,207.69);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00, 16.35) rectangle (303.53,207.69);
\end{scope}
\begin{scope}
\path[clip] ( 40.51, 47.03) rectangle (195.67,202.19);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 40.51, 47.03) rectangle (195.67,202.19);
\definecolor{drawColor}{gray}{0.87}

\path[draw=drawColor,line width= 0.1pt,line join=round] ( 40.51, 54.09) --
	(195.67, 54.09);

\path[draw=drawColor,line width= 0.1pt,line join=round] ( 40.51,107.30) --
	(195.67,107.30);

\path[draw=drawColor,line width= 0.1pt,line join=round] ( 40.51,160.52) --
	(195.67,160.52);

\path[draw=drawColor,line width= 0.1pt,line join=round] ( 47.56, 47.03) --
	( 47.56,202.19);

\path[draw=drawColor,line width= 0.1pt,line join=round] (100.78, 47.03) --
	(100.78,202.19);

\path[draw=drawColor,line width= 0.1pt,line join=round] (154.00, 47.03) --
	(154.00,202.19);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 40.51, 80.69) --
	(195.67, 80.69);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 40.51,133.91) --
	(195.67,133.91);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 40.51,187.13) --
	(195.67,187.13);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 74.17, 47.03) --
	( 74.17,202.19);

\path[draw=drawColor,line width= 0.3pt,line join=round] (127.39, 47.03) --
	(127.39,202.19);

\path[draw=drawColor,line width= 0.3pt,line join=round] (180.61, 47.03) --
	(180.61,202.19);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.06, 91.57) -- ( 95.91, 94.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.06, 94.42) -- ( 95.91, 91.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.47, 93.00) -- ( 96.50, 93.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.48, 90.98) -- ( 94.48, 95.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 90.38, 91.57) -- ( 93.23, 94.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 90.38, 94.42) -- ( 93.23, 91.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 89.79, 93.00) -- ( 93.82, 93.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 91.81, 90.98) -- ( 91.81, 95.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 90.38, 93.41) -- ( 93.23, 96.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 90.38, 96.26) -- ( 93.23, 93.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 89.79, 94.84) -- ( 93.82, 94.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 91.81, 92.82) -- ( 91.81, 96.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.87, 94.07) -- (103.72, 96.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.87, 96.92) -- (103.72, 94.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.28, 95.50) -- (104.31, 95.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (102.29, 93.48) -- (102.29, 97.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.65, 93.06) -- ( 95.51, 95.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.65, 95.92) -- ( 95.51, 93.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.06, 94.49) -- ( 96.10, 94.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.08, 92.47) -- ( 94.08, 96.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 90.63, 91.57) -- ( 93.48, 94.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 90.63, 94.42) -- ( 93.48, 91.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 90.04, 93.00) -- ( 94.07, 93.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.05, 90.98) -- ( 92.05, 95.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 90.87, 91.96) -- ( 93.72, 94.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 90.87, 94.82) -- ( 93.72, 91.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 90.28, 93.39) -- ( 94.32, 93.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.30, 91.37) -- ( 92.30, 95.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 91.57, 93.74) -- ( 94.43, 96.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 91.57, 96.60) -- ( 94.43, 93.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 90.98, 95.17) -- ( 95.02, 95.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.00, 93.15) -- ( 93.00, 97.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.23, 94.69) -- ( 95.09, 97.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.23, 97.55) -- ( 95.09, 94.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 91.64, 96.12) -- ( 95.68, 96.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.66, 94.10) -- ( 93.66, 98.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 91.80, 91.17) -- ( 94.65, 94.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 91.80, 94.02) -- ( 94.65, 91.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 91.21, 92.59) -- ( 95.24, 92.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.22, 90.57) -- ( 93.22, 94.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.44, 92.34) -- ( 95.30, 95.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.44, 95.19) -- ( 95.30, 92.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 91.85, 93.77) -- ( 95.89, 93.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.87, 91.75) -- ( 93.87, 95.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 91.11, 91.96) -- ( 93.96, 94.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 91.11, 94.82) -- ( 93.96, 91.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 90.52, 93.39) -- ( 94.55, 93.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.54, 91.37) -- ( 92.54, 95.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 91.80, 92.71) -- ( 94.65, 95.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 91.80, 95.56) -- ( 94.65, 92.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 91.21, 94.13) -- ( 95.24, 94.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.22, 92.12) -- ( 93.22, 96.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 90.13, 91.96) -- ( 92.98, 94.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 90.13, 94.82) -- ( 92.98, 91.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 89.53, 93.39) -- ( 93.57, 93.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 91.55, 91.37) -- ( 91.55, 95.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.26, 93.74) -- ( 96.11, 96.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.26, 96.60) -- ( 96.11, 93.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.66, 95.17) -- ( 96.70, 95.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.68, 93.15) -- ( 94.68, 97.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.83, 92.71) -- ( 96.68, 95.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.83, 95.56) -- ( 96.68, 92.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.24, 94.13) -- ( 97.27, 94.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.26, 92.12) -- ( 95.26, 96.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.72, 97.15) -- ( 98.57,100.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.72,100.00) -- ( 98.57, 97.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.13, 98.58) -- ( 99.16, 98.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.14, 96.56) -- ( 97.14,100.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.23, 95.57) -- ( 98.08, 98.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.23, 98.43) -- ( 98.08, 95.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.64, 97.00) -- ( 98.68, 97.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 96.66, 94.98) -- ( 96.66, 99.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.07, 94.39) -- ( 97.92, 97.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.07, 97.24) -- ( 97.92, 94.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.47, 95.81) -- ( 98.51, 95.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 96.49, 93.80) -- ( 96.49, 97.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.01, 94.99) -- ( 96.87, 97.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.01, 97.85) -- ( 96.87, 94.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.42, 96.42) -- ( 97.46, 96.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.44, 94.40) -- ( 95.44, 98.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.06, 96.65) -- ( 99.91, 99.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.06, 99.50) -- ( 99.91, 96.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 96.47, 98.07) -- (100.50, 98.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 98.49, 96.06) -- ( 98.49,100.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.45, 93.74) -- ( 96.30, 96.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.45, 96.60) -- ( 96.30, 93.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.86, 95.17) -- ( 96.89, 95.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.88, 93.15) -- ( 94.88, 97.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.35,108.64) -- (103.20,111.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.35,111.49) -- (103.20,108.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.76,110.06) -- (103.79,110.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.78,108.05) -- (101.78,112.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.64, 93.74) -- ( 96.49, 96.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.64, 96.60) -- ( 96.49, 93.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.05, 95.17) -- ( 97.09, 95.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.07, 93.15) -- ( 95.07, 97.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.20, 93.74) -- ( 97.05, 96.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.20, 96.60) -- ( 97.05, 93.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.60, 95.17) -- ( 97.64, 95.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.62, 93.15) -- ( 95.62, 97.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.45, 95.57) -- ( 96.30, 98.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.45, 98.43) -- ( 96.30, 95.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.86, 97.00) -- ( 96.89, 97.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.88, 94.98) -- ( 94.88, 99.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 98.26, 98.09) -- (101.12,100.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 98.26,100.95) -- (101.12, 98.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.67, 99.52) -- (101.71, 99.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.69, 97.50) -- ( 99.69,101.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.72, 94.69) -- ( 98.57, 97.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.72, 97.55) -- ( 98.57, 94.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.13, 96.12) -- ( 99.16, 96.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.14, 94.10) -- ( 97.14, 98.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.45, 93.06) -- ( 96.30, 95.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.45, 95.92) -- ( 96.30, 93.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.86, 94.49) -- ( 96.89, 94.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.88, 92.47) -- ( 94.88, 96.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 96.48, 95.29) -- ( 99.34, 98.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 96.48, 98.14) -- ( 99.34, 95.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.89, 96.71) -- ( 99.93, 96.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.91, 94.70) -- ( 97.91, 98.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.90, 93.06) -- ( 97.75, 95.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.90, 95.92) -- ( 97.75, 93.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.31, 94.49) -- ( 98.34, 94.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 96.32, 92.47) -- ( 96.32, 96.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.88, 97.39) -- (100.73,100.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.88,100.25) -- (100.73, 97.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.28, 98.82) -- (101.32, 98.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.30, 96.80) -- ( 99.30,100.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.00,106.33) -- (101.85,109.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.00,109.18) -- (101.85,106.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 98.41,107.76) -- (102.45,107.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.43,105.74) -- (100.43,109.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.88, 97.15) -- (100.73,100.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.88,100.00) -- (100.73, 97.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.28, 98.58) -- (101.32, 98.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.30, 96.56) -- ( 99.30,100.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 96.33, 96.12) -- ( 99.19, 98.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 96.33, 98.98) -- ( 99.19, 96.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.74, 97.55) -- ( 99.78, 97.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.76, 95.53) -- ( 97.76, 99.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.74, 96.90) -- (100.60, 99.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.74, 99.76) -- (100.60, 96.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.15, 98.33) -- (101.19, 98.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.17, 96.31) -- ( 99.17,100.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (106.89, 98.54) -- (109.74,101.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (106.89,101.39) -- (109.74, 98.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (106.30, 99.97) -- (110.34, 99.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.32, 97.95) -- (108.32,101.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 96.77, 95.29) -- ( 99.63, 98.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 96.77, 98.14) -- ( 99.63, 95.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 96.18, 96.71) -- (100.22, 96.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 98.20, 94.70) -- ( 98.20, 98.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.81,103.00) -- (102.66,105.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.81,105.86) -- (102.66,103.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.22,104.43) -- (103.25,104.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.23,102.41) -- (101.23,106.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.81, 99.38) -- (102.66,102.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.81,102.23) -- (102.66, 99.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.22,100.81) -- (103.25,100.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.23, 98.79) -- (101.23,102.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 98.26, 97.63) -- (101.12,100.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 98.26,100.49) -- (101.12, 97.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.67, 99.06) -- (101.71, 99.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.69, 97.04) -- ( 99.69,101.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.12,107.19) -- (101.97,110.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.12,110.04) -- (101.97,107.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 98.53,108.61) -- (102.56,108.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.55,106.59) -- (100.55,110.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (105.14, 97.15) -- (107.99,100.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (105.14,100.00) -- (107.99, 97.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (104.55, 98.58) -- (108.58, 98.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (106.57, 96.56) -- (106.57,100.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 96.03, 96.12) -- ( 98.88, 98.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 96.03, 98.98) -- ( 98.88, 96.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.44, 97.55) -- ( 99.47, 97.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.46, 95.53) -- ( 97.46, 99.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.20, 95.57) -- (100.05, 98.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.20, 98.43) -- (100.05, 95.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 96.61, 97.00) -- (100.64, 97.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 98.63, 94.98) -- ( 98.63, 99.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.48, 96.90) -- (100.33, 99.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.48, 99.76) -- (100.33, 96.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 96.88, 98.33) -- (100.92, 98.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 98.90, 96.31) -- ( 98.90,100.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.06, 96.90) -- ( 99.91, 99.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 97.06, 99.76) -- ( 99.91, 96.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 96.47, 98.33) -- (100.50, 98.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 98.49, 96.31) -- ( 98.49,100.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.46,100.16) -- (104.31,103.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.46,103.02) -- (104.31,100.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.87,101.59) -- (104.90,101.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (102.89, 99.57) -- (102.89,103.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.52,130.78) -- (138.37,133.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.52,133.63) -- (138.37,130.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.93,132.21) -- (138.96,132.21);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.94,130.19) -- (136.94,134.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.78,132.43) -- (138.63,135.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.78,135.28) -- (138.63,132.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.19,133.85) -- (139.22,133.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.20,131.84) -- (137.20,135.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.97,130.82) -- (135.82,133.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.97,133.67) -- (135.82,130.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.38,132.25) -- (136.41,132.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.40,130.23) -- (134.40,134.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.11,130.29) -- (135.96,133.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.11,133.14) -- (135.96,130.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.52,131.72) -- (136.55,131.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.53,129.70) -- (134.53,133.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.41,130.89) -- (138.27,133.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.41,133.74) -- (138.27,130.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.82,132.31) -- (138.86,132.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.84,130.30) -- (136.84,134.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.45,130.85) -- (138.30,133.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.45,133.70) -- (138.30,130.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.86,132.27) -- (138.89,132.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.88,130.26) -- (136.88,134.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.75,131.79) -- (137.61,134.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.75,134.65) -- (137.61,131.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.16,133.22) -- (138.20,133.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.18,131.20) -- (136.18,135.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.88,130.87) -- (137.74,133.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.88,133.73) -- (137.74,130.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.29,132.30) -- (138.33,132.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.31,130.28) -- (136.31,134.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.39,130.87) -- (137.24,133.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.39,133.73) -- (137.24,130.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.79,132.30) -- (137.83,132.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.81,130.28) -- (135.81,134.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.42,131.27) -- (136.27,134.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.42,134.12) -- (136.27,131.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.83,132.69) -- (136.87,132.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.85,130.68) -- (134.85,134.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.73,130.19) -- (135.58,133.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.73,133.04) -- (135.58,130.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.14,131.62) -- (136.17,131.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.15,129.60) -- (134.15,133.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.14,130.70) -- (138.99,133.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.14,133.55) -- (138.99,130.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.55,132.13) -- (139.58,132.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.56,130.11) -- (137.56,134.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.97,132.26) -- (139.83,135.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.97,135.12) -- (139.83,132.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.38,133.69) -- (140.42,133.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.40,131.67) -- (138.40,135.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.84,131.73) -- (138.69,134.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.84,134.59) -- (138.69,131.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.25,133.16) -- (139.28,133.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.27,131.14) -- (137.27,135.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.75,130.59) -- (135.60,133.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.75,133.45) -- (135.60,130.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.16,132.02) -- (136.19,132.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.17,130.00) -- (134.17,134.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.21,134.41) -- (139.07,137.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.21,137.26) -- (139.07,134.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.62,135.83) -- (139.66,135.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.64,133.82) -- (137.64,137.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.57,134.43) -- (139.42,137.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.57,137.28) -- (139.42,134.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.98,135.85) -- (140.01,135.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.00,133.84) -- (138.00,137.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.43,138.50) -- (138.28,141.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.43,141.35) -- (138.28,138.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.84,139.93) -- (138.87,139.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.86,137.91) -- (136.86,141.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.42,134.94) -- (142.27,137.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.42,137.80) -- (142.27,134.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.83,136.37) -- (142.86,136.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.84,134.35) -- (140.84,138.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.47,131.68) -- (138.32,134.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.47,134.54) -- (138.32,131.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.87,133.11) -- (138.91,133.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.89,131.09) -- (136.89,135.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.81,132.22) -- (140.66,135.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.81,135.07) -- (140.66,132.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.22,133.64) -- (141.25,133.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.23,131.62) -- (139.23,135.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.73,131.96) -- (141.59,134.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.73,134.82) -- (141.59,131.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.14,133.39) -- (142.18,133.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.16,131.37) -- (140.16,135.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.08,131.65) -- (137.93,134.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.08,134.50) -- (137.93,131.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.49,133.07) -- (138.52,133.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.50,131.05) -- (136.50,135.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.34,133.56) -- (140.19,136.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.34,136.42) -- (140.19,133.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.75,134.99) -- (140.78,134.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.77,132.97) -- (138.77,137.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.26,131.63) -- (138.11,134.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.26,134.49) -- (138.11,131.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.67,133.06) -- (138.70,133.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.69,131.04) -- (136.69,135.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.71,131.57) -- (138.56,134.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.71,134.42) -- (138.56,131.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.12,133.00) -- (139.15,133.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.13,130.98) -- (137.13,135.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.33,133.65) -- (139.18,136.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.33,136.50) -- (139.18,133.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.74,135.07) -- (139.78,135.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.76,133.06) -- (137.76,137.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.67,131.44) -- (140.53,134.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.67,134.30) -- (140.53,131.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.08,132.87) -- (141.12,132.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.10,130.85) -- (139.10,134.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.45,134.90) -- (139.31,137.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.45,137.75) -- (139.31,134.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.86,136.32) -- (139.90,136.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.88,134.30) -- (137.88,138.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.34,132.60) -- (139.20,135.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.34,135.45) -- (139.20,132.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.75,134.03) -- (139.79,134.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.77,132.01) -- (137.77,136.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.49,132.71) -- (140.34,135.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.49,135.57) -- (140.34,132.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.90,134.14) -- (140.93,134.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.92,132.12) -- (138.92,136.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.37,134.10) -- (141.23,136.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.37,136.95) -- (141.23,134.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.78,135.53) -- (141.82,135.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.80,133.51) -- (139.80,137.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.33,135.14) -- (142.18,137.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.33,137.99) -- (142.18,135.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.74,136.56) -- (142.77,136.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.75,134.55) -- (140.75,138.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.59,139.31) -- (142.44,142.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.59,142.16) -- (142.44,139.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.00,140.74) -- (143.03,140.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.02,138.72) -- (141.02,142.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.35,136.71) -- (142.21,139.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.35,139.56) -- (142.21,136.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.76,138.13) -- (142.80,138.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.78,136.11) -- (140.78,140.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.76,135.54) -- (143.62,138.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.76,138.40) -- (143.62,135.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.17,136.97) -- (144.21,136.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.19,134.95) -- (142.19,138.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.71,136.83) -- (141.57,139.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.71,139.68) -- (141.57,136.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.12,138.25) -- (142.16,138.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.14,136.23) -- (140.14,140.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.58,137.00) -- (140.43,139.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.58,139.85) -- (140.43,137.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.99,138.43) -- (141.02,138.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.01,136.41) -- (139.01,140.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.29,138.11) -- (142.14,140.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.29,140.96) -- (142.14,138.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.70,139.54) -- (142.73,139.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.71,137.52) -- (140.71,141.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.14,136.69) -- (141.99,139.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.14,139.54) -- (141.99,136.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.55,138.12) -- (142.58,138.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.56,136.10) -- (140.56,140.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.94,137.93) -- (141.79,140.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.94,140.78) -- (141.79,137.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.35,139.36) -- (142.38,139.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.37,137.34) -- (140.37,141.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.26,135.75) -- (142.11,138.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.26,138.61) -- (142.11,135.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.67,137.18) -- (142.70,137.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.68,135.16) -- (140.68,139.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.76,133.34) -- (140.61,136.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.76,136.19) -- (140.61,133.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (137.17,134.77) -- (141.20,134.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.19,132.75) -- (139.19,136.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.70,136.97) -- (145.55,139.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.70,139.82) -- (145.55,136.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.11,138.39) -- (146.14,138.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (144.12,136.38) -- (144.12,140.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.45,136.87) -- (149.30,139.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (146.45,139.72) -- (149.30,136.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (145.86,138.29) -- (149.89,138.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.88,136.27) -- (147.88,140.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.71,135.08) -- (144.56,137.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.71,137.93) -- (144.56,135.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.12,136.51) -- (145.15,136.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.13,134.49) -- (143.13,138.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.35,138.62) -- (143.20,141.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.35,141.47) -- (143.20,138.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.75,140.04) -- (143.79,140.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.77,138.02) -- (141.77,142.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.53,141.02) -- (143.38,143.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.53,143.88) -- (143.38,141.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (139.94,142.45) -- (143.97,142.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (141.96,140.43) -- (141.96,144.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.78,152.24) -- (160.63,155.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.78,155.10) -- (160.63,152.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.19,153.67) -- (161.22,153.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.21,151.65) -- (159.21,155.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.08,153.17) -- (160.93,156.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.08,156.02) -- (160.93,153.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.49,154.60) -- (161.52,154.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.50,152.58) -- (159.50,156.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.04,153.43) -- (164.90,156.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.04,156.28) -- (164.90,153.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.45,154.85) -- (165.49,154.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.47,152.83) -- (163.47,156.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.29,152.69) -- (163.14,155.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.29,155.54) -- (163.14,152.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.70,154.12) -- (163.74,154.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.72,152.10) -- (161.72,156.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.10,155.89) -- (160.96,158.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.10,158.75) -- (160.96,155.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.51,157.32) -- (161.55,157.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.53,155.30) -- (159.53,159.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.68,154.48) -- (163.54,157.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.68,157.33) -- (163.54,154.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.09,155.91) -- (164.13,155.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.11,153.89) -- (162.11,157.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.45,154.01) -- (162.31,156.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.45,156.87) -- (162.31,154.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.86,155.44) -- (162.90,155.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.88,153.42) -- (160.88,157.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.88,155.04) -- (161.74,157.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.88,157.89) -- (161.74,155.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.29,156.47) -- (162.33,156.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.31,154.45) -- (160.31,158.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.49,154.48) -- (163.34,157.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.49,157.33) -- (163.34,154.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.90,155.91) -- (163.93,155.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.91,153.89) -- (161.91,157.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.31,155.12) -- (164.16,157.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.31,157.97) -- (164.16,155.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.72,156.55) -- (164.75,156.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.74,154.53) -- (162.74,158.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.50,156.09) -- (161.36,158.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.50,158.94) -- (161.36,156.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.91,157.51) -- (161.95,157.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.93,155.49) -- (159.93,159.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.93,155.61) -- (163.78,158.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.93,158.47) -- (163.78,155.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.34,157.04) -- (164.38,157.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.36,155.02) -- (162.36,159.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.39,152.31) -- (163.24,155.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.39,155.16) -- (163.24,152.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.80,153.73) -- (163.83,153.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.81,151.72) -- (161.81,155.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.39,155.76) -- (160.24,158.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.39,158.61) -- (160.24,155.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (156.80,157.18) -- (160.83,157.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.82,155.17) -- (158.82,159.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.58,152.55) -- (160.44,155.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.58,155.41) -- (160.44,152.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (156.99,153.98) -- (161.03,153.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.01,151.96) -- (159.01,156.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.05,154.41) -- (163.91,157.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.05,157.27) -- (163.91,154.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.46,155.84) -- (164.50,155.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.48,153.82) -- (162.48,157.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.64,154.02) -- (162.49,156.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.64,156.88) -- (162.49,154.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.04,155.45) -- (163.08,155.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.06,153.43) -- (161.06,157.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.03,154.63) -- (164.88,157.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.03,157.49) -- (164.88,154.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.44,156.06) -- (165.48,156.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.46,154.04) -- (163.46,158.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.39,157.93) -- (163.25,160.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.39,160.78) -- (163.25,157.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.80,159.35) -- (163.84,159.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.82,157.34) -- (161.82,161.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.55,154.86) -- (166.40,157.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.55,157.71) -- (166.40,154.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.96,156.29) -- (166.99,156.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.98,154.27) -- (164.98,158.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.06,153.97) -- (162.91,156.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.06,156.82) -- (162.91,153.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.47,155.40) -- (163.50,155.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.49,153.38) -- (161.49,157.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.34,155.60) -- (162.19,158.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.34,158.45) -- (162.19,155.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.75,157.03) -- (162.78,157.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.76,155.01) -- (160.76,159.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.63,154.33) -- (162.49,157.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.63,157.19) -- (162.49,154.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.04,155.76) -- (163.08,155.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.06,153.74) -- (161.06,157.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.23,155.00) -- (164.08,157.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.23,157.86) -- (164.08,155.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.64,156.43) -- (164.67,156.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.65,154.41) -- (162.65,158.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.51,153.98) -- (164.37,156.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.51,156.83) -- (164.37,153.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.92,155.41) -- (164.96,155.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.94,153.39) -- (162.94,157.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.79,156.26) -- (164.64,159.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.79,159.11) -- (164.64,156.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.20,157.69) -- (165.23,157.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.22,155.67) -- (163.22,159.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.69,154.01) -- (161.54,156.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.69,156.86) -- (161.54,154.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.10,155.44) -- (162.13,155.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.11,153.42) -- (160.11,157.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.58,155.99) -- (162.43,158.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.58,158.85) -- (162.43,155.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.98,157.42) -- (163.02,157.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.00,155.40) -- (161.00,159.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.56,158.81) -- (167.41,161.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.56,161.66) -- (167.41,158.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.97,160.24) -- (168.00,160.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.98,158.22) -- (165.98,162.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.74,157.81) -- (167.59,160.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.74,160.67) -- (167.59,157.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.15,159.24) -- (168.18,159.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.16,157.22) -- (166.16,161.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.23,156.22) -- (163.08,159.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.23,159.07) -- (163.08,156.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.64,157.64) -- (163.68,157.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.66,155.63) -- (161.66,159.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.49,157.22) -- (163.34,160.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.49,160.07) -- (163.34,157.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.90,158.64) -- (163.93,158.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.92,156.63) -- (161.92,160.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.16,156.73) -- (164.01,159.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.16,159.58) -- (164.01,156.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.57,158.16) -- (164.61,158.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.59,156.14) -- (162.59,160.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.73,158.71) -- (163.58,161.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.73,161.56) -- (163.58,158.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.14,160.13) -- (164.17,160.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.16,158.12) -- (162.16,162.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.91,155.82) -- (166.76,158.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.91,158.68) -- (166.76,155.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.32,157.25) -- (167.36,157.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.34,155.23) -- (165.34,159.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.02,157.37) -- (168.87,160.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.02,160.23) -- (168.87,157.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.43,158.80) -- (169.46,158.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.45,156.78) -- (167.45,160.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.64,156.78) -- (168.49,159.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.64,159.64) -- (168.49,156.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.05,158.21) -- (169.08,158.21);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (167.06,156.19) -- (167.06,160.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.67,162.47) -- (165.53,165.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.67,165.32) -- (165.53,162.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.08,163.90) -- (166.12,163.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.10,161.88) -- (164.10,165.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.47,157.62) -- (168.32,160.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.47,160.47) -- (168.32,157.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.88,159.04) -- (168.91,159.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.90,157.03) -- (166.90,161.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.32,156.93) -- (166.17,159.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.32,159.78) -- (166.17,156.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.73,158.35) -- (166.77,158.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.75,156.34) -- (164.75,160.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.67,155.33) -- (164.53,158.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.67,158.18) -- (164.53,155.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.08,156.76) -- (165.12,156.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.10,154.74) -- (163.10,158.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.75,155.50) -- (166.61,158.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.75,158.36) -- (166.61,155.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.16,156.93) -- (167.20,156.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.18,154.91) -- (165.18,158.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.27,156.76) -- (165.13,159.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.27,159.61) -- (165.13,156.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.68,158.18) -- (165.72,158.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.70,156.16) -- (163.70,160.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.82,155.54) -- (166.67,158.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.82,158.39) -- (166.67,155.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.23,156.96) -- (167.26,156.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.24,154.95) -- (165.24,158.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.43,158.45) -- (168.29,161.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (165.43,161.30) -- (168.29,158.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.84,159.88) -- (168.88,159.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (166.86,157.86) -- (166.86,161.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.41,158.61) -- (164.26,161.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.41,161.46) -- (164.26,158.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.82,160.03) -- (164.85,160.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.84,158.02) -- (162.84,162.05);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.82,160.33) -- (165.67,163.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.82,163.19) -- (165.67,160.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.23,161.76) -- (166.26,161.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (164.25,159.74) -- (164.25,163.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.41,156.49) -- (165.26,159.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (162.41,159.35) -- (165.26,156.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (161.82,157.92) -- (165.85,157.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (163.84,155.90) -- (163.84,159.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,175.49) -- (190.04,178.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,178.34) -- (190.04,175.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,176.92) -- (190.63,176.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,174.90) -- (188.62,178.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,174.87) -- (190.04,177.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,177.72) -- (190.04,174.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,176.29) -- (190.63,176.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,174.28) -- (188.62,178.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,177.44) -- (190.04,180.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,180.30) -- (190.04,177.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,178.87) -- (190.63,178.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,176.85) -- (188.62,180.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.50,173.07) -- (181.36,175.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.50,175.92) -- (181.36,173.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.91,174.50) -- (181.95,174.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.93,172.48) -- (179.93,176.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,174.53) -- (190.04,177.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,177.39) -- (190.04,174.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,175.96) -- (190.63,175.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,173.94) -- (188.62,177.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,174.38) -- (190.04,177.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,177.24) -- (190.04,174.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,175.81) -- (190.63,175.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,173.79) -- (188.62,177.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,174.07) -- (190.04,176.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,176.92) -- (190.04,174.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,175.49) -- (190.63,175.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,173.48) -- (188.62,177.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,177.73) -- (190.04,180.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,180.58) -- (190.04,177.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,179.15) -- (190.63,179.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,177.14) -- (188.62,181.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.40,175.18) -- (182.25,178.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.40,178.04) -- (182.25,175.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.81,176.61) -- (182.85,176.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.83,174.59) -- (180.83,178.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.77,171.83) -- (180.62,174.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.77,174.68) -- (180.62,171.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.18,173.25) -- (181.21,173.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.20,171.24) -- (179.20,175.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.80,176.79) -- (182.65,179.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.80,179.64) -- (182.65,176.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.21,178.22) -- (183.24,178.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.22,176.20) -- (181.22,180.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,174.49) -- (190.04,177.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,177.35) -- (190.04,174.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,175.92) -- (190.63,175.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,173.90) -- (188.62,177.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,174.48) -- (190.04,177.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,177.33) -- (190.04,174.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,175.90) -- (190.63,175.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,173.89) -- (188.62,177.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,174.03) -- (190.04,176.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,176.88) -- (190.04,174.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,175.45) -- (190.63,175.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,173.44) -- (188.62,177.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,175.55) -- (190.04,178.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,178.40) -- (190.04,175.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,176.98) -- (190.63,176.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,174.96) -- (188.62,178.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.15,176.75) -- (181.00,179.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.15,179.60) -- (181.00,176.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (177.56,178.17) -- (181.59,178.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.57,176.16) -- (179.57,180.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,176.62) -- (190.04,179.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,179.47) -- (190.04,176.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,178.04) -- (190.63,178.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,176.03) -- (188.62,180.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,173.64) -- (190.04,176.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,176.50) -- (190.04,173.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,175.07) -- (190.63,175.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,173.05) -- (188.62,177.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,176.52) -- (190.04,179.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,179.37) -- (190.04,176.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,177.94) -- (190.63,177.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,175.93) -- (188.62,179.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,175.51) -- (190.04,178.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,178.36) -- (190.04,175.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,176.94) -- (190.63,176.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,174.92) -- (188.62,178.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,180.37) -- (190.04,183.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,183.22) -- (190.04,180.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,181.80) -- (190.63,181.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,179.78) -- (188.62,183.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,176.54) -- (190.04,179.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,179.39) -- (190.04,176.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,177.96) -- (190.63,177.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,175.95) -- (188.62,179.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,175.70) -- (190.04,178.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,178.56) -- (190.04,175.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,177.13) -- (190.63,177.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,175.11) -- (188.62,179.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,175.95) -- (190.04,178.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,178.81) -- (190.04,175.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,177.38) -- (190.63,177.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,175.36) -- (188.62,179.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,176.22) -- (190.04,179.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,179.08) -- (190.04,176.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,177.65) -- (190.63,177.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,175.63) -- (188.62,179.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,173.55) -- (190.04,176.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,176.40) -- (190.04,173.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,174.98) -- (190.63,174.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,172.96) -- (188.62,176.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.55,176.58) -- (182.41,179.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.55,179.44) -- (182.41,176.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (178.96,178.01) -- (183.00,178.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.98,175.99) -- (180.98,180.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,174.91) -- (190.04,177.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,177.77) -- (190.04,174.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,176.34) -- (190.63,176.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,174.32) -- (188.62,178.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,175.45) -- (190.04,178.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,178.31) -- (190.04,175.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,176.88) -- (190.63,176.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,174.86) -- (188.62,178.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,174.52) -- (190.04,177.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,177.37) -- (190.04,174.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,175.95) -- (190.63,175.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,173.93) -- (188.62,177.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.95,175.78) -- (182.81,178.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.95,178.63) -- (182.81,175.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.36,177.20) -- (183.40,177.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.38,175.19) -- (181.38,179.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.13,177.35) -- (182.99,180.21);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.13,180.21) -- (182.99,177.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.54,178.78) -- (183.58,178.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.56,176.76) -- (181.56,180.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,179.65) -- (190.04,182.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,182.50) -- (190.04,179.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,181.07) -- (190.63,181.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,179.06) -- (188.62,183.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,178.96) -- (190.04,181.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,181.81) -- (190.04,178.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,180.39) -- (190.63,180.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,178.37) -- (188.62,182.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,179.64) -- (190.04,182.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,182.49) -- (190.04,179.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,181.06) -- (190.63,181.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,179.05) -- (188.62,183.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,176.26) -- (190.04,179.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,179.11) -- (190.04,176.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,177.68) -- (190.63,177.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,175.67) -- (188.62,179.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,181.05) -- (190.04,183.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,183.90) -- (190.04,181.05);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,182.48) -- (190.63,182.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,180.46) -- (188.62,184.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,177.15) -- (190.04,180.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,180.00) -- (190.04,177.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,178.57) -- (190.63,178.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,176.56) -- (188.62,180.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.20,174.13) -- (183.05,176.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.20,176.98) -- (183.05,174.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.61,175.55) -- (183.64,175.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.62,173.54) -- (181.62,177.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.04,178.39) -- (182.89,181.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (180.04,181.24) -- (182.89,178.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (179.45,179.81) -- (183.49,179.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (181.47,177.80) -- (181.47,181.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,179.61) -- (190.04,182.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,182.47) -- (190.04,179.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,181.04) -- (190.63,181.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,179.02) -- (188.62,183.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,175.42) -- (190.04,178.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,178.27) -- (190.04,175.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,176.85) -- (190.63,176.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,174.83) -- (188.62,178.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,176.00) -- (190.04,178.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,178.86) -- (190.04,176.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,177.43) -- (190.63,177.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,175.41) -- (188.62,179.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,176.14) -- (190.04,178.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,178.99) -- (190.04,176.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,177.57) -- (190.63,177.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,175.55) -- (188.62,179.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,176.19) -- (190.04,179.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,179.04) -- (190.04,176.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,177.61) -- (190.63,177.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,175.60) -- (188.62,179.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,178.31) -- (190.04,181.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,181.16) -- (190.04,178.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,179.74) -- (190.63,179.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,177.72) -- (188.62,181.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,178.43) -- (190.04,181.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,181.29) -- (190.04,178.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,179.86) -- (190.63,179.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,177.84) -- (188.62,181.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,182.26) -- (190.04,185.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,185.11) -- (190.04,182.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,183.68) -- (190.63,183.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,181.66) -- (188.62,185.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,186.74) -- (190.04,189.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,189.60) -- (190.04,186.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,188.17) -- (190.63,188.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,186.15) -- (188.62,190.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,193.71) -- (190.04,196.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,196.56) -- (190.04,193.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,195.14) -- (190.63,195.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,193.12) -- (188.62,197.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,186.98) -- (190.04,189.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,189.83) -- (190.04,186.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,188.40) -- (190.63,188.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,186.39) -- (188.62,190.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,182.60) -- (190.04,185.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,185.45) -- (190.04,182.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,184.03) -- (190.63,184.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,182.01) -- (188.62,186.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,184.24) -- (190.04,187.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,187.09) -- (190.04,184.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,185.66) -- (190.63,185.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,183.65) -- (188.62,187.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,184.44) -- (190.04,187.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,187.30) -- (190.04,184.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,185.87) -- (190.63,185.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,183.85) -- (188.62,187.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,182.02) -- (190.04,184.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,184.88) -- (190.04,182.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,183.45) -- (190.63,183.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,181.43) -- (188.62,185.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,180.15) -- (190.04,183.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,183.01) -- (190.04,180.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,181.58) -- (190.63,181.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,179.56) -- (188.62,183.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,180.07) -- (190.04,182.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,182.92) -- (190.04,180.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,181.50) -- (190.63,181.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,179.48) -- (188.62,183.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,185.11) -- (190.04,187.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,187.96) -- (190.04,185.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,186.54) -- (190.63,186.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,184.52) -- (188.62,188.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,183.02) -- (190.04,185.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,185.87) -- (190.04,183.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,184.45) -- (190.63,184.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,182.43) -- (188.62,186.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,180.80) -- (190.04,183.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,183.66) -- (190.04,180.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,182.23) -- (190.63,182.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,180.21) -- (188.62,184.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,186.88) -- (190.04,189.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,189.73) -- (190.04,186.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,188.30) -- (190.63,188.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,186.29) -- (188.62,190.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,182.23) -- (190.04,185.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,185.08) -- (190.04,182.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,183.66) -- (190.63,183.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,181.64) -- (188.62,185.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,185.96) -- (190.04,188.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,188.82) -- (190.04,185.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,187.39) -- (190.63,187.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,185.37) -- (188.62,189.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,186.54) -- (190.04,189.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,189.39) -- (190.04,186.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,187.96) -- (190.63,187.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,185.94) -- (188.62,189.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,186.47) -- (190.04,189.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,189.32) -- (190.04,186.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,187.90) -- (190.63,187.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,185.88) -- (188.62,189.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,193.71) -- (190.04,196.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,196.56) -- (190.04,193.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,195.14) -- (190.63,195.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,193.12) -- (188.62,197.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,193.71) -- (190.04,196.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,196.56) -- (190.04,193.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,195.14) -- (190.63,195.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,193.12) -- (188.62,197.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,193.71) -- (190.04,196.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,196.56) -- (190.04,193.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,195.14) -- (190.63,195.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,193.12) -- (188.62,197.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,193.71) -- (190.04,196.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,196.56) -- (190.04,193.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,195.14) -- (190.63,195.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,193.12) -- (188.62,197.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,193.71) -- (190.04,196.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,196.56) -- (190.04,193.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,195.14) -- (190.63,195.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,193.12) -- (188.62,197.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,185.27) -- (190.04,188.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,188.12) -- (190.04,185.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,186.70) -- (190.63,186.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,184.68) -- (188.62,188.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,181.12) -- (190.04,183.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,183.97) -- (190.04,181.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,182.54) -- (190.63,182.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,180.53) -- (188.62,184.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,184.12) -- (190.04,186.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,186.97) -- (190.04,184.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,185.54) -- (190.63,185.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,183.52) -- (188.62,187.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,181.58) -- (190.04,184.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,184.43) -- (190.04,181.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,183.00) -- (190.63,183.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,180.99) -- (188.62,185.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,184.34) -- (190.04,187.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,187.19) -- (190.04,184.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,185.77) -- (190.63,185.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,183.75) -- (188.62,187.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,186.33) -- (190.04,189.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,189.18) -- (190.04,186.33);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,187.75) -- (190.63,187.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,185.74) -- (188.62,189.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,193.71) -- (190.04,196.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,196.56) -- (190.04,193.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,195.14) -- (190.63,195.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,193.12) -- (188.62,197.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,187.19) -- (190.04,190.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,190.04) -- (190.04,187.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,188.62) -- (190.63,188.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,186.60) -- (188.62,190.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,193.71) -- (190.04,196.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,196.56) -- (190.04,193.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,195.14) -- (190.63,195.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,193.12) -- (188.62,197.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,186.66) -- (190.04,189.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,189.51) -- (190.04,186.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,188.09) -- (190.63,188.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,186.07) -- (188.62,190.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,193.71) -- (190.04,196.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,196.56) -- (190.04,193.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,195.14) -- (190.63,195.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,193.12) -- (188.62,197.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,186.14) -- (190.04,189.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,189.00) -- (190.04,186.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,187.57) -- (190.63,187.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,185.55) -- (188.62,189.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,185.40) -- (190.04,188.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,188.25) -- (190.04,185.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,186.82) -- (190.63,186.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,184.80) -- (188.62,188.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,185.28) -- (190.04,188.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,188.13) -- (190.04,185.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,186.71) -- (190.63,186.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,184.69) -- (188.62,188.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,187.34) -- (190.04,190.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,190.19) -- (190.04,187.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,188.77) -- (190.63,188.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,186.75) -- (188.62,190.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,187.16) -- (190.04,190.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,190.01) -- (190.04,187.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,188.59) -- (190.63,188.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,186.57) -- (188.62,190.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,185.56) -- (190.04,188.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,188.42) -- (190.04,185.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,186.99) -- (190.63,186.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,184.97) -- (188.62,189.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,182.02) -- (190.04,184.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,184.87) -- (190.04,182.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,183.44) -- (190.63,183.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,181.43) -- (188.62,185.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,186.24) -- (190.04,189.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,189.10) -- (190.04,186.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,187.67) -- (190.63,187.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,185.65) -- (188.62,189.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,181.26) -- (190.04,184.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,184.11) -- (190.04,181.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,182.68) -- (190.63,182.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,180.66) -- (188.62,184.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,185.10) -- (190.04,187.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,187.95) -- (190.04,185.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,186.53) -- (190.63,186.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,184.51) -- (188.62,188.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,182.82) -- (190.04,185.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,185.67) -- (190.04,182.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,184.25) -- (190.63,184.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,182.23) -- (188.62,186.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,193.71) -- (190.04,196.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,196.56) -- (190.04,193.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,195.14) -- (190.63,195.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,193.12) -- (188.62,197.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,185.71) -- (190.04,188.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,188.56) -- (190.04,185.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,187.14) -- (190.63,187.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,185.12) -- (188.62,189.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,188.79) -- (190.04,191.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,191.64) -- (190.04,188.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,190.22) -- (190.63,190.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,188.20) -- (188.62,192.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,188.04) -- (190.04,190.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,190.89) -- (190.04,188.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,189.46) -- (190.63,189.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,187.45) -- (188.62,191.48);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor,fill opacity=0.50] ( 99.58, 99.18) --
	(102.44, 99.18) --
	(102.44,102.03) --
	( 99.58,102.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (101.65,100.72) --
	(104.50,100.72) --
	(104.50,103.57) --
	(101.65,103.57) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (101.65,108.55) --
	(104.50,108.55) --
	(104.50,111.40) --
	(101.65,111.40) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 98.64, 97.15) --
	(101.49, 97.15) --
	(101.49,100.00) --
	( 98.64,100.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (100.97, 98.97) --
	(103.82, 98.97) --
	(103.82,101.82) --
	(100.97,101.82) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (100.56,105.52) --
	(103.41,105.52) --
	(103.41,108.38) --
	(100.56,108.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (102.30, 99.78) --
	(105.15, 99.78) --
	(105.15,102.63) --
	(102.30,102.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 99.92, 96.90) --
	(102.77, 96.90) --
	(102.77, 99.76) --
	( 99.92, 99.76) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 99.58, 96.90) --
	(102.44, 96.90) --
	(102.44, 99.76) --
	( 99.58, 99.76) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (102.12, 98.75) --
	(104.97, 98.75) --
	(104.97,101.61) --
	(102.12,101.61) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (102.38,101.25) --
	(105.24,101.25) --
	(105.24,104.10) --
	(102.38,104.10) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (100.87, 99.18) --
	(103.72, 99.18) --
	(103.72,102.03) --
	(100.87,102.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (100.45,108.27) --
	(103.31,108.27) --
	(103.31,111.12) --
	(100.45,111.12) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (100.77, 99.78) --
	(103.62, 99.78) --
	(103.62,102.63) --
	(100.77,102.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (101.93, 99.38) --
	(104.78, 99.38) --
	(104.78,102.23) --
	(101.93,102.23) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (100.87, 98.75) --
	(103.72, 98.75) --
	(103.72,101.61) --
	(100.87,101.61) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.43,117.01) --
	(120.28,117.01) --
	(120.28,119.86) --
	(117.43,119.86) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.78,111.31) --
	(117.63,111.31) --
	(117.63,114.16) --
	(114.78,114.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.97,108.91) --
	(115.83,108.91) --
	(115.83,111.76) --
	(112.97,111.76) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.76,110.09) --
	(115.61,110.09) --
	(115.61,112.94) --
	(112.76,112.94) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.65,109.85) --
	(115.50,109.85) --
	(115.50,112.70) --
	(112.65,112.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.25,110.48) --
	(117.10,110.48) --
	(117.10,113.34) --
	(114.25,113.34) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.99,110.56) --
	(116.85,110.56) --
	(116.85,113.41) --
	(113.99,113.41) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.32,109.51) --
	(116.18,109.51) --
	(116.18,112.37) --
	(113.32,112.37) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.29,111.66) --
	(118.14,111.66) --
	(118.14,114.52) --
	(115.29,114.52) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.81,118.99) --
	(117.66,118.99) --
	(117.66,121.84) --
	(114.81,121.84) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.46,114.23) --
	(118.31,114.23) --
	(118.31,117.08) --
	(115.46,117.08) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.74,113.94) --
	(118.60,113.94) --
	(118.60,116.80) --
	(115.74,116.80) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.50,113.71) --
	(117.36,113.71) --
	(117.36,116.56) --
	(114.50,116.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.12,118.30) --
	(121.97,118.30) --
	(121.97,121.15) --
	(119.12,121.15) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.84,117.14) --
	(123.69,117.14) --
	(123.69,119.99) --
	(120.84,119.99) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.49,113.59) --
	(118.34,113.59) --
	(118.34,116.45) --
	(115.49,116.45) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.52,141.63) --
	(143.37,141.63) --
	(143.37,144.49) --
	(140.52,144.49) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.07,128.41) --
	(131.93,128.41) --
	(131.93,131.27) --
	(129.07,131.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (132.53,123.69) --
	(135.38,123.69) --
	(135.38,126.54) --
	(132.53,126.54) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.34,123.81) --
	(132.20,123.81) --
	(132.20,126.66) --
	(129.34,126.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (125.97,126.41) --
	(128.83,126.41) --
	(128.83,129.26) --
	(125.97,129.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (126.73,125.38) --
	(129.59,125.38) --
	(129.59,128.24) --
	(126.73,128.24) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (127.37,121.87) --
	(130.23,121.87) --
	(130.23,124.72) --
	(127.37,124.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.53,119.46) --
	(126.38,119.46) --
	(126.38,122.31) --
	(123.53,122.31) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (126.08,123.95) --
	(128.93,123.95) --
	(128.93,126.81) --
	(126.08,126.81) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.70,118.38) --
	(125.55,118.38) --
	(125.55,121.23) --
	(122.70,121.23) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.88,120.77) --
	(127.74,120.77) --
	(127.74,123.63) --
	(124.88,123.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (138.00,140.25) --
	(140.85,140.25) --
	(140.85,143.11) --
	(138.00,143.11) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (135.64,142.29) --
	(138.50,142.29) --
	(138.50,145.14) --
	(135.64,145.14) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.88,139.69) --
	(140.73,139.69) --
	(140.73,142.54) --
	(137.88,142.54) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.33,143.87) --
	(145.18,143.87) --
	(145.18,146.72) --
	(142.33,146.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.28,142.80) --
	(139.14,142.80) --
	(139.14,145.65) --
	(136.28,145.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (109.03,102.24) --
	(111.88,102.24) --
	(111.88,105.09) --
	(109.03,105.09) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (101.93, 99.97) --
	(104.78, 99.97) --
	(104.78,102.83) --
	(101.93,102.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (102.30, 99.78) --
	(105.15, 99.78) --
	(105.15,102.63) --
	(102.30,102.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (102.21,107.89) --
	(105.06,107.89) --
	(105.06,110.74) --
	(102.21,110.74) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (103.32,101.07) --
	(106.18,101.07) --
	(106.18,103.93) --
	(103.32,103.93) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (108.98, 99.97) --
	(111.83, 99.97) --
	(111.83,102.83) --
	(108.98,102.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (104.86,101.59) --
	(107.71,101.59) --
	(107.71,104.44) --
	(104.86,104.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (103.16,101.59) --
	(106.01,101.59) --
	(106.01,104.44) --
	(103.16,104.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (102.91, 99.58) --
	(105.76, 99.58) --
	(105.76,102.43) --
	(102.91,102.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (105.42,103.72) --
	(108.27,103.72) --
	(108.27,106.58) --
	(105.42,106.58) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (103.57,101.75) --
	(106.42,101.75) --
	(106.42,104.61) --
	(103.57,104.61) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (103.88,101.42) --
	(106.74,101.42) --
	(106.74,104.27) --
	(103.88,104.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (108.36,103.58) --
	(111.21,103.58) --
	(111.21,106.44) --
	(108.36,106.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (105.35,103.00) --
	(108.20,103.00) --
	(108.20,105.86) --
	(105.35,105.86) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (108.46,103.58) --
	(111.32,103.58) --
	(111.32,106.44) --
	(108.46,106.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (105.82,103.00) --
	(108.67,103.00) --
	(108.67,105.86) --
	(105.82,105.86) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (126.31,138.57) --
	(129.17,138.57) --
	(129.17,141.42) --
	(126.31,141.42) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.74,112.80) --
	(120.59,112.80) --
	(120.59,115.65) --
	(117.74,115.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.26,112.98) --
	(120.11,112.98) --
	(120.11,115.84) --
	(117.26,115.84) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.05,116.37) --
	(121.91,116.37) --
	(121.91,119.22) --
	(119.05,119.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.04,113.59) --
	(119.89,113.59) --
	(119.89,116.45) --
	(117.04,116.45) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.14,113.94) --
	(119.99,113.94) --
	(119.99,116.80) --
	(117.14,116.80) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.27,115.25) --
	(121.12,115.25) --
	(121.12,118.10) --
	(118.27,118.10) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.88,114.50) --
	(120.74,114.50) --
	(120.74,117.36) --
	(117.88,117.36) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.53,118.42) --
	(119.38,118.42) --
	(119.38,121.27) --
	(116.53,121.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.34,112.73) --
	(119.20,112.73) --
	(119.20,115.59) --
	(116.34,115.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.49,111.80) --
	(118.34,111.80) --
	(118.34,114.65) --
	(115.49,114.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.71,127.09) --
	(126.56,127.09) --
	(126.56,129.94) --
	(123.71,129.94) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (125.48,122.97) --
	(128.33,122.97) --
	(128.33,125.82) --
	(125.48,125.82) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.12,122.84) --
	(124.97,122.84) --
	(124.97,125.69) --
	(122.12,125.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.37,123.51) --
	(125.23,123.51) --
	(125.23,126.36) --
	(122.37,126.36) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.91,126.07) --
	(127.76,126.07) --
	(127.76,128.92) --
	(124.91,128.92) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.32,116.46) --
	(116.18,116.46) --
	(116.18,119.32) --
	(113.32,119.32) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (108.57,104.53) --
	(111.42,104.53) --
	(111.42,107.38) --
	(108.57,107.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (109.03,106.10) --
	(111.88,106.10) --
	(111.88,108.96) --
	(109.03,108.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (107.25,104.53) --
	(110.10,104.53) --
	(110.10,107.38) --
	(107.25,107.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (107.93,106.55) --
	(110.78,106.55) --
	(110.78,109.40) --
	(107.93,109.40) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.47,104.13) --
	(117.33,104.13) --
	(117.33,106.99) --
	(114.47,106.99) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (107.13,105.88) --
	(109.98,105.88) --
	(109.98,108.73) --
	(107.13,108.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (108.83,105.04) --
	(111.68,105.04) --
	(111.68,107.89) --
	(108.83,107.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.72,105.04) --
	(117.57,105.04) --
	(117.57,107.89) --
	(114.72,107.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (108.20,107.59) --
	(111.05,107.59) --
	(111.05,110.44) --
	(108.20,110.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.08,105.04) --
	(112.93,105.04) --
	(112.93,107.89) --
	(110.08,107.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.39,108.27) --
	(113.25,108.27) --
	(113.25,111.12) --
	(110.39,111.12) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.04,112.21) --
	(113.90,112.21) --
	(113.90,115.06) --
	(111.04,115.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (107.59,107.98) --
	(110.45,107.98) --
	(110.45,110.84) --
	(107.59,110.84) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (109.23,106.10) --
	(112.08,106.10) --
	(112.08,108.96) --
	(109.23,108.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.28,112.14) --
	(115.13,112.14) --
	(115.13,115.00) --
	(112.28,115.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.47,132.27) --
	(127.32,132.27) --
	(127.32,135.13) --
	(124.47,135.13) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.26,106.66) --
	(113.11,106.66) --
	(113.11,109.51) --
	(110.26,109.51) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (108.78,108.17) --
	(111.63,108.17) --
	(111.63,111.03) --
	(108.78,111.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (109.76,107.69) --
	(112.61,107.69) --
	(112.61,110.54) --
	(109.76,110.54) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.54,109.26) --
	(115.39,109.26) --
	(115.39,112.11) --
	(112.54,112.11) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (109.85,108.73) --
	(112.70,108.73) --
	(112.70,111.58) --
	(109.85,111.58) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (108.62,106.10) --
	(111.47,106.10) --
	(111.47,108.96) --
	(108.62,108.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.16,109.51) --
	(115.02,109.51) --
	(115.02,112.37) --
	(112.16,112.37) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.33,107.79) --
	(114.19,107.79) --
	(114.19,110.64) --
	(111.33,110.64) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.45,114.72) --
	(119.30,114.72) --
	(119.30,117.57) --
	(116.45,117.57) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (109.76,109.51) --
	(112.61,109.51) --
	(112.61,112.37) --
	(109.76,112.37) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.51,121.84) --
	(122.36,121.84) --
	(122.36,124.69) --
	(119.51,124.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.53,122.94) --
	(120.38,122.94) --
	(120.38,125.80) --
	(117.53,125.80) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.73,126.21) --
	(122.58,126.21) --
	(122.58,129.06) --
	(119.73,129.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.25,126.27) --
	(121.10,126.27) --
	(121.10,129.12) --
	(118.25,129.12) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.18,128.28) --
	(124.03,128.28) --
	(124.03,131.13) --
	(121.18,131.13) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (187.19,193.71) --
	(190.04,193.71) --
	(190.04,196.56) --
	(187.19,196.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (187.19,183.07) --
	(190.04,183.07) --
	(190.04,185.92) --
	(187.19,185.92) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (155.27,171.74) --
	(158.12,171.74) --
	(158.12,174.60) --
	(155.27,174.60) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (154.23,154.12) --
	(157.08,154.12) --
	(157.08,156.97) --
	(154.23,156.97) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (157.70,160.65) --
	(160.56,160.65) --
	(160.56,163.50) --
	(157.70,163.50) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (150.75,173.50) --
	(153.60,173.50) --
	(153.60,176.36) --
	(150.75,176.36) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.78,113.53) --
	(121.63,113.53) --
	(121.63,116.39) --
	(118.78,116.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.99,116.04) --
	(121.84,116.04) --
	(121.84,118.89) --
	(118.99,118.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.95,114.99) --
	(120.80,114.99) --
	(120.80,117.84) --
	(117.95,117.84) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.87,116.56) --
	(124.73,116.56) --
	(124.73,119.41) --
	(121.87,119.41) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (187.19,193.71) --
	(190.04,193.71) --
	(190.04,196.56) --
	(187.19,196.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (187.19,193.71) --
	(190.04,193.71) --
	(190.04,196.56) --
	(187.19,196.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (187.19,193.71) --
	(190.04,193.71) --
	(190.04,196.56) --
	(187.19,196.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (187.19,193.71) --
	(190.04,193.71) --
	(190.04,196.56) --
	(187.19,196.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (187.19,193.71) --
	(190.04,193.71) --
	(190.04,196.56) --
	(187.19,196.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (187.19,193.71) --
	(190.04,193.71) --
	(190.04,196.56) --
	(187.19,196.56) --
	cycle;
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 76.84, 86.83) -- ( 80.87, 86.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 78.86, 84.81) -- ( 78.86, 88.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 75.19, 82.80) -- ( 79.22, 82.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 77.20, 80.78) -- ( 77.20, 84.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 73.26, 81.80) -- ( 77.29, 81.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 75.27, 79.78) -- ( 75.27, 83.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 75.19, 83.73) -- ( 79.22, 83.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 77.20, 81.71) -- ( 77.20, 85.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 76.04, 83.73) -- ( 80.08, 83.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 78.06, 81.71) -- ( 78.06, 85.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 76.04, 84.58) -- ( 80.08, 84.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 78.06, 82.56) -- ( 78.06, 86.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 74.26, 82.80) -- ( 78.30, 82.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 76.28, 80.78) -- ( 76.28, 84.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 75.19, 83.73) -- ( 79.22, 83.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 77.20, 81.71) -- ( 77.20, 85.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 76.84, 83.73) -- ( 80.87, 83.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 78.86, 81.71) -- ( 78.86, 85.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 76.84, 83.73) -- ( 80.87, 83.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 78.86, 81.71) -- ( 78.86, 85.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 75.19, 82.80) -- ( 79.22, 82.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 77.20, 80.78) -- ( 77.20, 84.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 75.19, 82.80) -- ( 79.22, 82.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 77.20, 80.78) -- ( 77.20, 84.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 76.04, 83.73) -- ( 80.08, 83.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 78.06, 81.71) -- ( 78.06, 85.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 75.19, 81.80) -- ( 79.22, 81.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 77.20, 79.78) -- ( 77.20, 83.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 76.04, 83.73) -- ( 80.08, 83.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 78.06, 81.71) -- ( 78.06, 85.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 76.84, 84.58) -- ( 80.87, 84.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 78.86, 82.56) -- ( 78.86, 86.60);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.45,101.25) -- (103.31,104.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.45,104.10) -- (103.31,101.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.86,102.67) -- (103.90,102.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.88,100.66) -- (101.88,104.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.35,100.35) -- (103.20,103.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.35,103.20) -- (103.20,100.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.76,101.78) -- (103.79,101.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.78, 99.76) -- (101.78,103.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.46,100.16) -- (104.31,103.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.46,103.02) -- (104.31,100.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.87,101.59) -- (104.90,101.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (102.89, 99.57) -- (102.89,103.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.46,100.90) -- (104.31,103.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.46,103.75) -- (104.31,100.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.87,102.32) -- (104.90,102.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (102.89,100.31) -- (102.89,104.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.46,100.54) -- (104.31,103.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.46,103.39) -- (104.31,100.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.87,101.96) -- (104.90,101.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (102.89, 99.95) -- (102.89,103.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.56,101.07) -- (104.41,103.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.56,103.93) -- (104.41,101.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.96,102.50) -- (105.00,102.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (102.98,100.48) -- (102.98,104.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 98.64, 98.75) -- (101.49,101.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 98.64,101.61) -- (101.49, 98.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 98.05,100.18) -- (102.08,100.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.06, 98.16) -- (100.06,102.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.56, 99.38) -- (103.41,102.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.56,102.23) -- (103.41, 99.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.97,100.81) -- (104.00,100.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.99, 98.79) -- (101.99,102.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.45,100.72) -- (103.31,103.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.45,103.57) -- (103.31,100.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.86,102.14) -- (103.90,102.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.88,100.13) -- (101.88,104.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.56, 99.78) -- (103.41,102.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.56,102.63) -- (103.41, 99.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.97,101.20) -- (104.00,101.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.99, 99.19) -- (101.99,103.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.45, 99.97) -- (103.31,102.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.45,102.83) -- (103.31, 99.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.86,101.40) -- (103.90,101.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.88, 99.38) -- (101.88,103.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.84,100.16) -- (104.69,103.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.84,103.02) -- (104.69,100.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.25,101.59) -- (105.28,101.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (103.27, 99.57) -- (103.27,103.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.35, 99.97) -- (103.20,102.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.35,102.83) -- (103.20, 99.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.76,101.40) -- (103.79,101.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.78, 99.38) -- (101.78,103.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.84,100.35) -- (104.69,103.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.84,103.20) -- (104.69,100.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.25,101.78) -- (105.28,101.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (103.27, 99.76) -- (103.27,103.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.45,100.35) -- (103.31,103.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.45,103.20) -- (103.31,100.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.86,101.78) -- (103.90,101.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.88, 99.76) -- (101.88,103.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.97,100.54) -- (103.82,103.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.97,103.39) -- (103.82,100.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.38,101.96) -- (104.41,101.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (102.39, 99.95) -- (102.39,103.98);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (156.05,157.19) -- (160.08,157.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.06,155.18) -- (158.06,159.21);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.48,156.54) -- (159.52,156.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.50,154.52) -- (157.50,158.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.05,157.32) -- (161.09,157.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.07,155.30) -- (159.07,159.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.64,158.13) -- (161.67,158.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.66,156.11) -- (159.66,160.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.35,158.10) -- (161.38,158.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.37,156.09) -- (159.37,160.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.04,158.54) -- (162.08,158.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.06,156.53) -- (160.06,160.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.31,157.92) -- (161.34,157.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.32,155.90) -- (159.32,159.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.30,158.07) -- (161.34,158.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.32,156.05) -- (159.32,160.09);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.11,158.28) -- (161.15,158.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.13,156.26) -- (159.13,160.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (156.02,156.44) -- (160.06,156.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.04,154.42) -- (158.04,158.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.73,156.06) -- (159.77,156.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.75,154.04) -- (157.75,158.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (156.45,157.17) -- (160.48,157.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.46,155.15) -- (158.46,159.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (156.04,157.55) -- (160.07,157.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (158.05,155.53) -- (158.05,159.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.77,159.16) -- (161.80,159.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.78,157.14) -- (159.78,161.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.99,159.34) -- (162.03,159.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (160.01,157.33) -- (160.01,161.36);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.84,159.31) -- (161.87,159.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (159.85,157.29) -- (159.85,161.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 91.85, 96.12) -- ( 95.89, 96.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.87, 94.10) -- ( 93.87, 98.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.05, 96.71) -- ( 97.09, 96.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.07, 94.70) -- ( 95.07, 98.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.96, 97.28) -- ( 98.00, 97.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.98, 95.26) -- ( 95.98, 99.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.60, 97.28) -- ( 97.64, 97.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.62, 95.26) -- ( 95.62, 99.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.78, 97.82) -- ( 97.82, 97.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.80, 95.80) -- ( 95.80, 99.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.60, 97.28) -- ( 97.64, 97.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.62, 95.26) -- ( 95.62, 99.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.42, 97.00) -- ( 97.46, 97.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.44, 94.98) -- ( 95.44, 99.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.60, 97.28) -- ( 97.64, 97.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.62, 95.26) -- ( 95.62, 99.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.78, 98.82) -- ( 97.82, 98.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.80, 96.80) -- ( 95.80,100.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 91.85, 96.12) -- ( 95.89, 96.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.87, 94.10) -- ( 93.87, 98.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.86, 97.28) -- ( 96.89, 97.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.88, 95.26) -- ( 94.88, 99.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.24, 96.71) -- ( 97.27, 96.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.26, 94.70) -- ( 95.26, 98.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.05, 97.28) -- ( 97.09, 97.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.07, 95.26) -- ( 95.07, 99.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.96, 97.28) -- ( 98.00, 97.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.98, 95.26) -- ( 95.98, 99.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.60, 97.82) -- ( 97.64, 97.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.62, 95.80) -- ( 95.62, 99.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 91.85, 96.42) -- ( 95.89, 96.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.87, 94.40) -- ( 93.87, 98.44);

\path[fill=fillColor,fill opacity=0.50] (102.30, 99.58) --
	(105.15, 99.58) --
	(105.15,102.43) --
	(102.30,102.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (103.88,101.25) --
	(106.74,101.25) --
	(106.74,104.10) --
	(103.88,104.10) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (101.84,100.16) --
	(104.69,100.16) --
	(104.69,103.02) --
	(101.84,103.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (102.99,101.42) --
	(105.84,101.42) --
	(105.84,104.27) --
	(102.99,104.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (101.93,101.07) --
	(104.78,101.07) --
	(104.78,103.93) --
	(101.93,103.93) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (103.49,101.07) --
	(106.34,101.07) --
	(106.34,103.93) --
	(103.49,103.93) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (103.49,101.59) --
	(106.34,101.59) --
	(106.34,104.44) --
	(103.49,104.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (102.91,102.24) --
	(105.76,102.24) --
	(105.76,105.09) --
	(102.91,105.09) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (102.91,102.24) --
	(105.76,102.24) --
	(105.76,105.09) --
	(102.91,105.09) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (102.38, 99.58) --
	(105.24, 99.58) --
	(105.24,102.43) --
	(102.38,102.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (104.64,101.25) --
	(107.49,101.25) --
	(107.49,104.10) --
	(104.64,104.10) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (103.32,102.08) --
	(106.18,102.08) --
	(106.18,104.93) --
	(103.32,104.93) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (102.21, 99.97) --
	(105.06, 99.97) --
	(105.06,102.83) --
	(102.21,102.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (104.27,102.55) --
	(107.12,102.55) --
	(107.12,105.40) --
	(104.27,105.40) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (103.08,102.24) --
	(105.93,102.24) --
	(105.93,105.09) --
	(103.08,105.09) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (103.73,102.40) --
	(106.58,102.40) --
	(106.58,105.25) --
	(103.73,105.25) --
	cycle;

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.63,175.48) -- (174.66,175.48);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.65,173.47) -- (172.65,177.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.23,174.94) -- (174.27,174.94);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.25,172.92) -- (172.25,176.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.77,174.86) -- (173.81,174.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.79,172.84) -- (171.79,176.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.20,174.16) -- (174.24,174.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.22,172.15) -- (172.22,176.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.45,174.04) -- (174.49,174.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (172.47,172.02) -- (172.47,176.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.38,173.64) -- (173.41,173.64);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.39,171.63) -- (171.39,175.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.69,174.82) -- (173.73,174.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.71,172.80) -- (171.71,176.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.71,175.39) -- (173.75,175.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.73,173.37) -- (171.73,177.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.30,174.39) -- (172.34,174.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.32,172.37) -- (170.32,176.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.44,174.16) -- (173.48,174.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.46,172.15) -- (171.46,176.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.70,174.72) -- (172.73,174.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.72,172.70) -- (170.72,176.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.66,173.83) -- (172.69,173.83);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.67,171.81) -- (170.67,175.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.08,173.77) -- (173.12,173.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.10,171.75) -- (171.10,175.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.03,174.85) -- (173.06,174.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.05,172.83) -- (171.05,176.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.70,172.88) -- (172.73,172.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (170.71,170.87) -- (170.71,174.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (169.75,175.10) -- (173.79,175.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (171.77,173.08) -- (171.77,177.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.36,147.82) -- (142.39,147.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.37,145.80) -- (140.37,149.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.88,123.69) -- (126.91,123.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.90,121.67) -- (124.90,125.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.44,113.02) -- (122.48,113.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.46,111.00) -- (120.46,115.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (116.42,112.29) -- (120.46,112.29);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.44,110.27) -- (118.44,114.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.37,120.19) -- (126.40,120.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.38,118.17) -- (124.38,122.21);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.49,114.90) -- (123.52,114.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.50,112.88) -- (121.50,116.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (124.62,124.21) -- (128.65,124.21);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (126.64,122.19) -- (126.64,126.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.08,117.51) -- (123.11,117.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.09,115.50) -- (121.09,119.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (115.67,110.42) -- (119.71,110.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.69,108.41) -- (117.69,112.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.81,113.02) -- (123.84,113.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.83,111.00) -- (121.83,115.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (128.35,131.87) -- (132.38,131.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (130.37,129.85) -- (130.37,133.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (140.20,147.26) -- (144.23,147.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (142.21,145.24) -- (142.21,149.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.87,159.25) -- (155.90,159.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.89,157.23) -- (153.89,161.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.87,158.47) -- (156.91,158.47);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.89,156.46) -- (154.89,160.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.09,160.65) -- (159.12,160.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.10,158.63) -- (157.10,162.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.97,157.40) -- (152.01,157.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.99,155.38) -- (149.99,159.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.16,133.69) -- (154.20,133.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.18,131.67) -- (152.18,135.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.83,132.43) -- (155.86,132.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.85,130.42) -- (153.85,134.45);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.80,132.17) -- (154.84,132.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.82,130.15) -- (152.82,134.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,131.17) -- (190.63,131.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,129.15) -- (188.62,133.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (147.82,130.52) -- (151.85,130.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.84,128.51) -- (149.84,132.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.22,131.39) -- (153.26,131.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.24,129.37) -- (151.24,133.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.47,130.78) -- (153.51,130.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.49,128.77) -- (151.49,132.80);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.66,130.49) -- (152.70,130.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.68,128.48) -- (150.68,132.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.48,131.74) -- (154.51,131.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.50,129.73) -- (152.50,133.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.80,131.08) -- (153.84,131.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.82,129.07) -- (151.82,133.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.98,131.66) -- (154.01,131.66);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.00,129.64) -- (152.00,133.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.55,134.58) -- (154.58,134.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.57,132.57) -- (152.57,136.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,136.65) -- (190.63,136.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,134.64) -- (188.62,138.67);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.33,134.77) -- (154.36,134.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.35,132.75) -- (152.35,136.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.27,134.74) -- (154.30,134.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.28,132.72) -- (152.28,136.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.44,134.41) -- (154.47,134.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.45,132.39) -- (152.45,136.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.06,132.11) -- (154.09,132.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.08,130.10) -- (152.08,134.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.36,132.60) -- (155.39,132.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.38,130.59) -- (153.38,134.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.24,132.52) -- (157.28,132.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.26,130.51) -- (155.26,134.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.61,132.74) -- (155.65,132.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.63,130.73) -- (153.63,134.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.78,132.30) -- (155.81,132.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.80,130.28) -- (153.80,134.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.98,130.24) -- (153.01,130.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.99,128.22) -- (150.99,132.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (149.35,130.35) -- (153.39,130.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.37,128.34) -- (151.37,132.37);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.12,131.63) -- (155.15,131.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.13,129.61) -- (153.13,133.65);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.21,131.26) -- (155.25,131.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.23,129.24) -- (153.23,133.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (148.80,129.59) -- (152.83,129.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.81,127.57) -- (150.81,131.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.16,131.77) -- (155.19,131.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.17,129.75) -- (153.17,133.79);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.22,133.74) -- (155.26,133.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.24,131.72) -- (153.24,135.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,133.58) -- (190.63,133.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,131.56) -- (188.62,135.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.66,133.03) -- (154.69,133.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.68,131.02) -- (152.68,135.05);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.04,132.14) -- (154.08,132.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.06,130.12) -- (152.06,134.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.51,132.56) -- (155.54,132.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.52,130.55) -- (153.52,134.58);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.61,137.82) -- (155.65,137.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.63,135.81) -- (153.63,139.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.52,137.14) -- (156.56,137.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.54,135.12) -- (154.54,139.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.54,135.70) -- (154.58,135.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.56,133.68) -- (152.56,137.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.30,135.89) -- (156.34,135.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.32,133.87) -- (154.32,137.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (186.60,135.01) -- (190.63,135.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (188.62,132.99) -- (188.62,137.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.15,137.25) -- (156.18,137.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.17,135.23) -- (154.17,139.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.14,137.84) -- (155.18,137.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.16,135.82) -- (153.16,139.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.75,134.57) -- (156.79,134.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.77,132.56) -- (154.77,136.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.09,132.85) -- (156.13,132.85);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.11,130.83) -- (154.11,134.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (151.64,132.42) -- (155.68,132.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.66,130.40) -- (153.66,134.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.36,133.82) -- (156.39,133.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.37,131.80) -- (154.37,135.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.16,141.44) -- (157.19,141.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.17,139.43) -- (155.17,143.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (153.04,138.96) -- (157.08,138.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (155.06,136.94) -- (155.06,140.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (150.29,137.54) -- (154.33,137.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.31,135.52) -- (152.31,139.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.35,140.05) -- (156.38,140.05);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.37,138.03) -- (154.37,142.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (152.60,140.72) -- (156.63,140.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (154.61,138.70) -- (154.61,142.74);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.08,128.95) -- (138.11,128.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.09,126.93) -- (136.09,130.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.13,128.61) -- (137.16,128.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.15,126.59) -- (135.15,130.63);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.93,128.88) -- (138.96,128.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.94,126.86) -- (136.94,130.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.90,128.97) -- (137.94,128.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.92,126.95) -- (135.92,130.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.59,126.42) -- (135.63,126.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.61,124.40) -- (133.61,128.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (131.90,126.57) -- (135.93,126.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.91,124.56) -- (133.91,128.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.39,128.41) -- (137.43,128.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.41,126.39) -- (135.41,130.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.66,128.91) -- (137.70,128.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.68,126.90) -- (135.68,130.93);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.03,131.72) -- (138.07,131.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.05,129.70) -- (136.05,133.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.89,129.00) -- (137.93,129.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.91,126.99) -- (135.91,131.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.50,128.22) -- (137.54,128.22);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.52,126.20) -- (135.52,130.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.02,128.88) -- (138.05,128.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.03,126.86) -- (136.03,130.90);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (132.86,128.44) -- (136.90,128.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.88,126.42) -- (134.88,130.46);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (134.10,128.70) -- (138.13,128.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (136.12,126.68) -- (136.12,130.72);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.20,127.60) -- (137.24,127.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.22,125.58) -- (135.22,129.61);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (133.63,128.57) -- (137.67,128.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (135.65,126.55) -- (135.65,130.59);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.77,121.06) -- (104.81,121.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (102.79,119.05) -- (102.79,123.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.07,105.15) -- (104.11,105.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (102.09,103.13) -- (102.09,107.17);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.76,103.98) -- (103.79,103.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.78,101.96) -- (101.78,106.00);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 98.88, 99.75) -- (102.91, 99.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.89, 97.73) -- (100.89,101.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 98.53, 99.97) -- (102.56, 99.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.55, 97.95) -- (100.55,101.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.86,105.96) -- (103.90,105.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.88,103.94) -- (101.88,107.97);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.97,100.18) -- (104.00,100.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.99, 98.16) -- (101.99,102.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.06,100.60) -- (105.10,100.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (103.08, 98.58) -- (103.08,102.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (106.83,100.39) -- (110.86,100.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (108.85, 98.38) -- (108.85,102.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.67,103.18) -- (104.71,103.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (102.69,101.16) -- (102.69,105.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (100.17, 99.52) -- (104.21, 99.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (102.19, 97.50) -- (102.19,101.54);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.79,107.98) -- (105.83,107.98);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (103.81,105.96) -- (103.81,109.99);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (102.23,107.30) -- (106.27,107.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (104.25,105.28) -- (104.25,109.32);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (112.28,108.19) -- (116.31,108.19);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (114.29,106.17) -- (114.29,110.21);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 99.44,100.18) -- (103.47,100.18);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (101.45, 98.16) -- (101.45,102.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (102.06,108.51) -- (106.09,108.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (104.08,106.49) -- (104.08,110.53);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 87.55, 87.84) rectangle ( 90.40, 90.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 87.55, 87.84) -- ( 90.40, 90.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 87.55, 90.69) -- ( 90.40, 87.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.76, 89.86) rectangle ( 91.62, 92.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.76, 89.86) -- ( 91.62, 92.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.76, 92.71) -- ( 91.62, 89.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.47, 88.89) rectangle ( 91.33, 91.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.47, 88.89) -- ( 91.33, 91.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.47, 91.75) -- ( 91.33, 88.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 87.55, 94.69) rectangle ( 90.40, 97.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 87.55, 94.69) -- ( 90.40, 97.55);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 87.55, 97.55) -- ( 90.40, 94.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 89.33, 88.38) rectangle ( 92.18, 91.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 89.33, 88.38) -- ( 92.18, 91.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 89.33, 91.23) -- ( 92.18, 88.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 89.33, 88.38) rectangle ( 92.18, 91.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 89.33, 88.38) -- ( 92.18, 91.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 89.33, 91.23) -- ( 92.18, 88.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 87.86, 87.84) rectangle ( 90.72, 90.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 87.86, 87.84) -- ( 90.72, 90.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 87.86, 90.69) -- ( 90.72, 87.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 86.19, 87.28) rectangle ( 89.04, 90.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 86.19, 87.28) -- ( 89.04, 90.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 86.19, 90.13) -- ( 89.04, 87.28);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.47, 88.38) rectangle ( 91.33, 91.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.47, 88.38) -- ( 91.33, 91.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.47, 91.23) -- ( 91.33, 88.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 89.05, 88.89) rectangle ( 91.90, 91.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 89.05, 88.89) -- ( 91.90, 91.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 89.05, 91.75) -- ( 91.90, 88.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.76, 89.38) rectangle ( 91.62, 92.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.76, 89.38) -- ( 91.62, 92.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.76, 92.24) -- ( 91.62, 89.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.47, 89.38) rectangle ( 91.33, 92.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.47, 89.38) -- ( 91.33, 92.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.47, 92.24) -- ( 91.33, 89.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.17, 91.57) rectangle ( 91.03, 94.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.17, 91.57) -- ( 91.03, 94.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.17, 94.42) -- ( 91.03, 91.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 89.33, 90.31) rectangle ( 92.18, 93.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 89.33, 90.31) -- ( 92.18, 93.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 89.33, 93.16) -- ( 92.18, 90.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.17, 88.89) rectangle ( 91.03, 91.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.17, 88.89) -- ( 91.03, 91.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 88.17, 91.75) -- ( 91.03, 88.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 87.55, 90.31) rectangle ( 90.40, 93.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 87.55, 90.31) -- ( 90.40, 93.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 87.55, 93.16) -- ( 90.40, 90.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.01, 93.41) rectangle ( 96.87, 96.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.01, 93.41) -- ( 96.87, 96.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.01, 96.26) -- ( 96.87, 93.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.01, 93.41) rectangle ( 96.87, 96.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.01, 93.41) -- ( 96.87, 96.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.01, 96.26) -- ( 96.87, 93.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.73, 94.07) rectangle ( 97.58, 96.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.73, 94.07) -- ( 97.58, 96.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.73, 96.92) -- ( 97.58, 94.07);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.07, 93.41) rectangle ( 97.92, 96.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.07, 93.41) -- ( 97.92, 96.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 95.07, 96.26) -- ( 97.92, 93.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.01, 93.06) rectangle ( 96.87, 95.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.01, 93.06) -- ( 96.87, 95.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.01, 95.92) -- ( 96.87, 93.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.55, 93.06) rectangle ( 97.40, 95.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.55, 93.06) -- ( 97.40, 95.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.55, 95.92) -- ( 97.40, 93.06);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.73, 92.71) rectangle ( 97.58, 95.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.73, 92.71) -- ( 97.58, 95.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.73, 95.56) -- ( 97.58, 92.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.01, 92.71) rectangle ( 96.87, 95.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.01, 92.71) -- ( 96.87, 95.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.01, 95.56) -- ( 96.87, 92.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.06, 91.57) rectangle ( 95.91, 94.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.06, 91.57) -- ( 95.91, 94.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.06, 94.42) -- ( 95.91, 91.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.02, 90.31) rectangle ( 94.87, 93.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.02, 90.31) -- ( 94.87, 93.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.02, 93.16) -- ( 94.87, 90.31);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.90, 92.71) rectangle ( 97.75, 95.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.90, 92.71) -- ( 97.75, 95.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.90, 95.56) -- ( 97.75, 92.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.55, 93.41) rectangle ( 97.40, 96.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.55, 93.41) -- ( 97.40, 96.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 94.55, 96.26) -- ( 97.40, 93.41);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.45, 92.71) rectangle ( 96.30, 95.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.45, 92.71) -- ( 96.30, 95.56);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.45, 95.56) -- ( 96.30, 92.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 96.03,101.25) rectangle ( 98.88,104.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 96.03,101.25) -- ( 98.88,104.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 96.03,104.10) -- ( 98.88,101.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.02, 91.96) rectangle ( 94.87, 94.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.02, 91.96) -- ( 94.87, 94.82);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.02, 94.82) -- ( 94.87, 91.96);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.64, 93.06) rectangle ( 96.49, 95.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.64, 93.06) -- ( 96.49, 95.92);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 93.64, 95.92) -- ( 96.49, 93.06);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.18,108.71) -- (123.21,108.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.19,106.70) -- (121.19,110.73);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.49,109.50) -- (123.52,109.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.50,107.49) -- (121.50,111.52);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.96,111.11) -- (123.99,111.11);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.98,109.09) -- (121.98,113.13);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.01,109.60) -- (124.05,109.60);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.03,107.58) -- (122.03,111.62);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.47,110.51) -- (123.50,110.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.49,108.49) -- (121.49,112.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.10,109.02) -- (124.14,109.02);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.12,107.00) -- (122.12,111.03);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.42,108.08) -- (122.46,108.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.44,106.07) -- (120.44,110.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (117.88,107.87) -- (121.92,107.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.90,105.85) -- (119.90,109.88);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.75,109.69) -- (123.79,109.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.77,107.68) -- (121.77,111.71);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.90,107.76) -- (122.93,107.76);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.91,105.74) -- (120.91,109.77);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.66,108.51) -- (123.69,108.51);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (121.68,106.49) -- (121.68,110.53);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.90,111.68) -- (124.93,111.68);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.91,109.66) -- (122.91,113.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.83,110.24) -- (122.87,110.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.85,108.23) -- (120.85,112.26);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.32,115.08) -- (124.36,115.08);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.34,113.06) -- (122.34,117.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.99,112.14) -- (124.03,112.14);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.01,110.12) -- (122.01,114.16);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (120.53,109.69) -- (124.57,109.69);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (122.55,107.68) -- (122.55,111.71);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 71.53, 79.27) rectangle ( 74.38, 82.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 71.53, 79.27) -- ( 74.38, 82.12);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 71.53, 82.12) -- ( 74.38, 79.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 81.86, 83.95) rectangle ( 84.71, 86.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 81.86, 83.95) -- ( 84.71, 86.81);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 81.86, 86.81) -- ( 84.71, 83.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (105.55, 99.38) rectangle (108.41,102.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (105.55, 99.38) -- (108.41,102.23);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (105.55,102.23) -- (108.41, 99.38);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.77,118.30) rectangle (130.62,121.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.77,118.30) -- (130.62,121.15);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.77,121.15) -- (130.62,118.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.80,131.42) rectangle (146.65,134.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.80,131.42) -- (146.65,134.27);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (143.80,134.27) -- (146.65,131.42);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.70,145.40) rectangle (160.56,148.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.70,145.40) -- (160.56,148.25);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (157.70,148.25) -- (160.56,145.40);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.63,155.04) rectangle (171.48,157.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.63,155.04) -- (171.48,157.89);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (168.63,157.89) -- (171.48,155.04);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,167.44) rectangle (190.04,170.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,167.44) -- (190.04,170.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,170.30) -- (190.04,167.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,177.50) rectangle (190.04,180.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,177.50) -- (190.04,180.35);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,180.35) -- (190.04,177.50);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,185.01) rectangle (190.04,187.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,185.01) -- (190.04,187.86);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,187.86) -- (190.04,185.01);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,186.06) rectangle (190.04,188.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,186.06) -- (190.04,188.91);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (187.19,188.91) -- (190.04,186.06);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor,fill opacity=0.50] (188.62,153.94) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,136.00) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,138.50) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,158.63) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,161.09) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,153.48) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,157.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,140.33) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,139.63) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,148.24) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,154.05) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,170.95) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,155.41) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,159.97) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,155.03) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,161.93) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,195.14) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,195.14) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,195.14) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,179.40) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,150.04) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,121.24) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (136.01,139.89) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,118.82) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (176.40,140.88) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (170.18,132.65) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,155.03) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,122.97) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (150.03,133.50) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (188.62,129.33) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (101.34, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 84.29, 86.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 89.29, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 82.75, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 81.59, 95.50) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 86.07, 87.49) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 80.69) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 88.31, 90.81) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 91.74) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 92.17) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 80.69) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 94.84) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 98.76, 94.49) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 84.76, 88.11) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.96, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 88.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 87.25, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 95.17) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 95.81) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 80.69) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20,100.60) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (105.54, 94.49) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 82.18, 90.81) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.96, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 87.97, 91.74) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 84.76, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (103.81, 95.17) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 85.65, 86.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 86.07, 90.32) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (113.55, 92.59) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 83.80, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 82.18, 99.06) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 94.48, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (104.67, 97.55) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 97.61, 99.97) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 81.59, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 86.07, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (112.84,103.34) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (113.78,114.29) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 85.65, 91.28) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.96, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (128.14,118.78) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 99.04, 98.82) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 85.21, 89.27) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 98.06, 97.55) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (101.99,101.59) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 91.55, 86.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 94.49) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 80.69) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 88.11) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 90.81) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 98.34, 93.00) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 82.18,105.01) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 86.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (101.56, 92.17) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 75.27, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (105.39, 88.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.96, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06,103.98) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.96, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (104.75,101.96) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (101.45, 88.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 83.28, 87.49) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (100.43, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 94.28, 88.70) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 89.90, 94.84) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 81.59, 86.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 81.59, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 83.80, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 83.28, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.96, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 95.44, 97.55) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (127.68,116.88) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.26, 96.71) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 95.62, 98.07) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 93.44, 94.49) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 85.21, 90.32) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (133.86,125.45) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.70, 96.42) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (102.98, 96.42) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (118.46,110.42) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (107.38, 97.00) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (114.26,110.06) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 99.82,113.02) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 83.80,100.60) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (112.88, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 84.29, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (111.41,115.25) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (102.89, 99.97) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 85.21, 86.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (112.64,111.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (106.64,108.30) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 83.80, 86.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 82.75, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 81.59, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 88.31, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 88.31, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 95.07, 89.27) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06,100.81) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 90.81) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.96, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] (108.79, 89.81) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 86.87, 86.83) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 82.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 80.96, 83.73) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 85.38) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 81.80) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 86.13) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 84.58) circle (  1.43);

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 83.73) circle (  1.43);
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 84.01) --
	( 80.78, 80.69) --
	( 76.94, 80.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 81.59, 87.60) --
	( 83.51, 84.27) --
	( 79.67, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 84.01) --
	( 81.52, 80.69) --
	( 77.68, 80.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 82.75, 85.94) --
	( 84.67, 82.62) --
	( 80.82, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 80.96, 85.94) --
	( 82.89, 82.62) --
	( 79.04, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 82.91) --
	( 79.12, 79.58) --
	( 75.28, 79.58) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 82.91) --
	( 79.98, 79.58) --
	( 76.14, 79.58) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 81.59, 86.80) --
	( 83.51, 83.47) --
	( 79.67, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 85.02) --
	( 80.78, 81.69) --
	( 76.94, 81.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 82.18, 87.60) --
	( 84.10, 84.27) --
	( 80.26, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 84.29, 89.04) --
	( 86.21, 85.72) --
	( 82.37, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 82.18, 85.02) --
	( 84.10, 81.69) --
	( 80.26, 81.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 84.76, 88.34) --
	( 86.68, 85.02) --
	( 82.84, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 88.65, 89.04) --
	( 90.57, 85.72) --
	( 86.73, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 86.47, 90.33) --
	( 88.40, 87.00) --
	( 84.55, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 83.28, 85.94) --
	( 85.20, 82.62) --
	( 81.36, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 82.18, 86.80) --
	( 84.10, 83.47) --
	( 80.26, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 85.21, 90.33) --
	( 87.13, 87.00) --
	( 83.29, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 96.98,101.28) --
	( 98.90, 97.95) --
	( 95.06, 97.95) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 83.80, 88.34) --
	( 85.72, 85.02) --
	( 81.87, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 91.29, 90.92) --
	( 93.21, 87.59) --
	( 89.37, 87.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (103.72, 97.05) --
	(105.64, 93.73) --
	(101.80, 93.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 98.34,103.23) --
	(100.27, 99.90) --
	( 96.42, 99.90) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (103.45,110.93) --
	(105.37,107.61) --
	(101.53,107.61) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 86.07, 87.60) --
	( 87.99, 84.27) --
	( 84.15, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 94.88, 95.61) --
	( 96.80, 92.28) --
	( 92.96, 92.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (101.34, 94.39) --
	(103.26, 91.06) --
	( 99.42, 91.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (100.31, 97.05) --
	(102.23, 93.73) --
	( 98.39, 93.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 90.48, 90.33) --
	( 92.40, 87.00) --
	( 88.56, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 93.44, 95.22) --
	( 95.36, 91.89) --
	( 91.52, 91.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 94.88,100.55) --
	( 96.80, 97.22) --
	( 92.96, 97.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 93.66,100.55) --
	( 95.58, 97.22) --
	( 91.74, 97.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (109.02, 98.64) --
	(110.94, 95.31) --
	(107.10, 95.31) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (101.12,110.19) --
	(103.04,106.87) --
	( 99.20,106.87) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 87.97, 89.04) --
	( 89.89, 85.72) --
	( 86.05, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (102.09, 99.22) --
	(104.01, 95.89) --
	(100.17, 95.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 96.32,101.28) --
	( 98.24, 97.95) --
	( 94.40, 97.95) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 97.30,115.72) --
	( 99.22,112.39) --
	( 95.38,112.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (102.89,119.04) --
	(104.81,115.72) --
	(100.97,115.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (101.23,103.23) --
	(103.15, 99.90) --
	( 99.31, 99.90) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (102.98,104.00) --
	(104.90,100.67) --
	(101.06,100.67) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (104.25,122.25) --
	(106.17,118.93) --
	(102.33,118.93) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.81,122.33) --
	(120.73,119.00) --
	(116.89,119.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (103.63,112.37) --
	(105.55,109.05) --
	(101.71,109.05) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.14,139.55) --
	(121.07,136.22) --
	(117.22,136.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.40,122.41) --
	(115.32,119.08) --
	(111.48,119.08) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.82,109.05) --
	(113.74,105.72) --
	(109.90,105.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 96.15, 97.39) --
	( 98.07, 94.06) --
	( 94.23, 94.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 99.17,110.41) --
	(101.09,107.08) --
	( 97.25,107.08) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (107.18,111.91) --
	(109.10,108.59) --
	(105.26,108.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (105.62,124.86) --
	(107.54,121.53) --
	(103.70,121.53) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.55,121.46) --
	(112.48,118.13) --
	(108.63,118.13) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (105.62,109.64) --
	(107.54,106.31) --
	(103.70,106.31) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.51,130.07) --
	(118.43,126.74) --
	(114.59,126.74) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.78,103.02) --
	(115.70, 99.70) --
	(111.86, 99.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (107.05,122.85) --
	(108.97,119.53) --
	(105.13,119.53) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.23,114.66) --
	(113.15,111.33) --
	(109.31,111.33) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.99,124.42) --
	(117.91,121.09) --
	(114.07,121.09) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.92,124.70) --
	(116.84,121.37) --
	(113.00,121.37) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.05,127.18) --
	(135.97,123.85) --
	(132.12,123.85) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.36,147.84) --
	(119.28,144.51) --
	(115.44,144.51) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.88,127.21) --
	(121.80,123.88) --
	(117.95,123.88) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.99,120.78) --
	(116.91,117.45) --
	(113.07,117.45) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.81,118.31) --
	(122.73,114.98) --
	(118.89,114.98) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (127.50,127.79) --
	(129.42,124.46) --
	(125.58,124.46) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.87,126.16) --
	(136.79,122.83) --
	(132.95,122.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.18,123.80) --
	(113.10,120.47) --
	(109.26,120.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.54,124.92) --
	(122.46,121.59) --
	(118.62,121.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (107.31,117.53) --
	(109.23,114.20) --
	(105.39,114.20) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.50,130.40) --
	(112.43,127.07) --
	(108.58,127.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.79,124.03) --
	(121.71,120.70) --
	(117.87,120.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.30,131.27) --
	(125.22,127.95) --
	(121.38,127.95) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.67,134.92) --
	(122.59,131.60) --
	(118.75,131.60) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.83,124.23) --
	(122.75,120.90) --
	(118.91,120.90) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.05,146.39) --
	(131.97,143.07) --
	(128.13,143.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (152.47,137.19) --
	(154.39,133.86) --
	(150.54,133.86) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.68,131.54) --
	(118.60,128.21) --
	(114.76,128.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.12,139.34) --
	(132.04,136.01) --
	(128.20,136.01) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.08,119.69) --
	(120.00,116.36) --
	(116.16,116.36) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.99,132.01) --
	(131.91,128.68) --
	(128.06,128.68) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (132.82,144.76) --
	(134.74,141.43) --
	(130.90,141.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.06,140.30) --
	(134.98,136.97) --
	(131.14,136.97) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (158.09,147.36) --
	(160.01,144.03) --
	(156.17,144.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.31,135.59) --
	(139.23,132.26) --
	(135.39,132.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.34,153.77) --
	(135.26,150.45) --
	(131.41,150.45) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.70,162.28) --
	(146.62,158.96) --
	(142.78,158.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.54,135.72) --
	(144.47,132.39) --
	(140.62,132.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (132.99,146.84) --
	(134.91,143.51) --
	(131.07,143.51) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.71,137.16) --
	(142.63,133.83) --
	(138.78,133.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (160.15,126.85) --
	(162.07,123.52) --
	(158.23,123.52) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 85.02) --
	( 80.78, 81.69) --
	( 76.94, 81.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 84.01) --
	( 79.12, 80.69) --
	( 75.28, 80.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 87.60) --
	( 82.22, 84.27) --
	( 78.38, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 84.01) --
	( 79.98, 80.69) --
	( 76.14, 80.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 90.33) --
	( 81.52, 87.00) --
	( 77.68, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 84.01) --
	( 79.12, 80.69) --
	( 75.28, 80.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 87.60) --
	( 82.22, 84.27) --
	( 78.38, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 86.80) --
	( 78.20, 83.47) --
	( 74.36, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 85.02) --
	( 78.20, 81.69) --
	( 74.36, 81.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 85.02) --
	( 81.52, 81.69) --
	( 77.68, 81.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 81.59, 88.34) --
	( 83.51, 85.02) --
	( 79.67, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 85.94) --
	( 79.98, 82.62) --
	( 76.14, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 82.75, 88.34) --
	( 84.67, 85.02) --
	( 80.82, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 83.80, 90.92) --
	( 85.72, 87.59) --
	( 81.87, 87.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 82.18, 88.34) --
	( 84.10, 85.02) --
	( 80.26, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 85.02) --
	( 81.52, 81.69) --
	( 77.68, 81.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 82.75, 87.60) --
	( 84.67, 84.27) --
	( 80.82, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 85.94) --
	( 81.52, 82.62) --
	( 77.68, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 82.75, 87.60) --
	( 84.67, 84.27) --
	( 80.82, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 81.59, 88.34) --
	( 83.51, 85.02) --
	( 79.67, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 80.96, 86.80) --
	( 82.89, 83.47) --
	( 79.04, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 83.28, 88.34) --
	( 85.20, 85.02) --
	( 81.36, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 82.75, 86.80) --
	( 84.67, 83.47) --
	( 80.82, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 85.65, 89.70) --
	( 87.57, 86.38) --
	( 83.73, 86.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 86.07, 90.33) --
	( 87.99, 87.00) --
	( 84.15, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 85.21, 90.92) --
	( 87.13, 87.59) --
	( 83.29, 87.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 85.21, 93.50) --
	( 87.13, 90.17) --
	( 83.29, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 84.76,101.28) --
	( 86.68, 97.95) --
	( 82.84, 97.95) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 87.61, 92.02) --
	( 89.53, 88.70) --
	( 85.69, 88.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 87.25, 95.99) --
	( 89.17, 92.66) --
	( 85.32, 92.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 84.76, 91.49) --
	( 86.68, 88.16) --
	( 82.84, 88.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 88.97, 88.34) --
	( 90.89, 85.02) --
	( 87.05, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 86.47, 89.04) --
	( 88.40, 85.72) --
	( 84.55, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 83.80, 86.80) --
	( 85.72, 83.47) --
	( 81.87, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 83.80, 87.60) --
	( 85.72, 84.27) --
	( 81.87, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 86.47, 91.49) --
	( 88.40, 88.16) --
	( 84.55, 88.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 83.80, 89.04) --
	( 85.72, 85.72) --
	( 81.87, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 86.47, 93.50) --
	( 88.40, 90.17) --
	( 84.55, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 88.65, 90.92) --
	( 90.57, 87.59) --
	( 86.73, 87.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 89.29, 90.92) --
	( 91.21, 87.59) --
	( 87.37, 87.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 86.87, 89.70) --
	( 88.79, 86.38) --
	( 84.95, 86.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 88.97, 90.33) --
	( 90.89, 87.00) --
	( 87.05, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 90.75, 92.02) --
	( 92.68, 88.70) --
	( 88.83, 88.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 89.29, 97.71) --
	( 91.21, 94.39) --
	( 87.37, 94.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 91.55, 92.54) --
	( 93.47, 89.21) --
	( 89.63, 89.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 93.66, 89.04) --
	( 95.58, 85.72) --
	( 91.74, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 88.97, 93.03) --
	( 90.89, 89.70) --
	( 87.05, 89.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 92.05, 92.54) --
	( 93.98, 89.21) --
	( 90.13, 89.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 89.90, 97.05) --
	( 91.82, 93.73) --
	( 87.98, 93.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 88.97, 92.54) --
	( 90.89, 89.21) --
	( 87.05, 89.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 90.75, 92.54) --
	( 92.68, 89.21) --
	( 88.83, 89.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 89.90, 93.50) --
	( 91.82, 90.17) --
	( 87.98, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 89.90, 95.61) --
	( 91.82, 92.28) --
	( 87.98, 92.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 94.48, 95.22) --
	( 96.41, 91.89) --
	( 92.56, 91.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 90.48, 90.33) --
	( 92.40, 87.00) --
	( 88.56, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 88.65, 89.70) --
	( 90.57, 86.38) --
	( 86.73, 86.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 88.31, 93.50) --
	( 90.23, 90.17) --
	( 86.39, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 92.54, 94.39) --
	( 94.46, 91.06) --
	( 90.61, 91.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 90.48, 93.50) --
	( 92.40, 90.17) --
	( 88.56, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 92.77, 98.93) --
	( 94.69, 95.60) --
	( 90.85, 95.60) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 93.87,100.03) --
	( 95.79, 96.71) --
	( 91.95, 96.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 94.08, 95.61) --
	( 96.00, 92.28) --
	( 92.16, 92.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 97.91,100.80) --
	( 99.83, 97.47) --
	( 95.99, 97.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (103.17, 98.03) --
	(105.09, 94.70) --
	(101.25, 94.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 91.29, 93.50) --
	( 93.21, 90.17) --
	( 89.37, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 92.30, 93.50) --
	( 94.22, 90.17) --
	( 90.38, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 93.44,108.17) --
	( 95.36,104.85) --
	( 91.52,104.85) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 93.22, 95.22) --
	( 95.14, 91.89) --
	( 91.30, 91.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 98.34,103.42) --
	(100.27,100.10) --
	( 96.42,100.10) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (101.88, 94.39) --
	(103.80, 91.06) --
	( 99.96, 91.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 97.46,114.59) --
	( 99.38,111.26) --
	( 95.53,111.26) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 93.66,108.17) --
	( 95.58,104.85) --
	( 91.74,104.85) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (108.01,102.18) --
	(109.93, 98.86) --
	(106.09, 98.86) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 96.49,106.50) --
	( 98.41,103.17) --
	( 94.57,103.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 94.28,100.55) --
	( 96.21, 97.22) --
	( 92.36, 97.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 95.98, 95.22) --
	( 97.90, 91.89) --
	( 94.06, 91.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 94.88,100.80) --
	( 96.80, 97.47) --
	( 92.96, 97.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 95.80,110.52) --
	( 97.72,107.19) --
	( 93.88,107.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (106.21,109.17) --
	(108.13,105.84) --
	(104.29,105.84) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 94.48, 95.22) --
	( 96.41, 91.89) --
	( 92.56, 91.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (115.26,124.73) --
	(117.18,121.41) --
	(113.34,121.41) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.46,122.82) --
	(139.38,119.49) --
	(135.54,119.49) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 97.91,124.45) --
	( 99.83,121.12) --
	( 95.99,121.12) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 95.26, 96.71) --
	( 97.18, 93.38) --
	( 93.33, 93.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.36, 98.03) --
	(121.28, 94.70) --
	(117.43, 94.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (100.55,104.72) --
	(102.47,101.39) --
	( 98.62,101.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 96.82,107.51) --
	( 98.74,104.18) --
	( 94.90,104.18) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (100.43,102.40) --
	(102.35, 99.07) --
	( 98.51, 99.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.35,103.81) --
	(112.28,100.48) --
	(108.43,100.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (103.08,110.30) --
	(105.00,106.97) --
	(101.16,106.97) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.25,110.30) --
	(112.17,106.97) --
	(108.33,106.97) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (111.64,101.74) --
	(113.56, 98.41) --
	(109.72, 98.41) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (108.01,127.79) --
	(109.93,124.46) --
	(106.09,124.46) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (120.40,110.83) --
	(122.32,107.50) --
	(118.47,107.50) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (104.91,108.04) --
	(106.83,104.72) --
	(102.99,104.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 96.98,108.04) --
	( 98.90,104.72) --
	( 95.06,104.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.33,131.40) --
	(125.25,128.07) --
	(121.41,128.07) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (109.57,125.48) --
	(111.49,122.15) --
	(107.65,122.15) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (102.79,117.06) --
	(104.71,113.73) --
	(100.87,113.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 96.82,126.46) --
	( 98.74,123.13) --
	( 94.90,123.13) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (100.43,122.25) --
	(102.35,118.93) --
	( 98.51,118.93) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.29,122.10) --
	(116.21,118.77) --
	(112.37,118.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.82,133.11) --
	(115.74,129.78) --
	(111.90,129.78) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.14,130.03) --
	(126.06,126.70) --
	(122.22,126.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.17,142.14) --
	(132.09,138.81) --
	(128.25,138.81) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (103.54,121.42) --
	(105.46,118.09) --
	(101.62,118.09) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.05,134.31) --
	(122.97,130.98) --
	(119.13,130.98) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.66,123.28) --
	(120.58,119.95) --
	(116.74,119.95) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (121.31,118.63) --
	(123.23,115.30) --
	(119.39,115.30) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.89,118.42) --
	(115.81,115.09) --
	(111.97,115.09) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.12,121.66) --
	(118.04,118.34) --
	(114.19,118.34) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.51,125.45) --
	(114.43,122.13) --
	(110.59,122.13) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (132.31,114.44) --
	(134.23,111.11) --
	(130.39,111.11) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (119.05,137.77) --
	(120.97,134.45) --
	(117.13,134.45) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.64,124.45) --
	(125.56,121.12) --
	(121.72,121.12) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.93,132.82) --
	(115.85,129.49) --
	(112.01,129.49) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (109.41,120.65) --
	(111.33,117.32) --
	(107.49,117.32) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.02,145.01) --
	(145.95,141.68) --
	(142.10,141.68) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (124.28,129.07) --
	(126.20,125.74) --
	(122.36,125.74) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (107.95,113.33) --
	(109.87,110.00) --
	(106.03,110.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (146.12,143.05) --
	(148.05,139.72) --
	(144.20,139.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.26,159.10) --
	(131.18,155.77) --
	(127.34,155.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (132.83,115.02) --
	(134.75,111.70) --
	(130.91,111.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (147.57,132.20) --
	(149.50,128.88) --
	(145.65,128.88) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (113.20,126.54) --
	(115.12,123.21) --
	(111.28,123.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.80,156.09) --
	(132.72,152.76) --
	(128.87,152.76) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.46,157.29) --
	(146.38,153.96) --
	(142.54,153.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (126.88,120.52) --
	(128.80,117.19) --
	(124.96,117.19) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.93,129.55) --
	(120.85,126.22) --
	(117.01,126.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (118.98,116.75) --
	(120.90,113.43) --
	(117.06,113.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (116.91,122.06) --
	(118.84,118.73) --
	(114.99,118.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (153.37,154.01) --
	(155.29,150.68) --
	(151.45,150.68) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.28,128.48) --
	(119.20,125.16) --
	(115.36,125.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.57,138.18) --
	(146.50,134.85) --
	(142.65,134.85) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (146.27,127.60) --
	(148.19,124.27) --
	(144.35,124.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (142.16,158.89) --
	(144.08,155.56) --
	(140.24,155.56) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.78,168.92) --
	(147.70,165.60) --
	(143.86,165.60) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.33,144.80) --
	(132.25,141.47) --
	(128.41,141.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.99,168.96) --
	(136.91,165.63) --
	(133.07,165.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.59,137.56) --
	(131.51,134.23) --
	(127.67,134.23) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (153.06,135.72) --
	(154.98,132.39) --
	(151.14,132.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.32,151.91) --
	(138.24,148.58) --
	(134.40,148.58) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (153.70,157.32) --
	(155.62,153.99) --
	(151.78,153.99) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (156.91,154.24) --
	(158.83,150.91) --
	(154.99,150.91) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.88,135.52) --
	(135.80,132.20) --
	(131.96,132.20) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (146.77,152.77) --
	(148.69,149.44) --
	(144.85,149.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (139.83,165.16) --
	(141.75,161.83) --
	(137.91,161.83) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.16,125.54) --
	(124.08,122.21) --
	(120.24,122.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (151.15,139.83) --
	(153.07,136.50) --
	(149.23,136.50) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (134.45,149.10) --
	(136.37,145.77) --
	(132.53,145.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 86.80) --
	( 81.52, 83.47) --
	( 77.68, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 85.94) --
	( 79.98, 82.62) --
	( 76.14, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 76.28, 85.94) --
	( 78.20, 82.62) --
	( 74.36, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 93.03) --
	( 80.78, 89.70) --
	( 76.94, 89.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 77.20, 85.94) --
	( 79.12, 82.62) --
	( 75.28, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 85.94) --
	( 79.98, 82.62) --
	( 76.14, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 85.94) --
	( 79.98, 82.62) --
	( 76.14, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 85.02) --
	( 79.98, 81.69) --
	( 76.14, 81.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 78.86, 85.94) --
	( 80.78, 82.62) --
	( 76.94, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 83.80, 85.94) --
	( 85.72, 82.62) --
	( 81.87, 82.62) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 87.60) --
	( 82.22, 84.27) --
	( 78.38, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 80.96, 86.80) --
	( 82.89, 83.47) --
	( 79.04, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 79.60, 86.80) --
	( 81.52, 83.47) --
	( 77.68, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 80.96, 94.81) --
	( 82.89, 91.48) --
	( 79.04, 91.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 99.22) --
	( 82.22, 95.89) --
	( 78.38, 95.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 78.06, 86.80) --
	( 79.98, 83.47) --
	( 76.14, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 80.30, 86.80) --
	( 82.22, 83.47) --
	( 78.38, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 81.59, 86.80) --
	( 83.51, 83.47) --
	( 79.67, 83.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 80.96, 87.60) --
	( 82.89, 84.27) --
	( 79.04, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 80.96, 87.60) --
	( 82.89, 84.27) --
	( 79.04, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 83.80, 89.04) --
	( 85.72, 85.72) --
	( 81.87, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 86.87, 93.50) --
	( 88.79, 90.17) --
	( 84.95, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 89.90, 88.34) --
	( 91.82, 85.02) --
	( 87.98, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 82.18, 89.04) --
	( 84.10, 85.72) --
	( 80.26, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 81.59, 87.60) --
	( 83.51, 84.27) --
	( 79.67, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 83.80, 90.92) --
	( 85.72, 87.59) --
	( 81.87, 87.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 82.75, 88.34) --
	( 84.67, 85.02) --
	( 80.82, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 84.76, 89.04) --
	( 86.68, 85.72) --
	( 82.84, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 82.18, 89.04) --
	( 84.10, 85.72) --
	( 80.26, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 82.18, 87.60) --
	( 84.10, 84.27) --
	( 80.26, 84.27) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 84.76, 88.34) --
	( 86.68, 85.02) --
	( 82.84, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 86.07, 92.02) --
	( 87.99, 88.70) --
	( 84.15, 88.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 85.21, 89.70) --
	( 87.13, 86.38) --
	( 83.29, 86.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 86.07, 89.70) --
	( 87.99, 86.38) --
	( 84.15, 86.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 83.80, 89.04) --
	( 85.72, 85.72) --
	( 81.87, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 83.80, 89.04) --
	( 85.72, 85.72) --
	( 81.87, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 84.76, 90.33) --
	( 86.68, 87.00) --
	( 82.84, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 89.90, 92.02) --
	( 91.82, 88.70) --
	( 87.98, 88.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 84.29, 92.02) --
	( 86.21, 88.70) --
	( 82.37, 88.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 82.18, 88.34) --
	( 84.10, 85.02) --
	( 80.26, 85.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 87.97, 92.54) --
	( 89.89, 89.21) --
	( 86.05, 89.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 85.21, 89.70) --
	( 87.13, 86.38) --
	( 83.29, 86.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 84.29, 90.33) --
	( 86.21, 87.00) --
	( 82.37, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 88.97, 94.81) --
	( 90.89, 91.48) --
	( 87.05, 91.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 85.21, 89.04) --
	( 87.13, 85.72) --
	( 83.29, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 86.47, 90.33) --
	( 88.40, 87.00) --
	( 84.55, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 85.21, 89.04) --
	( 87.13, 85.72) --
	( 83.29, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 91.29,100.80) --
	( 93.21, 97.47) --
	( 89.37, 97.47) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 85.65, 89.04) --
	( 87.57, 85.72) --
	( 83.73, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 86.47, 89.04) --
	( 88.40, 85.72) --
	( 84.55, 85.72) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 86.47, 98.93) --
	( 88.40, 95.60) --
	( 84.55, 95.60) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 86.07, 90.33) --
	( 87.99, 87.00) --
	( 84.15, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 87.25, 90.33) --
	( 89.17, 87.00) --
	( 85.32, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 87.61, 94.39) --
	( 89.53, 91.06) --
	( 85.69, 91.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 86.87, 94.39) --
	( 88.79, 91.06) --
	( 84.95, 91.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 86.87, 90.33) --
	( 88.79, 87.00) --
	( 84.95, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 87.25, 90.92) --
	( 89.17, 87.59) --
	( 85.32, 87.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 92.05, 91.49) --
	( 93.98, 88.16) --
	( 90.13, 88.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 87.61, 90.33) --
	( 89.53, 87.00) --
	( 85.69, 87.00) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 87.97, 91.49) --
	( 89.89, 88.16) --
	( 86.05, 88.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 91.03, 91.49) --
	( 92.95, 88.16) --
	( 89.11, 88.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 88.31, 92.02) --
	( 90.23, 88.70) --
	( 86.39, 88.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 87.61, 90.92) --
	( 89.53, 87.59) --
	( 85.69, 87.59) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 89.60,104.89) --
	( 91.52,101.57) --
	( 87.68,101.57) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 87.61, 93.50) --
	( 89.53, 90.17) --
	( 85.69, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 88.31, 92.02) --
	( 90.23, 88.70) --
	( 86.39, 88.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 88.65, 91.49) --
	( 90.57, 88.16) --
	( 86.73, 88.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 87.25, 92.02) --
	( 89.17, 88.70) --
	( 85.32, 88.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 87.25, 91.49) --
	( 89.17, 88.16) --
	( 85.32, 88.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 95.26, 91.49) --
	( 97.18, 88.16) --
	( 93.33, 88.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 90.48, 93.03) --
	( 92.40, 89.70) --
	( 88.56, 89.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 90.75, 93.03) --
	( 92.68, 89.70) --
	( 88.83, 89.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 90.75, 93.50) --
	( 92.68, 90.17) --
	( 88.83, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 88.97, 91.49) --
	( 90.89, 88.16) --
	( 87.05, 88.16) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 87.97, 92.02) --
	( 89.89, 88.70) --
	( 86.05, 88.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 91.55, 93.03) --
	( 93.47, 89.70) --
	( 89.63, 89.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 89.90, 92.54) --
	( 91.82, 89.21) --
	( 87.98, 89.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 89.90, 92.54) --
	( 91.82, 89.21) --
	( 87.98, 89.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 88.97, 93.50) --
	( 90.89, 90.17) --
	( 87.05, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 90.19, 93.95) --
	( 92.11, 90.63) --
	( 88.27, 90.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 90.75, 93.03) --
	( 92.68, 89.70) --
	( 88.83, 89.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 91.55,100.55) --
	( 93.47, 97.22) --
	( 89.63, 97.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 92.54,104.18) --
	( 94.46,100.85) --
	( 90.61,100.85) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 97.30, 93.03) --
	( 99.22, 89.70) --
	( 95.38, 89.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 90.48, 93.95) --
	( 92.40, 90.63) --
	( 88.56, 90.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 92.05, 94.39) --
	( 93.98, 91.06) --
	( 90.13, 91.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 90.48, 93.03) --
	( 92.40, 89.70) --
	( 88.56, 89.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 91.29, 93.03) --
	( 93.21, 89.70) --
	( 89.37, 89.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 91.29, 93.95) --
	( 93.21, 90.63) --
	( 89.37, 90.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 92.54, 93.95) --
	( 94.46, 90.63) --
	( 90.61, 90.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 93.22, 94.81) --
	( 95.14, 91.48) --
	( 91.30, 91.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 93.44, 95.22) --
	( 95.36, 91.89) --
	( 91.52, 91.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 94.48, 94.81) --
	( 96.41, 91.48) --
	( 92.56, 91.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 94.28, 94.39) --
	( 96.21, 91.06) --
	( 92.36, 91.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 92.05, 93.50) --
	( 93.98, 90.17) --
	( 90.13, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 94.68,102.61) --
	( 96.60, 99.28) --
	( 92.76, 99.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 91.81, 93.50) --
	( 93.73, 90.17) --
	( 89.88, 90.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 94.48, 94.81) --
	( 96.41, 91.48) --
	( 92.56, 91.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 92.05, 93.03) --
	( 93.98, 89.70) --
	( 90.13, 89.70) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 94.08, 93.95) --
	( 96.00, 90.63) --
	( 92.16, 90.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 92.54, 95.61) --
	( 94.46, 92.28) --
	( 90.61, 92.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 96.15, 94.81) --
	( 98.07, 91.48) --
	( 94.23, 91.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 94.48, 95.61) --
	( 96.41, 92.28) --
	( 92.56, 92.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 95.07, 95.99) --
	( 96.99, 92.66) --
	( 93.15, 92.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 94.08, 94.39) --
	( 96.00, 91.06) --
	( 92.16, 91.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 98.06, 96.35) --
	( 99.98, 93.03) --
	( 96.13, 93.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 94.48, 93.95) --
	( 96.41, 90.63) --
	( 92.56, 90.63) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 95.26, 97.39) --
	( 97.18, 94.06) --
	( 93.33, 94.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 96.32, 94.81) --
	( 98.24, 91.48) --
	( 94.40, 91.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 94.88, 94.81) --
	( 96.80, 91.48) --
	( 92.96, 91.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 95.80, 96.71) --
	( 97.72, 93.38) --
	( 93.88, 93.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 95.98, 97.05) --
	( 97.90, 93.73) --
	( 94.06, 93.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (103.63,106.04) --
	(105.55,102.71) --
	(101.71,102.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 95.07, 95.22) --
	( 96.99, 91.89) --
	( 93.15, 91.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 95.26, 95.61) --
	( 97.18, 92.28) --
	( 93.33, 92.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 95.98, 96.71) --
	( 97.90, 93.38) --
	( 94.06, 93.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 94.88, 98.34) --
	( 96.80, 95.01) --
	( 92.96, 95.01) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 96.82, 95.99) --
	( 98.74, 92.66) --
	( 94.90, 92.66) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 97.30, 95.61) --
	( 99.22, 92.28) --
	( 95.38, 92.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 94.68,100.55) --
	( 96.60, 97.22) --
	( 92.76, 97.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 97.91, 95.61) --
	( 99.83, 92.28) --
	( 95.99, 92.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 96.98, 96.71) --
	( 98.90, 93.38) --
	( 95.06, 93.38) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 96.32, 97.05) --
	( 98.24, 93.73) --
	( 94.40, 93.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 97.76, 97.39) --
	( 99.68, 94.06) --
	( 95.84, 94.06) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 97.46,100.03) --
	( 99.38, 96.71) --
	( 95.53, 96.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 97.14, 97.71) --
	( 99.06, 94.39) --
	( 95.22, 94.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 97.46, 99.22) --
	( 99.38, 95.89) --
	( 95.53, 95.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 97.30, 95.61) --
	( 99.22, 92.28) --
	( 95.38, 92.28) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 97.14, 97.71) --
	( 99.06, 94.39) --
	( 95.22, 94.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 97.30, 97.05) --
	( 99.22, 93.73) --
	( 95.38, 93.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 98.76, 97.71) --
	(100.69, 94.39) --
	( 96.84, 94.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (101.01, 99.22) --
	(102.93, 95.89) --
	( 99.09, 95.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 96.82, 97.05) --
	( 98.74, 93.73) --
	( 94.90, 93.73) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (100.43,100.55) --
	(102.35, 97.22) --
	( 98.51, 97.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 99.17, 99.22) --
	(101.09, 95.89) --
	( 97.25, 95.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (100.06,104.18) --
	(101.99,100.85) --
	( 98.14,100.85) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] ( 98.90, 97.71) --
	(100.82, 94.39) --
	( 96.98, 94.39) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (102.98,117.76) --
	(104.90,114.43) --
	(101.06,114.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (100.19,102.18) --
	(102.11, 98.86) --
	( 98.27, 98.86) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (104.42,131.31) --
	(106.34,127.98) --
	(102.50,127.98) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (100.55, 99.50) --
	(102.47, 96.17) --
	( 98.62, 96.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.51,109.17) --
	(116.43,105.84) --
	(112.58,105.84) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (101.88, 99.50) --
	(103.80, 96.17) --
	( 99.96, 96.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (100.55,102.18) --
	(102.47, 98.86) --
	( 98.62, 98.86) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (103.81,103.81) --
	(105.73,100.48) --
	(101.89,100.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (101.56,101.51) --
	(103.48, 98.18) --
	( 99.64, 98.18) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (101.78,108.17) --
	(103.70,104.85) --
	( 99.85,104.85) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (106.21, 99.50) --
	(108.13, 96.17) --
	(104.29, 96.17) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (102.29,107.64) --
	(104.22,104.31) --
	(100.37,104.31) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (100.06,100.03) --
	(101.99, 96.71) --
	( 98.14, 96.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (107.38,116.44) --
	(109.30,113.11) --
	(105.46,113.11) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (125.67,112.55) --
	(127.59,109.22) --
	(123.75,109.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (125.55,107.37) --
	(127.47,104.04) --
	(123.63,104.04) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (128.92,123.46) --
	(130.84,120.13) --
	(127.00,120.13) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (104.75,105.56) --
	(106.67,102.24) --
	(102.83,102.24) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (130.09,139.26) --
	(132.01,135.93) --
	(128.17,135.93) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (107.76,109.29) --
	(109.68,105.96) --
	(105.84,105.96) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (109.08,106.04) --
	(111.00,102.71) --
	(107.15,102.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.98,137.64) --
	(124.90,134.32) --
	(121.06,134.32) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (106.77,132.56) --
	(108.70,129.23) --
	(104.85,129.23) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (128.29,143.14) --
	(130.21,139.82) --
	(126.37,139.82) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.44,112.10) --
	(116.36,108.77) --
	(112.51,108.77) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.99,122.78) --
	(116.91,119.45) --
	(113.07,119.45) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.72,110.30) --
	(114.64,106.97) --
	(110.80,106.97) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.55,108.81) --
	(112.48,105.48) --
	(108.63,105.48) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (110.00,119.15) --
	(111.92,115.82) --
	(108.07,115.82) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (117.00,132.22) --
	(118.92,128.89) --
	(115.08,128.89) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (123.33,150.73) --
	(125.25,147.40) --
	(121.41,147.40) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (112.55,118.53) --
	(114.47,115.20) --
	(110.63,115.20) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (114.85,114.88) --
	(116.77,111.55) --
	(112.93,111.55) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (136.95,147.34) --
	(138.87,144.02) --
	(135.03,144.02) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,160.17) --
	(190.54,156.84) --
	(186.69,156.84) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (131.35,124.76) --
	(133.27,121.44) --
	(129.43,121.44) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (144.32,153.26) --
	(146.24,149.93) --
	(142.40,149.93) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (133.49,131.82) --
	(135.41,128.50) --
	(131.56,128.50) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (128.38,141.19) --
	(130.31,137.87) --
	(126.46,137.87) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (171.86,182.15) --
	(173.78,178.82) --
	(169.94,178.82) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (125.05,141.73) --
	(126.97,138.40) --
	(123.13,138.40) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (122.95,143.67) --
	(124.87,140.35) --
	(121.03,140.35) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (125.12,117.76) --
	(127.04,114.43) --
	(123.20,114.43) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (149.96,166.04) --
	(151.89,162.71) --
	(148.04,162.71) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (167.16,180.65) --
	(169.08,177.32) --
	(165.23,177.32) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (129.50,127.55) --
	(131.42,124.22) --
	(127.57,124.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (151.27,132.63) --
	(153.19,129.31) --
	(149.35,129.31) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,197.36) --
	(190.54,194.03) --
	(186.69,194.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (150.10,149.98) --
	(152.02,146.65) --
	(148.17,146.65) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (151.08,145.91) --
	(153.00,142.58) --
	(149.16,142.58) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,144.41) --
	(190.54,141.08) --
	(186.69,141.08) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (160.49,197.36) --
	(162.41,194.03) --
	(158.57,194.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (171.53,169.55) --
	(173.45,166.22) --
	(169.60,166.22) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (145.61,176.67) --
	(147.53,173.34) --
	(143.69,173.34) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (169.57,138.70) --
	(171.49,135.37) --
	(167.65,135.37) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (153.98,141.47) --
	(155.90,138.14) --
	(152.06,138.14) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (172.69,175.25) --
	(174.61,171.92) --
	(170.77,171.92) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (154.60,163.06) --
	(156.52,159.74) --
	(152.68,159.74) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,197.36) --
	(190.54,194.03) --
	(186.69,194.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (178.40,181.08) --
	(180.32,177.75) --
	(176.48,177.75) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,197.36) --
	(190.54,194.03) --
	(186.69,194.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (163.96,171.27) --
	(165.88,167.94) --
	(162.04,167.94) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (172.39,197.36) --
	(174.31,194.03) --
	(170.47,194.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,197.36) --
	(190.54,194.03) --
	(186.69,194.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,188.01) --
	(190.54,184.69) --
	(186.69,184.69) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,197.36) --
	(190.54,194.03) --
	(186.69,194.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (137.08,178.99) --
	(139.00,175.67) --
	(135.16,175.67) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (159.87,197.36) --
	(161.79,194.03) --
	(157.95,194.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,183.54) --
	(190.54,180.21) --
	(186.69,180.21) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,197.36) --
	(190.54,194.03) --
	(186.69,194.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (140.40,158.67) --
	(142.32,155.34) --
	(138.48,155.34) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (157.40,197.36) --
	(159.32,194.03) --
	(155.47,194.03) --
	cycle;

\path[fill=fillColor,fill opacity=0.50] (188.62,197.36) --
	(190.54,194.03) --
	(186.69,194.03) --
	cycle;

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 90.87, 94.39) rectangle ( 93.72, 97.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 90.87, 94.39) -- ( 93.72, 97.24);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 90.87, 97.24) -- ( 93.72, 94.39);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.44, 95.57) rectangle ( 95.30, 98.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.44, 95.57) -- ( 95.30, 98.43);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] ( 92.44, 98.43) -- ( 95.30, 95.57);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.68,150.84) rectangle (141.54,153.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.68,150.84) -- (141.54,153.70);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (138.68,153.70) -- (141.54,150.84);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.70,119.10) rectangle (113.56,121.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.70,119.10) -- (113.56,121.95);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (110.70,121.95) -- (113.56,119.10);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.56,161.49) rectangle (121.41,164.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.56,161.49) -- (121.41,164.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (118.56,164.34) -- (121.41,161.49);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.28,115.20) rectangle (122.14,118.05);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.28,115.20) -- (122.14,118.05);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (119.28,118.05) -- (122.14,115.20);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.10,122.49) rectangle (129.96,125.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.10,122.49) -- (129.96,125.34);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (127.10,125.34) -- (129.96,122.49);
\definecolor{drawColor}{RGB}{152,152,152}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.51, 47.03) -- (195.67,202.19);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.66, 47.03) -- ( 41.66, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 43.44, 47.03) -- ( 43.44, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 44.98, 47.03) -- ( 44.98, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 46.35, 47.03) -- ( 46.35, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 47.56, 47.03) -- ( 47.56, 55.57);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 55.57, 47.03) -- ( 55.57, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 60.26, 47.03) -- ( 60.26, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 63.58, 47.03) -- ( 63.58, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 66.16, 47.03) -- ( 66.16, 52.72);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 68.27, 47.03) -- ( 68.27, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 70.05, 47.03) -- ( 70.05, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 71.59, 47.03) -- ( 71.59, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 72.95, 47.03) -- ( 72.95, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 74.17, 47.03) -- ( 74.17, 55.57);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 82.18, 47.03) -- ( 82.18, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 86.87, 47.03) -- ( 86.87, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 90.19, 47.03) -- ( 90.19, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 92.77, 47.03) -- ( 92.77, 52.72);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 94.88, 47.03) -- ( 94.88, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 96.66, 47.03) -- ( 96.66, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 98.20, 47.03) -- ( 98.20, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 99.56, 47.03) -- ( 99.56, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (100.78, 47.03) -- (100.78, 55.57);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (108.79, 47.03) -- (108.79, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (113.48, 47.03) -- (113.48, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (116.80, 47.03) -- (116.80, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (119.38, 47.03) -- (119.38, 52.72);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (121.49, 47.03) -- (121.49, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (123.27, 47.03) -- (123.27, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (124.81, 47.03) -- (124.81, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (126.17, 47.03) -- (126.17, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (127.39, 47.03) -- (127.39, 55.57);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (135.40, 47.03) -- (135.40, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (140.08, 47.03) -- (140.08, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (143.41, 47.03) -- (143.41, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (145.99, 47.03) -- (145.99, 52.72);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (148.09, 47.03) -- (148.09, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (149.87, 47.03) -- (149.87, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (151.42, 47.03) -- (151.42, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (152.78, 47.03) -- (152.78, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (154.00, 47.03) -- (154.00, 55.57);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (162.01, 47.03) -- (162.01, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (166.69, 47.03) -- (166.69, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (170.02, 47.03) -- (170.02, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (172.60, 47.03) -- (172.60, 52.72);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (174.70, 47.03) -- (174.70, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (176.48, 47.03) -- (176.48, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (178.03, 47.03) -- (178.03, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (179.39, 47.03) -- (179.39, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (180.61, 47.03) -- (180.61, 55.57);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (188.62, 47.03) -- (188.62, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (193.30, 47.03) -- (193.30, 49.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 48.18) -- ( 43.36, 48.18);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 49.96) -- ( 43.36, 49.96);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 51.51) -- ( 43.36, 51.51);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 52.87) -- ( 43.36, 52.87);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 54.09) -- ( 49.05, 54.09);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 62.10) -- ( 43.36, 62.10);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 66.78) -- ( 43.36, 66.78);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 70.11) -- ( 43.36, 70.11);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 72.68) -- ( 46.20, 72.68);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 74.79) -- ( 43.36, 74.79);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 76.57) -- ( 43.36, 76.57);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 78.12) -- ( 43.36, 78.12);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 79.48) -- ( 43.36, 79.48);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 80.69) -- ( 49.05, 80.69);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 88.70) -- ( 43.36, 88.70);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 93.39) -- ( 43.36, 93.39);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 96.71) -- ( 43.36, 96.71);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 99.29) -- ( 46.20, 99.29);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,101.40) -- ( 43.36,101.40);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,103.18) -- ( 43.36,103.18);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,104.72) -- ( 43.36,104.72);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,106.08) -- ( 43.36,106.08);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,107.30) -- ( 49.05,107.30);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,115.31) -- ( 43.36,115.31);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,120.00) -- ( 43.36,120.00);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,123.32) -- ( 43.36,123.32);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,125.90) -- ( 46.20,125.90);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,128.01) -- ( 43.36,128.01);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,129.79) -- ( 43.36,129.79);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,131.33) -- ( 43.36,131.33);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,132.69) -- ( 43.36,132.69);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,133.91) -- ( 49.05,133.91);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,141.92) -- ( 43.36,141.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,146.61) -- ( 43.36,146.61);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,149.93) -- ( 43.36,149.93);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,152.51) -- ( 46.20,152.51);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,154.62) -- ( 43.36,154.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,156.40) -- ( 43.36,156.40);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,157.94) -- ( 43.36,157.94);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,159.30) -- ( 43.36,159.30);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,160.52) -- ( 49.05,160.52);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,168.53) -- ( 43.36,168.53);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,173.21) -- ( 43.36,173.21);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,176.54) -- ( 43.36,176.54);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,179.12) -- ( 46.20,179.12);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,181.22) -- ( 43.36,181.22);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,183.01) -- ( 43.36,183.01);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,184.55) -- ( 43.36,184.55);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,185.91) -- ( 43.36,185.91);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,187.13) -- ( 49.05,187.13);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,195.14) -- ( 43.36,195.14);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51,199.82) -- ( 43.36,199.82);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 47.03) rectangle (195.67,202.19);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.56, 77.66) {0.1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.56,130.88) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.56,184.10) {1000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 37.76, 80.69) --
	( 40.51, 80.69);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 37.76,133.91) --
	( 40.51,133.91);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 37.76,187.13) --
	( 40.51,187.13);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 74.17, 44.28) --
	( 74.17, 47.03);

\path[draw=drawColor,line width= 0.3pt,line join=round] (127.39, 44.28) --
	(127.39, 47.03);

\path[draw=drawColor,line width= 0.3pt,line join=round] (180.61, 44.28) --
	(180.61, 47.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 74.17, 36.02) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (127.39, 36.02) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (180.61, 36.02) {1000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (118.09, 23.99) {\textsf{DPMC} + \texttt{bklm16} time (s)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,124.61) {\textsf{DPMC} + \texttt{bklm16++} time (s)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (206.67, 68.14) rectangle (298.03,181.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (212.17,145.91) rectangle (226.62,160.37);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor,fill opacity=0.50] (219.39,153.14) circle (  1.43);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (212.17,131.46) rectangle (226.62,145.91);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor,fill opacity=0.50] (219.39,140.90) --
	(221.32,137.58) --
	(217.47,137.58) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (212.17,117.00) rectangle (226.62,131.46);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor,fill opacity=0.50] (217.97,122.80) --
	(220.82,122.80) --
	(220.82,125.66) --
	(217.97,125.66) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (212.17,102.55) rectangle (226.62,117.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (217.38,109.78) -- (221.41,109.78);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (219.39,107.76) -- (219.39,111.79);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (212.17, 88.10) rectangle (226.62,102.55);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{102,166,30}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (217.97, 93.90) rectangle (220.82, 96.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (217.97, 93.90) -- (220.82, 96.75);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (217.97, 96.75) -- (220.82, 93.90);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (212.17, 73.64) rectangle (226.62, 88.10);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{230,171,2}

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (217.97, 79.44) -- (220.82, 82.30);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (217.97, 82.30) -- (220.82, 79.44);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (217.38, 80.87) -- (221.41, 80.87);

\path[draw=drawColor,draw opacity=0.50,line width= 0.4pt,line join=round,line cap=round] (219.39, 78.85) -- (219.39, 82.89);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (232.12,150.11) {DQMR};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (232.12,135.65) {Grid};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (232.12,121.20) {Mastermind};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (232.12,106.75) {Non-binary};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (232.12, 92.29) {Other binary};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (232.12, 77.84) {Random Blocks};
\end{scope}
\end{tikzpicture}