### 2021-06-14 doc/SAT_long_talk/box.tex
Some TikZ

% Created by tikzDevice version 0.12.3.1 on 2021-06-14 12:12:55
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (303.53,224.04);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (303.53,224.04);
\end{scope}
\begin{scope}
\path[clip] ( 40.51, 33.38) rectangle (298.03,218.54);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 40.51, 33.38) rectangle (298.03,218.54);
\definecolor{drawColor}{gray}{0.87}

\path[draw=drawColor,line width= 0.1pt,line join=round] ( 40.51, 73.27) --
	(298.03, 73.27);

\path[draw=drawColor,line width= 0.1pt,line join=round] ( 40.51,136.86) --
	(298.03,136.86);

\path[draw=drawColor,line width= 0.1pt,line join=round] ( 40.51,200.45) --
	(298.03,200.45);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 40.51, 41.48) --
	(298.03, 41.48);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 40.51,105.07) --
	(298.03,105.07);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 40.51,168.66) --
	(298.03,168.66);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 70.22, 33.38) --
	( 70.22,218.54);

\path[draw=drawColor,line width= 0.3pt,line join=round] (119.75, 33.38) --
	(119.75,218.54);

\path[draw=drawColor,line width= 0.3pt,line join=round] (169.27, 33.38) --
	(169.27,218.54);

\path[draw=drawColor,line width= 0.3pt,line join=round] (218.80, 33.38) --
	(218.80,218.54);

\path[draw=drawColor,line width= 0.3pt,line join=round] (268.32, 33.38) --
	(268.32,218.54);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 60.94,178.58) -- ( 60.94,327.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 60.94, 52.60) -- ( 60.94, 42.30);
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 52.58,178.58) --
	( 52.58, 52.60) --
	( 69.30, 52.60) --
	( 69.30,178.58) --
	( 52.58,178.58) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 52.58, 89.33) -- ( 69.30, 89.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 79.51, 92.86) -- ( 79.51,158.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 79.51, 45.93) -- ( 79.51, 41.79);
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 71.15, 92.86) --
	( 71.15, 45.93) --
	( 87.87, 45.93) --
	( 87.87, 92.86) --
	( 71.15, 92.86) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 71.15, 53.43) -- ( 87.87, 53.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (110.46,185.67) -- (110.46,366.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (110.46, 54.83) -- (110.46, 42.49);
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (102.11,185.67) --
	(102.11, 54.83) --
	(118.82, 54.83) --
	(118.82,185.67) --
	(102.11,185.67) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (102.11, 92.03) -- (118.82, 92.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (129.03,160.26) -- (129.03,292.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (129.03, 50.38) -- (129.03, 42.21);
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (120.68,160.26) --
	(120.68, 50.38) --
	(137.39, 50.38) --
	(137.39,160.26) --
	(120.68,160.26) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (120.68, 66.91) -- (137.39, 66.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.99,185.77) -- (159.99,366.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.99, 54.83) -- (159.99, 42.49);
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (151.63,185.77) --
	(151.63, 54.83) --
	(168.34, 54.83) --
	(168.34,185.77) --
	(151.63,185.77) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (151.63, 92.03) -- (168.34, 92.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (178.56,160.26) -- (178.56,292.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (178.56, 50.38) -- (178.56, 42.21);
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (170.20,160.26) --
	(170.20, 50.38) --
	(186.91, 50.38) --
	(186.91,160.26) --
	(170.20,160.26) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (170.20, 66.91) -- (186.91, 66.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.51,164.59) -- (209.51,306.11);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.51, 50.38) -- (209.51, 42.18);
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (201.15,164.59) --
	(201.15, 50.38) --
	(217.87, 50.38) --
	(217.87,164.59) --
	(201.15,164.59) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (201.15, 75.53) -- (217.87, 75.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.08, 99.15) -- (228.08,158.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.08, 45.93) -- (228.08, 41.79);
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (219.72, 99.15) --
	(219.72, 45.93) --
	(236.44, 45.93) --
	(236.44, 99.15) --
	(219.72, 99.15) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (219.72, 54.19) -- (236.44, 54.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] (268.32,132.06) -- (268.32,244.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (268.32, 48.15) -- (268.32, 41.98);
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (251.61,132.06) --
	(251.61, 48.15) --
	(285.03, 48.15) --
	(285.03,132.06) --
	(251.61,132.06) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (251.61, 69.27) -- (285.03, 69.27);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.51, 33.38) rectangle (298.03,218.54);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.56, 38.45) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.56,102.04) {2000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.56,165.63) {4000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 37.76, 41.48) --
	( 40.51, 41.48);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 37.76,105.07) --
	( 40.51,105.07);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 37.76,168.66) --
	( 40.51,168.66);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 70.22, 30.63) --
	( 70.22, 33.38);

\path[draw=drawColor,line width= 0.3pt,line join=round] (119.75, 30.63) --
	(119.75, 33.38);

\path[draw=drawColor,line width= 0.3pt,line join=round] (169.27, 30.63) --
	(169.27, 33.38);

\path[draw=drawColor,line width= 0.3pt,line join=round] (218.80, 30.63) --
	(218.80, 33.38);

\path[draw=drawColor,line width= 0.3pt,line join=round] (268.32, 30.63) --
	(268.32, 33.38);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 70.22, 22.37) {\texttt{bklm16}};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (119.75, 22.37) {\texttt{cd05}};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (169.27, 22.37) {\texttt{cd06}};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (218.80, 22.37) {\texttt{d02}};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (268.32, 22.37) {\texttt{sbk05}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,125.96) {Variables};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (123.14,  5.50) rectangle (215.40, -2.81);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (128.64,  5.50) rectangle (143.10, 19.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (135.87,  6.95) --
	(135.87,  9.11);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (135.87, 16.34) --
	(135.87, 18.51);
\definecolor{fillColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (130.45,  9.11) rectangle (141.29, 16.34);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (130.45, 12.73) --
	(141.29, 12.73);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (177.58,  5.50) rectangle (192.04, 19.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (184.81,  6.95) --
	(184.81,  9.11);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (184.81, 16.34) --
	(184.81, 18.51);
\definecolor{fillColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (179.39,  9.11) rectangle (190.23, 16.34);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (179.39, 12.73) --
	(190.23, 12.73);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (148.60,  9.70) {before};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (303.53,224.04);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (197.54,  9.70) {after};
\end{scope}
\end{tikzpicture}

\end{document}