### 2021-06-19 doc/SAT_long_talk/talk.tex
Update some details

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{tikz}
\usepackage{listings}
\usepackage{complexity}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{xcolor}

\beamertemplatenavigationsymbolsempty
\usetheme{boxes}
\usecolortheme{rose}

\usetikzlibrary{arrows}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{shapes}
\usetikzlibrary{positioning}
\usetikzlibrary{tikzmark}
\usetikzmarklibrary{listings}

\tikzstyle{every picture}+=[remember picture]

\newcounter{tmkcount}
\tikzset{
  use tikzmark/.style={
    remember picture,
    overlay,
    execute at end picture={
      \stepcounter{tmkcount}
    },
  },
  tikzmark suffix={-\thetmkcount}
}

\definecolor{color1}{HTML}{1b9e77}
\def\speciallstcolor{\begingroup\color{color1}}
\def\endspeciallstcolor{\endgroup}

\definecolor{color2}{HTML}{d95f02}
\def\speciallstcolortwo{\begingroup\color{color2}}
\def\endspeciallstcolortwo{\endgroup}

\definecolor{color3}{HTML}{7570b3}
\def\speciallstcolorthree{\begingroup\color{color3}}
\def\endspeciallstcolorthree{\endgroup}

\definecolor{color4}{HTML}{e7298a}
\def\speciallstcolorfour{\begingroup\color{color4}}
\def\endspeciallstcolorfour{\endgroup}

\definecolor{color5}{HTML}{66a61e}
\def\speciallstcolorfive{\begingroup\color{color5}}
\def\endspeciallstcolorfive{\endgroup}

\definecolor{color6}{HTML}{e6ab02}
\def\speciallstcolorsix{\begingroup\color{color6}}
\def\endspeciallstcolorsix{\endgroup}

\definecolor{color7}{HTML}{a6761d}
\def\speciallstcolorseven{\begingroup\color{color7}}
\def\endspeciallstcolorseven{\endgroup}

\author[P. Dilkas, V. Belle]{\textbf{Paulius Dilkas} \and Vaishak Belle}
\title{Weighted Model Counting Without Parameter Variables}
\date{SAT 2021}
\institute[University of Edinburgh]{University of Edinburgh, Edinburgh, UK}

\AtBeginSection[]
{
  \begin{frame}
    \frametitle{Outline}
    \tableofcontents[currentsection]
  \end{frame}
}

\begin{document}

\begin{frame}
  \tikz[remember picture,overlay]{
    \node at ([yshift=20pt,xshift=30pt]current page.south)
    {\includegraphics[height=40pt]{../poster/logo_inf.png}};
    \node at ([yshift=20pt,xshift=75pt]current page.south)
    {\includegraphics[height=40pt]{../poster/logo_ecr.png}};
    \node at ([yshift=15pt,xshift=140pt]current page.south)
    {\includegraphics[height=20pt]{../poster/logo_ukri.png}};
  }
  \titlepage
\end{frame}

\begin{frame}[fragile]{The Computational Problem of Probabilistic Inference}
  \vspace{-0.75cm}
  \begin{columns}[t]
    \begin{column}{0.65\textwidth}
      \centering
      \begin{block}{ProbLog}
        \vspace{-0.2cm}
        \begin{lstlisting}[basicstyle=\tiny]
0.001 :: burglary.
0.002 :: earthquake.
0.95  :: alarm     :- burglary, earthquake.
0.94  :: alarm     :- burglary, \+ earthquake.
0.29  :: alarm     :- \+ burglary, earthquake.
0.001 :: alarm     :- \+ burglary, \+ earthquake.
0.9   :: johnCalls :- alarm.
0.05  :: johnCalls :- \+ alarm.
0.7   :: maryCalls :- alarm.
0.01  :: maryCalls :- \+ alarm.
        \end{lstlisting}
        \vspace{-0.2cm}
      \end{block}
      \begin{block}{BLOG}
        \vspace{-0.2cm}
        \begin{lstlisting}[escapeinside={(*}{*)},basicstyle=\tiny]
random Boolean Burglary (*$\sim$*) BooleanDistrib(0.001);
random Boolean Earthquake (*$\sim$*) BooleanDistrib(0.002);
random Boolean Alarm (*$\sim$*)
  if Burglary then
    if Earthquake then BooleanDistrib(0.95)
    else BooleanDistrib(0.94)
  else
    if Earthquake then BooleanDistrib(0.29)
    else BooleanDistrib(0.001);
random Boolean JohnCalls (*$\sim$*)
  if Alarm then BooleanDistrib(0.9)
  else BooleanDistrib(0.05);
random Boolean MaryCalls (*$\sim$*)
  if Alarm then BooleanDistrib(0.7)
  else BooleanDistrib(0.01);
        \end{lstlisting}
        \vspace{-0.2cm}
      \end{block}
    \end{column}
    \begin{column}{0.35\textwidth}
      \begin{block}{Bayesian Network}
        \centering
        \begin{tikzpicture}[node distance=2cm,scale=0.5,every node/.style={scale=0.5}]
          \node[draw,ellipse] (alarm) {Alarm};
          \node[draw,ellipse,above left of=alarm] (burglary) {Burglary};
          \node[draw,ellipse,above right of=alarm] (earthquake) {Earthquake};
          \node[draw,ellipse,below left of=alarm] (johnCalls) {JohnCalls};
          \node[draw,ellipse,below right of=alarm] (maryCalls) {MaryCalls};
          \draw[-Latex] (burglary) -- (alarm);
          \draw[-Latex] (earthquake) -- (alarm);
          \draw[-Latex] (alarm) -- (johnCalls);
          \draw[-Latex] (alarm) -- (maryCalls);
        \end{tikzpicture}
      \end{block}
      \vspace{1cm}
      \begin{block}{Markov Random Field}
        \centering
        \begin{tikzpicture}[node distance=2cm,scale=0.5,every node/.style={scale=0.5}]
          \node[draw,ellipse] (alarm) {Alarm};
          \node[draw,ellipse,above left of=alarm] (burglary) {Burglary};
          \node[draw,ellipse,above right of=alarm] (earthquake) {Earthquake};
          \node[draw,ellipse,below left of=alarm] (johnCalls) {JohnCalls};
          \node[draw,ellipse,below right of=alarm] (maryCalls) {MaryCalls};
          \draw (burglary) -- (earthquake);
          \draw (burglary) -- (alarm);
          \draw (earthquake) -- (alarm);
          \draw (alarm) -- (johnCalls);
          \draw (alarm) -- (maryCalls);
        \end{tikzpicture}
      \end{block}
    \end{column}
  \end{columns}
  \onslide<2>{
    \begin{tikzpicture}[remember picture,overlay]
      \node[draw,star,fill=red!10] (wmc) at (current page.center) {WMC};
      \coordinate[xshift=-0.25\linewidth,yshift=-0.25\textheight] (p1) at (current page.center);
      \coordinate[xshift=0.25\linewidth,yshift=-0.25\textheight] (p2) at (current page.center);
      \coordinate[xshift=-0.25\linewidth,yshift=0.25\textheight] (p3) at (current page.center);
      \coordinate[xshift=0.25\linewidth,yshift=0.25\textheight] (p4) at (current page.center);
      \draw[-latex,line width=2pt,color=red!50] (p1) -- (wmc);
      \draw[-latex,line width=2pt,color=red!50] (p2) -- (wmc);
      \draw[-latex,line width=2pt,color=red!50] (p3) -- (wmc);
      \draw[-latex,line width=2pt,color=red!50] (p4) -- (wmc);
    \end{tikzpicture}
  }
\end{frame}

\begin{frame}[fragile]{Weighted Model Counting (WMC)}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item Generalises propositional model counting ($\#\SAT{}$)
      \item Applications:
        \begin{itemize}
        \item graphical models
        \item probabilistic programming
        \item neural-symbolic artificial intelligence
        \end{itemize}
      \item Main types of algorithms:
        \begin{itemize}
        \item using knowledge compilation
        \item using a \SAT{} solver
        \item manipulating pseudo-Boolean functions
        \end{itemize}
      \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{example}
      $w(x) = 0.3$, $w(\neg x) = 0.7$, $w(y) = 0.2$, $w(\neg y) = 0.8$
      \vspace{1cm}

      $\mathsf{WMC}(\alert{x \lor y}) = w(x)w(y) + w(x)w(\neg y) + w(\neg x)w(y)
      = 0.44$
      \end{example}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[fragile]{The Problem with Assigning Weights to Literals}
  \begin{columns}[t]
    \begin{column}{0.5\textwidth}
      \begin{block}{A Simple Bayesian Network}
        \centering
        \begin{tikzpicture}[edge from parent/.style={draw,-{Latex}}]
          \node[draw,circle] at (0, 0) (a) {$A$}
          child {node[draw,circle] (b) {$B$}};
        \end{tikzpicture}
      \end{block}
      \begin{itemize}
      \item from \structure{2 binary} variables
      \item to \structure{8} variables and \structure{17} clauses
      \item with lots of redundancy
      \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{block}{Its WMC Encoding}
        \vspace{-0.2cm}
\begin{lstlisting}[basicstyle=\scriptsize,escapeinside={(*}{*)},name=codeL]
p cnf 8 17
(*\only<2->{\aftergroup\speciallstcolor}*)-2 -1 0(*\only<2->{\aftergroup\endspeciallstcolor}*)
(*\only<2->{\aftergroup\speciallstcolor}*)1 2 0(*\only<2->{\aftergroup\endspeciallstcolor}*)
(*\only<2->{\aftergroup\speciallstcolortwo}*)-3 1 0(*\only<2->{\aftergroup\endspeciallstcolortwo}*)
(*\only<2->{\aftergroup\speciallstcolortwo}*)-1 3 0(*\only<2->{\aftergroup\endspeciallstcolortwo}*)
(*\only<2->{\aftergroup\speciallstcolorthree}*)-5 -1 0(*\only<2->{\aftergroup\endspeciallstcolorthree}*)
(*\only<2->{\aftergroup\speciallstcolorthree}*)-5 -4 0(*\only<2->{\aftergroup\endspeciallstcolorthree}*)
(*\only<2->{\aftergroup\speciallstcolorthree}*)1 4 5 0(*\only<2->{\aftergroup\endspeciallstcolorthree}*)
(*\only<2->{\aftergroup\speciallstcolorfour}*)-6 -1 0(*\only<2->{\aftergroup\endspeciallstcolorfour}*)
(*\only<2->{\aftergroup\speciallstcolorfour}*)-6 4 0(*\only<2->{\aftergroup\endspeciallstcolorfour}*)
(*\only<2->{\aftergroup\speciallstcolorfour}*)-4 1 6 0(*\only<2->{\aftergroup\endspeciallstcolorfour}*)
(*\only<2->{\aftergroup\speciallstcolorfive}*)-7 1 0(*\only<2->{\aftergroup\endspeciallstcolorfive}*)
(*\only<2->{\aftergroup\speciallstcolorfive}*)-7 -4 0(*\only<2->{\aftergroup\endspeciallstcolorfive}*)
(*\only<2->{\aftergroup\speciallstcolorfive}*)-1 4 7 0(*\only<2->{\aftergroup\endspeciallstcolorfive}*)
(*\only<2->{\aftergroup\speciallstcolorsix}*)-8 1 0(*\only<2->{\aftergroup\endspeciallstcolorsix}*)
(*\only<2->{\aftergroup\speciallstcolorsix}*)-8 4 0(*\only<2->{\aftergroup\endspeciallstcolorsix}*)
(*\only<2->{\aftergroup\speciallstcolorsix}*)-4 -1 8 0(*\only<2->{\aftergroup\endspeciallstcolorsix}*)
(*\only<2->{\aftergroup\speciallstcolorseven}*)-4 0(*\only<2->{\aftergroup\endspeciallstcolorseven}*)
c weights 1.0 1.0 0.5 1.0 \
0.5 1.0 1.0 1.0 0.6 1.0 \
0.4 1.0 0.1 1.0 0.9 1.0
\end{lstlisting}
        \vspace{-0.2cm}
      \end{block}
    \end{column}
  \end{columns}
  \only<2->{
    \begin{tikzpicture}[overlay,use tikzmark]
      \node[above right = 0 and 15 mm of pic cs:line-codeL-3-end] {\color{color1}{$\neg x_1 \Leftrightarrow x_2$}};
      \node[above right = 0 and 15 mm of pic cs:line-codeL-5-end] {\color{color2}{$x_1 \Leftrightarrow x_3$}};
      \node[above right = 0.5 ex and 10 mm of pic cs:line-codeL-8-end] {\color{color3}{$\neg x_1 \land \neg x_4 \Leftrightarrow x_5$}};
      \node[above right = 0.5 ex and 10 mm of pic cs:line-codeL-11-end] {\color{color4}{$\neg x_1 \land x_4 \Leftrightarrow x_6$}};
      \node[above right = 0.5 ex and 10 mm of pic cs:line-codeL-14-end] {\color{color5}{$x_1 \land \neg x_4 \Leftrightarrow x_7$}};
      \node[above right = 0.5 ex and 10 mm of pic cs:line-codeL-17-end] {\color{color6}{$x_1 \land x_4 \Leftrightarrow x_8$}};
      \node[above right = -0.5 ex and 20 mm of pic cs:line-codeL-18-end] {\color{color7}{$\neg x_4$}};
    \end{tikzpicture}
  }
\end{frame}

\begin{frame}
  \frametitle{Outline}
  \tableofcontents
\end{frame}

\section{A More Expressive Alternative}

\begin{frame}{WMC, Formally}
  \begin{definition}
    A \alert{WMC instance} is a tuple \structure{$(\phi, X_I, X_P, w)$}, where
    \begin{itemize}
    \item \structure{$X_I$} is the set of \alert{indicator variables},
    \item \structure{$X_P$} is the set of \alert{parameter variables} (with
      \structure{$X_I \cap X_P = \emptyset$}),
    \item \structure{$\phi$} is a propositional formula in CNF over
      \structure{$X_I \cup X_P$},
    \item \structure{$w\colon X_I \cup X_P \cup \{\neg x \mid x \in X_I \cup
        X_P\} \to \mathbb{R}$} is a \alert{weight function}
      \begin{itemize}
      \item such that \structure{$w(x) = w(\neg x) = 1$} for all \structure{$x
          \in X_I$}.
      \end{itemize}
    \end{itemize}
  \end{definition}
  % \begin{definition}
  %   Let \structure{$\phi$} be a formula over a set of variables \structure{$X$}.
  %   Then \structure{$Y \subseteq X$} is a \alert{minimum-cardinality model} of
  %   \structure{$\phi$} if
  %   \begin{itemize}
  %   \item \structure{$Y \models \phi$},
  %   \item and \structure{$|Y| \le |Z|$} for all \structure{$Z \models \phi$}.
  %   \end{itemize}
  % \end{definition}
\end{frame}

% \begin{frame}{WMC and Minimum-Cardinality WMC}
%   The goal of \alert{WMC} is to compute
%   \[
%     \sum_{Y \models \phi} \prod_{Y \models l} w(l)
%   \]
%   whereas the goal of \alert{minimum-cardinality WMC} is to compute
%   \[
%     \sum_{Y \models \phi\text{, }\alert{|Y| = k}} \prod_{Y \models l} w(l),
%   \]
%   where
%   \[
%     k = \min_{Y \models \phi} |Y|.
%   \]
% \end{frame}

\begin{frame}{A More Expressive Alternative}
  \begin{definition}[Pseudo-Boolean Projection (PBP)]
    A \alert{PBP instance} is a tuple \structure{$(F, X, \omega)$}, where
    \structure{$X$} is the set of variables, \structure{$F$} is a set of
    two-valued pseudo-Boolean functions \structure{$2^X \to \mathbb{R}$}, and
    \structure{$\omega \in \mathbb{R}$} is the scaling factor.
  \end{definition}

  For any propositional formula \structure{$\phi$} over a set of variables
  \structure{$X$} and \structure{$p, q \in \mathbb{R}$}, let
  \structure{$[\phi]^p_q\colon 2^X \to \mathbb{R}$} be the pseudo-Boolean
  function defined as
  \[
    [\phi]^p_q(Y) \coloneqq
    \begin{cases}
      p & \text{if } Y \models \phi \\
      q & \text{otherwise}
    \end{cases}
  \]
  for any \structure{$Y \subseteq X$}.
\end{frame}

\begin{frame}{From WMC to PBP}
  \begin{example}
    \begin{itemize}
    \item Indicator variable: \structure{$x$}
    \item Parameter variables: \structure{$p$}, \structure{$q$}
    \item Weights: \structure{$w(p) = 0.2$}, \structure{$w(q) = 0.8$}, and \structure{$w(\neg p) = w(\neg q) = 1$}
    \end{itemize}
    \begin{center}
      \begin{tabular}{llll}
        \toprule
        WMC Clause & \onslide<2->{In CNF} & \onslide<3->{Pseudo-Boolean Function} & \\
        \midrule
        $\neg x \Rightarrow p$ & \onslide<2->{$x \lor p$} & \onslide<3->{$[\neg x]_1^{0.2}$}\tikz \coordinate (one); & \\
        $p \Rightarrow \neg x$ & \onslide<2->{$\neg x \lor \neg p$} & & \tikz \coordinate (two); \onslide<4->{$[x]^{0.8}_{0.2}$} \\
        $x \Rightarrow q$ & \onslide<2->{$\neg x \lor q$} & \onslide<3->{$[x]_1^{0.8}$} \tikz \coordinate (three); & \\
        $q \Rightarrow x$ & \onslide<2->{$x \lor \neg q$} & & \\
        $\neg x$ & \onslide<2->{$\neg x$} & \onslide<3->{$[\neg x]_0^1$} & \onslide<4->{$[\neg x]_0^1$} \\
        \bottomrule
      \end{tabular}
    \end{center}
  \end{example}
  \onslide<4->{
    \begin{tikzpicture}[remember picture,overlay]
      \draw[-Latex,blue] (one) -- (two);
      \draw[-Latex,blue] (three) -- (two);
    \end{tikzpicture}
  }
\end{frame}

\section{When Does This Transformation Work?}

\begin{frame}{Correctness Conditions (1/2)}
  For each parameter variable \structure{$p \in X_P$},
  \begin{itemize}
  \item \structure{$w(\neg p) = 1$},
  \item and the set of clauses that mention \structure{$p$} or \structure{$\neg
      p$} is
    \[
      \left\{p \lor \bigvee_{i=1}^n \neg l_i \right\} \cup \{ l_i \lor \neg p
      \mid i = 1, \dots, n \}
    \]
    for some non-empty family of \alert{indicator} literals
    \structure{$(l_i)_{i=1}^n$}.
  \end{itemize}
  In other words, \structure{$p$} is defined to be equivalent to
  \structure{$\bigwedge_{i=1}^n l_i$}.
\end{frame}

\begin{frame}{Correctness Conditions (2/2)}
  For each parameter variable \structure{$p \in X_P$},
  \begin{itemize}
  \item \structure{$w(p) + w(\neg p) = 1$},
  \item each clause has at most one parameter variable,
  \item there are no negative parameter literals,
  \item if \structure{$\{ p \} \in \phi$}, then this is the only clause that
    mentions \structure{$p$},
  \item and for any two clauses of the form \structure{$\chi \Rightarrow p$} and
    \structure{$\psi \Rightarrow p$}, \structure{$\chi \land \psi \equiv \bot$}.
  \end{itemize}
\end{frame}

% \begin{frame}{Additional Conditions for Minimum-Cardinality WMC}
%   \begin{itemize}
%   \item All models of \structure{$\{ c \in \phi \mid c \cap X_P = \emptyset \}$}
%     have the same number of positive indicator literals,
%   \item and
%     \[
%       \min_{Z \subseteq X_P} |Z| \quad \text{s.t. } \quad Y \cup Z \models \phi
%     \]
%     is the same for all \structure{$Y \models \{ c \in \phi \mid c \cap X_P =
%       \emptyset \}$}.
%   \end{itemize}
% \end{frame}

\section{How Good Is It?}

\begin{frame}{WMC/PBP Encodings for Bayesian Networks}
  \centering
  \includegraphics[width=\textwidth]{cumulative1}
\end{frame}

\begin{frame}{WMC/PBP Encodings for Bayesian Networks}
  \centering
  \includegraphics[width=\textwidth]{cumulative2}
\end{frame}

\begin{frame}{WMC/PBP Encodings for Bayesian Networks}
  \centering
  \includegraphics[width=\textwidth]{cumulative3}
\end{frame}

\begin{frame}{WMC/PBP Encodings for Bayesian Networks}
  \centering
  \includegraphics[width=\textwidth]{cumulative4}
\end{frame}

\begin{frame}{Compared to the Previous State of the Art}
  \centering
  \includegraphics[width=\textwidth]{scatter1}
\end{frame}

\begin{frame}{The Best Encoding for DPMC: Before and After}
  \centering
  \includegraphics[width=\textwidth]{scatter2}
\end{frame}

% \begin{frame}{Variable Reduction for Bayesian Network Encodings}
%   \begin{columns}
%     \begin{column}{0.4\textwidth}
%       \begin{itemize}
%       \item foo
%       \end{itemize}
%     \end{column}
%     \begin{column}{0.6\textwidth}
%       \centering
%       \includegraphics[width=\textwidth]{box}
%     \end{column}
%   \end{columns}
% \end{frame}

\section{Summary and Future Work}

\begin{frame}{Summary and Future Work}
  \begin{itemize}
  \item PBP is a more expressive alternative to WMC that works with
    state-of-the-art WMC algorithms based on pseudo-Boolean function
    manipulation.
  \item Many WMC encodings can be efficiently transformed into PBP while
    removing unnecessary variables and clauses.
  \item The identified conditions for this transformation to work help explain
    how WMC encodings for Bayesian networks operate.
  \item Performance improvements depend on the encoding.
    \begin{itemize}
    \item The very first encoding was virtually unaffected,
    \item whereas the state-of-the-art encoding was significantly improved.
    \end{itemize}
  \item Can the identified conditions be generalised further?
  \item Can the transformation be applied to WMC encodings for other application
    domains?
  \end{itemize}
\end{frame}

\end{document}